### 2023-05-04 paper/paperDraft.Rmd
Adds files without models

---
title: 'Taking uncertainty in word embedding bias estimation seriously: a  bayesian
  approach'
author: "Alicja Dobrzeniecka and Rafal Urbaniak"
output:
  pdf_document:
    number_sections: yes
    df_print: kable
    keep_tex: yes
    includes:
      in_header: Rafal_latex7.sty
  html_document:
    df_print: paged
classoption: dvipsnames,enabledeprecatedfontcommands
fontsize: 10pt
documentclass: scrartcl
urlcolor: blue
bibliography: ../references/cosineReferences1.bib
csl: ../references/transactions-on-computational-logic.csl
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')

library(ggplot2)
library(ggthemes)
library(gridExtra)
library(dplyr)
library(kableExtra)
library(tidyverse)
library(magrittr)
library(ggpubr)
library(ggExtra)
library(ggthemes)
library(ggforce)
library(latex2exp)
library(Hmisc)
library(grid)

```

# Introduction

Natural language processing (NLP) models are used to perform various language-related tasks such as providing email filters, intelligent assistants, search results,  translations, text analysis and so on. Such tools require input words to be represented by numbers. This can be done with word embeddings, which represent words as vectors of real numbers.  A common way to construct an embedding is to use  a large natural language corpus to train  a  neural network to assign vectors to words that optimize for co-occurrence prediction accuracy. The vectors can then be compared in terms of their similarity -- the usual measure is cosine similarity -- and the results of such comparisons can be used in downstream tasks. Roughly speaking, cosine similarity is an imperfect mathematical proxy for semantic similarity [@Mikolov2013efficient].

It has been suggested [@Bolukbasi2016man; @Caliskan2017semanticsBiases; @Gonen2019lipstick; @Lauscher2019multidimensional; @Garg2018years; @Manzini2019blackToCriminal]  that  such models can learn implicit biases that reflect harmful stereotypical thinking---for example,  the (vector corresponding to the) word \textit{she} might be  much closer in the vector space to  the word \textit{cooking} than  the word \textit{he}. Such phenomena may be undesirable at least  in some  downstream tasks, such as web search, recommendations, and so on. To investigate such issues, several measures of bias in word embeddings have been formulated and applied. Our goal is to use two prominent examples of such measures to argue that this approach oversimplifies the situation and to develop a Bayesian alternative.

One response to the raising of the issue of bias in natural language models might be to say that there is not much point to reflecting on such biases, as they are  unavoidable. This unavoidability might seem in line with the arguments to the effect that learning algorithms are always value-laden [@JohnsonValueFree]: they employ inductive methods that require design-, data-, or risk-related decisions that have to be guided by extra-algorithmic considerations. Such choices necessarily involve value judgments and have to do, for instance, with what simplifications or risks one finds acceptable. 
Admittedly, algorithmic decision making cannot fulfill the value-free ideal, but this only means that even more attention needs to be paid to the values underlying different techniques and decisions, and to the values being pursued in a particular use of an algorithm.

Another response might be to insist that there is no bias introduced by the use of machine learning methods here, since the algorithm is simply learning to correctly predict co-occurrences based on what "reality" looks like. However, this objection overlooks the fact that we, humans, are the ones who construct this linguistic reality, is shaped in part by the natural language processing tools we use on a massive scale. Sure, if there is unfairness and our goal is to diagnose it, we should do complete justice to learning it in the model used in a study thereof. One example of this approach is [@Garg2017hundredYears]. However, if our goal is to develop downstream tools that perform tasks that we care about well without further perpetuating  or exacerbating harmful stereotypes, we still have good reasons to try to minimize the negative impact, at least as long as this does not damage the overall performance of the tool. Moreover, it is often not the case that the corpora mirror reality---to give a trivial example, heads are spoken of more often than kidneys, but this does not mean that kidneys occur much less often in reality than heads. To give a more relevant example, the disproportionate association of female words with female occupations in a corpus actually greatly exaggerates the actual lower disproportion in the real distribution of occupations. [@gordon2012reporting].

In what follows, we focus on  two popular  measures of bias applicable to many existing word embeddings ( \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.} \emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.} and \emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}): \emph{Word Embedding Association Test}  (WEAT) [@Caliskan2017semanticsBiases],  and \emph{Mean Average Cosine Distance} (MAC)  [@Manzini2019blackToCriminal].  We first explain how these measures are supposed to work. Then we argue that they are problematic for various reasons, a key one being that by pre-averaging data they manufacture false confidence, which we illustrate in terms of simulations showing that the measures often suggest the existence of bias even if by design it is non-existent in a simulated dataset.

We  propose to replace them with a Bayesian data analysis (hierarchical models), which not only provides more modest and realistic assessment of the uncertainty involved, but also allows for inspection at various levels of granularity. Once we introduce the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of WEAT or MAC (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive. Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable. The tools that we have proposed, however, allow for a more fine-grained and multi-level evaluation of bias and debiasing in language models without abandoning modesty about the uncertainties involved.

# Two measures of bias: WEAT and MAC

Conceptually the idea is that if a particular harmful stereotype is learned in a particular embedding, then certain groups of words will be systematically closer to (or further from) each other. This gives rise to the idea of protected groups---for example, in guiding online search completion or recommendation, female words might require protection in that they should not be systematically closer to stereotypically female job names, and male words require protection in that they should not be systematically closer to toxic masculinity stereotypes, such as "tough", "never complaining" or "macho".^[However, for some research-related purposes, such as the study of stereotypes across history [@Garg2017hundredYears], embeddings which do not protect certain classes may also be useful.]

The key role in the measures to be discussed is played by the notion of cosine distance (or, symmetrically, by cosine similarity). These are defined as follows:^[Here, "$-$" stands for point-wise difference, "$\cdot$" stands for the dot product operation, and $\lVert a\rVert  = \sqrt{(a \cdot a)}$.]
\begin{align} \tag{Sim}
\mathsf{cosineSimilarity}(A,B) & = \frac{A \cdot B}{\lVert  A \rVert \,\lVert B \rVert}
\\
\tag{Distance}
\mathsf{cosineDistance}(A,B) &  = 1 - \mathsf{cosineSimilarity}(A,B).
\end{align}
Note  that this terminology is slightly misleading, as mathematically cosine distance is not a distance measure, because it does not satisfy the triangle inequality, as generally $\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)$. We will keep using this mainstream terminology.

One of the first measures of bias has been developed in [@Bolukbasi2016man].  The general idea is that a certain topic is associated with a vector of real numbers (the topic "direction"), and the bias of a word is investigated by considering the projection of its corresponding vector on this direction. For instance, in [@Bolukbasi2016man],   the gender direction $\mathsf{gd}$   is obtained by taking the differences of the vectors corresponding to ten different gendered pairs (such as $\overrightarrow{she} - \overrightarrow{he}$ or $\overrightarrow{girl} - \overrightarrow{boy}$), and then identifying their principal component.^[Roughly, the principal component  is the vector obtained by projecting the data points on their  linear combination in a way that maximizes the variance of the projections.]  The gender bias of a word $w$ is then understood as $w$'s projection on the gender direction: $\vec{w} \cdot gd$ (which, after normalizing by dividing by $\lVert  w \rVert \,\lVert \mathsf{gd} \rVert$, is the same as cosine similarity).  Given a list  $N$ of  supposedly gender neutral words,^[We follow the methodology used in the debate in assuming that there is a class of words  identified as more or less neutral, such as \emph{ballpark, eat, walk, sleep, table}, whose average similarity to the gender direction (or other protected words) is around 0. We will have more to say about this assumption when we describe our dataset construction.]  and the gender direction $\mathsf{gd}$,    the direct gender bias is defined as the average cosine similarity of the words in $N$ from $\mathsf{gd}$ ($c$ is a parameter determining how strict we want to be):
\begin{align*}
\mathsf{directBias_c(N,gd)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},\mathsf{gd})\vert^c}{\vert N \vert }
\end{align*}
\normalsize

The use of projections in bias estimation has been criticized for instance in [@Gonen2019lipstick], where it is pointed out that while a higher average similarity to the  gender direction might be an indicator of bias with respect to a given  class of words, but it is only one possible manifestation of it, and reducing the cosine similarity to such a projection may not be sufficient to eliminate bias. For instance, "math" and "delicate" might be equally similar to a pair of  opposed explicitly gendered words (\emph{she}, \emph{he}), while being closer to quite different stereotypical attribute words (such as \emph{scientific} or \emph{caring}). Further, it is observed in [@Gonen2019lipstick]  that most word pairs retain similarity under debiasing meant to minimize projection-based bias.^[In [@Bolukbasi2016man]  another method which involves analogies and their evaluations by human users on Mechanical Turk is also used. We do not discuss this method in this paper, see its  criticism in [@Nissim2020fair].]

A measure of bias in word embeddings which does not employ similarities to  pre-calculated protected-class-level directions, the Word Embedding Association Test (WEAT), has been proposed in  [@Caliskan2017semanticsBiases]. The idea here is that the   bias between two sets of target words, $X$ and $Y$ (we call them protected words), should be quantified in terms of the cosine similarity  between the protected words and attribute words coming from  two sets of stereotype attribute words, $A$ and $B$ (we will call them attributes). For instance, $X$ might be a set of male names, $Y$ a set of female names, $A$ might contain stereotypically male-related, and $B$ stereotypically female-related career words. The association difference for a particular word $t$ (belonging to either $X$ or $Y$) is:

\vspace{-2mm}

\begin{align*}
\mathsf{s}(t,A,B) & = \frac{\sum_{a\in A}\mathsf{cos}(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}\mathsf{cos}(t,b)}{\vert B\vert}
\end{align*}
\normalsize
\noindent then, the association difference between $A$ a $B$ is:

\begin{align*}
\mathsf{s}(X,Y,A,B) & = \sum_{x\in X} \mathsf{s}(x,A,B) -  \sum_{y\in Y} \mathsf{s}(y,A,B)
\end{align*}

\noindent The authors use it as a test statistic in some tests, and the assumption that $X$ and $Y$ are of the same size is necessary for the statistic to make sense. In the final measure (of effect size), WEAT, means are taken, which makes it insensitive to size differences between $X$ and $Y$:
\begin{align}
\mathsf{weat}(A,B) & = \frac{
\mu(\{\mathsf{s}(x,A,B)\}_{x\in X}) -\mu(\{\mathsf{s}(y,A,B)\}_{y\in Y}) 
}{
\sigma(\{\mathsf{s}(w,A,B)\}_{w\in X\cup Y})
}
\end{align}

 WEAT is inspired by  the Implicit Association Test (IAT) [@Nosek2002harvesting] used in psychology, and it uses almost the same word sets, allowing for a \emph{prima facie} sensible comparison with bias in humans. In  [@Caliskan2017semanticsBiases] the authors argue that significant biases---thus measured--- similar to the ones discovered by IAT can be discovered in word embeddings. In [@Lauscher2019multidimensional] the methodology is  extended to a multilingual and cross-lingual setting, arguing that using Euclidean distance instead of cosine similarity does not make much difference, while the bias effects vary greatly across embedding models (interestingly, with social media-text trained embeddings being less biased than those based on Wikipedia).
 A similar methodology is employed in [@Garg2018years]. The authors employ word embeddings trained on corpora from different decades to study the shifts in various biases.^[Strictly speaking, these authors use Euclidean distances and their differences, but the way they take averages and averages thereof is analogous, and so what we will have to say about pre-averaging leading to false confidence applies to this methodology as well.]

<!--  For instance, to compute the occupational embeddings bias for women the authors first compute the average vector of vector embeddings of words that represent women (e.g. \emph{she}, and \emph{female}), then calculate the Euclidean distance between this mean vector and words for occupations. Then they take the mean of these distances and subtract from it the analogously obtained mean for the average vector of vector embeddings of words that represent men. Formally they take the relative norm distance between $X$ and $Y$ to be: -->

<!-- \vspace{-2mm} -->

<!-- \footnotesize -->

<!-- The effect size is computed by normalizing the difference in means as follows: -->

<!-- \begin{align} -->
<!-- \textsf{relative norm distance} & = \sum_{v_m\in M} \vert \vert v_m - v_X\vert \vert_2 - \vert v_m - v_Y\vert \vert_2 -->
<!-- \end{align} -->

<!-- \normalsize  -->

<!-- \noindent where the norm used is Euclidean, and $v_X$ and $x_Y$ are average vectors for sets $X$ and $Y$ respectively.  -->
<!-- ] -->

<!-- \vspace{-2mm} -->

WEAT, however, has been developed to investigate biases corresponding to a pair of supposedly opposing stereotypes, and so the question arises as to how generalize the measure to contexts in which biases with respect to more than two stereotypical groups are to be measured. Such a generalization can be found in [@Manzini2019blackToCriminal]. The authors introduce Mean Average Cosine distance (MAC) as a measure of bias. Let $T = \{t_1, \dots, t_k\}$ be a set of protected words, and let each $A_j\in \mathcal{A}$ be a set of attributes stereotypically associated with a protected word where $\mathcal{A}$.
For instance, when biases related to religion are to be investigated, they use a dataset of the  format illustrated in Table \ref{tab:religionOriginal}. The measure is defined as follows:

```{r religionTableHeadEarly,echo=FALSE,eval=FALSE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
religion <- read.csv("datasets/outdated/religionReddit.csv")[-c(1,6,7)]
colnames(religion) <- c("protected words (T)","attributes","atrribute set (A~j)",
                        "cosine distance")

religionOriginal <- religion[!(religion$`atrtibute set (A~j)` %in% c("neutral","human")),]

set.seed(22)
religionSample <- sample_n(religionOriginal, size = 15)
religionSample %>%  kable(format = "latex",booktabs=T,
                          linesep = "",  escape = FALSE, 
                          caption = "Sample 12 rows of the religion dataset.") %>%
  kable_styling(latex_options=c("striped"))
```

\begin{table}
\footnotesize

\centering

\begin{tabular}[t]{lllr}
\toprule
protected words ($T$) & attributes & attribute set ($A_j$) & cosine distance\\
\midrule
\cellcolor{gray!15}{rabbi} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{1.0306175}\\
church & familial & christianStereotype & 0.7087424\\
\cellcolor{gray!15}{synagogue} & \cellcolor{gray!15}{liberal} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{0.7922607}\\
jew & familial & christianStereotype & 0.9783245\\
\cellcolor{gray!15}{quran} & \cellcolor{gray!15}{dirty} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{1.1207093}\\
muslim & uneducated & muslimStereotype & 0.5160429\\
\cellcolor{gray!15}{torah} & \cellcolor{gray!15}{terrorist} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{0.9341137}\\
quran & hairy & jewStereotype & 1.1764642\\
\cellcolor{gray!15}{synagogue} & \cellcolor{gray!15}{violent} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{0.9549743}\\
bible & cheap & jewStereotype & 1.2234364\\
\cellcolor{gray!15}{christianity} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{0.9728545}\\
muslim & hairy & jewStereotype & 0.8788219\\
\cellcolor{gray!15}{islam} & \cellcolor{gray!15}{critical} & \cellcolor{gray!15}{christianStereotype} & \cellcolor{gray!15}{0.7880706}\\
muslim & conservative & christianStereotype & 0.4453191\\
\cellcolor{gray!15}{mosque} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{1.1541524}\\
\bottomrule
\end{tabular}

\caption{Sample 15 rows of the religion dataset. The whole dataset has 15 unique protected words ($T$), and 11 unique attributes divided between 3 attribute sets ($A_1=\mathsf{jewStereotype}, A_2= \mathsf{christianStereotype}, A_3=\mathsf{muslimStereotype}$). $\mathcal{A}$ consists of these three sets, $\mathcal{A}= \{A_1, A_2, A_3\}$. The whole dataset has $15\times 11 = 165$ rows.}
\label{tab:religionOriginal}
\normalsize 
\end{table}

\begin{align*}
\mathsf{s}(t, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cosineDistance}(t,a) \\
\mathsf{mac}(T,\mathcal{A}) & = \frac{1}{\vert T \vert \,\vert \mathcal{A}\vert}\sum_{t \in T }\sum_{A_j \in \mathcal{A}}    \mathsf{s}(t,A_j)
\end{align*}

\noindent That is, for each  protected word $t\in T$, and each attribute set $A_j$, they first take the mean of distances for this protected word and all attributes in a given attribute class, and then take the mean of thus obtained means for all the protected words and all the protected classes.^[The authors'  code are in their github repository, [https://github.com/TManzini/DebiasMulticlassWordEmbedding](https://github.com/TManzini/DebiasMulticlassWordEmbedding).]

Having introduced the measures,  first, we will introduce a selection of general problems with this approach, and then we will move on to more specific but important problems related to the fact that the measures take  averages and averages of averages. Once this is done, we move to the development of our Bayesian alternative and the presentation of its deployment.

<!-- WEAT will be useful in our criticism of the statistical methods involved, as it is simpler, so the explanation and visualizations will be more transparent (and \emph{mutatis mutandis} this criticism will apply to MAC as well), and MAC will be useful in the development of our alternative method as its range of applicability is the widest. -->

# General methodological problems with cosine-based measures of bias

Consider the MAC results from Table REF. Are the initial MAC values lower than 1 indicative of the presence of bias? Of course, thinking abstractly, 1 is the ideal distance for unrelated words. But clearly, there is variation in distances, which might lead to non-biased lists also having MAC smaller than 1. How much smaller? While the original paper did not employ control lists for comparison in the paper. To obtain a more realistic gauge on how to understand MAC values (and in further development of our approach), we introduced control word lists. One of them is a list of words we intuitively considered neutral (see REF in the APPENDIX). Moreover, it might be the case that words that have to do with human activities are systematically closer to the protected words than merely neutral words, for which reason we also use a second list with intuitively non-stereotypical human attributes (see REF in the APPENDIX). Another important observation is that in their calculations the authors do not distinguish whether a given  attribute is associated with a given protected word (e.g. protected word \emph{jew} and the attribute \emph{greedy}) or whether it is associated with a different stereotype from the same list (e.g. protected word \emph{quran} and the attribute \emph{greedy}), simply averaging across all such groups.

In Figure  \ref{fig:empirical} We look at the empirical distributions, while paying attention to such divisions. Horizontal lines represent 1-MAC values that the authors considered indicative of bias for stereotypes corresponding to  given  word lists. For instance, in religion, MAC was .859, so we plot $0\pm (1-.859)\approx .14$ lines around similarity = 0 (that is, distance = 1). Notice that most distributions are quite wide, and the proportions of even neutral or human neutral words with similarities higher than the value of 1-MAC deserving debiasing according to the authors is quite high.

```{r cosineDistributions6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%"}

genderGoogle <- read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1]
genderReddit <-  read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1]
genderGlove <- read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1]

raceGlove <-  read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1]
raceGoogle <- read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1]
raceReddit <- read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1]

religionGlove <- read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1]
religionGoogle <- read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1]
religionReddit <- read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1]

debiasedGenderReddit <- read.csv("datasets/macWeatDatasets/debiased_gender_reddit.csv")[-1]
debiasedRaceReddit <- read.csv("datasets/macWeatDatasets/debiased_race_reddit.csv")[-1]
debiasedReligionReddit <- read.csv("datasets/macWeatDatasets/debiased_religion_reddit.csv")[-1]

violinReligionReddit <- ggplot(religionReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
  geom_violin(alpha = .5)+coord_flip()+
  geom_vline(xintercept = .14) + geom_hline(yintercept = -.14, color = "grey") +
  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
  labs(title =  "Religion (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
  geom_jitter(size = 0.5, alpha = .2, width=0.08)+
  stat_summary(fun.data = "mean_cl_boot", fun.args = list(
    conf.int = .95), geom = "pointrange",
               colour = "black", size  = .2  )+
  annotate(geom = "label", x = 4.2, y = 0, label = "56%")+
  annotate(geom = "label", x = 3.2, y = 0, label = "55%")+
  annotate(geom = "label", x = 2.2, y = 0, label = "43%")+
  annotate(geom = "label", x = 1.2, y = 0, label = "36%")+
  theme(plot.title.position = "plot",legend.position="top")

violinGenderReddit <- ggplot(genderReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
  geom_vline(xintercept = .38) + geom_hline(yintercept = -.38, color = "grey") +
  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .38, color = "grey")+
  ggtitle("Gender: empirical distribution of cosine similarity")+
  stat_summary(fun.data = "mean_cl_boot", fun.args = list(
    conf.int = .95), geom = "pointrange",
               colour = "black", size  = .2  )+
  annotate(geom = "label", x = 4.2, y = 0, label = "96%")+
  annotate(geom = "label", x = 3.2, y = 0, label = "87%")+
  annotate(geom = "label", x = 2.2, y = 0, label = "53%")+
  annotate(geom = "label", x = 1.2, y = 0, label = "47%")+
  labs(title =  "Gender (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
  theme(plot.title.position = "plot",legend.position='none')

violinRaceReddit <- ggplot(raceReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
  geom_vline(xintercept = .11) + geom_hline(yintercept = -.11, color = "grey") +
  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .11, color = "grey")+
   labs(title =  "Race (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
  stat_summary(fun.data = "mean_cl_boot", fun.args = list(
    conf.int = .95), geom = "pointrange",
               colour = "black", size  = .2  )+
  annotate(geom = "label", x = 4.2, y = 0, label = "46%")+
  annotate(geom = "label", x = 3.2, y = 0, label = "67%")+
  annotate(geom = "label", x = 2.2, y = 0, label = "39%")+
  annotate(geom = "label", x = 1.2, y = 0, label = "33%")+
  theme(plot.title.position = "plot",legend.position='none')

```
\pagebreak

\begin{figure}[H]
```{r fig:cosineDistributions8b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", fig.height = 16}

grid.arrange(violinReligionReddit, violinGenderReddit, violinRaceReddit, ncol = 1)

```

\caption{Empirical distributions of cosine similarities (1-distances) for three word lists used in  the original paper.  }

\label{fig:empirical}
\end{figure}

One issue to consider is  the selection of attributes for bias measurement. The word lists used in the literature are often fairly small (5-50).

While the papers do employ statistical tests to measure the uncertainty involved, we will later on argue that these method are not  proper for the goal at hand and  show that a more appropriate use of statistical methods leads to  estimates of uncertainty that are rather epistemologically pessimistic.  \todo{sensitivity to choice of words}

Let's think about MAC, using the  case of religion-related stereotypes. The full lists from [@Manzini2019blackToCriminal] can be found in \ref{appendix:manzini_word_lists}. In the original paper,  words from all three religions were compared against all of the stereotypes. One reason this is problematic is that  no distinction between cases in which the stereotype is associated with a given religion, as opposed to the situation in which it is associated with another one, is made. This is problematic, as  not all of the stereotypical words have to be considered as harmful for all of the religions. One should investigate the religions separately as some of them may have stronger harmful associations that others.

The interpretation of the results is also a challenge. In [@Manzini2019blackToCriminal] we can find  summaries of  average cosine distances per group (such as gender, race, or religion). For instance, for religion, here is the relevant fragment of table:

<!-- Religion Debiasing  | MAC | p-value
------------- | ------------- | -------
Biased        | 0.859 | N/A
Hard Debiased | 0.934 | 3.006e-07
Soft Debiased ($\lambda$ = 0.2) | 0.894 | 0.007 -->

\begin{table}
\footnotesize

\centering

\begin{tabular}[t]{lllr}
\toprule
Religion Debiasing & MAC & a p-value \\
\midrule
\cellcolor{gray!15}{Biased} & \cellcolor{gray!15}{0.859} & \cellcolor{gray!15}{N/A} \\
Hard Debiased & 0.934 & 3.006e-07\\
\cellcolor{gray!15}{Soft Debiased ($\lambda$ = 0.2)} & \cellcolor{gray!15}{0.894} & \cellcolor{gray!15}{0.007} \\
\bottomrule
\end{tabular}

\caption{The associated mean average cosine similarity
(MAC) and p-values for debiasing methods for religious bias}
\label{tab:religionOriginal2}
\normalsize 
\end{table}

\noindent (MAC stand for mean average cosine similarity, although in reality the the table contains mean cosine distances). What may attract attention is the fact that the value of cosine distance in "Biased" category is already quite high (i.e. close to 1) even before debiasing. High cosine distance indicates low cosine similarity between values. One could think that average cosine similarity equal to approximately 0.141 is not significant enough to consider it as bias. However,  the authors still aim to mitigate such "bias"  to make the distance even larger. Methodologically the question is, on what basis is this small similarity still considered as a proof of the presence of bias, and whether these small changes are meaningful.

The underlying problem here is that in the paper there is no  control group.  One should also include control groups to have a way of comparing the  results for a supposedly stereotyped group with the results for sets of neutral or human-related neutral words. In our approach later on, we distinguish between stereotypes associated with a given group, stereotypes associated with different groups, and introduce control groups: neutral words and stereotype-free human predicates.

<!-- One may consider the issue of the curse of dimensionality. In our case, the curse of dimensionality may take place when there is an increase in the volume of data that results in adding extra dimensions to the Euclidean space. As the number of features increases, it may be harder and harder to obtain useful information from the data using the available algorithms. Using cosine similarity in high dimensions in word embeddings may be prone to the curse of dimensionality. According to @Venkat2018Curse there are reasons to consider this phenomenon when searching for word similarities in higher dimensions. An experiment is conducted that aims at showing how the similarity values and variation change as the number of dimensions increases. The hypothesis made in the paper states that two things will happen as the number of dimensions increase. First, the effort required to measure cosine similarity will be greater, and two, the similarity between data will blur out and have less variation. The authors generate random points with increasing number of dimensions where each dimension of a data point is given a value between 0 and 1. Then they pick one vector at random from each dimension class and calculate the cosine similarity between the chosen vector and the rest of the data. Then they check how the variation of values changes as the number of dimensions increases. It seems like the more dimensions there are, the smaller the variance and therefore it is less obvious how to interpret the resulting cosine similarities. Maybe the scale should be adjusted to the number of dimensions and variance so that it still gives us sensible information about data. -->

## Metrics that pre-average are a bad guide

In contrast, statistical intervals will help us decide whether a given cosine similarity is high enough to consider the words to be more similar than if we chose them at random. We will use highest posterior density intervals, in line with Bayesian methodology.

Crucially, these approaches use  means of mean average cosine similarities to measure  similarity between protected word and harmful stereotypes. If one takes  a closer look at the individual values that are taken for the calculations, it turns out that there are quite a few  outliers and surprisingly dissimilar words. This problem becomes transparent when we examine the visualizations of the individual cosine distances, following the idea that one of the first steps in understanding data is to look at it.

With such a method the uncertainty involved is not really considered which makes it even more difficult to give reasonable interpretations of the results. We propose the use of Bayesian method to obtain some understanding of the influence the uncertainty has on the interpretation of final results.

\noindent
$s(X,Y,A,B)$ is the statistic used in the significance test, and the $p$-value is obtained by bootstrapping: it is the frequency of $s(X_i,Y_i,A,B)>s(X,Y,A,B)$ for all equally sized partitions $X_i, Y_i$ of $X\cup Y$. The effect size is computed by normalizing the difference in means as follows:

\vspace{-2mm}

\footnotesize

\begin{align}
bias(A,B) & = \frac{
\mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y}) 
}{
\sigma(\{s(w,A,B)\}_{w\in X\cup Y})
}
\end{align}

\normalsize

The t-tests they employ are run on average cosines used to calculate MAC.

#  Pre-averaging, bootstrapping and testing with the null model

The effect sizes reported by [@Caliskan2017semanticsBiases] for lists of words for which the embeddings are supposedly biased range from 2.06 to 1.81, and the reported $p$-values are in the range of $10^{-7}-10^{-2}$ with one exception for Math vs arts, where it is $.018$. The question is, are those results meaningful?

One way to answer this question is to think in terms of null models. If the words actually are samples from two populations with equal mean, how often would we see effect sizes in this range? How often would we reach the $p$-values that the authors reported?

We start our exploration with considering a list of 16 protected words, each class containing 8, with 16 attributes, 8 in each attribute class. In each particular sample we will investigate the consequence of the model being in fact null: we draw the cosine distances, every time, from the $\mathsf{Normal}(0,.08)$ distribution. $0.08$ is approximately the empirical standard deviation observed in fairly large samples of neutral words, and 16 is the sample size used in the WEAT7 word list, which is not much different from the other sample sizes in word lists used by [@Caliskan2017semanticsBiases]

```{r CaliskanCalc2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
vectorMean <- function(vector){ sum(vector)/length(vector)}
vectorSd <- function(vector) {
     vmean <- vectorMean(vector)
     return (sqrt(sum((vector - vmean)^2)/(length(vector))) )
}

set.seed(10232169)
count <- 16
randomizeDataset <- function(count = 16, mean = 0, sd = 0.08) {
    tlist <- list()
    for (t in 1:count) {
                  tlist[[t]] <- data.frame(t = rep(t,count/2), A = 1:(count/2),
                  Avalue  = rnorm(count/2,mean,sd), B = 1:(count/2),
                  Bvalue = rnorm(count/2,mean,sd))
                  }

  tDF <- do.call(rbind,tlist)
  tDF$target <- as.factor(ifelse(tDF$t <= (count/2), "X", "Y"))
  tDF$t <- as.factor(tDF$t)
  tDF$A <- as.factor(tDF$A)
  tDF$B <- as.factor(tDF$B)

  return(tDF)
}

tDF <- randomizeDataset()

svalues <- numeric(count)
for (term in 1:count){
  svalues[term] <- vectorMean(tDF[tDF$t== term,]$Avalue) - 
    vectorMean(tDF[tDF$t== term,]$Bvalue)
}

sDF <- data.frame(term = 1:count, s = svalues, 
                  attribute = as.factor(c(rep("X", (count/2)), rep("Y", (count/2)))))

testListX <- 1:(count/2)
testListY <- (count/2+1):16

testStatistic <- function (listX, listY){
  sum(sDF[listX,]$s) -  sum(sDF[listY,]$s)
      }

ourStatistic <- testStatistic(testListX, testListY)

effectSize <- function (listX, listY){
  ( vectorMean(sDF[listX,]$s) -  vectorMean(sDF[listY,]$s) ) /
     vectorSd(sDF$s)
}

ourEffectSize <- effectSize (testListX, testListY)

combinations <- as.data.frame(t(combn(1:count, (count/2))))
start_time <- Sys.time()
for (row in 1:nrow(combinations)){
  rowCodes <- as.integer(combinations[row,])
  remainderCodes <- as.integer(1:count)[ !(1:count %in% rowCodes)]
  combinations$statistic[row] <- testStatistic(rowCodes, remainderCodes)
  combinations$es[row] <- effectSize(rowCodes, remainderCodes)
}

plotStat <- ggplot(combinations) + geom_density(aes(x = statistic))+
geom_vline(xintercept = ourStatistic, linewidth = .5, 
           lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
  theme(plot.title.position = "plot", axis.text.y = element_blank(), 
        axis.ticks.y = element_blank()) +
  labs(title = "Boostrapped distribution of test statistics",
       subtitle = "Random null model with N(0,.008) and n = 16")+
   annotate("label", x = ourStatistic, y = 2,
            label =  paste(round(ourStatistic,2),
                           ", p=", 0.011, sep = ""), size = 3)+
  xlim(-2,2)+xlab("test statistic")+ylab("density")

plotEff <- ggplot(combinations) + geom_density(aes(x = es))+
  xlab("effect size")+ylab("density")+
  geom_vline(xintercept = ourEffectSize, linewidth = .5, 
             lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
  theme(plot.title.position = "plot", axis.text.y = element_blank(), 
        axis.ticks.y = element_blank()) +
  labs(title = "Boostrapped distribution of effect sizes",
       subtitle = "Random null model with N(0,.008) and n = 16")+
  annotate("label", x = ourEffectSize, y = .5,
           label = paste(round(ourEffectSize,2),
                         ", p=", 0.0086, sep = "" ), 
           size = 3)+xlim(-2,2)

plotCaliskanDist <- ggplot(tDF) + 
  geom_point(aes(x = Avalue, y = Bvalue, color = target, shape = target
                 ), size = 3, alpha = .8)+
  theme_tufte(base_size = 10)+
  ggtitle("Cosine distances by word group in  our null dataset")+
  theme(plot.title.position = "plot", 
        legend.position = c(0.08, 0.89),
        legend.background = element_rect(fill="lightblue", 
         size=0.01, linetype="solid"))

```

So let's draw one sample of this type. We calculate the s-values, the test statistics and the effect sizes. Then, following the original methodology, we obtain bootstrapped distributions of the effect sizes and the test statistics by calculating the same for each possible equal split of the initial random data set. The observed test statistic is `r ourStatistic` and `r ourEffectSize`. The bootstrapped distributions of the test statistics and effect sizes are illustrated in Figure \ref{fig:caliskanCalc}. Quite notably both (two-sided) $p$ values are uneven and rather low.

\begin{figure}[H]
```{r fig:caliskanCalcB2B,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
grid.arrange(plotStat,plotEff)
```

\caption{Bootstrapped distributions of test statistics and effect sizes in a random sample given the null hypothesis.}
\label{fig:caliskanCalc}
\end{figure}

At this point, we might think---right, so while drawing on the assumption of the null hypothesis we just stumbled into a data set that randomly happened to display strong bias. We decide to double-check this by visual inspection expecting exactly this: a strong, clearly visible bias (Figure \ref{fig:caliskanDistances}).

\begin{figure}[H]
```{r fig:caliskanDistC2B,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
plotCaliskanDist 
```

\caption{Cosine distances to two attribute sets by protected word groups. Observe nothing unusual except for a few outliers.}
\label{fig:caliskanDistances}
\end{figure}

\noindent In fact, while there might be some outliers here and there, saying that a clear bias on which one group is systematically closer to $A$s than another is definitely a stretch. What happened?

In the calculations of WEAT means are taken twice. The $s$-values themselves are means, and then means of $s$-values are compared between groups. The test statistic used itself is of this sort as well---while it compares sums, these are sums of sets of the same size, so they are really linear transformations of means of means. Statistical troubles start when we run statistical tests on sets of means, for at least two reasons  One, by pre-averaging data we throw away information about sample sizes. For the former point, think about proportions: 10 out of 20 and 2 out of 4 give the same mean, but you would obtain more information by making the former observation rather than by making the latter.  And especially in this context, in which the word lists are not huge, sample sizes should matter. Two, when we pre-average, we remove variation, and so pre-averaging  tend to manufacture false confidence. Group means display less variation than the raw data points, and the standard deviation of a set of means of sets of means is bound to be lower than the original standard deviation in the row data. Now, if you calculate your effect size by dividing by the pre-averaged standard deviation, your quite likely to get something that looks like a strong effect size, but the results of your calculations might not track anything interesting.

So let us think again about the question that we are  ultimately interested in. Are the $X$ terms systematically closer to (further from) the $A$ attributes ($B$ attributes) than the $Y$ words? But now let's use the raw data points to try to answer these questions. As the first stab, let us run two quick $t$-tests to gauge what the raw data illustrated in Figure \ref{fig:caliskanDistances} tell us.

First, distances to $A$ attributes for $X$ words and $Y$ words. Whoa, the result is significant! The $p$-value is $0.02$. So sure, the sample is in some sense unusual. But the 95\% confidence interval for the difference in means is $[.0052 .061]$, clearly nothing that a reader would expect given that the calculated effect size seemed quite large. How about the $B$ attributes? Here the $p$-value is  $.22$ and the 95\% confidence interval is $[-0.03, .009]$, even less of a reason to think a bias is present.

The difficulties are exacerbated by the fact that statistical tests are based on bootstrapping from a relatively small data sets, which is quite likely to underestimate the population variance.  To make our point clear, let us avoid bootstrapping and work with the null generative model with  $\mathsf{Norm}(0,.08)$ for both word groups. We keep the sizes the same: we have eight protected words in each group, sixteen in total, and for each we randomly draw 8 distances from hypothetical $A$ attributes, and $8$ distances from hypothetical $B$ attributes. Calculate the test statistic and effect size the way [@Caliskan2017semanticsBiases] did. Do this 10000 times and look at what the distributions of these values are on the assumption of the null model with realistic empirically motivated raw data point standard deviation.

```{r ourDistances,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}

s <- function (table){ mean(table$A) - mean(table$B)}

set.seed(123456)

biasesD0 <- numeric(10000)
svaluesD0 <- numeric(10000)
factorD0  <- numeric(10000)
numeratorD0 <- numeric(10000)

for(i in 1:10000){
  t1 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t2 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t3 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t4 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t5 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t6 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t7 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t8 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))

factorD0[i] <- sd(c(s(t1),s(t2),s(t3),s(t4),s(t5),
                s(t6),s(t7),s(t8),s(t9),             s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16)))

numeratorD0[i] <-  mean(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),
                s(t7),s(t8)) -    
mean(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))

svaluesD0[i] <-  sum(s(t1),s(t2),s(t3),s(t4),s(t5),
                     s(t6),s(t7),s(t8)) -   sum(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))

biasesD0[i] <- numeratorD0[i]/factorD0[i]
}

```

```{r ourDistances3B,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}

dataPLOT <- data.frame(biasesD0, svaluesD0)

ourESplot <- ggplot(dataPLOT )+geom_histogram(aes(x=biasesD0, y = ..density..), bins = 50, alpha = .6)+
  theme_tufte(base_size = 10)+
  labs(title="Distribution of effect sizes given a null model",
       subtitle = "Simulation with Norm(0,0.8), n=10k")+ xlab("effect size")+
    theme(plot.title.position = "plot", axis.text.y = element_blank(), 
        axis.ticks.y = element_blank()) +
  geom_vline(xintercept = ourEffectSize, linewidth = .5,
             lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
  annotate("label", x = ourEffectSize, y = .06,
           label =  paste(round(ourEffectSize,2),
                          ", p=", 0.38, sep = ""), size = 3)

ourStatPlot <- ggplot(dataPLOT )+ geom_histogram(aes(x=svaluesD0, y = ..density..), alpha = .6, bins = 50) +
  theme_tufte(base_size = 10)+
  labs(title="Distribution of test statistics given a null model",
       subtitle = "Simulation with Norm(0,0.8), n=10k")+ xlab("test statistic")+
    theme(plot.title.position = "plot", axis.text.y = element_blank(), 
        axis.ticks.y = element_blank()) +
  geom_vline(xintercept = ourStatistic, linewidth = .5,
             lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
  annotate("label", x = ourStatistic, y = 1.2,
           label =  paste(round(ourStatistic,2),
                          ", p=", 0.01, sep = ""), size = 3)

```

\begin{figure}[H]
```{r fig:ourDistancesPlot2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}

grid.arrange(ourESplot, ourStatPlot, ncol = 1)

```

\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08$.}
\label{fig:ourDistances}
\end{figure}

The first observation is that the supposedly large effect size we obtained is not that unusual even assuming a null model. Around 38\% of samples result in effect size at least as extreme. This illustrates the point that the effect size used does not track anything interesting. Second, the distribution of test statistics is much more narrow, which means that if we use it to calculate $p$-values, it is not too difficult to obtain a supposedly significant test statistic which nevertheless does not correspond to anything interesting happening in the data set.

Where do we stand now? We have seen  that seemingly high effect sizes  might arise even if the underlying processes actually have the same mean.  The uncertainty resulting from including the raw data point variance into considerations is more extensive than the one suggested by the low p-values obtained from taking means or means of means as data points. In this section, we discussed the performance of the WEAT measure, but since the [@Manzini2019blackToCriminal] one is a generalization thereof, including the method of running statistical tests on pre-averaged data, our remarks, \emph{mutatis nutandis}, apply. Moreover, a general problem is that studies such as [@Caliskan2017semanticsBiases] or [@Manzini2019blackToCriminal]  do not use  control groups, and this makes the interpretation of the analysis additionally difficult.

What is the alternative? As we already emphasized: focusing on what the real underlying question is, and trying to answer it using a statistical analysis of the raw data. Moreover, since the data sets are not too large and since multiple evaluations are to be made, we will pursue this method from the Bayesian perspective.

# Bayesian method

## Introduction and model definition

Bayesian data analysis takes prior probability distributions, a mathematical model structure and the data, and returns the posterior probability distributions over the parameters of interest, thus capturing our uncertainty about their actual values. One important difference between such  a result and the result of a classical statistical analysis is that classical confidence intervals (CIs) have a rather complicated and somewhat confusing interpretation, which has little to do with the posterior probability distribution.^[Here are a few usual problems. CIs are often mistakenly interpreted as providing  the probability that a resulting confidence interval contains the true value of a parameter. CIs bring confusion also with regard to  precision, it is a common mistake to interpret narrow intervals as the ones corresponding to a more precise knowledge. Another fallacy is to associate CIs with likelihood and stating that values within a given interval are more probable than the ones outside it. The theory of confidence intervals does not support the above interpretations. CIs should be plainly interpreted as a result of certain procedure (there are many ways to obtain CIs from a given set of data) that will in the long run contain the true value if the procedure is performed a fixed amount of times. For a nice survey and explanation of these misinterpretations, see  [@Morey2015confidenceFallacy]. For a psychological study of the occurrence of such  misinterpretations, see REF. In this study, 120 researchers and 442 students were asked to assess the truth value of six false statements involving different interpretations of a CI.  Both researchers and students endorsed, on average, more than three of these statements.] [@Hoekstra2014Misinterpretation]

In fact, Bayesian highest posterior density intervals  (HPDIs, the narrowest intervals containing a certain ratio of the area under the curve) and CIs end up being numerically the same only if the prior probabilities are uniform. This illustrates that (1) classical analysis is unable to incorporate non-trivial priors, and (2) is therefore more susceptible to over-fitting, unless regularization (equivalent to a more straightforward Bayesian approach) is used. In contrast with CIs, the posterior distributions are easily interpretable and have direct relevance for the question at hand. Moreover, Bayesian data analysis is better at handling hierarchical models and small datasets, which is exactly what we will we dealing with.

In a standard Bayesian analysis, the first step is to understand the data and select potential predictors and the outcome variable. Once the data is understood, the next step is to formulate a mathematical description of the generative model of the relationships between the predictors and the outcome variable. Prior distributions must then be chosen for the parameters used in the model. Next, Bayesian inference must be applied to find posterior probabilities over the possible parameter values. Finally, after finding the posterior probabilities, it is important to check how well the posterior predictions reflect the data with a posterior predictive check. In our analysis the outcome variable is  the cosine distances  between the protected words and attribute words. The predictor is a factor determining whether a given attribute word is a neutral word, a human predicate, is stereotypically associated with the protected word, or comes from a different stereotype connected with another protected word. The idea is, if bias is present in the embedding, distances to associated attribute words should be systematically lower than to other attribute words.

Furthermore, conceptually there are two levels of analysis in our approach. On the one hand, we are interested in the general question of whether related attributes are systematically closer across the dataset. On the other hand, we are interested in a more fine-grained picture of the role of the predictor for particular protected words. Learning in hierarchical Bayesian models involves using Bayesian inference to update the parameters of the model. This update is based on the observed data, and estimates are made at different levels of the data hierarchy. We use hierarchical Bayesian models in which we simultaneously estimate parameters at the protected word level and at the global level, assuming that all lower-level parameters are drawn from global distributions. Such models can be thought of as incorporating adaptive regularization, which avoids overfitting and leads to improved estimates for unbalanced datasets (and the datasets we need to use are unbalanced).

To be more specific, the underlying mathematical model is as follows. First, we assume that distances are normally distributed:
\begin{align*} \mathsf{distance_i} & \sim \mathsf{dnorm}(\mu_i,\sigma_i)
\end{align*}
\noindent Second, for each particular protected word $\mathsf{pw}$ there are four parameters to be estimated. Its mean distance to associated stereotypes $a[pw]$, its mean distance to attributes coming from different stereotypes, $d[pw]$, its mean distance to human attributes, $h[pw]$, and its mean distance to neutral attributes, $n[pw]$:
\begin{align*}
\mu_i & = d_{\mathsf{pw[i]}} \times \mathsf{different}_i  + a_{\mathsf{pw[i]}} \times \mathsf{associated}_i  + h_{\mathsf{pw[i]}} \times \mathsf{human}_i  + n_{\mathsf{pw[i]}}\times \mathsf{neutral}_i
\end{align*}
\noindent where $\mathsf{different}, \mathsf{associated},\mathsf{human}$ and $\mathsf{neutral}$ are binary variables.

This completes our description of the underlying mechanism. Now the priors and the hierarchy. We assume all the $a$ parameters come from one distribution, that is normal around a higher-level parameter $\bar{a}$ and so on for the other three groups of parameters. That is, $a_{\mathsf{pw[i]}}$ is the average distance of a given particular protected word to attributes stereotypically associated with it, while $\bar{a}$ is the overall average distance of protected words to attributes associated with them     [@kruschke2015bayesian; @statrethinkingbook2020].
\begin{align*}
d_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{d}, \overline{\sigma_d}) \\
a_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{a}, \overline{\sigma_a}) \\
h_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{h}, \overline{\sigma_h}) \\
n_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{n}, \overline{\sigma_n}) 
\end{align*}

According to our priors, the group means  $\bar{a}$, $\bar{d}$, $\bar{h}$ and $\bar{n}$  all come from one normal distribution with mean equal to $1$ and standard deviation equal to $.3$.  The standard deviations $\bar{\sigma_a}$, $\bar{\sigma_d}$, $\bar{\sigma_h}$ and $\bar{sigma_n}$ to be estimated, according to our prior,  come from one  distribution, exponential with rate parameter equal to $2$. Our priors are slightly skeptical. They do reflect our knowledge and intuition on the probable distribution of the cosine distances in the data. We know that the cosine distances lie in the range $0-2$, and we expect two randomly chosen  vectors  from the embedding to have rather small similarity, so we expect the distances to be centered around $1$. However, we use a rather wide standard deviation ($.3$) to easily account for cases where there is actually much higher similarity between two vectors (especially in cases where the embedding is supposed to be biased). Our priors for the standard deviations are also fairly weak.\footnote{To build a model one may use original \textit{Stan} which is a state-of-the-art platform for statistical modeling and high-performance statistical computation (https://mc-stan.org/). The simpler option is to use \textit{Rethinking} package from the \textit{Statistical rethinking} course. With one of the functions \textit{ulam} one may build RStan models from the formulas in an easy manner}

\begin{align*}
\bar{d}, \bar{a}, \bar{h}, \bar{n} &\sim \mathsf{Norm}(1, .3)\\ 
\overline{\sigma_d}, \overline{\sigma_a},  \overline{\sigma_h},  \overline{\sigma_n}  &\sim \mathsf{Exp}(2)
\end{align*}

## Posterior predictive check

A posterior predictive check is a technique used to evaluate the fit of a Bayesian model by comparing its predictions with observed data. The underlying principle is to generate simulated data from the posterior distribution of the model parameters and compare them with the observed data. If the model is a good fit to the data, the simulated data should resemble the observed data. A posterior predictive check is crucial because it provides a way to assess the validity of a Bayesian model. This is important because a model that fits the observed data well may still fail to generalize to new data. In the figure below we can see that the model fits the the data well.

```{r generatePosteriorPredictiveCheckA2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
source("functions/posteriorCheckPlot.R")

posteriorPlotsA <- list()

for (i in 1:length(predictions)){
  name <- names(predictions)[i]
  posteriorPlotsA[[i]] <- posteriorCheckPlotA(predictions[[i]], name = name) 
}
```

```{r generatePosteriorPredictiveCheckB6,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}

posteriorPlotsB <- list()

for (i in 1:length(predictions)){
  posteriorPlotsB[[i]] <- posteriorCheckPlotB(predictions[[i]], simulated[[i]],names(predictions)[i]) 
}

```

\begin{figure}[H]
```{r figposteriorPrCheckA6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", dpi = 300, warning = FALSE, message = FALSE}
grid.arrange(posteriorPlotsA[[3]], posteriorPlotsB[[3]], ncol = 2)
```
\caption{Example of a posterior predictive check. (Left) Actual cosine distances are plotted against mean predictions with 89\% highest posterior density intervals. Notice that 90\% of actual values fall within the 89\% HPDI and 55\% of actual values fall into 50\% HPDI, which indicates appropriate performance of the model. The left-right alignment of differrent colors coressponds to the fact that cosine differences between elements of different categories differ, to some extent systematically (this will be studied in the results section). (Right) Densities of predicted and observed distances.}
\label{fig:posteriorCheck1}
\end{figure}

\normalsize

# Results and discussion

\vspace{1mm}
\footnotesize
```{r generatePlots2,echo=FALSE, eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
source("functions/cleanDataset.R")
source("functions/plotFromPrecis.R")

weat7Google <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_7_google.csv")[,-1])
precisWeat7Google <- readRDS("resultsDFs/Weat7GoogleDF.rds")
resultsWeat7Google <- plotFromPrecis(precis = precisWeat7Google,
                                     dataset = weat7Google,
                                     list = "Weat 7",
                                     embedding = "Google",
                                     ylims = c(.75,1.1))

weat7Glove <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_7_glove.csv")[,-1])
precisWeat7Glove <- readRDS("resultsDFs/Weat7GloveDF.rds")
resultsWeat7Glove <- plotFromPrecis(precis = precisWeat7Glove,
                                     dataset = weat7Glove,
                                     list = "Weat 7",
                                     embedding = "Glove")

weat7Reddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_7_reddit.csv")[,-1])
precisWeat7Reddit <- readRDS("resultsDFs/Weat7RedditDF.rds")
resultsWeat7Reddit <- plotFromPrecis(precis = precisWeat7Reddit,
                                    dataset = weat7Reddit,
                                    list = "Weat 7",
                                    embedding = "Reddit", ylims = c(.5,1.3))

weat1Google <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_1_google.csv")[,-1])
precisWeat1Google <- readRDS("resultsDFs/Weat1GoogleDF.rds")
resultsWeat1Google <- plotFromPrecis(precis = precisWeat1Google,
                                     dataset = weat1Google,
                                     list = "Weat 1",
                                     embedding = "Google",
                                     ylims = c(.8,1.05))

weat1Glove <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_1_glove.csv")[,-1])
precisWeat1Glove <- readRDS("resultsDFs/Weat1GloveDF.rds")
resultsWeat1Glove <- plotFromPrecis(precis = precisWeat1Glove,
                                     dataset = weat1Glove,
                                     list = "Weat 1",
                                     embedding = "Glove",
                                     ylims = c(.75,1.2))

weat1Reddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_1_reddit.csv")[,-1])
precisWeat1Reddit <- readRDS("resultsDFs/Weat1RedditDF.rds")
resultsWeat1Reddit <- plotFromPrecis(precis = precisWeat1Reddit,
                                     dataset = weat1Reddit,
                                     list = "Weat 1",
                                     embedding = "Reddit",
                                      ylims = c(.5,1.1))

religionGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1])
precisReligionGoogle <- readRDS("resultsDFs/ReligionGoogleDF.rds")
resultsReligionGoogle <- plotFromPrecis(precis = precisReligionGoogle,
                                     dataset = religionGoogle,
                                     list = "Religion",
                                     embedding = "Google",
                                      ylims = c(.65,1.05))

religionGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1])
precisReligionGlove <- readRDS("resultsDFs/ReligionGloveDF.rds")
resultsReligionGlove <- plotFromPrecis(precis = precisReligionGlove,
                                        dataset = religionGlove,
                                        list = "Religion",
                                        embedding = "Glove",
                                        ylims = c(.7,1.1))

religionReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1])
precisReligionReddit <- readRDS("resultsDFs/ReligionRedditDF.rds")
resultsReligionReddit <- plotFromPrecis(precis = precisReligionReddit,
                                       dataset = religionReddit,
                                       list = "Religion",
                                       embedding = "Reddit")

raceGoogle <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1])
precisRaceGoogle <- readRDS("resultsDFs/RaceGoogleDF.rds")
resultsRaceGoogle <- plotFromPrecis(precis = precisRaceGoogle,
                                        dataset = raceGoogle,
                                        list = "Race",
                                        embedding = "Google",
                                     ylims = c(.8,1.1))

raceGlove <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1])
precisRaceGlove <- readRDS("resultsDFs/RaceGloveDF.rds")
resultsRaceGlove <- plotFromPrecis(precis = precisRaceGlove,
                                    dataset = raceGlove,
                                    list = "Race",
                                    embedding = "Glove",
                                    ylims = c(.7,1.1))

raceReddit <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1])
precisRaceReddit <- readRDS("resultsDFs/RaceRedditDF.rds")
resultsRaceReddit <- plotFromPrecis(precis = precisRaceReddit,
                                    dataset = raceReddit,
                                    list = "Race",
                                    embedding = "Reddit")

genderGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1])
precisGenderGoogle <- readRDS("resultsDFs/GenderGoogleDF.rds")
resultsGenderGoogle <- plotFromPrecis(precis = precisGenderGoogle,
                                    dataset = genderGoogle,
                                    list = "Gender",
                                    embedding = "Google",
                                     ylims = c(.6,1.05))

genderGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1])
precisGenderGlove <- readRDS("resultsDFs/GenderGloveDF.rds")
resultsGenderGlove <- plotFromPrecis(precis = precisGenderGlove,
                                      dataset = genderGlove,
                                      list = "Gender",
                                      embedding = "Glove",
                                      ylims = c(.5,1.1))

genderReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1])
precisGenderReddit <- readRDS("resultsDFs/GenderRedditDF.rds")

resultsGenderReddit <- plotFromPrecis(precis = precisGenderReddit,
                                     dataset = genderReddit,
                                     list = "Gender",
                                     embedding = "Reddit",  ylims = c(.4,1.1))

#now debiased
debiasedReligionReddit <- read.csv("./datasets/macWeatDatasets/debiased_religion_reddit.csv")[,-1]
debiasedRaceReddit <- read.csv("./datasets/macWeatDatasets/debiased_race_reddit.csv")[,-1]
debiasedGenderReddit <- read.csv("./datasets/macWeatDatasets/debiased_gender_reddit.csv")[,-1]

debiasedReligionReddit <- cleanDataset(debiasedReligionReddit)
debiasedRaceReddit <- cleanDataset(debiasedRaceReddit)
debiasedGenderReddit <- cleanDataset(debiasedGenderReddit)
precisDebiasedReligionReddit <- readRDS("resultsDFs/debiasedReligionRedditDF.rds")
precisDebiasedRaceReddit <- readRDS("resultsDFs/debiasedRaceRedditDF.rds")
precisDebiasedGenderReddit <- readRDS("resultsDFs/debiasedGenderRedditDF.rds")

resultsDebiasedReligionReddit <- plotFromPrecis(precis = precisDebiasedReligionReddit,
                                        dataset = debiasedReligionReddit,
                                        list = "Religion (MAC)",
                                        embedding = "Reddit (debiased)"
)

#resultsDebiasedReligionReddit$plotJoint
resultsDebiasedRaceReddit <- plotFromPrecis(precis = precisDebiasedRaceReddit,
                                                dataset = debiasedRaceReddit,
                                                list = "Race (MAC)",
                                                embedding = "Reddit (debiased)"
)

#resultsDebiasedRaceReddit$plotJoint
resultsDebiasedGenderReddit <- plotFromPrecis(precis = precisDebiasedGenderReddit,
                                            dataset = debiasedGenderReddit,
                                            list = "Gender (MAC)",
                                            embedding = "Reddit (debiased)",
                                            ylims = c(.4,1.1)
)

#resultsDebiasedGenderReddit$plotJoint
```

## Observations

In brief, despite one-number metrics suggesting otherwise, our Bayesian analysis reveals that insofar as the short word lists usually used in related research projects are involved, there usually are no strong reasons to claim the presence of systematic bias. Moreover, comparison between the groups (including control word lists) leads to the conclusion that the effect sizes (that is, the absolute differences between cosine distances between groups) tend to be rather small, with few exceptions.  Moreover, the choice of protected words is crucial --- as there is a lot of variance when it comes to the protected word-level analysis.

In a bit more detail, the visualizations in  Appendix \ref{appendix:visualizations} show that the situation is more complicated than merely looking at one-number summaries might suggest. Note that the axes are sometimes in different scales to increase visibility.

To start with, let us look at the association-type level coefficients (illustrated in the top parts of the plots). Depending on the corpus used and word class,  there is a large variety as to  the posterior densities. Quite aware of this being a crude approximation, let's compare the HPDIs and whether they overlap for different attribute groups.

- In Weat 7 (Reddit)    there is no reason to think there are systematic differences between cosine distances (recall that words from Weat 7 were mostly not available in other embeddings).

- In Weat 1 (Google, Glove and Reddit) associated words are somewhat closer, but the cosine distance differences from neutral words are very low, and surprisingly it is human attributes, not neutral predicates that are systematically the furthest.

- In Religion (Google, Glove, Reddit) and Race (Google, Glove), the associated attributes are not systematically closer than attributes belonging to different stereotypes, and the difference from neutral and human predicates is rather low, if noticeable. The situation is interestingly different in Race (Reddit) where both human and neutral predicates are systematically further than associated and different attributes - but even then, there is no clear difference between associated and different attributes.

- For Gender (Google, Glove), despite the superficially binary nature, associated and opposite attributes tend to be more or less in the same distances, much closer than neutral words (but not closer than human predicates in Glove). Reddit is an extreme example: both associated and opposite attributes are much closer than neutral and human (around .6 vs. .9), but even then, there seems to be no reason to think than cosine distances to associated predicates are much different from distances to opposite predicates.

Moreover, when we look at particular protected words, the situation is even less straightforward. We will just go over a few notable examples, leaving the visual inspection of particular results for other protected words to the reader. One general phenomenon is that--as we already pointed out--the word lists are quite short, which contributes to large uncertainty involved in some cases.

- For some protected words  the different attributes are somewhat closer than the associated attributes.

- For some protected words, associated and different attributes are closer than neutral attributes, but so are human attributes.

- In some cases, associated attributes are closer, but so are neutral and human predicates, which illustrates that just looking at average cosine similarity as compared to the theoretically expected value of 1, instead of running comparison to neutral and human attributes is misleading.

- The only group of protected words where differences are noticeable at the protected word level is Gender-related words-- as in Gender (Google) and in Gender (Reddit) --- note however that in the latter, for some words the opposite attributes seem to be a little bit closer than the associated ones.

## Rethinking debiasing

These visualizations can be also handy when it comes to the investigation of the effect that debiasing has on the embedding space. Below one may see chosen pair depicting the difference in means with 89% highest posterior density intervals before and after applying debiasing.

\begin{figure}[H]
```{r fig:debiasedCosinePair2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, out.width="85%"}

resultsGenderReddit$plotJoint

resultsDebiasedGenderReddit$plotJoint

```
\caption{Means with 89\% highest posterior density intervals for gender before and after debiasing.}
\label{fig:empiricalDebiasedPair}
\end{figure}

In Figure \ref{fig:empiricalDebiased} we look at the empirical distributions for the debiased embeddings. Comparing the results to the original embedding, one may notice that for the Religion group the neutral and human distribution has changed slightly. Before within the "correct" cosine similarity boundaries there were 56\% of neutral and 55\% of human word lists. After the debiasing the values changed to 59\% (for neutral) and 59\% (for human). The different and associated word lists were more influenced. The general shape of both distributions is less stretched. Before debiasing 43\% of the different word lists and 35\% of the associated word lists were within the accepted boundaries. After the embedding manipulation the percentage has increased for both lists to 63\%. Visualization for Gender group illustrates almost no change for the neutral and human word lists before and after debiasing. The values for different and associated word lists are also barely impacted by the embedding modification. In the Race group, the percentage within the boundaries for neutral and associated word lists have increased. The opposite happened for human and different word lists, where the percentage of "correct" cosine similarity dropped from 67\% to 55\% (human) and from 39\% to 36\% (different).

```{r debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "75%"}

debiasedGenderReddit <- read.csv("datasets/macWeatDatasets/debiased_gender_reddit.csv")[-1]
debiasedRaceReddit <- read.csv("datasets/macWeatDatasets/debiased_race_reddit.csv")[-1]
debiasedReligionReddit <- read.csv("datasets/macWeatDatasets/debiased_religion_reddit.csv")[-1]

debiasedViolinReligionReddit <- ggplot(debiasedReligionReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
  geom_vline(xintercept = .14) + geom_hline(yintercept = -.14, color = "grey") +
  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
  labs(title =  "Debiased Religion (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
    geom_jitter(size = 0.5, alpha = .2, width=0.08)+
  stat_summary(fun.data = "mean_cl_boot", fun.args = list(
    conf.int = .95), geom = "pointrange",
               colour = "black", size  = .2  )+
  annotate(geom = "label", x = 4.15, y = 0, label = "59%")+
  annotate(geom = "label", x = 3.15, y = 0, label = "59%")+
  annotate(geom = "label", x = 2.15, y = 0, label = "63%")+
  annotate(geom = "label", x = 1.15, y = 0, label = "63%")+
  theme(plot.title.position = "plot",legend.position="top")

debiasedViolinGenderReddit <- ggplot(debiasedGenderReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
  geom_vline(xintercept = .38) + geom_hline(yintercept = -.38, color = "grey") +
  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .38, color = "grey")+
  labs(title =  "Debiased Gender (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
    stat_summary(fun.data = "mean_cl_boot", fun.args = list(
    conf.int = .95), geom = "pointrange",
               colour = "black", size  = .2  )+
  annotate(geom = "label", x = 4.15, y = 0, label = "96%")+
  annotate(geom = "label", x = 3.15, y = 0, label = "92%")+
  annotate(geom = "label", x = 2.15, y = 0, label = "51%")+
  annotate(geom = "label", x = 1.15, y = 0, label = "51%")+
  theme(plot.title.position = "plot",legend.position='none')

debiasedViolinRaceReddit <- ggplot(debiasedRaceReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
  geom_vline(xintercept = .14) + geom_hline(yintercept = -.14, color = "grey") +
  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
  labs(title =  "Debiased Race (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
    stat_summary(fun.data = "mean_cl_boot", fun.args = list(
    conf.int = .95), geom = "pointrange",
               colour = "black", size  = .2  )+
  annotate(geom = "label", x = 4.15, y = 0, label = "48%")+
  annotate(geom = "label", x = 3.15, y = 0, label = "55%")+
  annotate(geom = "label", x = 2.15, y = 0, label = "36%")+
  annotate(geom = "label", x = 1.15, y = 0, label = "38%")+
  theme(plot.title.position = "plot")+geom_point(size = 1.2, alpha = .3)+
  theme(plot.title.position = "plot",legend.position='none')
```
\pagebreak

\begin{figure}[H]
```{r fig:debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "70%", fig.height = 16}

grid.arrange(violinReligionReddit, debiasedViolinReligionReddit, violinGenderReddit, debiasedViolinGenderReddit, violinRaceReddit, debiasedViolinRaceReddit, ncol = 2)

```
\caption{Empirical distributions of debiased cosine similarities (1-distances) for three word lists used in  the original paper.}
\label{fig:empiricalDebiased}
\end{figure}

# Related work and conclusions

It has been suggested [@Ethayarajh2020Bernstein] to use Bernstein bounds to state the uncertainty of the sample bias in a form of confidence interval. The interpretation of this interval is as follows, 95\% confidence interval contains such a range of values, that one can be 95% certain that it contains the population mean. As we will prove later in the text, confidence intervals are quite problematic for several reasons, among others the confusing interpretability. In [@Ethayarajh2020Bernstein] the current problem of bias estimation was correctly identified.  Indeed, the bias estimate should not be expressed as a single number without taking into account that the estimate is made on a sample of data and therefore has intrinsic uncertainty. However, we argue that Bernstein bounds do not provide the best solution to this problem. Applying their method to a popular WinoBias dataset leads to the conclusion that more than 11903 samples are needed to claim a 95\% confidence interval for a bias estimate. This amount exceeds existing datasets for bias measurement. We propose a more realistic method that can be applied to commonly used datasets to estimate the uncertainty of the bias measurement.

\todo{More references to similar papers}

A Bayesian data analysis with hierarchical models of cosine distances between protected words, control group words, and stereotypical attributes  provides more modest and realistic assessment of the uncertainty involved. It reveals that much complexity is hidden when one instead chases single bias metrics present in the literature.

After introducing the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of WEAT or MAC (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive. Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable. The tools that we propose, however, allow for a more fine-grained and multi-level evaluation of bias and debiasing in language models without losing modesty about the uncertainties involved.

\newpage
\appendix
# Appendix

## Visualizations

\label{appendix:visualizations}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsWeat7Google$plotJoint
```
\caption{dsds}
\label{fig:weat7google}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsWeat7Glove$plotJoint
```
\caption{dsds}
\label{fig:weat7glove}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsWeat7Reddit$plotJoint
```
\caption{dsds}
\label{fig:weat7reddit}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsWeat1Google$plotJoint
```
\caption{dsds}
\label{fig:weat1google}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsWeat1Glove$plotJoint
```
\caption{dsds}
\label{fig:weat1glove}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsWeat1Reddit$plotJoint
```
\caption{dsds}
\label{fig:weat1reddit}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsReligionGoogle$plotJoint
```
\caption{dsds}
\label{fig:religionGoogle}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsReligionGlove$plotJoint
```
\caption{dsds}
\label{fig:religionGlove}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsReligionReddit$plotJoint
```
\caption{dsds}
\label{fig:religionReddit}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsRaceGoogle$plotJoint
```
\caption{dsds}
\label{fig:raceGoogle}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsRaceGlove$plotJoint
```
\caption{dsds}
\label{fig:raceGlove}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsRaceReddit$plotJoint
```
\caption{dsds}
\label{fig:raceReddit}
\end{figure}

\begin{figure}[H]

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsGenderGoogle$plotJoint
```
\caption{dsds}
\label{fig:genderGoogle}
\end{figure}

\begin{figure}[H]
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsGenderGlove$plotJoint
```
\caption{dsds}
\label{fig:genderGlove}
\end{figure}

\begin{figure}[H]
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsGenderReddit$plotJoint
```
\caption{dsds}
\label{fig:genderReddit}
\end{figure}

\begin{figure}[H]
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsDebiasedReligionReddit$plotJoint
```
\caption{dsds}
\label{fig:debiasedReligion}
\end{figure}

\begin{figure}[H]
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsDebiasedRaceReddit$plotJoint
```
\caption{dsds}
\label{fig:debiasedRace}
\end{figure}

\begin{figure}[H]
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
resultsDebiasedGenderReddit$plotJoint
```
\caption{dsds}
\label{fig:debiasedGender}
\end{figure}

<!-- ## Examples of WEAT and MAC calculations -->
<!-- \label{appendix: calculations} -->

## Posterior predictive checks

```{r figposteriorPrCheckAppendixa,echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "85%", out.height= "100%", dpi = 300, warning = FALSE, message = FALSE}
grid.arrange(posteriorCheckPlotA(predictions[[1]], name = names(predictions)[1]),
             posteriorCheckPlotB(predictions[[1]], simulated[[1]],names(predictions)[1]),
             posteriorCheckPlotA(predictions[[2]], name = names(predictions)[2]),
             posteriorCheckPlotB(predictions[[2]], simulated[[2]],names(predictions)[2]),
             ncol = 2
             )

             # , posteriorPlotsB[[1]],
             # posteriorPlotsA[[2]], posteriorPlotsB[[2]],
             # ncol = 2)
```

```{r figposteriorPrCheckAppendixb,echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "85%", out.height= "100%", dpi = 300, warning = FALSE, message = FALSE}
grid.arrange(posteriorCheckPlotA(predictions[[4]], name = names(predictions)[4]),
             posteriorCheckPlotB(predictions[[4]], simulated[[4]],names(predictions)[4]),
             posteriorCheckPlotA(predictions[[5]], name = names(predictions)[5]),
             posteriorCheckPlotB(predictions[[5]], simulated[[5]],names(predictions)[5]),
             ncol = 2
             )

```

```{r figposteriorPrCheckAppendixc,echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "87%", out.height= "100%", dpi = 300, warning = FALSE, message = FALSE}
grid.arrange(posteriorCheckPlotA(predictions[[6]], name = names(predictions)[6]),
             posteriorCheckPlotB(predictions[[6]], simulated[[6]],names(predictions)[6]),
             posteriorCheckPlotA(predictions[[7]], name = names(predictions)[7]),
             posteriorCheckPlotB(predictions[[7]], simulated[[7]],names(predictions)[7]),
             ncol = 2
             )
```

```{r figposteriorPrCheckAppendixd,echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "89%", out.height= "100%", dpi = 300, warning = FALSE, message = FALSE}
grid.arrange(posteriorCheckPlotA(predictions[[8]], name = names(predictions)[8]),
             posteriorCheckPlotB(predictions[[8]], simulated[[8]],names(predictions)[8]),
             posteriorCheckPlotA(predictions[[9]], name = names(predictions)[9]),
             posteriorCheckPlotB(predictions[[9]], simulated[[9]],names(predictions)[9]),
             ncol = 2
             )
```

## Word lists

### Lists for [@Manzini2019blackToCriminal]:
\label{appendix:manzini_word_lists}
The lists are available here:

-  **Gender:** \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/gender_attributes_optm.json}

- **Race:** \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/race_attributes_optm.json}

- **Religion:** \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/religion_attributes_optm.json}

The lists are:

- **Gender:**

Pairs of protected attributes:
		["he", "she"],
		["his", "hers"],
		["son", "daughter"],
		["father", "mother"],
		["male", "female"],
		["boy", "girl"],
		["uncle", "aunt"]

Lists of stereotypes per gender:
		man: ["manager", "executive", "doctor", "lawyer", "programmer", "scientist", "soldier", "supervisor", "rancher", "janitor", "firefighter", "officer"],
		woman: ["secretary", "nurse", "clerk", "artist", "homemaker", "dancer", "singer", "librarian", "maid", "hairdresser", "stylist", "receptionist", "counselor"]
		}

- **Race:**

Sets of protected attributes:
		["black", "caucasian", "asian"],
		["african", "caucasian", "asian"],
		["black", "white", "asian"],
		["africa", "america", "asia"],
		["africa", "america", "china"],
		["africa", "europe", "asia"]

Lists of stereotypes per race:
			"caucasian": ["manager", "executive", "redneck", "hillbilly", "leader", "farmer"],
			"asian"    : ["doctor", "engineer", "laborer", "teacher"],
			"black"    : ["slave", "musician", "runner", "criminal", "homeless"]

- **Religion:**

Sets of protected attributes:
		["judaism", "christianity", "islam"],
		["jew", "christian", "muslim"],
    ["synagogue", "church", "mosque"],
    ["torah", "bible", "quran"],
    ["rabbi", "priest", "imam"]

Lists of stereotypes per race:
			"jew"       : ["greedy", "cheap", "hairy", "liberal"],
    	"christian" : ["judgemental", "conservative", "familial"],
    	"muslim"    : ["violent", "terrorist", "dirty", "uneducated"]

### Our custom word lists

- **Neutral:**

 ['ballpark', 'glitchy', 'billy', 'dallas', 'rip', 'called', 'outlooks',
                 'floater', 'rattlesnake', 'exports',
                 'recursion', 'shortfall', 'corrected', 'solutions', 'diagnostic', 'patently', 'flops', 'approx', 'percents',
                 'lox', 'hamburger', 'engulfed', 'households', 'north', 'playtest', 'replayability', 'glottal',
                 'parable', 'gingers', 'anachronism', 'organizing', 'reach', 'shtick', 'eleventh', 'cpu', 'ranked',
                 'irreversibly', 'ponce', 'velociraptor', 'defects', 'puzzle', 'smasher', 'northside', 'heft', 'observation', 'rectum',
                 'mystical', 'telltale', 'remnants', 'inquiry', 'indisputable', 'boatload', 'lessening', 'uselessness', 'observes', 'fictitious',
                 'repatriation', 'duh', 'attic', 'schilling', 'charges', 'chatter', 'pad', 'smurfing', 'worthiness', 'definitive', 'neat', 'homogenized',
                 'lexicon', 'nationalized', 'earpiece', 'specializations', 'lapse', 'concludes', 'weaving', 'apprentices', 'fri',
                 'militias', 'inscriptions', 'gouda', 'lift', 'laboring', 'adaptive', 'lecture', 'hogging', 'thorne', 'fud', 'skews',
                 'epistles', 'tagging', 'crud', 'two', 'rebalanced', 'payroll', 'damned', 'approve', 'reason', 'formally', 'releasing', 'muddled',
                 'mineral', 'shied', 'capital', 'nodded', 'escrow', 'disconnecting', 'marshals',
                 'winamp', 'forceful', 'lowes', 'sip', 'pencils', 'stomachs', 'goff', 'cg', 'backyard', 'uprooting', 'merging',
                 'helpful', 'eid', 'trenchcoat', 'airlift', 'frothing', 'pulls', 'volta', 'guinness', 'viewership',
                 'eruption', 'peeves', 'goat', 'goofy', 'disbanding', 'relented', 'ratings', 'disputed', 'vitamins', 'singled', 'hydroxide',
                 'telegraphed', 'mercantile', 'headache', 'muppets', 'petal', 'arrange', 'donovan', 'scrutinized', 'spoil', 'examiner', 'ironed',
                 'maia', 'condensation', 'receipt', 'solider', 'tattooing', 'encoded', 'compartmentalize', 'lain', 'gov',
                 'printers', 'hiked', 'resentment', 'revisionism', 'tavern', 'backpacking', 'pestering', 'acknowledges',
                 'testimonies', 'parlance', 'hallucinate', 'speeches', 'engaging', 'solder', 'perceptive', 'microbiology', 'reconnaissance', 'garlic',
                 'neutrals', 'width', 'literaly', 'guild', 'despicable', 'dion', 'option', 'transistors',
                 'chiropractic', 'tattered', 'consolidating', 'olds', 'garmin', 'shift', 'granted', 'intramural', 'allie', 'cylinders',
                 'wishlist', 'crank', 'wrongly', 'workshop', 'yesterday', 'wooden', 'without', 'wheel', 'weather', 'watch', 'version', 'usually', 'twice',                      'tomato', 'ticket', 'text', 'switch', 'studio', 'stick', 'soup', 'sometimes', 'signal', 'prior', 'plant', 'photo', 'path', 'park', 'near',
                 'menu', 'latter', 'grass', 'clock']

- **Human-related:**

['wear', 'walk', 'visitor', 'toy', 'tissue', 'throw', 'talk', 'sleep', 'eye',
               'enjoy', 'blogger', 'character', 'candidate', 'breakfast', 'supper', 'dinner',
               'eat', 'drink', "carry", "run", "cast", "ask", "awake", "ear", "nose", "lunch", "coalition",
               "policies", "restaurant", "stood", "assumed", "attend", "swimming", "trip", "door", "determine", "gets",
               "leg", "arrival", "translated", "eyes", "step", "whilst", "translation", "practices", "measure",
               "storage", "window", "journey", "interested", "tries", "suggests", "allied", "cinema", "finding",
               "restoration", "expression","visitors", "tell", "visiting",
               "appointment", "adults", "bringing", "camera", "deaths", "filmed", "annually", "plane", "speak",
               "meetings", "arm", "speaking", "touring", "weekend", "accept", "describe", "everyone", "ready",
               "recovered", "birthday", "seeing", "steps", "indicate", "anyone", "youtube"]

### 2023-05-04 paper/paperDraft.tex
Adds files without models

--- 2023-05-04 paper/paperDraft.Rmd
@@
----
-title: 'Taking uncertainty in word embedding bias estimation seriously: a  bayesian
-  approach'
-author: "Alicja Dobrzeniecka and Rafal Urbaniak"
-output:
-  pdf_document:
-    number_sections: yes
-    df_print: kable
-    keep_tex: yes
-    includes:
-      in_header: Rafal_latex7.sty
-  html_document:
-    df_print: paged
-classoption: dvipsnames,enabledeprecatedfontcommands
-fontsize: 10pt
-documentclass: scrartcl
-urlcolor: blue
-bibliography: ../references/cosineReferences1.bib
-csl: ../references/transactions-on-computational-logic.csl
----
-
-```{r, setup, include=FALSE}
-knitr::opts_knit$set(root.dir = '../')
-
-library(ggplot2)
-library(ggthemes)
-library(gridExtra)
-library(dplyr)
-library(kableExtra)
-library(tidyverse)
-library(magrittr)
-library(ggpubr)
-library(ggExtra)
-library(ggthemes)
-library(ggforce)
-library(latex2exp)
-library(Hmisc)
-library(grid)
-
-```
-
-# Introduction
-
-Natural language processing (NLP) models are used to perform various language-related tasks such as providing email filters, intelligent assistants, search results,  translations, text analysis and so on.
-Such tools require input words to be represented by numbers.
-This can be done with word embeddings, which represent words as vectors of real numbers.
-A common way to construct an embedding is to use  a large natural language corpus to train  a  neural network to assign vectors to words that optimize for co-occurrence prediction accuracy.
-The vectors can then be compared in terms of their similarity -- the usual measure is cosine similarity -- and the results of such comparisons can be used in downstream tasks.
-Roughly speaking, cosine similarity is an imperfect mathematical proxy for semantic similarity [@Mikolov2013efficient].
-
-It has been suggested [@Bolukbasi2016man; @Caliskan2017semanticsBiases; @Gonen2019lipstick; @Lauscher2019multidimensional; @Garg2018years; @Manzini2019blackToCriminal]  that  such models can learn implicit biases that reflect harmful stereotypical thinking---for example,  the (vector corresponding to the) word \textit{she} might be  much closer in the vector space to  the word \textit{cooking} than  the word \textit{he}.
-Such phenomena may be undesirable at least  in some  downstream tasks, such as web search, recommendations, and so on.
-To investigate such issues, several measures of bias in word embeddings have been formulated and applied.
-Our goal is to use two prominent examples of such measures to argue that this approach oversimplifies the situation and to develop a Bayesian alternative.
-
-One response to the raising of the issue of bias in natural language models might be to say that there is not much point to reflecting on such biases, as they are  unavoidable.
-This unavoidability might seem in line with the arguments to the effect that learning algorithms are always value-laden [@JohnsonValueFree]: they employ inductive methods that require design-, data-, or risk-related decisions that have to be guided by extra-algorithmic considerations.
-Such choices necessarily involve value judgments and have to do, for instance, with what simplifications or risks one finds acceptable.
-Admittedly, algorithmic decision making cannot fulfill the value-free ideal, but this only means that even more attention needs to be paid to the values underlying different techniques and decisions, and to the values being pursued in a particular use of an algorithm.
-
-Another response might be to insist that there is no bias introduced by the use of machine learning methods here, since the algorithm is simply learning to correctly predict co-occurrences based on what "reality" looks like.
-However, this objection overlooks the fact that we, humans, are the ones who construct this linguistic reality, is shaped in part by the natural language processing tools we use on a massive scale.
-Sure, if there is unfairness and our goal is to diagnose it, we should do complete justice to learning it in the model used in a study thereof.
-One example of this approach is [@Garg2017hundredYears].
-However, if our goal is to develop downstream tools that perform tasks that we care about well without further perpetuating  or exacerbating harmful stereotypes, we still have good reasons to try to minimize the negative impact, at least as long as this does not damage the overall performance of the tool.
-Moreover, it is often not the case that the corpora mirror reality---to give a trivial example, heads are spoken of more often than kidneys, but this does not mean that kidneys occur much less often in reality than heads.
-To give a more relevant example, the disproportionate association of female words with female occupations in a corpus actually greatly exaggerates the actual lower disproportion in the real distribution of occupations.
-[@gordon2012reporting].
-
-In what follows, we focus on  two popular  measures of bias applicable to many existing word embeddings ( \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.} \emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.} and \emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}): \emph{Word Embedding Association Test}  (WEAT) [@Caliskan2017semanticsBiases],  and \emph{Mean Average Cosine Distance} (MAC)  [@Manzini2019blackToCriminal].
-We first explain how these measures are supposed to work.
-Then we argue that they are problematic for various reasons, a key one being that by pre-averaging data they manufacture false confidence, which we illustrate in terms of simulations showing that the measures often suggest the existence of bias even if by design it is non-existent in a simulated dataset.
-
-We  propose to replace them with a Bayesian data analysis (hierarchical models), which not only provides more modest and realistic assessment of the uncertainty involved, but also allows for inspection at various levels of granularity.
-Once we introduce the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of WEAT or MAC (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
-Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable.
-The tools that we have proposed, however, allow for a more fine-grained and multi-level evaluation of bias and debiasing in language models without abandoning modesty about the uncertainties involved.
-
-# Two measures of bias: WEAT and MAC
-
-Conceptually the idea is that if a particular harmful stereotype is learned in a particular embedding, then certain groups of words will be systematically closer to (or further from) each other.
-This gives rise to the idea of protected groups---for example, in guiding online search completion or recommendation, female words might require protection in that they should not be systematically closer to stereotypically female job names, and male words require protection in that they should not be systematically closer to toxic masculinity stereotypes, such as "tough", "never complaining" or "macho".^[However, for some research-related purposes, such as the study of stereotypes across history [@Garg2017hundredYears], embeddings which do not protect certain classes may also be useful.]
-
-The key role in the measures to be discussed is played by the notion of cosine distance (or, symmetrically, by cosine similarity).
-These are defined as follows:^[Here, "$-$" stands for point-wise difference, "$\cdot$" stands for the dot product operation, and $\lVert a\rVert  = \sqrt{(a \cdot a)}$.]
-\begin{align} \tag{Sim}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  10pt,
+  dvipsnames,enabledeprecatedfontcommands]{scrartcl}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+%\documentclass{article}
+
+% %packages
+\usepackage{booktabs}
+%\usepackage[left]{showlabels}
+\usepackage{multirow}
+\usepackage{subcaption}
+
+\usepackage{graphicx}
+%\usepackage{longtable}
+\usepackage{supertabular, booktabs}
+\usepackage{ragged2e}
+\usepackage{etex}
+%\usepackage{yfonts}
+\usepackage{marvosym}
+\usepackage[notextcomp]{kpfonts}
+\usepackage{nicefrac}
+\newcommand*{\QED}{\hfill \footnotesize {\sc Q.e.d.}}
+\usepackage{floatrow}
+
+\usepackage[textsize=footnotesize]{todonotes}
+\newcommand{\ali}[1]{\todo[color=gray!40]{\textbf{Alicja:} #1}}
+\newcommand{\mar}[1]{\todo[color=blue!40]{#1}}
+\newcommand{\raf}[1]{\todo[color=olive!40]{#1}}
+
+%\linespread{1.5}
+\newcommand{\indep}{\!\perp \!\!\! \perp\!}
+
+\setlength{\parindent}{5mm}
+\setlength{\parskip}{1pt}
+
+%language
+%\usepackage{times}
+\usepackage{mathptmx}
+\usepackage[scaled=0.86]{helvet}
+\usepackage{t1enc}
+%\usepackage[utf8x]{inputenc}
+%\usepackage[polish]{babel}
+%\usepackage{polski}
+
+%AMS
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{amsmath}
+\usepackage{mathtools}
+
+\usepackage{geometry}
+ \geometry{a4paper,left=35mm,top=20mm,}
+
+%environments
+\newtheorem{fact}{Fact}
+
+%abbreviations
+\newcommand{\ra}{\rangle}
+\newcommand{\la}{\langle}
+\newcommand{\n}{\neg}
+\newcommand{\et}{\wedge}
+\newcommand{\jt}{\rightarrow}
+\newcommand{\ko}[1]{\forall  #1\,}
+\newcommand{\ro}{\leftrightarrow}
+\newcommand{\exi}[1]{\exists\, {_{#1}}}
+\newcommand{\pr}[1]{\ensuremath{\mathsf{P}(#1)}}
+\newcommand{\cost}{\mathsf{cost}}
+\newcommand{\benefit}{\mathsf{benefit}}
+\newcommand{\ut}{\mathsf{ut}}
+
+\newcommand{\odds}{\mathsf{Odds}}
+\newcommand{\ind}{\mathsf{Ind}}
+\newcommand{\nf}[2]{\nicefrac{#1\,}{#2}}
+\newcommand{\R}[1]{\texttt{#1}}
+\newcommand{\prr}[1]{\mbox{$\mathtt{P}_{prior}(#1)$}}
+\newcommand{\prp}[1]{\mbox{$\mathtt{P}_{posterior}(#1)$}}
+
+\newtheorem{q}{\color{blue}Question}
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}[fact]
+
+%technical intermezzo
+%---------------------
+
+\newcommand{\intermezzoa}{
+	\begin{minipage}[c]{13cm}
+	\begin{center}\rule{10cm}{0.4pt}
+
+	\tiny{\sc Optional Content Starts}
+
+	\vspace{-1mm}
+
+	\rule{10cm}{0.4pt}\end{center}
+	\end{minipage}\nopagebreak 
+	}
+
+\newcommand{\intermezzob}{\nopagebreak 
+	\begin{minipage}[c]{13cm}
+	\begin{center}\rule{10cm}{0.4pt}
+
+	\tiny{\sc Optional Content Ends}
+
+	\vspace{-1mm}
+
+	\rule{10cm}{0.4pt}\end{center}
+	\end{minipage}
+	}
+
+%--------------------
+
+\newtheorem*{reply*}{Reply}
+\usepackage{enumitem}
+\newcommand{\question}[1]{\begin{enumerate}[resume,leftmargin=0cm,labelsep=0cm,align=left]
+\item #1
+\end{enumerate}}
+
+\usepackage{float}
+
+\usepackage{sectsty}
+\sectionfont{\fontsize{11}{11}\selectfont}
+
+% \setbeamertemplate{blocks}[rounded][shadow=true]
+% \setbeamertemplate{itemize items}[ball]
+% \AtBeginPart{}
+% \AtBeginSection{}
+% \AtBeginSubsection{}
+% \AtBeginSubsubsection{}
+% \setlength{\emergencystretch}{0em}
+% \setlength{\parskip}{0pt}
+
+\usepackage[authoryear]{natbib}
+
+%\bibliographystyle{apalike}
+
+\usepackage{tikz}
+\usetikzlibrary{positioning,shapes,arrows}
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Taking uncertainty in word embedding bias estimation seriously: a bayesian approach},
+  pdfauthor={Alicja Dobrzeniecka and Rafal Urbaniak},
+  colorlinks=true,
+  linkcolor={Maroon},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={blue},
+  pdfcreator={LaTeX via pandoc}}
+
+\title{Taking uncertainty in word embedding bias estimation seriously: a
+bayesian approach}
+\author{Alicja Dobrzeniecka and Rafal Urbaniak}
+\date{}
+
+\begin{document}
+\maketitle
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Natural language processing (NLP) models are used to perform various
+language-related tasks such as providing email filters, intelligent
+assistants, search results, translations, text analysis and so on. Such
+tools require input words to be represented by numbers. This can be done
+with word embeddings, which represent words as vectors of real numbers.
+A common way to construct an embedding is to use a large natural
+language corpus to train a neural network to assign vectors to words
+that optimize for co-occurrence prediction accuracy. The vectors can
+then be compared in terms of their similarity -- the usual measure is
+cosine similarity -- and the results of such comparisons can be used in
+downstream tasks. Roughly speaking, cosine similarity is an imperfect
+mathematical proxy for semantic similarity {[}14{]}.
+
+It has been suggested {[}1,4,5,8,11,12{]} that such models can learn
+implicit biases that reflect harmful stereotypical thinking---for
+example, the (vector corresponding to the) word \textit{she} might be
+much closer in the vector space to the word \textit{cooking} than the
+word \textit{he}. Such phenomena may be undesirable at least in some
+downstream tasks, such as web search, recommendations, and so on. To
+investigate such issues, several measures of bias in word embeddings
+have been formulated and applied. Our goal is to use two prominent
+examples of such measures to argue that this approach oversimplifies the
+situation and to develop a Bayesian alternative.
+
+One response to the raising of the issue of bias in natural language
+models might be to say that there is not much point to reflecting on
+such biases, as they are unavoidable. This unavoidability might seem in
+line with the arguments to the effect that learning algorithms are
+always value-laden {[}9{]}: they employ inductive methods that require
+design-, data-, or risk-related decisions that have to be guided by
+extra-algorithmic considerations. Such choices necessarily involve value
+judgments and have to do, for instance, with what simplifications or
+risks one finds acceptable. Admittedly, algorithmic decision making
+cannot fulfill the value-free ideal, but this only means that even more
+attention needs to be paid to the values underlying different techniques
+and decisions, and to the values being pursued in a particular use of an
+algorithm.
+
+Another response might be to insist that there is no bias introduced by
+the use of machine learning methods here, since the algorithm is simply
+learning to correctly predict co-occurrences based on what ``reality''
+looks like. However, this objection overlooks the fact that we, humans,
+are the ones who construct this linguistic reality, is shaped in part by
+the natural language processing tools we use on a massive scale. Sure,
+if there is unfairness and our goal is to diagnose it, we should do
+complete justice to learning it in the model used in a study thereof.
+One example of this approach is {[}3{]}. However, if our goal is to
+develop downstream tools that perform tasks that we care about well
+without further perpetuating or exacerbating harmful stereotypes, we
+still have good reasons to try to minimize the negative impact, at least
+as long as this does not damage the overall performance of the tool.
+Moreover, it is often not the case that the corpora mirror reality---to
+give a trivial example, heads are spoken of more often than kidneys, but
+this does not mean that kidneys occur much less often in reality than
+heads. To give a more relevant example, the disproportionate association
+of female words with female occupations in a corpus actually greatly
+exaggerates the actual lower disproportion in the real distribution of
+occupations. {[}6{]}.
+
+In what follows, we focus on two popular measures of bias applicable to
+many existing word embeddings (
+\emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.}
+\emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.}
+and
+\emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}):
+\emph{Word Embedding Association Test} (WEAT) {[}8{]}, and
+\emph{Mean Average Cosine Distance} (MAC) {[}12{]}. We first explain how
+these measures are supposed to work. Then we argue that they are
+problematic for various reasons, a key one being that by pre-averaging
+data they manufacture false confidence, which we illustrate in terms of
+simulations showing that the measures often suggest the existence of
+bias even if by design it is non-existent in a simulated dataset.
+
+We propose to replace them with a Bayesian data analysis (hierarchical
+models), which not only provides more modest and realistic assessment of
+the uncertainty involved, but also allows for inspection at various
+levels of granularity. Once we introduce the method, we apply it to
+multiple word embeddings and results of supposed debiasing, putting
+forward some general observations that are not exactly in line with the
+usual picture painted in terms of WEAT or MAC (and the problem
+generalizes to any approach that focuses on chasing a single numeric
+metric): the word list sizes and sample sizes used in the studies are
+usually
+small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}
+posterior density intervals are fairly wide, often the differences
+between associated, different and neutral human predicates, are not very
+impressive. Also, a preliminary inspection suggests that the
+desirability of changes obtained by the usual debiasing methods is
+debatable. The tools that we have proposed, however, allow for a more
+fine-grained and multi-level evaluation of bias and debiasing in
+language models without abandoning modesty about the uncertainties
+involved.
+
+\hypertarget{two-measures-of-bias-weat-and-mac}{%
+\section{Two measures of bias: WEAT and
+MAC}\label{two-measures-of-bias-weat-and-mac}}
+
+Conceptually the idea is that if a particular harmful stereotype is
+learned in a particular embedding, then certain groups of words will be
+systematically closer to (or further from) each other. This gives rise
+to the idea of protected groups---for example, in guiding online search
+completion or recommendation, female words might require protection in
+that they should not be systematically closer to stereotypically female
+job names, and male words require protection in that they should not be
+systematically closer to toxic masculinity stereotypes, such as
+``tough'', ``never complaining'' or ``macho''.\footnote{However, for
+  some research-related purposes, such as the study of stereotypes
+  across history {[}3{]}, embeddings which do not protect certain
+  classes may also be useful.}
+
+The key role in the measures to be discussed is played by the notion of
+cosine distance (or, symmetrically, by cosine similarity). These are
+defined as follows:\footnote{Here, ``\(-\)'' stands for point-wise
+  difference, ``\(\cdot\)'' stands for the dot product operation, and
+  \(\lVert a\rVert = \sqrt{(a \cdot a)}\).} \begin{align} \tag{Sim}
 \mathsf{cosineSimilarity}(A,B) & = \frac{A \cdot B}{\lVert  A \rVert \,\lVert B \rVert}
@@
 \mathsf{cosineDistance}(A,B) &  = 1 - \mathsf{cosineSimilarity}(A,B).
-\end{align}
-Note  that this terminology is slightly misleading, as mathematically cosine distance is not a distance measure, because it does not satisfy the triangle inequality, as generally $\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)$.
+\end{align} Note that this terminology is slightly misleading, as
+mathematically cosine distance is not a distance measure, because it
+does not satisfy the triangle inequality, as generally
+\(\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)\).
 We will keep using this mainstream terminology.
 
-One of the first measures of bias has been developed in [@Bolukbasi2016man].
-The general idea is that a certain topic is associated with a vector of real numbers (the topic "direction"), and the bias of a word is investigated by considering the projection of its corresponding vector on this direction.
-For instance, in [@Bolukbasi2016man],   the gender direction $\mathsf{gd}$   is obtained by taking the differences of the vectors corresponding to ten different gendered pairs (such as $\overrightarrow{she} - \overrightarrow{he}$ or $\overrightarrow{girl} - \overrightarrow{boy}$), and then identifying their principal component.^[Roughly, the principal component  is the vector obtained by projecting the data points on their  linear combination in a way that maximizes the variance of the projections.]  The gender bias of a word $w$ is then understood as $w$'s projection on the gender direction: $\vec{w} \cdot gd$ (which, after normalizing by dividing by $\lVert  w \rVert \,\lVert \mathsf{gd} \rVert$, is the same as cosine similarity).
-Given a list  $N$ of  supposedly gender neutral words,^[We follow the methodology used in the debate in assuming that there is a class of words  identified as more or less neutral, such as \emph{ballpark, eat, walk, sleep, table}, whose average similarity to the gender direction (or other protected words) is around 0.
-We will have more to say about this assumption when we describe our dataset construction.]  and the gender direction $\mathsf{gd}$,    the direct gender bias is defined as the average cosine similarity of the words in $N$ from $\mathsf{gd}$ ($c$ is a parameter determining how strict we want to be):
-\begin{align*}
+One of the first measures of bias has been developed in {[}1{]}. The
+general idea is that a certain topic is associated with a vector of real
+numbers (the topic ``direction''), and the bias of a word is
+investigated by considering the projection of its corresponding vector
+on this direction. For instance, in {[}1{]}, the gender direction
+\(\mathsf{gd}\) is obtained by taking the differences of the vectors
+corresponding to ten different gendered pairs (such as
+\(\overrightarrow{she} - \overrightarrow{he}\) or
+\(\overrightarrow{girl} - \overrightarrow{boy}\)), and then identifying
+their principal component.\footnote{Roughly, the principal component is
+  the vector obtained by projecting the data points on their linear
+  combination in a way that maximizes the variance of the projections.}
+The gender bias of a word \(w\) is then understood as \(w\)'s projection
+on the gender direction: \(\vec{w} \cdot gd\) (which, after normalizing
+by dividing by \(\lVert w \rVert \,\lVert \mathsf{gd} \rVert\), is the
+same as cosine similarity). Given a list \(N\) of supposedly gender
+neutral words,\footnote{We follow the methodology used in the debate in
+  assuming that there is a class of words identified as more or less
+  neutral, such as \emph{ballpark, eat, walk, sleep, table}, whose
+  average similarity to the gender direction (or other protected words)
+  is around 0. We will have more to say about this assumption when we
+  describe our dataset construction.} and the gender direction
+\(\mathsf{gd}\), the direct gender bias is defined as the average cosine
+similarity of the words in \(N\) from \(\mathsf{gd}\) (\(c\) is a
+parameter determining how strict we want to be): \begin{align*}
 \mathsf{directBias_c(N,gd)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},\mathsf{gd})\vert^c}{\vert N \vert }
-\end{align*}
-\normalsize
-
-The use of projections in bias estimation has been criticized for instance in [@Gonen2019lipstick], where it is pointed out that while a higher average similarity to the  gender direction might be an indicator of bias with respect to a given  class of words, but it is only one possible manifestation of it, and reducing the cosine similarity to such a projection may not be sufficient to eliminate bias.
-For instance, "math" and "delicate" might be equally similar to a pair of  opposed explicitly gendered words (\emph{she}, \emph{he}), while being closer to quite different stereotypical attribute words (such as \emph{scientific} or \emph{caring}).
-Further, it is observed in [@Gonen2019lipstick]  that most word pairs retain similarity under debiasing meant to minimize projection-based bias.^[In [@Bolukbasi2016man]  another method which involves analogies and their evaluations by human users on Mechanical Turk is also used.
-We do not discuss this method in this paper, see its  criticism in [@Nissim2020fair].]
-
-A measure of bias in word embeddings which does not employ similarities to  pre-calculated protected-class-level directions, the Word Embedding Association Test (WEAT), has been proposed in  [@Caliskan2017semanticsBiases].
-The idea here is that the   bias between two sets of target words, $X$ and $Y$ (we call them protected words), should be quantified in terms of the cosine similarity  between the protected words and attribute words coming from  two sets of stereotype attribute words, $A$ and $B$ (we will call them attributes).
-For instance, $X$ might be a set of male names, $Y$ a set of female names, $A$ might contain stereotypically male-related, and $B$ stereotypically female-related career words.
-The association difference for a particular word $t$ (belonging to either $X$ or $Y$) is:
+\end{align*} \normalsize
+
+The use of projections in bias estimation has been criticized for
+instance in {[}5{]}, where it is pointed out that while a higher average
+similarity to the gender direction might be an indicator of bias with
+respect to a given class of words, but it is only one possible
+manifestation of it, and reducing the cosine similarity to such a
+projection may not be sufficient to eliminate bias. For instance,
+``math'' and ``delicate'' might be equally similar to a pair of opposed
+explicitly gendered words (\emph{she}, \emph{he}), while being closer to
+quite different stereotypical attribute words (such as \emph{scientific}
+or \emph{caring}). Further, it is observed in {[}5{]} that most word
+pairs retain similarity under debiasing meant to minimize
+projection-based bias.\footnote{In {[}1{]} another method which involves
+  analogies and their evaluations by human users on Mechanical Turk is
+  also used. We do not discuss this method in this paper, see its
+  criticism in {[}16{]}.}
+
+A measure of bias in word embeddings which does not employ similarities
+to pre-calculated protected-class-level directions, the Word Embedding
+Association Test (WEAT), has been proposed in {[}8{]}. The idea here is
+that the bias between two sets of target words, \(X\) and \(Y\) (we call
+them protected words), should be quantified in terms of the cosine
+similarity between the protected words and attribute words coming from
+two sets of stereotype attribute words, \(A\) and \(B\) (we will call
+them attributes). For instance, \(X\) might be a set of male names,
+\(Y\) a set of female names, \(A\) might contain stereotypically
+male-related, and \(B\) stereotypically female-related career words. The
+association difference for a particular word \(t\) (belonging to either
+\(X\) or \(Y\)) is:
 
@@
 \mathsf{s}(t,A,B) & = \frac{\sum_{a\in A}\mathsf{cos}(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}\mathsf{cos}(t,b)}{\vert B\vert}
-\end{align*}
-\normalsize
-\noindent then, the association difference between $A$ a $B$ is:
+\end{align*} \normalsize \noindent then, the association difference
+between \(A\) a \(B\) is:
 
@@
 
-\noindent The authors use it as a test statistic in some tests, and the assumption that $X$ and $Y$ are of the same size is necessary for the statistic to make sense.
-In the final measure (of effect size), WEAT, means are taken, which makes it insensitive to size differences between $X$ and $Y$:
-\begin{align}
+\noindent The authors use it as a test statistic in some tests, and the
+assumption that \(X\) and \(Y\) are of the same size is necessary for
+the statistic to make sense. In the final measure (of effect size),
+WEAT, means are taken, which makes it insensitive to size differences
+between \(X\) and \(Y\): \begin{align}
 \mathsf{weat}(A,B) & = \frac{
@@
 
- WEAT is inspired by  the Implicit Association Test (IAT) [@Nosek2002harvesting] used in psychology, and it uses almost the same word sets, allowing for a \emph{prima facie} sensible comparison with bias in humans.
-In  [@Caliskan2017semanticsBiases] the authors argue that significant biases---thus measured--- similar to the ones discovered by IAT can be discovered in word embeddings.
-In [@Lauscher2019multidimensional] the methodology is  extended to a multilingual and cross-lingual setting, arguing that using Euclidean distance instead of cosine similarity does not make much difference, while the bias effects vary greatly across embedding models (interestingly, with social media-text trained embeddings being less biased than those based on Wikipedia).
- A similar methodology is employed in [@Garg2018years].
-The authors employ word embeddings trained on corpora from different decades to study the shifts in various biases.^[Strictly speaking, these authors use Euclidean distances and their differences, but the way they take averages and averages thereof is analogous, and so what we will have to say about pre-averaging leading to false confidence applies to this methodology as well.]
-
-<!--  For instance, to compute the occupational embeddings bias for women the authors first compute the average vector of vector embeddings of words that represent women (e.g.
-\emph{she}, and \emph{female}), then calculate the Euclidean distance between this mean vector and words for occupations.
-Then they take the mean of these distances and subtract from it the analogously obtained mean for the average vector of vector embeddings of words that represent men.
-Formally they take the relative norm distance between $X$ and $Y$ to be: -->
-
-<!-- \vspace{-2mm} -->
-
-<!-- \footnotesize -->
-
-<!-- The effect size is computed by normalizing the difference in means as follows: -->
-
-<!-- \begin{align} -->
-<!-- \textsf{relative norm distance} & = \sum_{v_m\in M} \vert \vert v_m - v_X\vert \vert_2 - \vert v_m - v_Y\vert \vert_2 -->
-<!-- \end{align} -->
-
-<!-- \normalsize  -->
-
-<!-- \noindent where the norm used is Euclidean, and $v_X$ and $x_Y$ are average vectors for sets $X$ and $Y$ respectively.  -->
-<!-- ] -->
-
-<!-- \vspace{-2mm} -->
-
-WEAT, however, has been developed to investigate biases corresponding to a pair of supposedly opposing stereotypes, and so the question arises as to how generalize the measure to contexts in which biases with respect to more than two stereotypical groups are to be measured.
-Such a generalization can be found in [@Manzini2019blackToCriminal].
-The authors introduce Mean Average Cosine distance (MAC) as a measure of bias.
-Let $T = \{t_1, \dots, t_k\}$ be a set of protected words, and let each $A_j\in \mathcal{A}$ be a set of attributes stereotypically associated with a protected word where $\mathcal{A}$.
-For instance, when biases related to religion are to be investigated, they use a dataset of the  format illustrated in Table \ref{tab:religionOriginal}. The measure is defined as follows:
-
-```{r religionTableHeadEarly,echo=FALSE,eval=FALSE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
-religion <- read.csv("datasets/outdated/religionReddit.csv")[-c(1,6,7)]
-colnames(religion) <- c("protected words (T)","attributes","atrribute set (A~j)",
-                        "cosine distance")
-
-religionOriginal <- religion[!(religion$`atrtibute set (A~j)` %in% c("neutral","human")),]
-
-set.seed(22)
-religionSample <- sample_n(religionOriginal, size = 15)
-religionSample %>%  kable(format = "latex",booktabs=T,
-                          linesep = "",  escape = FALSE, 
-                          caption = "Sample 12 rows of the religion dataset.") %>%
-  kable_styling(latex_options=c("striped"))
-```
+WEAT is inspired by the Implicit Association Test (IAT) {[}17{]} used in
+psychology, and it uses almost the same word sets, allowing for a
+\emph{prima facie} sensible comparison with bias in humans. In {[}8{]}
+the authors argue that significant biases---thus measured--- similar to
+the ones discovered by IAT can be discovered in word embeddings. In
+{[}11{]} the methodology is extended to a multilingual and cross-lingual
+setting, arguing that using Euclidean distance instead of cosine
+similarity does not make much difference, while the bias effects vary
+greatly across embedding models (interestingly, with social media-text
+trained embeddings being less biased than those based on Wikipedia). A
+similar methodology is employed in {[}4{]}. The authors employ word
+embeddings trained on corpora from different decades to study the shifts
+in various biases.\footnote{Strictly speaking, these authors use
+  Euclidean distances and their differences, but the way they take
+  averages and averages thereof is analogous, and so what we will have
+  to say about pre-averaging leading to false confidence applies to this
+  methodology as well.}
+
+WEAT, however, has been developed to investigate biases corresponding to
+a pair of supposedly opposing stereotypes, and so the question arises as
+to how generalize the measure to contexts in which biases with respect
+to more than two stereotypical groups are to be measured. Such a
+generalization can be found in {[}12{]}. The authors introduce Mean
+Average Cosine distance (MAC) as a measure of bias. Let
+\(T = \{t_1, \dots, t_k\}\) be a set of protected words, and let each
+\(A_j\in \mathcal{A}\) be a set of attributes stereotypically associated
+with a protected word where \(\mathcal{A}\). For instance, when biases
+related to religion are to be investigated, they use a dataset of the
+format illustrated in Table \ref{tab:religionOriginal}. The measure is
+defined as follows:
 
@@
 
-\noindent That is, for each  protected word $t\in T$, and each attribute set $A_j$, they first take the mean of distances for this protected word and all attributes in a given attribute class, and then take the mean of thus obtained means for all the protected words and all the protected classes.^[The authors'  code are in their github repository, [https://github.com/TManzini/DebiasMulticlassWordEmbedding](https://github.com/TManzini/DebiasMulticlassWordEmbedding).]
-
-Having introduced the measures,  first, we will introduce a selection of general problems with this approach, and then we will move on to more specific but important problems related to the fact that the measures take  averages and averages of averages.
-Once this is done, we move to the development of our Bayesian alternative and the presentation of its deployment.
-
-<!-- WEAT will be useful in our criticism of the statistical methods involved, as it is simpler, so the explanation and visualizations will be more transparent (and \emph{mutatis mutandis} this criticism will apply to MAC as well), and MAC will be useful in the development of our alternative method as its range of applicability is the widest.
--->
-
-# General methodological problems with cosine-based measures of bias
-
-Consider the MAC results from Table REF.
-Are the initial MAC values lower than 1 indicative of the presence of bias?
-Of course, thinking abstractly, 1 is the ideal distance for unrelated words.
-But clearly, there is variation in distances, which might lead to non-biased lists also having MAC smaller than 1.
-How much smaller?
-While the original paper did not employ control lists for comparison in the paper.
-To obtain a more realistic gauge on how to understand MAC values (and in further development of our approach), we introduced control word lists.
-One of them is a list of words we intuitively considered neutral (see REF in the APPENDIX).
-Moreover, it might be the case that words that have to do with human activities are systematically closer to the protected words than merely neutral words, for which reason we also use a second list with intuitively non-stereotypical human attributes (see REF in the APPENDIX).
-Another important observation is that in their calculations the authors do not distinguish whether a given  attribute is associated with a given protected word (e.g.
-protected word \emph{jew} and the attribute \emph{greedy}) or whether it is associated with a different stereotype from the same list (e.g.
-protected word \emph{quran} and the attribute \emph{greedy}), simply averaging across all such groups.
-
-In Figure  \ref{fig:empirical} We look at the empirical distributions, while paying attention to such divisions.
-Horizontal lines represent 1-MAC values that the authors considered indicative of bias for stereotypes corresponding to  given  word lists.
-For instance, in religion, MAC was .859, so we plot $0\pm (1-.859)\approx .14$ lines around similarity = 0 (that is, distance = 1).
-Notice that most distributions are quite wide, and the proportions of even neutral or human neutral words with similarities higher than the value of 1-MAC deserving debiasing according to the authors is quite high.
-
-```{r cosineDistributions6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%"}
-
-genderGoogle <- read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1]
-genderReddit <-  read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1]
-genderGlove <- read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1]
-
-raceGlove <-  read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1]
-raceGoogle <- read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1]
-raceReddit <- read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1]
-
-religionGlove <- read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1]
-religionGoogle <- read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1]
-religionReddit <- read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1]
-
-debiasedGenderReddit <- read.csv("datasets/macWeatDatasets/debiased_gender_reddit.csv")[-1]
-debiasedRaceReddit <- read.csv("datasets/macWeatDatasets/debiased_race_reddit.csv")[-1]
-debiasedReligionReddit <- read.csv("datasets/macWeatDatasets/debiased_religion_reddit.csv")[-1]
-
-violinReligionReddit <- ggplot(religionReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
-  geom_violin(alpha = .5)+coord_flip()+
-  geom_vline(xintercept = .14) + geom_hline(yintercept = -.14, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
-  labs(title =  "Religion (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
-  geom_jitter(size = 0.5, alpha = .2, width=0.08)+
-  stat_summary(fun.data = "mean_cl_boot", fun.args = list(
-    conf.int = .95), geom = "pointrange",
-               colour = "black", size  = .2  )+
-  annotate(geom = "label", x = 4.2, y = 0, label = "56%")+
-  annotate(geom = "label", x = 3.2, y = 0, label = "55%")+
-  annotate(geom = "label", x = 2.2, y = 0, label = "43%")+
-  annotate(geom = "label", x = 1.2, y = 0, label = "36%")+
-  theme(plot.title.position = "plot",legend.position="top")
-
-violinGenderReddit <- ggplot(genderReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
-  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
-  geom_vline(xintercept = .38) + geom_hline(yintercept = -.38, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .38, color = "grey")+
-  ggtitle("Gender: empirical distribution of cosine similarity")+
-  stat_summary(fun.data = "mean_cl_boot", fun.args = list(
-    conf.int = .95), geom = "pointrange",
-               colour = "black", size  = .2  )+
-  annotate(geom = "label", x = 4.2, y = 0, label = "96%")+
-  annotate(geom = "label", x = 3.2, y = 0, label = "87%")+
-  annotate(geom = "label", x = 2.2, y = 0, label = "53%")+
-  annotate(geom = "label", x = 1.2, y = 0, label = "47%")+
-  labs(title =  "Gender (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
-  theme(plot.title.position = "plot",legend.position='none')
-
-violinRaceReddit <- ggplot(raceReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
-  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
-  geom_vline(xintercept = .11) + geom_hline(yintercept = -.11, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .11, color = "grey")+
-   labs(title =  "Race (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
-  stat_summary(fun.data = "mean_cl_boot", fun.args = list(
-    conf.int = .95), geom = "pointrange",
-               colour = "black", size  = .2  )+
-  annotate(geom = "label", x = 4.2, y = 0, label = "46%")+
-  annotate(geom = "label", x = 3.2, y = 0, label = "67%")+
-  annotate(geom = "label", x = 2.2, y = 0, label = "39%")+
-  annotate(geom = "label", x = 1.2, y = 0, label = "33%")+
-  theme(plot.title.position = "plot",legend.position='none')
-
-```
+\noindent That is, for each protected word \(t\in T\), and each
+attribute set \(A_j\), they first take the mean of distances for this
+protected word and all attributes in a given attribute class, and then
+take the mean of thus obtained means for all the protected words and all
+the protected classes.\footnote{The authors' code are in their github
+  repository,
+  \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding}.}
+
+Having introduced the measures, first, we will introduce a selection of
+general problems with this approach, and then we will move on to more
+specific but important problems related to the fact that the measures
+take averages and averages of averages. Once this is done, we move to
+the development of our Bayesian alternative and the presentation of its
+deployment.
+
+\hypertarget{general-methodological-problems-with-cosine-based-measures-of-bias}{%
+\section{General methodological problems with cosine-based measures of
+bias}\label{general-methodological-problems-with-cosine-based-measures-of-bias}}
+
+Consider the MAC results from Table REF. Are the initial MAC values
+lower than 1 indicative of the presence of bias? Of course, thinking
+abstractly, 1 is the ideal distance for unrelated words. But clearly,
+there is variation in distances, which might lead to non-biased lists
+also having MAC smaller than 1. How much smaller? While the original
+paper did not employ control lists for comparison in the paper. To
+obtain a more realistic gauge on how to understand MAC values (and in
+further development of our approach), we introduced control word lists.
+One of them is a list of words we intuitively considered neutral (see
+REF in the APPENDIX). Moreover, it might be the case that words that
+have to do with human activities are systematically closer to the
+protected words than merely neutral words, for which reason we also use
+a second list with intuitively non-stereotypical human attributes (see
+REF in the APPENDIX). Another important observation is that in their
+calculations the authors do not distinguish whether a given attribute is
+associated with a given protected word (e.g.~protected word \emph{jew}
+and the attribute \emph{greedy}) or whether it is associated with a
+different stereotype from the same list (e.g.~protected word
+\emph{quran} and the attribute \emph{greedy}), simply averaging across
+all such groups.
+
+In Figure \ref{fig:empirical} We look at the empirical distributions,
+while paying attention to such divisions. Horizontal lines represent
+1-MAC values that the authors considered indicative of bias for
+stereotypes corresponding to given word lists. For instance, in
+religion, MAC was .859, so we plot \(0\pm (1-.859)\approx .14\) lines
+around similarity = 0 (that is, distance = 1). Notice that most
+distributions are quite wide, and the proportions of even neutral or
+human neutral words with similarities higher than the value of 1-MAC
+deserving debiasing according to the authors is quite high.
+
 \pagebreak
@@
 \begin{figure}[H]
-```{r fig:cosineDistributions8b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", fig.height = 16}
-
-grid.arrange(violinReligionReddit, violinGenderReddit, violinRaceReddit, ncol = 1)
-
-```
+
+\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8b-1} \end{center}
 
@@
 
-One issue to consider is  the selection of attributes for bias measurement. The word lists used in the literature are often fairly small (5-50).
-
-While the papers do employ statistical tests to measure the uncertainty involved, we will later on argue that these method are not  proper for the goal at hand and  show that a more appropriate use of statistical methods leads to  estimates of uncertainty that are rather epistemologically pessimistic.
-\todo{sensitivity to choice of words}
+One issue to consider is the selection of attributes for bias
+measurement. The word lists used in the literature are often fairly
+small (5-50).
+
+While the papers do employ statistical tests to measure the uncertainty
+involved, we will later on argue that these method are not proper for
+the goal at hand and show that a more appropriate use of statistical
+methods leads to estimates of uncertainty that are rather
+epistemologically pessimistic. \todo{sensitivity to choice of words}
 
 Let's think about MAC, using the case of religion-related stereotypes.
-The full lists from [@Manzini2019blackToCriminal] can be found in \ref{appendix:manzini_word_lists}.
-In the original paper,  words from all three religions were compared against all of the stereotypes.
-One reason this is problematic is that  no distinction between cases in which the stereotype is associated with a given religion, as opposed to the situation in which it is associated with another one, is made.
-This is problematic, as  not all of the stereotypical words have to be considered as harmful for all of the religions.
-One should investigate the religions separately as some of them may have stronger harmful associations that others.
-
-The interpretation of the results is also a challenge.
-In [@Manzini2019blackToCriminal] we can find  summaries of  average cosine distances per group (such as gender, race, or religion).
-For instance, for religion, here is the relevant fragment of table:
-
-<!-- Religion Debiasing  | MAC | p-value
-------------- | ------------- | -------
-Biased        | 0.859 | N/A
-Hard Debiased | 0.934 | 3.006e-07
-Soft Debiased ($\lambda$ = 0.2) | 0.894 | 0.007 -->
+The full lists from {[}12{]} can be found in
+\ref{appendix:manzini_word_lists}. In the original paper, words from all
+three religions were compared against all of the stereotypes. One reason
+this is problematic is that no distinction between cases in which the
+stereotype is associated with a given religion, as opposed to the
+situation in which it is associated with another one, is made. This is
+problematic, as not all of the stereotypical words have to be considered
+as harmful for all of the religions. One should investigate the
+religions separately as some of them may have stronger harmful
+associations that others.
+
+The interpretation of the results is also a challenge. In {[}12{]} we
+can find summaries of average cosine distances per group (such as
+gender, race, or religion). For instance, for religion, here is the
+relevant fragment of table:
 
@@
 
-\noindent (MAC stand for mean average cosine similarity, although in reality the the table contains mean cosine distances).
-What may attract attention is the fact that the value of cosine distance in "Biased" category is already quite high (i.e.
-close to 1) even before debiasing.
-High cosine distance indicates low cosine similarity between values.
-One could think that average cosine similarity equal to approximately 0.141 is not significant enough to consider it as bias.
-However,  the authors still aim to mitigate such "bias"  to make the distance even larger.
-Methodologically the question is, on what basis is this small similarity still considered as a proof of the presence of bias, and whether these small changes are meaningful.
-
-The underlying problem here is that in the paper there is no  control group.
-One should also include control groups to have a way of comparing the  results for a supposedly stereotyped group with the results for sets of neutral or human-related neutral words.
-In our approach later on, we distinguish between stereotypes associated with a given group, stereotypes associated with different groups, and introduce control groups: neutral words and stereotype-free human predicates.
-
-<!-- One may consider the issue of the curse of dimensionality.
-In our case, the curse of dimensionality may take place when there is an increase in the volume of data that results in adding extra dimensions to the Euclidean space.
-As the number of features increases, it may be harder and harder to obtain useful information from the data using the available algorithms.
-Using cosine similarity in high dimensions in word embeddings may be prone to the curse of dimensionality.
-According to @Venkat2018Curse there are reasons to consider this phenomenon when searching for word similarities in higher dimensions.
-An experiment is conducted that aims at showing how the similarity values and variation change as the number of dimensions increases.
-The hypothesis made in the paper states that two things will happen as the number of dimensions increase.
-First, the effort required to measure cosine similarity will be greater, and two, the similarity between data will blur out and have less variation.
-The authors generate random points with increasing number of dimensions where each dimension of a data point is given a value between 0 and 1.
-Then they pick one vector at random from each dimension class and calculate the cosine similarity between the chosen vector and the rest of the data.
-Then they check how the variation of values changes as the number of dimensions increases.
-It seems like the more dimensions there are, the smaller the variance and therefore it is less obvious how to interpret the resulting cosine similarities.
-Maybe the scale should be adjusted to the number of dimensions and variance so that it still gives us sensible information about data.
--->
-
-## Metrics that pre-average are a bad guide
-
-In contrast, statistical intervals will help us decide whether a given cosine similarity is high enough to consider the words to be more similar than if we chose them at random.
-We will use highest posterior density intervals, in line with Bayesian methodology.
-
-Crucially, these approaches use  means of mean average cosine similarities to measure  similarity between protected word and harmful stereotypes.
-If one takes  a closer look at the individual values that are taken for the calculations, it turns out that there are quite a few  outliers and surprisingly dissimilar words.
-This problem becomes transparent when we examine the visualizations of the individual cosine distances, following the idea that one of the first steps in understanding data is to look at it.
-
-With such a method the uncertainty involved is not really considered which makes it even more difficult to give reasonable interpretations of the results.
-We propose the use of Bayesian method to obtain some understanding of the influence the uncertainty has on the interpretation of final results.
-
-\noindent
-$s(X,Y,A,B)$ is the statistic used in the significance test, and the $p$-value is obtained by bootstrapping: it is the frequency of $s(X_i,Y_i,A,B)>s(X,Y,A,B)$ for all equally sized partitions $X_i, Y_i$ of $X\cup Y$.
-The effect size is computed by normalizing the difference in means as follows:
+\noindent (MAC stand for mean average cosine similarity, although in
+reality the the table contains mean cosine distances). What may attract
+attention is the fact that the value of cosine distance in ``Biased''
+category is already quite high (i.e.~close to 1) even before debiasing.
+High cosine distance indicates low cosine similarity between values. One
+could think that average cosine similarity equal to approximately 0.141
+is not significant enough to consider it as bias. However, the authors
+still aim to mitigate such ``bias'' to make the distance even larger.
+Methodologically the question is, on what basis is this small similarity
+still considered as a proof of the presence of bias, and whether these
+small changes are meaningful.
+
+The underlying problem here is that in the paper there is no control
+group. One should also include control groups to have a way of comparing
+the results for a supposedly stereotyped group with the results for sets
+of neutral or human-related neutral words. In our approach later on, we
+distinguish between stereotypes associated with a given group,
+stereotypes associated with different groups, and introduce control
+groups: neutral words and stereotype-free human predicates.
+
+\hypertarget{metrics-that-pre-average-are-a-bad-guide}{%
+\subsection{Metrics that pre-average are a bad
+guide}\label{metrics-that-pre-average-are-a-bad-guide}}
+
+In contrast, statistical intervals will help us decide whether a given
+cosine similarity is high enough to consider the words to be more
+similar than if we chose them at random. We will use highest posterior
+density intervals, in line with Bayesian methodology.
+
+Crucially, these approaches use means of mean average cosine
+similarities to measure similarity between protected word and harmful
+stereotypes. If one takes a closer look at the individual values that
+are taken for the calculations, it turns out that there are quite a few
+outliers and surprisingly dissimilar words. This problem becomes
+transparent when we examine the visualizations of the individual cosine
+distances, following the idea that one of the first steps in
+understanding data is to look at it.
+
+With such a method the uncertainty involved is not really considered
+which makes it even more difficult to give reasonable interpretations of
+the results. We propose the use of Bayesian method to obtain some
+understanding of the influence the uncertainty has on the interpretation
+of final results.
+
+\noindent \(s(X,Y,A,B)\) is the statistic used in the significance test,
+and the \(p\)-value is obtained by bootstrapping: it is the frequency of
+\(s(X_i,Y_i,A,B)>s(X,Y,A,B)\) for all equally sized partitions
+\(X_i, Y_i\) of \(X\cup Y\). The effect size is computed by normalizing
+the difference in means as follows:
 
@@
 
-The t-tests they employ are run on average cosines used to calculate MAC.
-
-#  Pre-averaging, bootstrapping and testing with the null model
-
-The effect sizes reported by [@Caliskan2017semanticsBiases] for lists of words for which the embeddings are supposedly biased range from 2.06 to 1.81, and the reported $p$-values are in the range of $10^{-7}-10^{-2}$ with one exception for Math vs arts, where it is $.018$.
-The question is, are those results meaningful?
-
-One way to answer this question is to think in terms of null models.
-If the words actually are samples from two populations with equal mean, how often would we see effect sizes in this range?
-How often would we reach the $p$-values that the authors reported?
-
-We start our exploration with considering a list of 16 protected words, each class containing 8, with 16 attributes, 8 in each attribute class.
-In each particular sample we will investigate the consequence of the model being in fact null: we draw the cosine distances, every time, from the $\mathsf{Normal}(0,.08)$ distribution.
-$0.08$ is approximately the empirical standard deviation observed in fairly large samples of neutral words, and 16 is the sample size used in the WEAT7 word list, which is not much different from the other sample sizes in word lists used by [@Caliskan2017semanticsBiases]
-
-```{r CaliskanCalc2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
-vectorMean <- function(vector){ sum(vector)/length(vector)}
-vectorSd <- function(vector) {
-     vmean <- vectorMean(vector)
-     return (sqrt(sum((vector - vmean)^2)/(length(vector))) )
-}
-
-set.seed(10232169)
-count <- 16
-randomizeDataset <- function(count = 16, mean = 0, sd = 0.08) {
-    tlist <- list()
-    for (t in 1:count) {
-                  tlist[[t]] <- data.frame(t = rep(t,count/2), A = 1:(count/2),
-                  Avalue  = rnorm(count/2,mean,sd), B = 1:(count/2),
-                  Bvalue = rnorm(count/2,mean,sd))
-                  }
-
-  tDF <- do.call(rbind,tlist)
-  tDF$target <- as.factor(ifelse(tDF$t <= (count/2), "X", "Y"))
-  tDF$t <- as.factor(tDF$t)
-  tDF$A <- as.factor(tDF$A)
-  tDF$B <- as.factor(tDF$B)
-
-  return(tDF)
-}
-
-tDF <- randomizeDataset()
-
-svalues <- numeric(count)
-for (term in 1:count){
-  svalues[term] <- vectorMean(tDF[tDF$t== term,]$Avalue) - 
-    vectorMean(tDF[tDF$t== term,]$Bvalue)
-}
-
-sDF <- data.frame(term = 1:count, s = svalues, 
-                  attribute = as.factor(c(rep("X", (count/2)), rep("Y", (count/2)))))
-
-testListX <- 1:(count/2)
-testListY <- (count/2+1):16
-
-testStatistic <- function (listX, listY){
-  sum(sDF[listX,]$s) -  sum(sDF[listY,]$s)
-      }
-
-ourStatistic <- testStatistic(testListX, testListY)
-
-effectSize <- function (listX, listY){
-  ( vectorMean(sDF[listX,]$s) -  vectorMean(sDF[listY,]$s) ) /
-     vectorSd(sDF$s)
-}
-
-ourEffectSize <- effectSize (testListX, testListY)
-
-combinations <- as.data.frame(t(combn(1:count, (count/2))))
-start_time <- Sys.time()
-for (row in 1:nrow(combinations)){
-  rowCodes <- as.integer(combinations[row,])
-  remainderCodes <- as.integer(1:count)[ !(1:count %in% rowCodes)]
-  combinations$statistic[row] <- testStatistic(rowCodes, remainderCodes)
-  combinations$es[row] <- effectSize(rowCodes, remainderCodes)
-}
-
-plotStat <- ggplot(combinations) + geom_density(aes(x = statistic))+
-geom_vline(xintercept = ourStatistic, linewidth = .5, 
-           lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
-  theme(plot.title.position = "plot", axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank()) +
-  labs(title = "Boostrapped distribution of test statistics",
-       subtitle = "Random null model with N(0,.008) and n = 16")+
-   annotate("label", x = ourStatistic, y = 2,
-            label =  paste(round(ourStatistic,2),
-                           ", p=", 0.011, sep = ""), size = 3)+
-  xlim(-2,2)+xlab("test statistic")+ylab("density")
-
-plotEff <- ggplot(combinations) + geom_density(aes(x = es))+
-  xlab("effect size")+ylab("density")+
-  geom_vline(xintercept = ourEffectSize, linewidth = .5, 
-             lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
-  theme(plot.title.position = "plot", axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank()) +
-  labs(title = "Boostrapped distribution of effect sizes",
-       subtitle = "Random null model with N(0,.008) and n = 16")+
-  annotate("label", x = ourEffectSize, y = .5,
-           label = paste(round(ourEffectSize,2),
-                         ", p=", 0.0086, sep = "" ), 
-           size = 3)+xlim(-2,2)
-
-plotCaliskanDist <- ggplot(tDF) + 
-  geom_point(aes(x = Avalue, y = Bvalue, color = target, shape = target
-                 ), size = 3, alpha = .8)+
-  theme_tufte(base_size = 10)+
-  ggtitle("Cosine distances by word group in  our null dataset")+
-  theme(plot.title.position = "plot", 
-        legend.position = c(0.08, 0.89),
-        legend.background = element_rect(fill="lightblue", 
-         size=0.01, linetype="solid"))
-
-```
-
-So let's draw one sample of this type.
-We calculate the s-values, the test statistics and the effect sizes.
-Then, following the original methodology, we obtain bootstrapped distributions of the effect sizes and the test statistics by calculating the same for each possible equal split of the initial random data set.
-The observed test statistic is `r ourStatistic` and `r ourEffectSize`.
-The bootstrapped distributions of the test statistics and effect sizes are illustrated in Figure \ref{fig:caliskanCalc}.
-Quite notably both (two-sided) $p$ values are uneven and rather low.
-
-\begin{figure}[H]
-```{r fig:caliskanCalcB2B,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
-grid.arrange(plotStat,plotEff)
-```
+The t-tests they employ are run on average cosines used to calculate
+MAC.
+
+\hypertarget{pre-averaging-bootstrapping-and-testing-with-the-null-model}{%
+\section{Pre-averaging, bootstrapping and testing with the null
+model}\label{pre-averaging-bootstrapping-and-testing-with-the-null-model}}
+
+The effect sizes reported by {[}8{]} for lists of words for which the
+embeddings are supposedly biased range from 2.06 to 1.81, and the
+reported \(p\)-values are in the range of \(10^{-7}-10^{-2}\) with one
+exception for Math vs arts, where it is \(.018\). The question is, are
+those results meaningful?
+
+One way to answer this question is to think in terms of null models. If
+the words actually are samples from two populations with equal mean, how
+often would we see effect sizes in this range? How often would we reach
+the \(p\)-values that the authors reported?
+
+We start our exploration with considering a list of 16 protected words,
+each class containing 8, with 16 attributes, 8 in each attribute class.
+In each particular sample we will investigate the consequence of the
+model being in fact null: we draw the cosine distances, every time, from
+the \(\mathsf{Normal}(0,.08)\) distribution. \(0.08\) is approximately
+the empirical standard deviation observed in fairly large samples of
+neutral words, and 16 is the sample size used in the WEAT7 word list,
+which is not much different from the other sample sizes in word lists
+used by {[}8{]}
+
+So let's draw one sample of this type. We calculate the s-values, the
+test statistics and the effect sizes. Then, following the original
+methodology, we obtain bootstrapped distributions of the effect sizes
+and the test statistics by calculating the same for each possible equal
+split of the initial random data set. The observed test statistic is
+0.386222 and 1.2671948. The bootstrapped distributions of the test
+statistics and effect sizes are illustrated in Figure
+\ref{fig:caliskanCalc}. Quite notably both (two-sided) \(p\) values are
+uneven and rather low.
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanCalcB2B-1} \end{center}
 
@@
 
-At this point, we might think---right, so while drawing on the assumption of the null hypothesis we just stumbled into a data set that randomly happened to display strong bias.
-We decide to double-check this by visual inspection expecting exactly this: a strong, clearly visible bias (Figure \ref{fig:caliskanDistances}).
-
-\begin{figure}[H]
-```{r fig:caliskanDistC2B,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
-plotCaliskanDist 
-```
+At this point, we might think---right, so while drawing on the
+assumption of the null hypothesis we just stumbled into a data set that
+randomly happened to display strong bias. We decide to double-check this
+by visual inspection expecting exactly this: a strong, clearly visible
+bias (Figure \ref{fig:caliskanDistances}).
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanDistC2B-1} \end{center}
 
@@
 
-\noindent In fact, while there might be some outliers here and there, saying that a clear bias on which one group is systematically closer to $A$s than another is definitely a stretch. What happened?
-
-In the calculations of WEAT means are taken twice.
-The $s$-values themselves are means, and then means of $s$-values are compared between groups.
-The test statistic used itself is of this sort as well---while it compares sums, these are sums of sets of the same size, so they are really linear transformations of means of means.
-Statistical troubles start when we run statistical tests on sets of means, for at least two reasons  One, by pre-averaging data we throw away information about sample sizes.
-For the former point, think about proportions: 10 out of 20 and 2 out of 4 give the same mean, but you would obtain more information by making the former observation rather than by making the latter.
-And especially in this context, in which the word lists are not huge, sample sizes should matter.
-Two, when we pre-average, we remove variation, and so pre-averaging  tend to manufacture false confidence.
-Group means display less variation than the raw data points, and the standard deviation of a set of means of sets of means is bound to be lower than the original standard deviation in the row data.
-Now, if you calculate your effect size by dividing by the pre-averaged standard deviation, your quite likely to get something that looks like a strong effect size, but the results of your calculations might not track anything interesting.
-
-So let us think again about the question that we are  ultimately interested in.
-Are the $X$ terms systematically closer to (further from) the $A$ attributes ($B$ attributes) than the $Y$ words?
-But now let's use the raw data points to try to answer these questions.
-As the first stab, let us run two quick $t$-tests to gauge what the raw data illustrated in Figure \ref{fig:caliskanDistances} tell us.
-
-First, distances to $A$ attributes for $X$ words and $Y$ words.
-Whoa, the result is significant!
-The $p$-value is $0.02$.
-So sure, the sample is in some sense unusual.
-But the 95\% confidence interval for the difference in means is $[.0052 .061]$, clearly nothing that a reader would expect given that the calculated effect size seemed quite large.
-How about the $B$ attributes?
-Here the $p$-value is  $.22$ and the 95\% confidence interval is $[-0.03, .009]$, even less of a reason to think a bias is present.
-
-The difficulties are exacerbated by the fact that statistical tests are based on bootstrapping from a relatively small data sets, which is quite likely to underestimate the population variance.
-To make our point clear, let us avoid bootstrapping and work with the null generative model with  $\mathsf{Norm}(0,.08)$ for both word groups.
-We keep the sizes the same: we have eight protected words in each group, sixteen in total, and for each we randomly draw 8 distances from hypothetical $A$ attributes, and $8$ distances from hypothetical $B$ attributes.
-Calculate the test statistic and effect size the way [@Caliskan2017semanticsBiases] did.
-Do this 10000 times and look at what the distributions of these values are on the assumption of the null model with realistic empirically motivated raw data point standard deviation.
-
-```{r ourDistances,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
-
-s <- function (table){ mean(table$A) - mean(table$B)}
-
-set.seed(123456)
-
-biasesD0 <- numeric(10000)
-svaluesD0 <- numeric(10000)
-factorD0  <- numeric(10000)
-numeratorD0 <- numeric(10000)
-
-for(i in 1:10000){
-  t1 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t2 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t3 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t4 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t5 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t6 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t7 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t8 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-  t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
-
-factorD0[i] <- sd(c(s(t1),s(t2),s(t3),s(t4),s(t5),
-                s(t6),s(t7),s(t8),s(t9),             s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16)))
-
-numeratorD0[i] <-  mean(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),
-                s(t7),s(t8)) -    
-mean(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))
-
-svaluesD0[i] <-  sum(s(t1),s(t2),s(t3),s(t4),s(t5),
-                     s(t6),s(t7),s(t8)) -   sum(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))
-
-biasesD0[i] <- numeratorD0[i]/factorD0[i]
-}
-
-```
-
-```{r ourDistances3B,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
-
-dataPLOT <- data.frame(biasesD0, svaluesD0)
-
-ourESplot <- ggplot(dataPLOT )+geom_histogram(aes(x=biasesD0, y = ..density..), bins = 50, alpha = .6)+
-  theme_tufte(base_size = 10)+
-  labs(title="Distribution of effect sizes given a null model",
-       subtitle = "Simulation with Norm(0,0.8), n=10k")+ xlab("effect size")+
-    theme(plot.title.position = "plot", axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank()) +
-  geom_vline(xintercept = ourEffectSize, linewidth = .5,
-             lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
-  annotate("label", x = ourEffectSize, y = .06,
-           label =  paste(round(ourEffectSize,2),
-                          ", p=", 0.38, sep = ""), size = 3)
-
-ourStatPlot <- ggplot(dataPLOT )+ geom_histogram(aes(x=svaluesD0, y = ..density..), alpha = .6, bins = 50) +
-  theme_tufte(base_size = 10)+
-  labs(title="Distribution of test statistics given a null model",
-       subtitle = "Simulation with Norm(0,0.8), n=10k")+ xlab("test statistic")+
-    theme(plot.title.position = "plot", axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank()) +
-  geom_vline(xintercept = ourStatistic, linewidth = .5,
-             lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
-  annotate("label", x = ourStatistic, y = 1.2,
-           label =  paste(round(ourStatistic,2),
-                          ", p=", 0.01, sep = ""), size = 3)
-
-```
-
-\begin{figure}[H]
-```{r fig:ourDistancesPlot2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
-
-grid.arrange(ourESplot, ourStatPlot, ncol = 1)
-
-```
+\noindent In fact, while there might be some outliers here and there,
+saying that a clear bias on which one group is systematically closer to
+\(A\)s than another is definitely a stretch. What happened?
+
+In the calculations of WEAT means are taken twice. The \(s\)-values
+themselves are means, and then means of \(s\)-values are compared
+between groups. The test statistic used itself is of this sort as
+well---while it compares sums, these are sums of sets of the same size,
+so they are really linear transformations of means of means. Statistical
+troubles start when we run statistical tests on sets of means, for at
+least two reasons One, by pre-averaging data we throw away information
+about sample sizes. For the former point, think about proportions: 10
+out of 20 and 2 out of 4 give the same mean, but you would obtain more
+information by making the former observation rather than by making the
+latter. And especially in this context, in which the word lists are not
+huge, sample sizes should matter. Two, when we pre-average, we remove
+variation, and so pre-averaging tend to manufacture false confidence.
+Group means display less variation than the raw data points, and the
+standard deviation of a set of means of sets of means is bound to be
+lower than the original standard deviation in the row data. Now, if you
+calculate your effect size by dividing by the pre-averaged standard
+deviation, your quite likely to get something that looks like a strong
+effect size, but the results of your calculations might not track
+anything interesting.
+
+So let us think again about the question that we are ultimately
+interested in. Are the \(X\) terms systematically closer to (further
+from) the \(A\) attributes (\(B\) attributes) than the \(Y\) words? But
+now let's use the raw data points to try to answer these questions. As
+the first stab, let us run two quick \(t\)-tests to gauge what the raw
+data illustrated in Figure \ref{fig:caliskanDistances} tell us.
+
+First, distances to \(A\) attributes for \(X\) words and \(Y\) words.
+Whoa, the result is significant! The \(p\)-value is \(0.02\). So sure,
+the sample is in some sense unusual. But the 95\% confidence interval
+for the difference in means is \([.0052 .061]\), clearly nothing that a
+reader would expect given that the calculated effect size seemed quite
+large. How about the \(B\) attributes? Here the \(p\)-value is \(.22\)
+and the 95\% confidence interval is \([-0.03, .009]\), even less of a
+reason to think a bias is present.
+
+The difficulties are exacerbated by the fact that statistical tests are
+based on bootstrapping from a relatively small data sets, which is quite
+likely to underestimate the population variance. To make our point
+clear, let us avoid bootstrapping and work with the null generative
+model with \(\mathsf{Norm}(0,.08)\) for both word groups. We keep the
+sizes the same: we have eight protected words in each group, sixteen in
+total, and for each we randomly draw 8 distances from hypothetical \(A\)
+attributes, and \(8\) distances from hypothetical \(B\) attributes.
+Calculate the test statistic and effect size the way {[}8{]} did. Do
+this 10000 times and look at what the distributions of these values are
+on the assumption of the null model with realistic empirically motivated
+raw data point standard deviation.
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2-1} \end{center}
 
@@
 
-The first observation is that the supposedly large effect size we obtained is not that unusual even assuming a null model.
-Around 38\% of samples result in effect size at least as extreme.
-This illustrates the point that the effect size used does not track anything interesting.
-Second, the distribution of test statistics is much more narrow, which means that if we use it to calculate $p$-values, it is not too difficult to obtain a supposedly significant test statistic which nevertheless does not correspond to anything interesting happening in the data set.
-
-Where do we stand now?
-We have seen  that seemingly high effect sizes  might arise even if the underlying processes actually have the same mean.
-The uncertainty resulting from including the raw data point variance into considerations is more extensive than the one suggested by the low p-values obtained from taking means or means of means as data points.
-In this section, we discussed the performance of the WEAT measure, but since the [@Manzini2019blackToCriminal] one is a generalization thereof, including the method of running statistical tests on pre-averaged data, our remarks, \emph{mutatis nutandis}, apply.
-Moreover, a general problem is that studies such as [@Caliskan2017semanticsBiases] or [@Manzini2019blackToCriminal]  do not use  control groups, and this makes the interpretation of the analysis additionally difficult.
-
-What is the alternative?
-As we already emphasized: focusing on what the real underlying question is, and trying to answer it using a statistical analysis of the raw data.
-Moreover, since the data sets are not too large and since multiple evaluations are to be made, we will pursue this method from the Bayesian perspective.
-
-# Bayesian method
-
-## Introduction and model definition
-
-Bayesian data analysis takes prior probability distributions, a mathematical model structure and the data, and returns the posterior probability distributions over the parameters of interest, thus capturing our uncertainty about their actual values.
-One important difference between such  a result and the result of a classical statistical analysis is that classical confidence intervals (CIs) have a rather complicated and somewhat confusing interpretation, which has little to do with the posterior probability distribution.^[Here are a few usual problems.
-CIs are often mistakenly interpreted as providing  the probability that a resulting confidence interval contains the true value of a parameter.
-CIs bring confusion also with regard to  precision, it is a common mistake to interpret narrow intervals as the ones corresponding to a more precise knowledge.
-Another fallacy is to associate CIs with likelihood and stating that values within a given interval are more probable than the ones outside it.
-The theory of confidence intervals does not support the above interpretations.
-CIs should be plainly interpreted as a result of certain procedure (there are many ways to obtain CIs from a given set of data) that will in the long run contain the true value if the procedure is performed a fixed amount of times.
-For a nice survey and explanation of these misinterpretations, see  [@Morey2015confidenceFallacy].
-For a psychological study of the occurrence of such  misinterpretations, see REF.
-In this study, 120 researchers and 442 students were asked to assess the truth value of six false statements involving different interpretations of a CI.
-Both researchers and students endorsed, on average, more than three of these statements.] [@Hoekstra2014Misinterpretation]
-
-In fact, Bayesian highest posterior density intervals  (HPDIs, the narrowest intervals containing a certain ratio of the area under the curve) and CIs end up being numerically the same only if the prior probabilities are uniform.
-This illustrates that (1) classical analysis is unable to incorporate non-trivial priors, and (2) is therefore more susceptible to over-fitting, unless regularization (equivalent to a more straightforward Bayesian approach) is used.
-In contrast with CIs, the posterior distributions are easily interpretable and have direct relevance for the question at hand.
-Moreover, Bayesian data analysis is better at handling hierarchical models and small datasets, which is exactly what we will we dealing with.
-
-In a standard Bayesian analysis, the first step is to understand the data and select potential predictors and the outcome variable.
-Once the data is understood, the next step is to formulate a mathematical description of the generative model of the relationships between the predictors and the outcome variable.
-Prior distributions must then be chosen for the parameters used in the model.
-Next, Bayesian inference must be applied to find posterior probabilities over the possible parameter values.
-Finally, after finding the posterior probabilities, it is important to check how well the posterior predictions reflect the data with a posterior predictive check.
-In our analysis the outcome variable is  the cosine distances  between the protected words and attribute words.
-The predictor is a factor determining whether a given attribute word is a neutral word, a human predicate, is stereotypically associated with the protected word, or comes from a different stereotype connected with another protected word.
-The idea is, if bias is present in the embedding, distances to associated attribute words should be systematically lower than to other attribute words.
-
-Furthermore, conceptually there are two levels of analysis in our approach.
-On the one hand, we are interested in the general question of whether related attributes are systematically closer across the dataset.
-On the other hand, we are interested in a more fine-grained picture of the role of the predictor for particular protected words.
-Learning in hierarchical Bayesian models involves using Bayesian inference to update the parameters of the model.
-This update is based on the observed data, and estimates are made at different levels of the data hierarchy.
-We use hierarchical Bayesian models in which we simultaneously estimate parameters at the protected word level and at the global level, assuming that all lower-level parameters are drawn from global distributions.
-Such models can be thought of as incorporating adaptive regularization, which avoids overfitting and leads to improved estimates for unbalanced datasets (and the datasets we need to use are unbalanced).
-
-To be more specific, the underlying mathematical model is as follows. First, we assume that distances are normally distributed:
+The first observation is that the supposedly large effect size we
+obtained is not that unusual even assuming a null model. Around 38\% of
+samples result in effect size at least as extreme. This illustrates the
+point that the effect size used does not track anything interesting.
+Second, the distribution of test statistics is much more narrow, which
+means that if we use it to calculate \(p\)-values, it is not too
+difficult to obtain a supposedly significant test statistic which
+nevertheless does not correspond to anything interesting happening in
+the data set.
+
+Where do we stand now? We have seen that seemingly high effect sizes
+might arise even if the underlying processes actually have the same
+mean. The uncertainty resulting from including the raw data point
+variance into considerations is more extensive than the one suggested by
+the low p-values obtained from taking means or means of means as data
+points. In this section, we discussed the performance of the WEAT
+measure, but since the {[}12{]} one is a generalization thereof,
+including the method of running statistical tests on pre-averaged data,
+our remarks, \emph{mutatis nutandis}, apply. Moreover, a general problem
+is that studies such as {[}8{]} or {[}12{]} do not use control groups,
+and this makes the interpretation of the analysis additionally
+difficult.
+
+What is the alternative? As we already emphasized: focusing on what the
+real underlying question is, and trying to answer it using a statistical
+analysis of the raw data. Moreover, since the data sets are not too
+large and since multiple evaluations are to be made, we will pursue this
+method from the Bayesian perspective.
+
+\hypertarget{bayesian-method}{%
+\section{Bayesian method}\label{bayesian-method}}
+
+\hypertarget{introduction-and-model-definition}{%
+\subsection{Introduction and model
+definition}\label{introduction-and-model-definition}}
+
+Bayesian data analysis takes prior probability distributions, a
+mathematical model structure and the data, and returns the posterior
+probability distributions over the parameters of interest, thus
+capturing our uncertainty about their actual values. One important
+difference between such a result and the result of a classical
+statistical analysis is that classical confidence intervals (CIs) have a
+rather complicated and somewhat confusing interpretation, which has
+little to do with the posterior probability distribution.\footnote{Here
+  are a few usual problems. CIs are often mistakenly interpreted as
+  providing the probability that a resulting confidence interval
+  contains the true value of a parameter. CIs bring confusion also with
+  regard to precision, it is a common mistake to interpret narrow
+  intervals as the ones corresponding to a more precise knowledge.
+  Another fallacy is to associate CIs with likelihood and stating that
+  values within a given interval are more probable than the ones outside
+  it. The theory of confidence intervals does not support the above
+  interpretations. CIs should be plainly interpreted as a result of
+  certain procedure (there are many ways to obtain CIs from a given set
+  of data) that will in the long run contain the true value if the
+  procedure is performed a fixed amount of times. For a nice survey and
+  explanation of these misinterpretations, see {[}15{]}. For a
+  psychological study of the occurrence of such misinterpretations, see
+  REF. In this study, 120 researchers and 442 students were asked to
+  assess the truth value of six false statements involving different
+  interpretations of a CI. Both researchers and students endorsed, on
+  average, more than three of these statements.} {[}7{]}
+
+In fact, Bayesian highest posterior density intervals (HPDIs, the
+narrowest intervals containing a certain ratio of the area under the
+curve) and CIs end up being numerically the same only if the prior
+probabilities are uniform. This illustrates that (1) classical analysis
+is unable to incorporate non-trivial priors, and (2) is therefore more
+susceptible to over-fitting, unless regularization (equivalent to a more
+straightforward Bayesian approach) is used. In contrast with CIs, the
+posterior distributions are easily interpretable and have direct
+relevance for the question at hand. Moreover, Bayesian data analysis is
+better at handling hierarchical models and small datasets, which is
+exactly what we will we dealing with.
+
+In a standard Bayesian analysis, the first step is to understand the
+data and select potential predictors and the outcome variable. Once the
+data is understood, the next step is to formulate a mathematical
+description of the generative model of the relationships between the
+predictors and the outcome variable. Prior distributions must then be
+chosen for the parameters used in the model. Next, Bayesian inference
+must be applied to find posterior probabilities over the possible
+parameter values. Finally, after finding the posterior probabilities, it
+is important to check how well the posterior predictions reflect the
+data with a posterior predictive check. In our analysis the outcome
+variable is the cosine distances between the protected words and
+attribute words. The predictor is a factor determining whether a given
+attribute word is a neutral word, a human predicate, is stereotypically
+associated with the protected word, or comes from a different stereotype
+connected with another protected word. The idea is, if bias is present
+in the embedding, distances to associated attribute words should be
+systematically lower than to other attribute words.
+
+Furthermore, conceptually there are two levels of analysis in our
+approach. On the one hand, we are interested in the general question of
+whether related attributes are systematically closer across the dataset.
+On the other hand, we are interested in a more fine-grained picture of
+the role of the predictor for particular protected words. Learning in
+hierarchical Bayesian models involves using Bayesian inference to update
+the parameters of the model. This update is based on the observed data,
+and estimates are made at different levels of the data hierarchy. We use
+hierarchical Bayesian models in which we simultaneously estimate
+parameters at the protected word level and at the global level, assuming
+that all lower-level parameters are drawn from global distributions.
+Such models can be thought of as incorporating adaptive regularization,
+which avoids overfitting and leads to improved estimates for unbalanced
+datasets (and the datasets we need to use are unbalanced).
+
+To be more specific, the underlying mathematical model is as follows.
+First, we assume that distances are normally distributed:
 \begin{align*} \mathsf{distance_i} & \sim \mathsf{dnorm}(\mu_i,\sigma_i)
-\end{align*}
-\noindent Second, for each particular protected word $\mathsf{pw}$ there are four parameters to be estimated.
-Its mean distance to associated stereotypes $a[pw]$, its mean distance to attributes coming from different stereotypes, $d[pw]$, its mean distance to human attributes, $h[pw]$, and its mean distance to neutral attributes, $n[pw]$:
-\begin{align*}
+\end{align*} \noindent Second, for each particular protected word
+\(\mathsf{pw}\) there are four parameters to be estimated. Its mean
+distance to associated stereotypes \(a[pw]\), its mean distance to
+attributes coming from different stereotypes, \(d[pw]\), its mean
+distance to human attributes, \(h[pw]\), and its mean distance to
+neutral attributes, \(n[pw]\): \begin{align*}
 \mu_i & = d_{\mathsf{pw[i]}} \times \mathsf{different}_i  + a_{\mathsf{pw[i]}} \times \mathsf{associated}_i  + h_{\mathsf{pw[i]}} \times \mathsf{human}_i  + n_{\mathsf{pw[i]}}\times \mathsf{neutral}_i
-\end{align*}
-\noindent where $\mathsf{different}, \mathsf{associated},\mathsf{human}$ and $\mathsf{neutral}$ are binary variables.
-
-This completes our description of the underlying mechanism.
-Now the priors and the hierarchy.
-We assume all the $a$ parameters come from one distribution, that is normal around a higher-level parameter $\bar{a}$ and so on for the other three groups of parameters.
-That is, $a_{\mathsf{pw[i]}}$ is the average distance of a given particular protected word to attributes stereotypically associated with it, while $\bar{a}$ is the overall average distance of protected words to attributes associated with them     [@kruschke2015bayesian; @statrethinkingbook2020].
-\begin{align*}
+\end{align*} \noindent where
+\(\mathsf{different}, \mathsf{associated},\mathsf{human}\) and
+\(\mathsf{neutral}\) are binary variables.
+
+This completes our description of the underlying mechanism. Now the
+priors and the hierarchy. We assume all the \(a\) parameters come from
+one distribution, that is normal around a higher-level parameter
+\(\bar{a}\) and so on for the other three groups of parameters. That is,
+\(a_{\mathsf{pw[i]}}\) is the average distance of a given particular
+protected word to attributes stereotypically associated with it, while
+\(\bar{a}\) is the overall average distance of protected words to
+attributes associated with them {[}10,13{]}. \begin{align*}
 d_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{d}, \overline{\sigma_d}) \\
@@
 
-According to our priors, the group means  $\bar{a}$, $\bar{d}$, $\bar{h}$ and $\bar{n}$  all come from one normal distribution with mean equal to $1$ and standard deviation equal to $.3$.
-The standard deviations $\bar{\sigma_a}$, $\bar{\sigma_d}$, $\bar{\sigma_h}$ and $\bar{sigma_n}$ to be estimated, according to our prior,  come from one  distribution, exponential with rate parameter equal to $2$.
-Our priors are slightly skeptical.
-They do reflect our knowledge and intuition on the probable distribution of the cosine distances in the data.
-We know that the cosine distances lie in the range $0-2$, and we expect two randomly chosen  vectors  from the embedding to have rather small similarity, so we expect the distances to be centered around $1$.
-However, we use a rather wide standard deviation ($.3$) to easily account for cases where there is actually much higher similarity between two vectors (especially in cases where the embedding is supposed to be biased).
-Our priors for the standard deviations are also fairly weak.\footnote{To build a model one may use original \textit{Stan} which is a state-of-the-art platform for statistical modeling and high-performance statistical computation (https://mc-stan.org/).
+According to our priors, the group means \(\bar{a}\), \(\bar{d}\),
+\(\bar{h}\) and \(\bar{n}\) all come from one normal distribution with
+mean equal to \(1\) and standard deviation equal to \(.3\). The standard
+deviations \(\bar{\sigma_a}\), \(\bar{\sigma_d}\), \(\bar{\sigma_h}\)
+and \(\bar{sigma_n}\) to be estimated, according to our prior, come from
+one distribution, exponential with rate parameter equal to \(2\). Our
+priors are slightly skeptical. They do reflect our knowledge and
+intuition on the probable distribution of the cosine distances in the
+data. We know that the cosine distances lie in the range \(0-2\), and we
+expect two randomly chosen vectors from the embedding to have rather
+small similarity, so we expect the distances to be centered around
+\(1\). However, we use a rather wide standard deviation (\(.3\)) to
+easily account for cases where there is actually much higher similarity
+between two vectors (especially in cases where the embedding is supposed
+to be biased). Our priors for the standard deviations are also fairly
+weak.\footnote{To build a model one may use original \textit{Stan} which is a state-of-the-art platform for statistical modeling and high-performance statistical computation (https://mc-stan.org/).
 The simpler option is to use \textit{Rethinking} package from the \textit{Statistical rethinking} course.
@@
 
-## Posterior predictive check
-
-A posterior predictive check is a technique used to evaluate the fit of a Bayesian model by comparing its predictions with observed data.
-The underlying principle is to generate simulated data from the posterior distribution of the model parameters and compare them with the observed data.
-If the model is a good fit to the data, the simulated data should resemble the observed data.
-A posterior predictive check is crucial because it provides a way to assess the validity of a Bayesian model.
-This is important because a model that fits the observed data well may still fail to generalize to new data.
-In the figure below we can see that the model fits the the data well.
-
-```{r generatePosteriorPredictiveCheckA2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
-source("functions/posteriorCheckPlot.R")
-
-posteriorPlotsA <- list()
-
-for (i in 1:length(predictions)){
-  name <- names(predictions)[i]
-  posteriorPlotsA[[i]] <- posteriorCheckPlotA(predictions[[i]], name = name) 
-}
-```
-
-```{r generatePosteriorPredictiveCheckB6,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
-
-posteriorPlotsB <- list()
-
-for (i in 1:length(predictions)){
-  posteriorPlotsB[[i]] <- posteriorCheckPlotB(predictions[[i]], simulated[[i]],names(predictions)[i]) 
-}
-
-```
-
-\begin{figure}[H]
-```{r figposteriorPrCheckA6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", dpi = 300, warning = FALSE, message = FALSE}
-grid.arrange(posteriorPlotsA[[3]], posteriorPlotsB[[3]], ncol = 2)
-```
+\hypertarget{posterior-predictive-check}{%
+\subsection{Posterior predictive
+check}\label{posterior-predictive-check}}
+
+A posterior predictive check is a technique used to evaluate the fit of
+a Bayesian model by comparing its predictions with observed data. The
+underlying principle is to generate simulated data from the posterior
+distribution of the model parameters and compare them with the observed
+data. If the model is a good fit to the data, the simulated data should
+resemble the observed data. A posterior predictive check is crucial
+because it provides a way to assess the validity of a Bayesian model.
+This is important because a model that fits the observed data well may
+still fail to generalize to new data. In the figure below we can see
+that the model fits the the data well.
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/figposteriorPrCheckA6b-1} \end{center}
 \caption{Example of a posterior predictive check.
@@
 
-# Results and discussion
+\hypertarget{results-and-discussion}{%
+\section{Results and discussion}\label{results-and-discussion}}
 
@@
 \footnotesize
-```{r generatePlots2,echo=FALSE, eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
-source("functions/cleanDataset.R")
-source("functions/plotFromPrecis.R")
-
-weat7Google <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_7_google.csv")[,-1])
-precisWeat7Google <- readRDS("resultsDFs/Weat7GoogleDF.rds")
-resultsWeat7Google <- plotFromPrecis(precis = precisWeat7Google,
-                                     dataset = weat7Google,
-                                     list = "Weat 7",
-                                     embedding = "Google",
-                                     ylims = c(.75,1.1))
-
-weat7Glove <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_7_glove.csv")[,-1])
-precisWeat7Glove <- readRDS("resultsDFs/Weat7GloveDF.rds")
-resultsWeat7Glove <- plotFromPrecis(precis = precisWeat7Glove,
-                                     dataset = weat7Glove,
-                                     list = "Weat 7",
-                                     embedding = "Glove")
-
-weat7Reddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_7_reddit.csv")[,-1])
-precisWeat7Reddit <- readRDS("resultsDFs/Weat7RedditDF.rds")
-resultsWeat7Reddit <- plotFromPrecis(precis = precisWeat7Reddit,
-                                    dataset = weat7Reddit,
-                                    list = "Weat 7",
-                                    embedding = "Reddit", ylims = c(.5,1.3))
-
-weat1Google <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_1_google.csv")[,-1])
-precisWeat1Google <- readRDS("resultsDFs/Weat1GoogleDF.rds")
-resultsWeat1Google <- plotFromPrecis(precis = precisWeat1Google,
-                                     dataset = weat1Google,
-                                     list = "Weat 1",
-                                     embedding = "Google",
-                                     ylims = c(.8,1.05))
-
-weat1Glove <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_1_glove.csv")[,-1])
-precisWeat1Glove <- readRDS("resultsDFs/Weat1GloveDF.rds")
-resultsWeat1Glove <- plotFromPrecis(precis = precisWeat1Glove,
-                                     dataset = weat1Glove,
-                                     list = "Weat 1",
-                                     embedding = "Glove",
-                                     ylims = c(.75,1.2))
-
-weat1Reddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/weat_1_reddit.csv")[,-1])
-precisWeat1Reddit <- readRDS("resultsDFs/Weat1RedditDF.rds")
-resultsWeat1Reddit <- plotFromPrecis(precis = precisWeat1Reddit,
-                                     dataset = weat1Reddit,
-                                     list = "Weat 1",
-                                     embedding = "Reddit",
-                                      ylims = c(.5,1.1))
-
-religionGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1])
-precisReligionGoogle <- readRDS("resultsDFs/ReligionGoogleDF.rds")
-resultsReligionGoogle <- plotFromPrecis(precis = precisReligionGoogle,
-                                     dataset = religionGoogle,
-                                     list = "Religion",
-                                     embedding = "Google",
-                                      ylims = c(.65,1.05))
-
-religionGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1])
-precisReligionGlove <- readRDS("resultsDFs/ReligionGloveDF.rds")
-resultsReligionGlove <- plotFromPrecis(precis = precisReligionGlove,
-                                        dataset = religionGlove,
-                                        list = "Religion",
-                                        embedding = "Glove",
-                                        ylims = c(.7,1.1))
-
-religionReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1])
-precisReligionReddit <- readRDS("resultsDFs/ReligionRedditDF.rds")
-resultsReligionReddit <- plotFromPrecis(precis = precisReligionReddit,
-                                       dataset = religionReddit,
-                                       list = "Religion",
-                                       embedding = "Reddit")
-
-raceGoogle <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1])
-precisRaceGoogle <- readRDS("resultsDFs/RaceGoogleDF.rds")
-resultsRaceGoogle <- plotFromPrecis(precis = precisRaceGoogle,
-                                        dataset = raceGoogle,
-                                        list = "Race",
-                                        embedding = "Google",
-                                     ylims = c(.8,1.1))
-
-raceGlove <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1])
-precisRaceGlove <- readRDS("resultsDFs/RaceGloveDF.rds")
-resultsRaceGlove <- plotFromPrecis(precis = precisRaceGlove,
-                                    dataset = raceGlove,
-                                    list = "Race",
-                                    embedding = "Glove",
-                                    ylims = c(.7,1.1))
-
-raceReddit <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1])
-precisRaceReddit <- readRDS("resultsDFs/RaceRedditDF.rds")
-resultsRaceReddit <- plotFromPrecis(precis = precisRaceReddit,
-                                    dataset = raceReddit,
-                                    list = "Race",
-                                    embedding = "Reddit")
-
-genderGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1])
-precisGenderGoogle <- readRDS("resultsDFs/GenderGoogleDF.rds")
-resultsGenderGoogle <- plotFromPrecis(precis = precisGenderGoogle,
-                                    dataset = genderGoogle,
-                                    list = "Gender",
-                                    embedding = "Google",
-                                     ylims = c(.6,1.05))
-
-genderGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1])
-precisGenderGlove <- readRDS("resultsDFs/GenderGloveDF.rds")
-resultsGenderGlove <- plotFromPrecis(precis = precisGenderGlove,
-                                      dataset = genderGlove,
-                                      list = "Gender",
-                                      embedding = "Glove",
-                                      ylims = c(.5,1.1))
-
-genderReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1])
-precisGenderReddit <- readRDS("resultsDFs/GenderRedditDF.rds")
-
-resultsGenderReddit <- plotFromPrecis(precis = precisGenderReddit,
-                                     dataset = genderReddit,
-                                     list = "Gender",
-                                     embedding = "Reddit",  ylims = c(.4,1.1))
-
-#now debiased
-debiasedReligionReddit <- read.csv("./datasets/macWeatDatasets/debiased_religion_reddit.csv")[,-1]
-debiasedRaceReddit <- read.csv("./datasets/macWeatDatasets/debiased_race_reddit.csv")[,-1]
-debiasedGenderReddit <- read.csv("./datasets/macWeatDatasets/debiased_gender_reddit.csv")[,-1]
-
-debiasedReligionReddit <- cleanDataset(debiasedReligionReddit)
-debiasedRaceReddit <- cleanDataset(debiasedRaceReddit)
-debiasedGenderReddit <- cleanDataset(debiasedGenderReddit)
-precisDebiasedReligionReddit <- readRDS("resultsDFs/debiasedReligionRedditDF.rds")
-precisDebiasedRaceReddit <- readRDS("resultsDFs/debiasedRaceRedditDF.rds")
-precisDebiasedGenderReddit <- readRDS("resultsDFs/debiasedGenderRedditDF.rds")
-
-resultsDebiasedReligionReddit <- plotFromPrecis(precis = precisDebiasedReligionReddit,
-                                        dataset = debiasedReligionReddit,
-                                        list = "Religion (MAC)",
-                                        embedding = "Reddit (debiased)"
-)
-
-#resultsDebiasedReligionReddit$plotJoint
-resultsDebiasedRaceReddit <- plotFromPrecis(precis = precisDebiasedRaceReddit,
-                                                dataset = debiasedRaceReddit,
-                                                list = "Race (MAC)",
-                                                embedding = "Reddit (debiased)"
-)
-
-#resultsDebiasedRaceReddit$plotJoint
-resultsDebiasedGenderReddit <- plotFromPrecis(precis = precisDebiasedGenderReddit,
-                                            dataset = debiasedGenderReddit,
-                                            list = "Gender (MAC)",
-                                            embedding = "Reddit (debiased)",
-                                            ylims = c(.4,1.1)
-)
-
-#resultsDebiasedGenderReddit$plotJoint
-```
-
-## Observations
-
-In brief, despite one-number metrics suggesting otherwise, our Bayesian analysis reveals that insofar as the short word lists usually used in related research projects are involved, there usually are no strong reasons to claim the presence of systematic bias.
-Moreover, comparison between the groups (including control word lists) leads to the conclusion that the effect sizes (that is, the absolute differences between cosine distances between groups) tend to be rather small, with few exceptions.
-Moreover, the choice of protected words is crucial --- as there is a lot of variance when it comes to the protected word-level analysis.
-
-In a bit more detail, the visualizations in  Appendix \ref{appendix:visualizations} show that the situation is more complicated than merely looking at one-number summaries might suggest.
-Note that the axes are sometimes in different scales to increase visibility.
-
-To start with, let us look at the association-type level coefficients (illustrated in the top parts of the plots).
-Depending on the corpus used and word class,  there is a large variety as to  the posterior densities.
-Quite aware of this being a crude approximation, let's compare the HPDIs and whether they overlap for different attribute groups.
-
-- In Weat 7 (Reddit)    there is no reason to think there are systematic differences between cosine distances (recall that words from Weat 7 were mostly not available in other embeddings).
-
-- In Weat 1 (Google, Glove and Reddit) associated words are somewhat closer, but the cosine distance differences from neutral words are very low, and surprisingly it is human attributes, not neutral predicates that are systematically the furthest.
-
-- In Religion (Google, Glove, Reddit) and Race (Google, Glove), the associated attributes are not systematically closer than attributes belonging to different stereotypes, and the difference from neutral and human predicates is rather low, if noticeable.
-The situation is interestingly different in Race (Reddit) where both human and neutral predicates are systematically further than associated and different attributes - but even then, there is no clear difference between associated and different attributes.
-
-- For Gender (Google, Glove), despite the superficially binary nature, associated and opposite attributes tend to be more or less in the same distances, much closer than neutral words (but not closer than human predicates in Glove).
-Reddit is an extreme example: both associated and opposite attributes are much closer than neutral and human (around .6 vs.
-.9), but even then, there seems to be no reason to think than cosine distances to associated predicates are much different from distances to opposite predicates.
-
-Moreover, when we look at particular protected words, the situation is even less straightforward.
-We will just go over a few notable examples, leaving the visual inspection of particular results for other protected words to the reader.
-One general phenomenon is that--as we already pointed out--the word lists are quite short, which contributes to large uncertainty involved in some cases.
-
-- For some protected words  the different attributes are somewhat closer than the associated attributes.
-
-- For some protected words, associated and different attributes are closer than neutral attributes, but so are human attributes.
-
-- In some cases, associated attributes are closer, but so are neutral and human predicates, which illustrates that just looking at average cosine similarity as compared to the theoretically expected value of 1, instead of running comparison to neutral and human attributes is misleading.
-
-- The only group of protected words where differences are noticeable at the protected word level is Gender-related words-- as in Gender (Google) and in Gender (Reddit) --- note however that in the latter, for some words the opposite attributes seem to be a little bit closer than the associated ones.
-
-## Rethinking debiasing
-
-These visualizations can be also handy when it comes to the investigation of the effect that debiasing has on the embedding space.
-Below one may see chosen pair depicting the difference in means with 89% highest posterior density intervals before and after applying debiasing.
-
-\begin{figure}[H]
-```{r fig:debiasedCosinePair2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, out.width="85%"}
-
-resultsGenderReddit$plotJoint
-
-resultsDebiasedGenderReddit$plotJoint
-
-```
+
+\hypertarget{observations}{%
+\subsection{Observations}\label{observations}}
+
+In brief, despite one-number metrics suggesting otherwise, our Bayesian
+analysis reveals that insofar as the short word lists usually used in
+related research projects are involved, there usually are no strong
+reasons to claim the presence of systematic bias. Moreover, comparison
+between the groups (including control word lists) leads to the
+conclusion that the effect sizes (that is, the absolute differences
+between cosine distances between groups) tend to be rather small, with
+few exceptions. Moreover, the choice of protected words is crucial ---
+as there is a lot of variance when it comes to the protected word-level
+analysis.
+
+In a bit more detail, the visualizations in Appendix
+\ref{appendix:visualizations} show that the situation is more
+complicated than merely looking at one-number summaries might suggest.
+Note that the axes are sometimes in different scales to increase
+visibility.
+
+To start with, let us look at the association-type level coefficients
+(illustrated in the top parts of the plots). Depending on the corpus
+used and word class, there is a large variety as to the posterior
+densities. Quite aware of this being a crude approximation, let's
+compare the HPDIs and whether they overlap for different attribute
+groups.
+
+\begin{itemize}
+\item
+  In Weat 7 (Reddit) there is no reason to think there are systematic
+  differences between cosine distances (recall that words from Weat 7
+  were mostly not available in other embeddings).
+\item
+  In Weat 1 (Google, Glove and Reddit) associated words are somewhat
+  closer, but the cosine distance differences from neutral words are
+  very low, and surprisingly it is human attributes, not neutral
+  predicates that are systematically the furthest.
+\item
+  In Religion (Google, Glove, Reddit) and Race (Google, Glove), the
+  associated attributes are not systematically closer than attributes
+  belonging to different stereotypes, and the difference from neutral
+  and human predicates is rather low, if noticeable. The situation is
+  interestingly different in Race (Reddit) where both human and neutral
+  predicates are systematically further than associated and different
+  attributes - but even then, there is no clear difference between
+  associated and different attributes.
+\item
+  For Gender (Google, Glove), despite the superficially binary nature,
+  associated and opposite attributes tend to be more or less in the same
+  distances, much closer than neutral words (but not closer than human
+  predicates in Glove). Reddit is an extreme example: both associated
+  and opposite attributes are much closer than neutral and human (around
+  .6 vs.~.9), but even then, there seems to be no reason to think than
+  cosine distances to associated predicates are much different from
+  distances to opposite predicates.
+\end{itemize}
+
+Moreover, when we look at particular protected words, the situation is
+even less straightforward. We will just go over a few notable examples,
+leaving the visual inspection of particular results for other protected
+words to the reader. One general phenomenon is that--as we already
+pointed out--the word lists are quite short, which contributes to large
+uncertainty involved in some cases.
+
+\begin{itemize}
+\item
+  For some protected words the different attributes are somewhat closer
+  than the associated attributes.
+\item
+  For some protected words, associated and different attributes are
+  closer than neutral attributes, but so are human attributes.
+\item
+  In some cases, associated attributes are closer, but so are neutral
+  and human predicates, which illustrates that just looking at average
+  cosine similarity as compared to the theoretically expected value of
+  1, instead of running comparison to neutral and human attributes is
+  misleading.
+\item
+  The only group of protected words where differences are noticeable at
+  the protected word level is Gender-related words-- as in Gender
+  (Google) and in Gender (Reddit) --- note however that in the latter,
+  for some words the opposite attributes seem to be a little bit closer
+  than the associated ones.
+\end{itemize}
+
+\hypertarget{rethinking-debiasing}{%
+\subsection{Rethinking debiasing}\label{rethinking-debiasing}}
+
+These visualizations can be also handy when it comes to the
+investigation of the effect that debiasing has on the embedding space.
+Below one may see chosen pair depicting the difference in means with
+89\% highest posterior density intervals before and after applying
+debiasing.
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2-1} \includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2-2} \end{center}
 \caption{Means with 89\% highest posterior density intervals for gender before and after debiasing.}
@@
 
-In Figure \ref{fig:empiricalDebiased} we look at the empirical distributions for the debiased embeddings.
-Comparing the results to the original embedding, one may notice that for the Religion group the neutral and human distribution has changed slightly.
-Before within the "correct" cosine similarity boundaries there were 56\% of neutral and 55\% of human word lists.
-After the debiasing the values changed to 59\% (for neutral) and 59\% (for human).
-The different and associated word lists were more influenced.
-The general shape of both distributions is less stretched.
-Before debiasing 43\% of the different word lists and 35\% of the associated word lists were within the accepted boundaries.
-After the embedding manipulation the percentage has increased for both lists to 63\%.
-Visualization for Gender group illustrates almost no change for the neutral and human word lists before and after debiasing.
-The values for different and associated word lists are also barely impacted by the embedding modification.
-In the Race group, the percentage within the boundaries for neutral and associated word lists have increased.
-The opposite happened for human and different word lists, where the percentage of "correct" cosine similarity dropped from 67\% to 55\% (human) and from 39\% to 36\% (different).
-
-```{r debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "75%"}
-
-debiasedGenderReddit <- read.csv("datasets/macWeatDatasets/debiased_gender_reddit.csv")[-1]
-debiasedRaceReddit <- read.csv("datasets/macWeatDatasets/debiased_race_reddit.csv")[-1]
-debiasedReligionReddit <- read.csv("datasets/macWeatDatasets/debiased_religion_reddit.csv")[-1]
-
-debiasedViolinReligionReddit <- ggplot(debiasedReligionReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
-  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
-  geom_vline(xintercept = .14) + geom_hline(yintercept = -.14, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
-  labs(title =  "Debiased Religion (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
-    geom_jitter(size = 0.5, alpha = .2, width=0.08)+
-  stat_summary(fun.data = "mean_cl_boot", fun.args = list(
-    conf.int = .95), geom = "pointrange",
-               colour = "black", size  = .2  )+
-  annotate(geom = "label", x = 4.15, y = 0, label = "59%")+
-  annotate(geom = "label", x = 3.15, y = 0, label = "59%")+
-  annotate(geom = "label", x = 2.15, y = 0, label = "63%")+
-  annotate(geom = "label", x = 1.15, y = 0, label = "63%")+
-  theme(plot.title.position = "plot",legend.position="top")
-
-debiasedViolinGenderReddit <- ggplot(debiasedGenderReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
-  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
-  geom_vline(xintercept = .38) + geom_hline(yintercept = -.38, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .38, color = "grey")+
-  labs(title =  "Debiased Gender (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
-    stat_summary(fun.data = "mean_cl_boot", fun.args = list(
-    conf.int = .95), geom = "pointrange",
-               colour = "black", size  = .2  )+
-  annotate(geom = "label", x = 4.15, y = 0, label = "96%")+
-  annotate(geom = "label", x = 3.15, y = 0, label = "92%")+
-  annotate(geom = "label", x = 2.15, y = 0, label = "51%")+
-  annotate(geom = "label", x = 1.15, y = 0, label = "51%")+
-  theme(plot.title.position = "plot",legend.position='none')
-
-debiasedViolinRaceReddit <- ggplot(debiasedRaceReddit, aes(x = connection, y = cosine_similarity, fill = connection, color = connection))+
-  geom_violin(alpha = .5)+coord_flip()+geom_jitter(size = 0.5, alpha = .2, width=0.08)+
-  geom_vline(xintercept = .14) + geom_hline(yintercept = -.14, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
-  labs(title =  "Debiased Race (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
-    stat_summary(fun.data = "mean_cl_boot", fun.args = list(
-    conf.int = .95), geom = "pointrange",
-               colour = "black", size  = .2  )+
-  annotate(geom = "label", x = 4.15, y = 0, label = "48%")+
-  annotate(geom = "label", x = 3.15, y = 0, label = "55%")+
-  annotate(geom = "label", x = 2.15, y = 0, label = "36%")+
-  annotate(geom = "label", x = 1.15, y = 0, label = "38%")+
-  theme(plot.title.position = "plot")+geom_point(size = 1.2, alpha = .3)+
-  theme(plot.title.position = "plot",legend.position='none')
-```
+In Figure \ref{fig:empiricalDebiased} we look at the empirical
+distributions for the debiased embeddings. Comparing the results to the
+original embedding, one may notice that for the Religion group the
+neutral and human distribution has changed slightly. Before within the
+``correct'' cosine similarity boundaries there were 56\% of neutral and
+55\% of human word lists. After the debiasing the values changed to 59\%
+(for neutral) and 59\% (for human). The different and associated word
+lists were more influenced. The general shape of both distributions is
+less stretched. Before debiasing 43\% of the different word lists and
+35\% of the associated word lists were within the accepted boundaries.
+After the embedding manipulation the percentage has increased for both
+lists to 63\%. Visualization for Gender group illustrates almost no
+change for the neutral and human word lists before and after debiasing.
+The values for different and associated word lists are also barely
+impacted by the embedding modification. In the Race group, the
+percentage within the boundaries for neutral and associated word lists
+have increased. The opposite happened for human and different word
+lists, where the percentage of ``correct'' cosine similarity dropped
+from 67\% to 55\% (human) and from 39\% to 36\% (different).
+
 \pagebreak
@@
 \begin{figure}[H]
-```{r fig:debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "70%", fig.height = 16}
-
-grid.arrange(violinReligionReddit, debiasedViolinReligionReddit, violinGenderReddit, debiasedViolinGenderReddit, violinRaceReddit, debiasedViolinRaceReddit, ncol = 2)
-
-```
+
+\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine-1} \end{center}
 \caption{Empirical distributions of debiased cosine similarities (1-distances) for three word lists used in  the original paper.}
@@
 
-# Related work and conclusions
-
-It has been suggested [@Ethayarajh2020Bernstein] to use Bernstein bounds to state the uncertainty of the sample bias in a form of confidence interval.
-The interpretation of this interval is as follows, 95\% confidence interval contains such a range of values, that one can be 95% certain that it contains the population mean.
-As we will prove later in the text, confidence intervals are quite problematic for several reasons, among others the confusing interpretability.
-In [@Ethayarajh2020Bernstein] the current problem of bias estimation was correctly identified.
-Indeed, the bias estimate should not be expressed as a single number without taking into account that the estimate is made on a sample of data and therefore has intrinsic uncertainty.
-However, we argue that Bernstein bounds do not provide the best solution to this problem.
-Applying their method to a popular WinoBias dataset leads to the conclusion that more than 11903 samples are needed to claim a 95\% confidence interval for a bias estimate.
-This amount exceeds existing datasets for bias measurement.
-We propose a more realistic method that can be applied to commonly used datasets to estimate the uncertainty of the bias measurement.
+\hypertarget{related-work-and-conclusions}{%
+\section{Related work and
+conclusions}\label{related-work-and-conclusions}}
+
+It has been suggested {[}2{]} to use Bernstein bounds to state the
+uncertainty of the sample bias in a form of confidence interval. The
+interpretation of this interval is as follows, 95\% confidence interval
+contains such a range of values, that one can be 95\% certain that it
+contains the population mean. As we will prove later in the text,
+confidence intervals are quite problematic for several reasons, among
+others the confusing interpretability. In {[}2{]} the current problem of
+bias estimation was correctly identified. Indeed, the bias estimate
+should not be expressed as a single number without taking into account
+that the estimate is made on a sample of data and therefore has
+intrinsic uncertainty. However, we argue that Bernstein bounds do not
+provide the best solution to this problem. Applying their method to a
+popular WinoBias dataset leads to the conclusion that more than 11903
+samples are needed to claim a 95\% confidence interval for a bias
+estimate. This amount exceeds existing datasets for bias measurement. We
+propose a more realistic method that can be applied to commonly used
+datasets to estimate the uncertainty of the bias measurement.
 
@@
 
-A Bayesian data analysis with hierarchical models of cosine distances between protected words, control group words, and stereotypical attributes  provides more modest and realistic assessment of the uncertainty involved.
-It reveals that much complexity is hidden when one instead chases single bias metrics present in the literature.
-
-After introducing the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of WEAT or MAC (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
-Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable.
-The tools that we propose, however, allow for a more fine-grained and multi-level evaluation of bias and debiasing in language models without losing modesty about the uncertainties involved.
+A Bayesian data analysis with hierarchical models of cosine distances
+between protected words, control group words, and stereotypical
+attributes provides more modest and realistic assessment of the
+uncertainty involved. It reveals that much complexity is hidden when one
+instead chases single bias metrics present in the literature.
+
+After introducing the method, we apply it to multiple word embeddings
+and results of supposed debiasing, putting forward some general
+observations that are not exactly in line with the usual picture painted
+in terms of WEAT or MAC (and the problem generalizes to any approach
+that focuses on chasing a single numeric metric): the word list sizes
+and sample sizes used in the studies are usually
+small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}
+posterior density intervals are fairly wide, often the differences
+between associated, different and neutral human predicates, are not very
+impressive. Also, a preliminary inspection suggests that the
+desirability of changes obtained by the usual debiasing methods is
+debatable. The tools that we propose, however, allow for a more
+fine-grained and multi-level evaluation of bias and debiasing in
+language models without losing modesty about the uncertainties involved.
 
@@
 \appendix
-# Appendix
-
-## Visualizations
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+\hypertarget{visualizations}{%
+\subsection{Visualizations}\label{visualizations}}
 
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsWeat7Google$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-1-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsWeat7Glove$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-2-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsWeat7Reddit$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-3-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsWeat1Google$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-4-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsWeat1Glove$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-5-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsWeat1Reddit$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-6-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsReligionGoogle$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-7-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsReligionGlove$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-8-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsReligionReddit$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-9-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsRaceGoogle$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-10-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsRaceGlove$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-11-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsRaceReddit$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-12-1} \end{center}
 \caption{dsds}
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsGenderGoogle$plotJoint
-```
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-13-1} \end{center}
 \caption{dsds}
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsGenderGlove$plotJoint
-```
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-14-1} \end{center}
 \caption{dsds}
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsGenderReddit$plotJoint
-```
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-15-1} \end{center}
 \caption{dsds}
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsDebiasedReligionReddit$plotJoint
-```
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-16-1} \end{center}
 \caption{dsds}
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsDebiasedRaceReddit$plotJoint
-```
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-17-1} \end{center}
 \caption{dsds}
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
-resultsDebiasedGenderReddit$plotJoint
-```
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-18-1} \end{center}
 \caption{dsds}
@@
 
-<!-- ## Examples of WEAT and MAC calculations -->
-<!-- \label{appendix: calculations} -->
-
-## Posterior predictive checks
-
-```{r figposteriorPrCheckAppendixa,echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "85%", out.height= "100%", dpi = 300, warning = FALSE, message = FALSE}
-grid.arrange(posteriorCheckPlotA(predictions[[1]], name = names(predictions)[1]),
-             posteriorCheckPlotB(predictions[[1]], simulated[[1]],names(predictions)[1]),
-             posteriorCheckPlotA(predictions[[2]], name = names(predictions)[2]),
-             posteriorCheckPlotB(predictions[[2]], simulated[[2]],names(predictions)[2]),
-             ncol = 2
-             )
-
-             # , posteriorPlotsB[[1]],
-             # posteriorPlotsA[[2]], posteriorPlotsB[[2]],
-             # ncol = 2)
-```
-
-```{r figposteriorPrCheckAppendixb,echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "85%", out.height= "100%", dpi = 300, warning = FALSE, message = FALSE}
-grid.arrange(posteriorCheckPlotA(predictions[[4]], name = names(predictions)[4]),
-             posteriorCheckPlotB(predictions[[4]], simulated[[4]],names(predictions)[4]),
-             posteriorCheckPlotA(predictions[[5]], name = names(predictions)[5]),
-             posteriorCheckPlotB(predictions[[5]], simulated[[5]],names(predictions)[5]),
-             ncol = 2
-             )
-
-```
-
-```{r figposteriorPrCheckAppendixc,echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "87%", out.height= "100%", dpi = 300, warning = FALSE, message = FALSE}
-grid.arrange(posteriorCheckPlotA(predictions[[6]], name = names(predictions)[6]),
-             posteriorCheckPlotB(predictions[[6]], simulated[[6]],names(predictions)[6]),
-             posteriorCheckPlotA(predictions[[7]], name = names(predictions)[7]),
-             posteriorCheckPlotB(predictions[[7]], simulated[[7]],names(predictions)[7]),
-             ncol = 2
-             )
-```
-
-```{r figposteriorPrCheckAppendixd,echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "89%", out.height= "100%", dpi = 300, warning = FALSE, message = FALSE}
-grid.arrange(posteriorCheckPlotA(predictions[[8]], name = names(predictions)[8]),
-             posteriorCheckPlotB(predictions[[8]], simulated[[8]],names(predictions)[8]),
-             posteriorCheckPlotA(predictions[[9]], name = names(predictions)[9]),
-             posteriorCheckPlotB(predictions[[9]], simulated[[9]],names(predictions)[9]),
-             ncol = 2
-             )
-```
-
-## Word lists
-
-### Lists for [@Manzini2019blackToCriminal]:
-\label{appendix:manzini_word_lists}
-The lists are available here:
-
--  **Gender:** \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/gender_attributes_optm.json}
-
-- **Race:** \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/race_attributes_optm.json}
-
-- **Religion:** \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/religion_attributes_optm.json}
+\hypertarget{posterior-predictive-checks}{%
+\subsection{Posterior predictive
+checks}\label{posterior-predictive-checks}}
+
+\begin{center}\includegraphics[width=0.85\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixa-1} \end{center}
+
+\begin{center}\includegraphics[width=0.85\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixb-1} \end{center}
+
+\begin{center}\includegraphics[width=0.87\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixc-1} \end{center}
+
+\begin{center}\includegraphics[width=0.89\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixd-1} \end{center}
+
+\hypertarget{word-lists}{%
+\subsection{Word lists}\label{word-lists}}
+
+\hypertarget{lists-for-manzini2019blacktocriminal}{%
+\subsubsection{Lists for
+{[}12{]}:}\label{lists-for-manzini2019blacktocriminal}}
+
+\label{appendix:manzini_word_lists} The lists are available here:
+
+\begin{itemize}
+\item
+  \textbf{Gender:}
+  \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/gender_attributes_optm.json}
+\item
+  \textbf{Race:}
+  \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/race_attributes_optm.json}
+\item
+  \textbf{Religion:}
+  \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/religion_attributes_optm.json}
+\end{itemize}
 
@@
 
-- **Gender:**
-
-Pairs of protected attributes:
-		["he", "she"],
-		["his", "hers"],
-		["son", "daughter"],
-		["father", "mother"],
-		["male", "female"],
-		["boy", "girl"],
-		["uncle", "aunt"]
-
-Lists of stereotypes per gender:
-		man: ["manager", "executive", "doctor", "lawyer", "programmer", "scientist", "soldier", "supervisor", "rancher", "janitor", "firefighter", "officer"],
-		woman: ["secretary", "nurse", "clerk", "artist", "homemaker", "dancer", "singer", "librarian", "maid", "hairdresser", "stylist", "receptionist", "counselor"]
-		}
-
-- **Race:**
-
-Sets of protected attributes:
-		["black", "caucasian", "asian"],
-		["african", "caucasian", "asian"],
-		["black", "white", "asian"],
-		["africa", "america", "asia"],
-		["africa", "america", "china"],
-		["africa", "europe", "asia"]
-
-Lists of stereotypes per race:
-			"caucasian": ["manager", "executive", "redneck", "hillbilly", "leader", "farmer"],
-			"asian"    : ["doctor", "engineer", "laborer", "teacher"],
-			"black"    : ["slave", "musician", "runner", "criminal", "homeless"]
-
-- **Religion:**
-
-Sets of protected attributes:
-		["judaism", "christianity", "islam"],
-		["jew", "christian", "muslim"],
-    ["synagogue", "church", "mosque"],
-    ["torah", "bible", "quran"],
-    ["rabbi", "priest", "imam"]
-
-Lists of stereotypes per race:
-			"jew"       : ["greedy", "cheap", "hairy", "liberal"],
-    	"christian" : ["judgemental", "conservative", "familial"],
-    	"muslim"    : ["violent", "terrorist", "dirty", "uneducated"]
-
-### Our custom word lists
-
-- **Neutral:**
-
- ['ballpark', 'glitchy', 'billy', 'dallas', 'rip', 'called', 'outlooks',
-                 'floater', 'rattlesnake', 'exports',
-                 'recursion', 'shortfall', 'corrected', 'solutions', 'diagnostic', 'patently', 'flops', 'approx', 'percents',
-                 'lox', 'hamburger', 'engulfed', 'households', 'north', 'playtest', 'replayability', 'glottal',
-                 'parable', 'gingers', 'anachronism', 'organizing', 'reach', 'shtick', 'eleventh', 'cpu', 'ranked',
-                 'irreversibly', 'ponce', 'velociraptor', 'defects', 'puzzle', 'smasher', 'northside', 'heft', 'observation', 'rectum',
-                 'mystical', 'telltale', 'remnants', 'inquiry', 'indisputable', 'boatload', 'lessening', 'uselessness', 'observes', 'fictitious',
-                 'repatriation', 'duh', 'attic', 'schilling', 'charges', 'chatter', 'pad', 'smurfing', 'worthiness', 'definitive', 'neat', 'homogenized',
-                 'lexicon', 'nationalized', 'earpiece', 'specializations', 'lapse', 'concludes', 'weaving', 'apprentices', 'fri',
-                 'militias', 'inscriptions', 'gouda', 'lift', 'laboring', 'adaptive', 'lecture', 'hogging', 'thorne', 'fud', 'skews',
-                 'epistles', 'tagging', 'crud', 'two', 'rebalanced', 'payroll', 'damned', 'approve', 'reason', 'formally', 'releasing', 'muddled',
-                 'mineral', 'shied', 'capital', 'nodded', 'escrow', 'disconnecting', 'marshals',
-                 'winamp', 'forceful', 'lowes', 'sip', 'pencils', 'stomachs', 'goff', 'cg', 'backyard', 'uprooting', 'merging',
-                 'helpful', 'eid', 'trenchcoat', 'airlift', 'frothing', 'pulls', 'volta', 'guinness', 'viewership',
-                 'eruption', 'peeves', 'goat', 'goofy', 'disbanding', 'relented', 'ratings', 'disputed', 'vitamins', 'singled', 'hydroxide',
-                 'telegraphed', 'mercantile', 'headache', 'muppets', 'petal', 'arrange', 'donovan', 'scrutinized', 'spoil', 'examiner', 'ironed',
-                 'maia', 'condensation', 'receipt', 'solider', 'tattooing', 'encoded', 'compartmentalize', 'lain', 'gov',
-                 'printers', 'hiked', 'resentment', 'revisionism', 'tavern', 'backpacking', 'pestering', 'acknowledges',
-                 'testimonies', 'parlance', 'hallucinate', 'speeches', 'engaging', 'solder', 'perceptive', 'microbiology', 'reconnaissance', 'garlic',
-                 'neutrals', 'width', 'literaly', 'guild', 'despicable', 'dion', 'option', 'transistors',
-                 'chiropractic', 'tattered', 'consolidating', 'olds', 'garmin', 'shift', 'granted', 'intramural', 'allie', 'cylinders',
-                 'wishlist', 'crank', 'wrongly', 'workshop', 'yesterday', 'wooden', 'without', 'wheel', 'weather', 'watch', 'version', 'usually', 'twice',                      'tomato', 'ticket', 'text', 'switch', 'studio', 'stick', 'soup', 'sometimes', 'signal', 'prior', 'plant', 'photo', 'path', 'park', 'near',
-                 'menu', 'latter', 'grass', 'clock']
-
-- **Human-related:**
-
-['wear', 'walk', 'visitor', 'toy', 'tissue', 'throw', 'talk', 'sleep', 'eye',
-               'enjoy', 'blogger', 'character', 'candidate', 'breakfast', 'supper', 'dinner',
-               'eat', 'drink', "carry", "run", "cast", "ask", "awake", "ear", "nose", "lunch", "coalition",
-               "policies", "restaurant", "stood", "assumed", "attend", "swimming", "trip", "door", "determine", "gets",
-               "leg", "arrival", "translated", "eyes", "step", "whilst", "translation", "practices", "measure",
-               "storage", "window", "journey", "interested", "tries", "suggests", "allied", "cinema", "finding",
-               "restoration", "expression","visitors", "tell", "visiting",
-               "appointment", "adults", "bringing", "camera", "deaths", "filmed", "annually", "plane", "speak",
-               "meetings", "arm", "speaking", "touring", "weekend", "accept", "describe", "everyone", "ready",
-               "recovered", "birthday", "seeing", "steps", "indicate", "anyone", "youtube"]
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Gender:}
+\end{itemize}
+
+Pairs of protected attributes: {[}``he'', ``she''{]}, {[}``his'',
+``hers''{]}, {[}``son'', ``daughter''{]}, {[}``father'', ``mother''{]},
+{[}``male'', ``female''{]}, {[}``boy'', ``girl''{]}, {[}``uncle'',
+``aunt''{]}
+
+Lists of stereotypes per gender: man: {[}``manager'', ``executive'',
+``doctor'', ``lawyer'', ``programmer'', ``scientist'', ``soldier'',
+``supervisor'', ``rancher'', ``janitor'', ``firefighter'',
+``officer''{]}, woman: {[}``secretary'', ``nurse'', ``clerk'',
+``artist'', ``homemaker'', ``dancer'', ``singer'', ``librarian'',
+``maid'', ``hairdresser'', ``stylist'', ``receptionist'',
+``counselor''{]} \}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Race:}
+\end{itemize}
+
+Sets of protected attributes: {[}``black'', ``caucasian'', ``asian''{]},
+{[}``african'', ``caucasian'', ``asian''{]}, {[}``black'', ``white'',
+``asian''{]}, {[}``africa'', ``america'', ``asia''{]}, {[}``africa'',
+``america'', ``china''{]}, {[}``africa'', ``europe'', ``asia''{]}
+
+Lists of stereotypes per race: ``caucasian'': {[}``manager'',
+``executive'', ``redneck'', ``hillbilly'', ``leader'', ``farmer''{]},
+``asian'' : {[}``doctor'', ``engineer'', ``laborer'', ``teacher''{]},
+``black'' : {[}``slave'', ``musician'', ``runner'', ``criminal'',
+``homeless''{]}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Religion:}
+\end{itemize}
+
+Sets of protected attributes: {[}``judaism'', ``christianity'',
+``islam''{]}, {[}``jew'', ``christian'', ``muslim''{]},
+{[}``synagogue'', ``church'', ``mosque''{]}, {[}``torah'', ``bible'',
+``quran''{]}, {[}``rabbi'', ``priest'', ``imam''{]}
+
+Lists of stereotypes per race: ``jew'' : {[}``greedy'', ``cheap'',
+``hairy'', ``liberal''{]}, ``christian'' : {[}``judgemental'',
+``conservative'', ``familial''{]}, ``muslim'' : {[}``violent'',
+``terrorist'', ``dirty'', ``uneducated''{]}
+
+\hypertarget{our-custom-word-lists}{%
+\subsubsection{Our custom word lists}\label{our-custom-word-lists}}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Neutral:}
+\end{itemize}
+
+{[}`ballpark', `glitchy', `billy', `dallas', `rip', `called',
+`outlooks', `floater', `rattlesnake', `exports', `recursion',
+`shortfall', `corrected', `solutions', `diagnostic', `patently',
+`flops', `approx', `percents', `lox', `hamburger', `engulfed',
+`households', `north', `playtest', `replayability', `glottal',
+`parable', `gingers', `anachronism', `organizing', `reach', `shtick',
+`eleventh', `cpu', `ranked', `irreversibly', `ponce', `velociraptor',
+`defects', `puzzle', `smasher', `northside', `heft', `observation',
+`rectum', `mystical', `telltale', `remnants', `inquiry', `indisputable',
+`boatload', `lessening', `uselessness', `observes', `fictitious',
+`repatriation', `duh', `attic', `schilling', `charges', `chatter',
+`pad', `smurfing', `worthiness', `definitive', `neat', `homogenized',
+`lexicon', `nationalized', `earpiece', `specializations', `lapse',
+`concludes', `weaving', `apprentices', `fri', `militias',
+`inscriptions', `gouda', `lift', `laboring', `adaptive', `lecture',
+`hogging', `thorne', `fud', `skews', `epistles', `tagging', `crud',
+`two', `rebalanced', `payroll', `damned', `approve', `reason',
+`formally', `releasing', `muddled', `mineral', `shied', `capital',
+`nodded', `escrow', `disconnecting', `marshals', `winamp', `forceful',
+`lowes', `sip', `pencils', `stomachs', `goff', `cg', `backyard',
+`uprooting', `merging', `helpful', `eid', `trenchcoat', `airlift',
+`frothing', `pulls', `volta', `guinness', `viewership', `eruption',
+`peeves', `goat', `goofy', `disbanding', `relented', `ratings',
+`disputed', `vitamins', `singled', `hydroxide', `telegraphed',
+`mercantile', `headache', `muppets', `petal', `arrange', `donovan',
+`scrutinized', `spoil', `examiner', `ironed', `maia', `condensation',
+`receipt', `solider', `tattooing', `encoded', `compartmentalize',
+`lain', `gov', `printers', `hiked', `resentment', `revisionism',
+`tavern', `backpacking', `pestering', `acknowledges', `testimonies',
+`parlance', `hallucinate', `speeches', `engaging', `solder',
+`perceptive', `microbiology', `reconnaissance', `garlic', `neutrals',
+`width', `literaly', `guild', `despicable', `dion', `option',
+`transistors', `chiropractic', `tattered', `consolidating', `olds',
+`garmin', `shift', `granted', `intramural', `allie', `cylinders',
+`wishlist', `crank', `wrongly', `workshop', `yesterday', `wooden',
+`without', `wheel', `weather', `watch', `version', `usually', `twice',
+`tomato', `ticket', `text', `switch', `studio', `stick', `soup',
+`sometimes', `signal', `prior', `plant', `photo', `path', `park',
+`near', `menu', `latter', `grass', `clock'{]}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Human-related:}
+\end{itemize}
+
+{[}`wear', `walk', `visitor', `toy', `tissue', `throw', `talk', `sleep',
+`eye', `enjoy', `blogger', `character', `candidate', `breakfast',
+`supper', `dinner', `eat', `drink', ``carry'', ``run'', ``cast'',
+``ask'', ``awake'', ``ear'', ``nose'', ``lunch'', ``coalition'',
+``policies'', ``restaurant'', ``stood'', ``assumed'', ``attend'',
+``swimming'', ``trip'', ``door'', ``determine'', ``gets'', ``leg'',
+``arrival'', ``translated'', ``eyes'', ``step'', ``whilst'',
+``translation'', ``practices'', ``measure'', ``storage'', ``window'',
+``journey'', ``interested'', ``tries'', ``suggests'', ``allied'',
+``cinema'', ``finding'', ``restoration'', ``expression'',``visitors'',
+``tell'', ``visiting'', ``appointment'', ``adults'', ``bringing'',
+``camera'', ``deaths'', ``filmed'', ``annually'', ``plane'', ``speak'',
+``meetings'', ``arm'', ``speaking'', ``touring'', ``weekend'',
+``accept'', ``describe'', ``everyone'', ``ready'', ``recovered'',
+``birthday'', ``seeing'', ``steps'', ``indicate'', ``anyone'',
+``youtube''{]}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Bolukbasi2016man}{}}%
+\CSLLeftMargin{{[}1{]} }%
+\CSLRightInline{Tolga Bolukbasi, Kai-Wei Chang, James Y. Zou, Venkatesh
+Saligrama, and Adam Kalai. 2016. Man is to computer programmer as woman
+is to homemaker? Debiasing word embeddings. \emph{CoRR} abs/1607.06520,
+(2016). Retrieved from \url{http://arxiv.org/abs/1607.06520}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ethayarajh2020Bernstein}{}}%
+\CSLLeftMargin{{[}2{]} }%
+\CSLRightInline{Kawin Ethayarajh. 2020. Is your classifier actually
+biased? Measuring fairness under uncertainty with bernstein bounds.
+\emph{CoRR} abs/2004.12332, (2020). Retrieved from
+\url{https://arxiv.org/abs/2004.12332}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Garg2017hundredYears}{}}%
+\CSLLeftMargin{{[}3{]} }%
+\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
+Zou. 2017. Word embeddings quantify 100 years of gender and ethnic
+stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
+(November 2017).
+DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Garg2018years}{}}%
+\CSLLeftMargin{{[}4{]} }%
+\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
+Zou. 2018. Word embeddings quantify 100 years of gender and ethnic
+stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
+16 (April 2018), E3635--E3644.
+DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gonen2019lipstick}{}}%
+\CSLLeftMargin{{[}5{]} }%
+\CSLRightInline{Hila Gonen and Yoav Goldberg. 2019. Lipstick on a pig:
+{D}ebiasing methods cover up systematic gender biases in word embeddings
+but do not remove them. In \emph{Proceedings of the 2019 conference of
+the north {A}merican chapter of the association for computational
+linguistics: Human language technologies, volume 1 (long and short
+papers)}, Association for Computational Linguistics, Minneapolis,
+Minnesota, 609--614.
+DOI:https://doi.org/\href{https://doi.org/10.18653/v1/N19-1061}{10.18653/v1/N19-1061}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gordon2012reporting}{}}%
+\CSLLeftMargin{{[}6{]} }%
+\CSLRightInline{Jonathan Gordon and Benjamin Durme. 2013. Reporting bias
+and knowledge acquisition. In \emph{AKBC 2013 - Proceedings of the 2013
+Workshop on Automated Knowledge Base Construction, Co-located with CIKM
+2013}, 25--30.
+DOI:https://doi.org/\href{https://doi.org/10.1145/2509558.2509563}{10.1145/2509558.2509563}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hoekstra2014Misinterpretation}{}}%
+\CSLLeftMargin{{[}7{]} }%
+\CSLRightInline{Rink Hoekstra, Richard D. Morey, Jeffrey N. Rouder, and
+Eric-Jan Wagenmakers. 2014. Robust misinterpretation of confidence
+intervals. \emph{Psychonomic Bulletin \& Review} 21, 5 (October 2014),
+1157--1164.
+DOI:https://doi.org/\href{https://doi.org/10.3758/s13423-013-0572-3}{10.3758/s13423-013-0572-3}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Caliskan2017semanticsBiases}{}}%
+\CSLLeftMargin{{[}8{]} }%
+\CSLRightInline{Aylin Caliskan Islam, Joanna J. Bryson, and Arvind
+Narayanan. 2016. Semantics derived automatically from language corpora
+necessarily contain human biases. \emph{CoRR} abs/1608.07187, (2016).
+Retrieved from \url{http://arxiv.org/abs/1608.07187}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-JohnsonValueFree}{}}%
+\CSLLeftMargin{{[}9{]} }%
+\CSLRightInline{Gabbrielle Johnson. forthcoming. Are algorithms
+value-free? Feminist theoretical virtues in machine learning.
+\emph{Journal Moral Philosophy} (forthcoming).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kruschke2015bayesian}{}}%
+\CSLLeftMargin{{[}10{]} }%
+\CSLRightInline{John Kruschke. 2015. \emph{Doing bayesian data analysis
+(second edition)}. Academic Press, Boston.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Lauscher2019multidimensional}{}}%
+\CSLLeftMargin{{[}11{]} }%
+\CSLRightInline{Anne Lauscher and Goran Glavas. 2019. Are we
+consistently biased? Multidimensional analysis of biases in
+distributional word vectors. \emph{CoRR} abs/1904.11783, (2019).
+Retrieved from \url{http://arxiv.org/abs/1904.11783}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Manzini2019blackToCriminal}{}}%
+\CSLLeftMargin{{[}12{]} }%
+\CSLRightInline{Thomas Manzini, Yao Chong Lim, Yulia Tsvetkov, and Alan
+W Black. 2019. Black is to criminal as caucasian is to police: Detecting
+and removing multiclass bias in word embeddings. Retrieved from
+\url{https://arxiv.org/abs/1904.04047}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-statrethinkingbook2020}{}}%
+\CSLLeftMargin{{[}13{]} }%
+\CSLRightInline{Richard McElreath. 2020. \emph{Statistical rethinking: A
+bayesian course with examples in r and stan, 2nd edition} (2nd ed.). CRC
+Press. Retrieved from
+\url{http://xcelab.net/rm/statistical-rethinking/}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Mikolov2013efficient}{}}%
+\CSLLeftMargin{{[}14{]} }%
+\CSLRightInline{Tomas Mikolov, Kai Chen, Greg Corrado, and Jeffrey Dean.
+2013. Efficient estimation of word representations in vector space.
+DOI:https://doi.org/\href{https://doi.org/10.48550/ARXIV.1301.3781}{10.48550/ARXIV.1301.3781}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Morey2015confidenceFallacy}{}}%
+\CSLLeftMargin{{[}15{]} }%
+\CSLRightInline{Richard Morey, Rink Hoekstra, Jeffrey Rouder, Michael
+Lee, and EJ Wagenmakers. 2015. The fallacy of placing confidence in
+confidence intervals. \emph{Psychonomic Bulletin \& Review} (September
+2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nissim2020fair}{}}%
+\CSLLeftMargin{{[}16{]} }%
+\CSLRightInline{Malvina Nissim, Rik van Noord, and Rob van der Goot.
+2020. Fair is better than sensational: Man is to doctor as woman is to
+doctor. \emph{Computational Linguistics} 46, 2 (June 2020), 487--497.
+DOI:https://doi.org/\href{https://doi.org/10.1162/coli_a_00379}{10.1162/coli\_a\_00379}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nosek2002harvesting}{}}%
+\CSLLeftMargin{{[}17{]} }%
+\CSLRightInline{Brian A. Nosek, Mahzarin R. Banaji, and Anthony G.
+Greenwald. 2002. Harvesting implicit group attitudes and beliefs from a
+demonstration web site. \emph{Group Dynamics: Theory, Research, and
+Practice} 6, 1 (2002), 101--115.
+DOI:https://doi.org/\href{https://doi.org/10.1037/1089-2699.6.1.101}{10.1037/1089-2699.6.1.101}}
+
+\end{CSLReferences}
+
+\end{document}

### 2023-05-04 functions/cleanDataset.R

@@ -1,0 +1,47 @@
+library(rethinking)
+library(ggplot2)
+library(ggthemes)
+library(tidyverse)
+library(gridExtra)
+library(grid)
+
+#dataset <- weat1Glove
+#nrow(dataset)
+
+cleanDataset <- function (dataset) {
+  
+  colnames(dataset) <- c("pw","word","stereotype", "distance", "similarity", "connection")
+  dataset$pw <- as.factor(dataset$pw)
+  dataset$word <- as.factor(dataset$word)
+  dataset$stereotype <- as.factor(dataset$stereotype)
+  dataset$connection <- as.factor(dataset$connection)
+  
+  
+  
+  ifelse(sum( dataset$similarity < -1 | dataset$distance < -1) != 0, 
+         print(
+           paste("WARNING:  ",  sum( dataset$similarity < -1 | dataset$distance < -1), 
+                 " out of ", nrow(dataset), " (", sum( dataset$similarity < -1 | dataset$distance < -1)/nrow(dataset) * 100, 
+                 "%) missing comparisons have been removed!", sep = "")
+         ),
+         print("No word removal needed.")
+  )
+  
+  
+  dataset <- dataset[  dataset$similarity >= -1 | dataset$distance >= -1,]
+  
+  dataset$pw <- droplevels(dataset$pw)
+  dataset$word <- droplevels(dataset$word)
+  dataset$stereotype <- droplevels(dataset$stereotype)
+  dataset$connection <- droplevels(dataset$connection)
+  
+  
+  dataset$associated <- as.integer(dataset$connection == "associated")
+  dataset$different <- as.integer(dataset$connection == "different")
+  dataset$human <- as.integer(dataset$connection == "human")
+  dataset$none <- as.integer(dataset$connection == "none")
+  dataset$pwi <- as.integer(dataset$pw)
+  
+
+  return(dataset)
+}

### 2023-05-04 functions/plotFromPrecis.R

@@ -1,0 +1,87 @@
+library(rethinking)
+library(ggplot2)
+library(ggthemes)
+library(tidyverse)
+library(gridExtra)
+library(grid)
+
+# precis
+
+
+plotFromPrecis <- function (precis, dataset, ylims = c(.5,1.2), list, embedding){
+  
+  
+  # precis = precisWeat7Google
+  # dataset = weat7Google
+  # list = "Weat 7"
+  # embedding = "Google"
+  
+  
+  results <- list()
+  DFoverall <-  precis[grep(c("^dbar|^abar|^hbar|^nbar"), rownames(precis)),]
+  DFindividual <- precis[grep(c("^d\\[|^a\\[|^h\\[|^n\\["), rownames(precis)),]
+  DFoverall$type <- c("different", "associated", "human", "none")
+  colnames(DFoverall) <- c("mean", "sd", "low", "high", "neff", "rhat", "type")
+  
+  nrow(DFindividual)
+    
+  dataset
+  
+  ifelse(nrow(DFindividual)/4 != nlevels(dataset$pw),
+         print("WARNING: dataset and precis don't match!"),
+         print("Dataset and precis match in size"))
+  
+  
+  DFindividual$pw <- as.factor(levels(dataset$pw)[rep((1:(nrow(DFindividual)/4)),4)])
+  DFindividual$type <- as.factor(c(rep("different", nrow(DFindividual)/4),rep("associated", 
+                                                                              nrow(DFindividual)/4),rep("human", nrow(DFindividual)/4), 
+                                   rep("none", nrow(DFindividual)/4)))
+  
+  
+  results$DFindividual <- DFindividual
+  results$DFoverall <- DFoverall
+  
+  
+  results$plotIndividual <- ggplot() + geom_point(data = DFindividual, aes( x = reorder(pw, desc(pw)), y = mean, color = type),
+                                  position = position_nudge((as.integer(DFindividual$type)-1)*0.15), size = 1)+
+    coord_flip()+theme_tufte()+expand_limits(x= c(-1, nrow(DFindividual)/4 +1))+
+    scale_color_manual(values = c("associated" ="orangered4", 
+                                  "different" = "chartreuse4",
+                                  "human" = "skyblue",
+                                  "none" = "grey")) +xlab("protected word")+
+    ylab("cosine distance")+
+    geom_segment(data = DFindividual,aes(x=reorder(pw, desc(pw)),xend=reorder(pw, desc(pw)), y  = `5.5%`, yend= `94.5%`,
[…]

### 2023-05-04 functions/posteriorCheckPlot.R

@@ -1,0 +1,59 @@
+library(ggplot2)
+library(ggthemes)
+library(gridExtra)
+library(grid)
+
+
+predictions <- readRDS(file = "resultsDFs/predictionsAll.RDS")
+simulated <- readRDS(file = "resultsDFs/simulatedAll.RDS")
+
+
+
+posteriorCheckPlotA <- function(dataset2, name){
+  rate89 <- mean(( dataset2$distance >= dataset2$low89 ) & (dataset2$distance <= dataset2$high89))
+  rate5 <- mean(( dataset2$distance >= dataset2$low5) & (dataset2$distance <= dataset2$high5))
+  
+  p1 <- ggplot(dataset2)+ geom_pointrange(aes(x = predictedMean,
+                                              y = distance, ymin = low89, ymax = high89,
+                                              color = connection),
+                                          size = .005, alpha = .2)+
+    scale_color_manual(values = c("associated" ="orangered4", 
+                                  "different" = "chartreuse4",
+                                  "human" = "skyblue",
+                                  "none" = "grey"))+
+    geom_abline(lty = 2, size =.2) +theme_tufte(base_size = 7)+
+    theme(plot.title.position = "plot",
+          legend.position =  c(.15,.9),
+          legend.title=element_blank(),
+          legend.key.height= unit(.2, 'cm'))+
+    xlab("predicted distance with 89% HPDI")+
+    ylab("actual distance")+
+    labs(title = paste("Predicted vs actual distances (", name , ")", sep = ""),
+         subtitle = paste(round(rate89,2)*100,
+                          "% within 89% HPDI, ",
+                          round(rate5,2)*100, "% within 50% HPDI", sep = ""))
+  
+  return(p1)
+}
+
+
+
+posteriorCheckPlotB <- function(dataset2,simulatedValues, name){
+  
+  distances <- c(dataset2$distance, c(simulatedValues))
+  
+  source <- c(rep("observed",length(dataset2$distance)),
+              rep("predicted", length(c(simulatedValues))))
+  densitiesData <- data.frame(distance = distances, 
+                              source = source)
+  
+  
+p2 <- ggplot(densitiesData)+geom_density(aes(x = distance, fill = source,
+                                             color = source), alpha = .4)+theme_tufte(base_size = 7)+
+  theme(legend.position =  c(.15,.9),
+        legend.title=element_blank(),
+        legend.key.height= unit(.2, 'cm'))+
+  ggtitle(paste("Densities (", name , ")", sep = ""))
+}
+
+

### 2023-05-14 functions/dagVis.R
Adds ALMOST working version

@@ -1,0 +1,58 @@
+library(dagitty)
+library(rethinking)
+
+dagBayesBias <- dagitty( 
+  "dag{
+  distances -> associated
+  distances -> different
+  distances -> human
+  distances -> neutral
+  
+  associated -> protectedWord1
+  associated -> protectedWord2
+  associated -> protectedWord3
+  
+  different -> protectedWord1
+  different -> protectedWord2
+  different -> protectedWord3
+  
+  human -> protectedWord1
+  human -> protectedWord2
+  human -> protectedWord3
+  
+  neutral -> protectedWord1
+  neutral -> protectedWord2
+  neutral -> protectedWord3
+  
+  protectedWord1 -> attribute1
+  protectedWord1 -> attribute2
+  protectedWord1 -> attribute3
+  protectedWord1 -> attribute4
+  
+  protectedWord2 -> attribute1
+  protectedWord2 -> attribute2
+  protectedWord2 -> attribute3
+  protectedWord2 -> attribute4
+  protectedWord3 -> attribute1
+  protectedWord3 -> attribute2
+  protectedWord3 -> attribute3
+  protectedWord3 -> attribute4
+  }
+  " 
+)
+
+coordinates(dagBayesBias) <- list(
+  x=c(
+    distances = 2.5, associated = 1, different = 2, human = 3, neutral = 4,
+    protectedWord1 = 1.5, protectedWord2 = 2.5, protectedWord3 = 3.5,
+    attribute1 = 1, attribute2 = 2, attribute3 = 3, attribute4 = 4
+  ) ,
+  y=c(
+    distances = .5, associated = 1, different = 1, human = 1, neutral = 1,
+    protectedWord1 = 2, protectedWord2 = 2, protectedWord3 = 2,
+    attribute1 = 3, attribute2 = 3, attribute3 = 3, attribute4 = 3
+  ) )
+
+
+drawdag(dagBayesBias)
+

### 2023-05-14 functions/generatePredictions.R

@@ -1,0 +1,97 @@
+
+source("functions/cleanDataset.R")
+source("functions/modelToPrediction.R")
+
+
+predictions <- list()
+
+
+
+genderGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1])
+genderReddit <-  cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1])
+genderGlove <- read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1]
+genderGlove <- cleanDataset(genderGlove)
+
+raceGlove <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1])
+raceGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1])
+raceReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1])
+
+religionGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1])
+religionGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1])
+religionReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1])
+
+
+
+modelList <- c(
+  "modelReligionGoogle.RDS", "modelReligionGlove.RDS", "modelReligionReddit.RDS",
+  "modelRaceReddit.RDS", "modelRaceGoogle.RDS", "modelRaceGlove.RDS", 
+  "modelGenderReddit.RDS", "modelGenderGoogle.RDS", "modelGenderGlove.RDS",
+ "modelWeat7Google.RDS", "modelWeat7Glove.RDS"
+)
+
+#"modelWeat1Reddit.RDS", #, "modelWeat1Glove.RDS
+# "modelWeat7Reddit.RDS",
+
+simulated <- list()
+
+
+model <- readRDS(file = "models/modelReligionGoogle.RDS")
+simulated$religionGoogle <- sim(model)
+
+predictions$religionGoogle <- modelToPredictions(model, religionGoogle, "religionGoogle")
+
+model <- readRDS(file = "models/modelReligionGlove.RDS")
+simulated$religionGlove <- sim(model)
+
[…]

### 2023-05-14 functions/modelToPrediction.R

@@ -1,0 +1,28 @@
+library(rethinking)
+
+
+modelToPredictions <- function(model, dataset, name){
+  predictions <- rethinking::sim(model )
+  predictedMean <- apply( predictions , 2 , mean )
+  predicted89 <- apply( predictions , 2 , HPDI, prob = .89 )
+  predicted5 <- apply( predictions , 2 , HPDI, prob = .5 )
+  
+  dataset2 <- dataset
+  dataset2$predictedMean <- predictedMean
+  
+  pr89 <- as.data.frame(t(predicted89))
+  pr5 <- as.data.frame(t(predicted5))
+  
+  colnames(pr89) <- c("low89", "high89")
+  colnames(pr5) <- c("low5", "high5")
+  
+  dataset2 <- cbind(dataset2, pr89, pr5)
+  
+  dataset2$set <- rep(name, nrow(dataset2))
+  
+  dataset2$within89 <-( dataset2$distance >= dataset2$low89 ) &
+                    (dataset2$distance <= dataset2$high89)
+  
+  
+  return(dataset2)
+}

### 2023-05-14 functions/visualisationTools.R

@@ -1,0 +1,64 @@
+library(ggplot2)
+library(ggthemes)
+
+#religion <- read.csv("datasets//religion_cosine_table.csv")
+
+#head(religion)
+
+
+#dataset <- religion
+#protected <- "judaism"
+
+set.seed(666)
+visualiseProtected <- function(dataset, protected){
+  bunch <- dataset[dataset$protectedWord == protected,]
+  bunch <- bunch[order(bunch$cosineDistance),]
+  bunch$xaxis <- rev(seq(0.01,sqrt(max(bunch$cosineDistance^2/40)),length.out = nrow(bunch)))
+  bunch$yaxis <-  sqrt(bunch$cosineDistance^2 - bunch$xaxis^2)
+  
+  #check if x and y yield the right distances
+  #check <- bunch[bunch$yaxis^2+bunch$xaxis^2 != bunch$cosine_distance^2,]
+  #approximate <- check$xaxis^2 + check$yaxis^2
+  #squaredDistance <- check$cosine_distance^2
+  #cbind(approximate,squaredDistance)
+  #ok, just rounding errors
+  
+  #closest to the right
+  
+  closestx <- bunch$xaxis[1]
+  closesty <- bunch$yaxis[1]
+  closestDistance <- bunch$cosineDistance[1]
+  
+  furthestx <- bunch$xaxis[nrow(bunch)]
+  furthesty <- bunch$yaxis[nrow(bunch)]
+  furthestDistance <- bunch$cosineDistance[nrow(bunch)]
+  
+  plot <- ggplot(bunch) + geom_point(aes(x=0, y=0), colour="black", size =10,alpha=0.02)+
+    geom_point(aes(x=xaxis,y=yaxis, color = connection, size = cosineSimilarity), alpha = 0.3)+ 
+    geom_segment(aes(x= 0, y = 0, xend = closestx,yend=closesty),alpha= 0.1, size = 0.2)+
+    geom_segment(aes(x= 0, y = 0, xend = furthestx,yend=furthesty),alpha=0.1,size = 0.2)+
+    annotate("text", x = 0.5*closestx, y = 0.5*closesty-0.05, label = round(closestDistance,3), size = 3)+
+    annotate("text", x = 1.5*furthestx, y = 0.5*furthesty, label = round(furthestDistance,3), size = 3)+
+    annotate("text", x = -0.02, y = -0.13, label = protected, size = 4)+
[…]

### 2023-05-14 paper/paperDraft.Rmd
Adds working version

@@
 classoption: dvipsnames,enabledeprecatedfontcommands
-fontsize: 10pt
+fontsize: 12pt
 documentclass: scrartcl
@@
 library(dplyr)
-library(kableExtra)
+library(dagitty)
 library(tidyverse)
 library(magrittr)
+library(kableExtra)
 library(ggpubr)
@@
 library(ggforce)
+library(rethinking)
 library(latex2exp)
@@
 Natural language processing (NLP) models are used to perform various language-related tasks such as providing email filters, intelligent assistants, search results,  translations, text analysis and so on.
-Such tools require input words to be represented by numbers.
-This can be done with word embeddings, which represent words as vectors of real numbers.
+Such tools require input words to be represented by  vectors of real numbers.
+This can usually be done with word embeddings.
 A common way to construct an embedding is to use  a large natural language corpus to train  a  neural network to assign vectors to words that optimize for co-occurrence prediction accuracy.
@@
 One example of this approach is [@Garg2017hundredYears].
-However, if our goal is to develop downstream tools that perform tasks that we care about well without further perpetuating  or exacerbating harmful stereotypes, we still have good reasons to try to minimize the negative impact, at least as long as this does not damage the overall performance of the tool.
+However, if our goal is to develop downstream tools that perform tasks that we care about without further perpetuating  or exacerbating harmful stereotypes, we still have good reasons to try to minimize the negative impact, at least as long as this does not damage the overall performance of the tool.
 Moreover, it is often not the case that the corpora mirror reality---to give a trivial example, heads are spoken of more often than kidneys, but this does not mean that kidneys occur much less often in reality than heads.
@@
 
-The use of projections in bias estimation has been criticized for instance in [@Gonen2019lipstick], where it is pointed out that while a higher average similarity to the  gender direction might be an indicator of bias with respect to a given  class of words, but it is only one possible manifestation of it, and reducing the cosine similarity to such a projection may not be sufficient to eliminate bias.
+The use of projections in bias estimation has been criticized for instance in [@Gonen2019lipstick], where it is pointed out that while a higher average similarity to the  gender direction might be an indicator of bias with respect to a given  class of words,  it is only one possible manifestation of it, and reducing the cosine similarity to such a projection may not be sufficient to eliminate bias.
 For instance, "math" and "delicate" might be equally similar to a pair of  opposed explicitly gendered words (\emph{she}, \emph{he}), while being closer to quite different stereotypical attribute words (such as \emph{scientific} or \emph{caring}).
@@
 
-\begin{align*}
+\begin{align}
+\label{eq:stAB}
 \mathsf{s}(t,A,B) & = \frac{\sum_{a\in A}\mathsf{cos}(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}\mathsf{cos}(t,b)}{\vert B\vert}
-\end{align*}
+\end{align}
 \normalsize
@@
 
-\begin{align*}
+\begin{align}
+\label{eq:sXYAB}
 \mathsf{s}(X,Y,A,B) & = \sum_{x\in X} \mathsf{s}(x,A,B) -  \sum_{y\in Y} \mathsf{s}(y,A,B)
-\end{align*}
+\end{align}
 
@@
 
-<!--  For instance, to compute the occupational embeddings bias for women the authors first compute the average vector of vector embeddings of words that represent women (e.g.
-\emph{she}, and \emph{female}), then calculate the Euclidean distance between this mean vector and words for occupations.
-Then they take the mean of these distances and subtract from it the analogously obtained mean for the average vector of vector embeddings of words that represent men.
-Formally they take the relative norm distance between $X$ and $Y$ to be: -->
-
-<!-- \vspace{-2mm} -->
-
-<!-- \footnotesize -->
-
-<!-- The effect size is computed by normalizing the difference in means as follows: -->
-
-<!-- \begin{align} -->
-<!-- \textsf{relative norm distance} & = \sum_{v_m\in M} \vert \vert v_m - v_X\vert \vert_2 - \vert v_m - v_Y\vert \vert_2 -->
-<!-- \end{align} -->
-
-<!-- \normalsize  -->
-
-<!-- \noindent where the norm used is Euclidean, and $v_X$ and $x_Y$ are average vectors for sets $X$ and $Y$ respectively.  -->
-<!-- ] -->
-
-<!-- \vspace{-2mm} -->
-
-WEAT, however, has been developed to investigate biases corresponding to a pair of supposedly opposing stereotypes, and so the question arises as to how generalize the measure to contexts in which biases with respect to more than two stereotypical groups are to be measured.
+\begin{figure}[H]
+```{r WEATexample,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "55%", warning = FALSE, message = FALSE}
+words <- c("man", "he", "boss", "architect","nurse","receptionist","woman","she")
+x <- c(.1,.1,.5,.5,.7,.7,1.2,1.2)
+y <- c(.4,.3,.3,.2,.5,.4,.4,.3)
+weat <- data.frame(words,x,y)
+
+ggplot(weat, aes(x=x,y=y, label = words))+geom_point()+geom_text(check_overlap = TRUE,
+                                                                 nudge_y = 0.018, size = 7 )+
+    geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.5, yend=.2),  linetype =2, arrow=arrow(), size=.05, color="blue") +
+  geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.5, yend=.3), linetype =2, arrow=arrow(), size=.05, color="blue")+
+  geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.7, yend=.4), linetype =2, arrow=arrow(), size=.05, color="orangered")+
+  geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.7, yend=.5), linetype =2, arrow=arrow(), size=.05, color="orangered")+
+    annotate("text", x = c(.32,.32,.32,.32), y = c(.23,.29,.35,.39), 
+             label = c(".7", ".6",".1", ".2") , color="black", 
+             size=7,  fontface="bold")+
+    annotate("text", x = c(.1,.5,.7,1.2), y = c(.55,.55,.55,.55), 
+           label = c("X", "A","B", "Y") , color=c("black", "blue","orangered","black"), 
+           size=10,  fontface="bold")+xlim(0,1.3)+
+  theme_void(base_size = 25)
+```
+\caption{Example WEAT calculations: $s_1 = s(he,A,B) =  \frac{.6+.7}{2}  - \frac{.2+.1}{2} = .65-.15= .5$,  $s_2 = s(man,A,B) = .3$,  $s_3 = s(woman,A,B) = -.6, s_4 = s(she, A, B) = -.3$, $\mathsf{WEAT}(A,B)  = \frac{\nicefrac{(s_1+s_2)}{2} - \nicefrac{(s_3+s_4)}{2}}{sd(\{s_1,s_2,s_3,s_4\})} \approx 1.93$.}
+\label{eq:WEATexample}
+\end{figure}
+
+\footnotesize
+
+- $s_1 = s(he,A,B) =  \frac{.6+.7}{2}  - \frac{.2+.1}{2} = .65-.15= .5$
+
+- $s_2 = s(man,A,B) = .3$, \linebreak  $s_3 = s(woman,A,B) = -.6, s_4 = s(she, A, B) = -.3$
+
+\vspace{-4mm}
+
+\normalsize 
+\begin{align*}
+\mathsf{WEAT}(A,B) & = \frac{\nicefrac{(s_1+s_2)}{2} - \nicefrac{(s_3+s_4)}{2}}{sd(\{s_1,s_2,s_3,s_4\})} \approx 1.93
+\end{align*}
+
+WEAT, however, has been developed to investigate biases corresponding to a pair of supposedly opposing stereotypes, and so the question arises as to how to generalize the measure to contexts in which biases with respect to more than two stereotypical groups are to be measured.
 Such a generalization can be found in [@Manzini2019blackToCriminal].
@@
 
-```{r religionTableHeadEarly,echo=FALSE,eval=FALSE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
+```{r religionTableHeadEarly,echo=FALSE,eval=FALSE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
 religion <- read.csv("datasets/outdated/religionReddit.csv")[-c(1,6,7)]
@@
 
+\begin{figure}[H]
+```{r MACexample,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "50%", warning = FALSE, message = FALSE}
+words <- c("jew", "muslim", "christian", "terrorist","dirty","conservative","familiar","greedy","cheap")
+x <- c(.1,.1,.1,.4,.4,.7,.7,1.1,1.1)
+y <- c(.4,.3,.2,.1,.2,.3,.4,.1,.2)
+mac <- data.frame(words,x)
+
+ggplot(mac, aes(x=x,y=y, label = words))+geom_point()+geom_text(check_overlap = TRUE,
+                                                                 nudge_y = 0.02, size = 9 )+
+  geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.4, yend=.1),  linetype =2, arrow=arrow(), size=.03, color="blue")+geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.4, yend=.2), linetype =4, arrow=arrow(), size=.0, color="blue") + geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.7, yend=.3), linetype =2, arrow=arrow(), size=.05, color="orangered")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.7, yend=.4), linetype =2, arrow=arrow(), size=.05, color="orangered")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=1.1, yend=.1), linetype =3, arrow=arrow(), size=.05, color="chartreuse4")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=1.1, yend=.2), linetype =2, arrow=arrow(), size=.05, color="chartreuse4")+ annotate("text", x = c(.1,.4,.7,1.2), y = c(.45,.45,.45,.45),  label = c("T", "A[1]","A[2]", "A[3]") ,parse = TRUE, color=c("black", "blue","orangered","chartreuse4"),            size=10,  fontface="bold")+xlim(0,1.3)+theme_void(base_size = 25)
+
+```
+\caption{Example of MAC calculations.  $s_1  = s(muslim,A_1) = \frac{cos(muslim,dirty)+cos(muslim,terrorist)}{2}$, 
+$s_2  = s(muslim,A_2) = \frac{cos(muslim,familiar)+cos(muslim,conservative)}{2}$, \dots. Then  $\mathsf{MAC}(T,A)  = \mathsf{mean}(\{s_i \vert i \in 1, \dots, k\})$.}
+\label{fig:MACexample}
+\end{figure}
+
 Having introduced the measures,  first, we will introduce a selection of general problems with this approach, and then we will move on to more specific but important problems related to the fact that the measures take  averages and averages of averages.
@@
 
-# General methodological problems with cosine-based measures of bias
-
-Consider the MAC results from Table REF.
-Are the initial MAC values lower than 1 indicative of the presence of bias?
+# Methodological problems with cosine-based bias metrics
+
+Here is an example of results of debiasing that can be found in [@Manzini2019blackToCriminal] (Table \ref{tab:religionOriginal2}).
+It presents the results of applying the procedure with respect to the Religion dataset.
+MAC scores are given for the original datasets, and two resulting from two debiasing methods applied to the Reddit embedding.
+The scores for debiasing come with p-values for the comparison to the original MAC.
+
+\begin{table}[H]
+\footnotesize
+
+\centering
+
+\begin{tabular}[t]{lllr}
+\toprule
+Religion Debiasing & MAC & a p-value \\
+\midrule
+\cellcolor{gray!15}{Biased} & \cellcolor{gray!15}{0.859} & \cellcolor{gray!15}{N/A} \\
+Hard Debiased & 0.934 & 3.006e-07\\
+\cellcolor{gray!15}{Soft Debiased ($\lambda$ = 0.2)} & \cellcolor{gray!15}{0.894} & \cellcolor{gray!15}{0.007} \\
+\bottomrule
+\end{tabular}
+
+\caption{The associated mean average cosine similarity
+(MAC) and p-values for debiasing methods for religious bias.}
+\label{tab:religionOriginal2}
+\normalsize 
+\end{table}
+
+The first conceptual question we should ask is whether   the initial MAC values lower than 1 indeed are  indicative of the presence of bias?
 Of course, thinking abstractly, 1 is the ideal distance for unrelated words.
@@
 How much smaller?
-While the original paper did not employ control lists for comparison in the paper.
-To obtain a more realistic gauge on how to understand MAC values (and in further development of our approach), we introduced control word lists.
-One of them is a list of words we intuitively considered neutral (see REF in the APPENDIX).
-Moreover, it might be the case that words that have to do with human activities are systematically closer to the protected words than merely neutral words, for which reason we also use a second list with intuitively non-stereotypical human attributes (see REF in the APPENDIX).
-Another important observation is that in their calculations the authors do not distinguish whether a given  attribute is associated with a given protected word (e.g.
-protected word \emph{jew} and the attribute \emph{greedy}) or whether it is associated with a different stereotype from the same list (e.g.
-protected word \emph{quran} and the attribute \emph{greedy}), simply averaging across all such groups.
-
-In Figure  \ref{fig:empirical} We look at the empirical distributions, while paying attention to such divisions.
-Horizontal lines represent 1-MAC values that the authors considered indicative of bias for stereotypes corresponding to  given  word lists.
-For instance, in religion, MAC was .859, so we plot $0\pm (1-.859)\approx .14$ lines around similarity = 0 (that is, distance = 1).
-Notice that most distributions are quite wide, and the proportions of even neutral or human neutral words with similarities higher than the value of 1-MAC deserving debiasing according to the authors is quite high.
-
-```{r cosineDistributions6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%"}
+What may attract attention is the fact that the value of cosine distance in "Biased" category is already quite high (i.e.
+close to 1) even before debiasing.
+High cosine distance indicates low cosine similarity between values.
+One could think that average cosine similarity equal to approximately 0.141 is not large enough to consider it as bias.
+The authors, though, still aim to mitigate such "bias" to make the distance even larger.
+The question is, on what basis is this small similarity still considered as a proof of the presence of bias, and whether these small changes are meaningful.
+
+The problem is that the original paper did not employ any control group of neutral attributes for comparison to obtain a more realistic gauge on how to understand MAC values.
+Later on, in our approach, we introduce such control word lists.
+One of them is a list of words we intuitively considered neutral (see the Appendix for word lists).
+
+Moreover, it might be the case that words that have to do with human activities in general, even if unbiased, are systematically closer to the protected words than merely neutral words.
+This, again, casts doubt on whether comparing MAC to the abstractly ideal value of 1 is a methodologically sound idea.
+For this reason we also use a second list with intuitively non-stereotypical human attributes (again, see the Appendix for a word list).
+
+Another important observation is that  MAC calculations  do not distinguish whether a given  attribute is associated with a given protected word, simply averaging across all such groups.
+Let us use the  case of religion-related stereotypes to illustrate.
+The full lists from [@Manzini2019blackToCriminal] can be found in Appendix \ref{appendix:manzini_word_lists}.
+In the original paper,  words from all three religions were compared against all of the stereotypes.
+No distinction between cases in which the stereotype is associated with a given religion, as opposed to the situation in which it is associated with another one, is made.
+For example, the protected word \emph{jew} is supposed to be stereotypically  connected with  the attribute \emph{greedy}, while  from the protected word \emph{quran} the  attribute \emph{greedy} comes from a different stereotype, and yet the distances between these pairs contribute equally to the final \textsf{MAC} score.
+This is problematic, as  not all of the stereotypical words have to be considered as harmful for all of the religions.
+To avoid the masking effect, one should  pay attention to how protected words and attributes are paired by stereotypes.
+
+In Figure  \ref{fig:empirical} we look at the empirical distributions, while paying attention to such divisions.
+The horizontal lines represent the values of 1-\textsf{MAC} (that is, we now talk in terms of cosine similarity rather than cosine distance; this is a trivial linear transformation) that the authors considered indicative of bias for stereotypes corresponding to  given  word lists.
+For instance, in religion, MAC was .859 and considered a sign of bias, so we plot $0\pm (1-.859)\approx .14$ lines around similarity = 0 (that is, distance = 1).
+Notice that most distributions are quite wide, and the proportions of even neutral or human neutral words with similarities higher than the value of 1-\textsf{MAC} deserving debiasing according to the authors is quite high.
+
+```{r cosineDistributions6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
 
@@
 \begin{figure}[H]
-```{r fig:cosineDistributions8b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", fig.height = 16}
+```{r fig:cosineDistributions8b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", fig.height = 16, warning = FALSE, message = FALSE}
 
@@
 
-One issue to consider is  the selection of attributes for bias measurement. The word lists used in the literature are often fairly small (5-50).
-
-While the papers do employ statistical tests to measure the uncertainty involved, we will later on argue that these method are not  proper for the goal at hand and  show that a more appropriate use of statistical methods leads to  estimates of uncertainty that are rather epistemologically pessimistic.
-\todo{sensitivity to choice of words}
-
-Let's think about MAC, using the  case of religion-related stereotypes.
-The full lists from [@Manzini2019blackToCriminal] can be found in \ref{appendix:manzini_word_lists}.
-In the original paper,  words from all three religions were compared against all of the stereotypes.
-One reason this is problematic is that  no distinction between cases in which the stereotype is associated with a given religion, as opposed to the situation in which it is associated with another one, is made.
-This is problematic, as  not all of the stereotypical words have to be considered as harmful for all of the religions.
-One should investigate the religions separately as some of them may have stronger harmful associations that others.
-
-The interpretation of the results is also a challenge.
-In [@Manzini2019blackToCriminal] we can find  summaries of  average cosine distances per group (such as gender, race, or religion).
-For instance, for religion, here is the relevant fragment of table:
-
-<!-- Religion Debiasing  | MAC | p-value
-------------- | ------------- | -------
-Biased        | 0.859 | N/A
-Hard Debiased | 0.934 | 3.006e-07
-Soft Debiased ($\lambda$ = 0.2) | 0.894 | 0.007 -->
-
-\begin{table}
-\footnotesize
-
-\centering
-
-\begin{tabular}[t]{lllr}
-\toprule
-Religion Debiasing & MAC & a p-value \\
-\midrule
-\cellcolor{gray!15}{Biased} & \cellcolor{gray!15}{0.859} & \cellcolor{gray!15}{N/A} \\
-Hard Debiased & 0.934 & 3.006e-07\\
-\cellcolor{gray!15}{Soft Debiased ($\lambda$ = 0.2)} & \cellcolor{gray!15}{0.894} & \cellcolor{gray!15}{0.007} \\
-\bottomrule
-\end{tabular}
-
-\caption{The associated mean average cosine similarity
-(MAC) and p-values for debiasing methods for religious bias}
-\label{tab:religionOriginal2}
-\normalsize 
-\end{table}
-
-\noindent (MAC stand for mean average cosine similarity, although in reality the the table contains mean cosine distances).
-What may attract attention is the fact that the value of cosine distance in "Biased" category is already quite high (i.e.
-close to 1) even before debiasing.
-High cosine distance indicates low cosine similarity between values.
-One could think that average cosine similarity equal to approximately 0.141 is not significant enough to consider it as bias.
-However,  the authors still aim to mitigate such "bias"  to make the distance even larger.
-Methodologically the question is, on what basis is this small similarity still considered as a proof of the presence of bias, and whether these small changes are meaningful.
-
-The underlying problem here is that in the paper there is no  control group.
-One should also include control groups to have a way of comparing the  results for a supposedly stereotyped group with the results for sets of neutral or human-related neutral words.
-In our approach later on, we distinguish between stereotypes associated with a given group, stereotypes associated with different groups, and introduce control groups: neutral words and stereotype-free human predicates.
+Another issue to consider is  the selection of attributes for bias measurement.
+The word lists used in the literature are often fairly small (5-50).
+The papers in the field do employ statistical tests to measure the uncertainty involved and do make claims of statistical significance.
+Yet, we will later on argue that these method are not  proper for the goal at hand.
+By using Bayesian methods we   will show that a more appropriate use of statistical methods leads to  estimates of uncertainty which suggest that larger word lists would be advisable.
+
+To avoid such problems, later on, we employ control groups and in line with Bayesian methodology, use posterior distributions and  highest posterior density intervals instead of chasing single-point metrics based on pre-averaged data.
+But before we do so, let us spend some time explaining why pre-averaging is a sub-optimal strategy.
 
@@
 
-## Metrics that pre-average are a bad guide
-
-In contrast, statistical intervals will help us decide whether a given cosine similarity is high enough to consider the words to be more similar than if we chose them at random.
-We will use highest posterior density intervals, in line with Bayesian methodology.
-
-Crucially, these approaches use  means of mean average cosine similarities to measure  similarity between protected word and harmful stereotypes.
-If one takes  a closer look at the individual values that are taken for the calculations, it turns out that there are quite a few  outliers and surprisingly dissimilar words.
+#  Pre-averaging, bootstrapping and testing with the null model
+
+The approaches we have been describing use  means of mean average cosine similarities to measure  similarity between protected words and attributes coming from harmful stereotypes.
+However, a closer inspection look at the individual values that are taken for the calculations, it turns out that there are quite a few  outliers and surprisingly dissimilar words.
 This problem becomes transparent when we examine the visualizations of the individual cosine distances, following the idea that one of the first steps in understanding data is to look at it.
+
+\begin{figure}[H]
+```{r visMuslim,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%",message =FALSE, warning = FALSE}
+religion <- read.csv("datasets/cleanedDatasets/religionRedditCleaned.csv")
+muslimWords <- c("imam","islam","mosque","muslim","quran")
+muslim <- religion %>% filter(protectedWord %in% muslimWords)
+muslimClass <- muslim %>% filter(protectedWord == "muslim")
+neutralSample <- sample_n(filter(muslimClass,connection == "none"), 5)
+humanSample <- sample_n(filter(muslimClass,connection == "human"), 5)
+muslimVis <- muslimClass %>% filter(connection != "none" & connection !="human")
+muslimVis <- rbind(muslimVis,neutralSample,humanSample)
+
+source("functions/visualisationTools.R")
+
+visualiseProtected(muslimVis,"muslim")+theme_void(base_size = 20)
+```
+
+```{r visPriest,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
+religion <- read.csv("datasets/cleanedDatasets/religionRedditCleaned.csv")
+priestClass <- religion %>% filter(protectedWord == "priest")
+neutralSample <- sample_n(filter(priestClass,connection == "none"), 5)
+humanSample <- sample_n(filter(priestClass,connection == "human"), 5)
+priestVis <- priestClass %>% filter(connection != "none" & connection !="human")
+priestVis <- rbind(priestVis,neutralSample,humanSample)
+
+source("functions/visualisationTools.R")
+
+visualiseProtected(priestVis,"priest")+theme_void(base_size = 20)
+```
+
+\caption{Actual distances for two protected words.
+For the protected word "muslim", the most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates.
+For the word "priest", the situation is even less as expected: the nearest attributes are human attributes, and all there seems to be no clear pattern to the distances to other attributes.}
+\end{figure}
 
@@
 The t-tests they employ are run on average cosines used to calculate MAC.
-
-#  Pre-averaging, bootstrapping and testing with the null model
 
@@
 
-Furthermore, conceptually there are two levels of analysis in our approach.
+Furthermore, conceptually there are two levels of analysis in our approach (see Figure \ref{fig:visDag}).
 On the one hand, we are interested in the general question of whether related attributes are systematically closer across the dataset.
@@
 Such models can be thought of as incorporating adaptive regularization, which avoids overfitting and leads to improved estimates for unbalanced datasets (and the datasets we need to use are unbalanced).
+
+\begin{figure}[H]
+```{r dagVis,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", message = FALSE, warning = FALSE, results = FALSE}
+dagBayesBias <- dagitty( 
+  "dag{
+  distances -> associated
+  distances -> different
+  distances -> human
+  distances -> neutral
+
+  associated -> protectedWord1
+  associated -> protectedWord2
+  associated -> protectedWord3
+
+  different -> protectedWord1
+  different -> protectedWord2
+  different -> protectedWord3
+
+  human -> protectedWord1
+  human -> protectedWord2
+  human -> protectedWord3
+
+  neutral -> protectedWord1
+  neutral -> protectedWord2
+  neutral -> protectedWord3
+
+  protectedWord1 -> attribute1
+  protectedWord1 -> attribute2
+  protectedWord1 -> attribute3
+  protectedWord1 -> attribute4
+
+  protectedWord2 -> attribute1
+  protectedWord2 -> attribute2
+  protectedWord2 -> attribute3
+  protectedWord2 -> attribute4
+  protectedWord3 -> attribute1
+  protectedWord3 -> attribute2
+  protectedWord3 -> attribute3
+  protectedWord3 -> attribute4
+  }
+  " 
+)
+
+coordinates(dagBayesBias) <- list(
+  x=c(
+    distances = 2.5, associated = 1, different = 2, human = 3, neutral = 4,
+    protectedWord1 = 1.5, protectedWord2 = 2.5, protectedWord3 = 3.5,
+    attribute1 = 1, attribute2 = 2, attribute3 = 3, attribute4 = 4
+    ) ,
+  y=c(
+    distances = .5, associated = 1, different = 1, human = 1, neutral = 1,
+    protectedWord1 = 2, protectedWord2 = 2, protectedWord3 = 2,
+    attribute1 = 3, attribute2 = 3, attribute3 = 3, attribute4 = 3
+  ) )
+
+drawdag(dagBayesBias)
+```
+
+\caption{At a general level, we will be estimating the coefficients for distances as grouped by whether they are between protected words and attributes coming from their respective associated/different/human/neutral attribute groups.
+At a more fine-grained level, for each protected word we will be estimating the proximity of that word to attributes that are associated with its respective stereotype, come from a different stereotype, or come from the human/neutral attribute lists.}
+\label{fig:visDag}
+\end{figure}
 
@@
 \begin{figure}[H]
-```{r fig:debiasedCosinePair2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, out.width="85%"}
+```{r fig:debiasedCosinePair2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, out.width="85%", warning = FALSE, message = FALSE}
 
@@
 
-```{r debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "75%"}
+```{r debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
 
@@
 \begin{figure}[H]
-```{r fig:debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "70%", fig.height = 16}
+```{r fig:debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "70%", fig.height = 16, warning = FALSE, message = FALSE}
 
@@
 
-After introducing the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of WEAT or MAC (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
+After introducing the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of WEAT or MAC (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,^[Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25.]  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
 Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable.
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsWeat7Google$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsWeat7Glove$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsWeat7Reddit$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsWeat1Google$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsWeat1Glove$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsWeat1Reddit$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsReligionGoogle$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsReligionGlove$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsReligionReddit$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsRaceGoogle$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsRaceGlove$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsRaceReddit$plotJoint
@@
 
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsGenderGoogle$plotJoint
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsGenderGlove$plotJoint
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsGenderReddit$plotJoint
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsDebiasedReligionReddit$plotJoint
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsDebiasedRaceReddit$plotJoint
@@
 \begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300}
+```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
 resultsDebiasedGenderReddit$plotJoint
@@
 
-             # , posteriorPlotsB[[1]],
-             # posteriorPlotsA[[2]], posteriorPlotsB[[2]],
-             # ncol = 2)
 ```

### 2023-05-14 functions/generatePredictions.R

@@ -1,97 +1,0 @@
-
-source("functions/cleanDataset.R")
-source("functions/modelToPrediction.R")
-
-
-predictions <- list()
-
-
-
-genderGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1])
-genderReddit <-  cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1])
-genderGlove <- read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1]
-genderGlove <- cleanDataset(genderGlove)
-
-raceGlove <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1])
-raceGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1])
-raceReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1])
-
-religionGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1])
-religionGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1])
-religionReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1])
-
-
-
-modelList <- c(
-  "modelReligionGoogle.RDS", "modelReligionGlove.RDS", "modelReligionReddit.RDS",
-  "modelRaceReddit.RDS", "modelRaceGoogle.RDS", "modelRaceGlove.RDS", 
-  "modelGenderReddit.RDS", "modelGenderGoogle.RDS", "modelGenderGlove.RDS",
- "modelWeat7Google.RDS", "modelWeat7Glove.RDS"
-)
-
-#"modelWeat1Reddit.RDS", #, "modelWeat1Glove.RDS
-# "modelWeat7Reddit.RDS",
-
-simulated <- list()
-
-
-model <- readRDS(file = "models/modelReligionGoogle.RDS")
-simulated$religionGoogle <- sim(model)
-
-predictions$religionGoogle <- modelToPredictions(model, religionGoogle, "religionGoogle")
-
-model <- readRDS(file = "models/modelReligionGlove.RDS")
-simulated$religionGlove <- sim(model)
-
[…]

### 2023-05-17 functions/allModels.R
cleaned up md and access instructions

@@ -1,0 +1,128 @@
+library(rethinking)
+
+
+getwd()
+source("functions/cleanDataset.R")
+source("functions/newModelBuilding.R")
+
+
+# GENDER
+genderGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1])
+genderReddit <-  cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1])
+
+
+
+modelGenderGoogle <- buildModel(genderGoogle)
+saveRDS(modelGenderGoogle, file = "models/modelGenderGoogle.RDS")
+precisGenderGoogle <- extractPrecis(modelGenderGoogle, name = "GenderGoogle")
+
+
+modelGenderReddit <- buildModel(genderReddit)
+saveRDS(modelGenderReddit, file = "models/modelGenderReddit.RDS")
+precisGenderReddit  <- extractPrecis(modelGenderReddit, name = "GenderReddit")
+
+
+genderGlove <- read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1]
+genderGlove <- cleanDataset(genderGlove)
+modelGenderGlove <- buildModel(genderGlove)
+save(modelGenderGlove, file = "models/modelGenderGlove.rda")
+
+# RACE
+raceGlove <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1])
+raceGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1])
+raceReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1])
+
+
+modelRaceGlove <- buildModel(raceGlove)
+saveRDS(modelRaceGlove, file = "models/modelRaceGlove.RDS")
+precisRaceGlove  <- extractPrecis(modelRaceGlove, name = "RaceGlove")
+
+
+
+modelRaceGoogle <- buildModel(raceGoogle)
+saveRDS(modelRaceGoogle, file = "models/modelRaceGoogle.RDS")
+precisRaceGoogle <- extractPrecis(modelRaceGoogle, name = "RaceGoogle")
+
+
+
+modelRaceReddit <- buildModel(raceReddit)
+saveRDS(modelRaceReddit, file = "models/modelRaceReddit.RDS")
[…]

### 2023-05-17 functions/generatePredictions.R

@@ -1,0 +1,94 @@
+
+source("functions/cleanDataset.R")
+source("functions/modelToPrediction.R")
+
+
+predictions <- list()
+
+
+genderGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1])
+genderReddit <-  cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1])
+genderGlove <- read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1]
+genderGlove <- cleanDataset(genderGlove)
+
+raceGlove <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1])
+raceGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1])
+raceReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1])
+
+religionGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1])
+religionGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1])
+religionReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1])
+
+
+
+modelList <- c(
+  "modelReligionGoogle.RDS", "modelReligionGlove.RDS", "modelReligionReddit.RDS",
+  "modelRaceReddit.RDS", "modelRaceGoogle.RDS", "modelRaceGlove.RDS", 
+  "modelGenderReddit.RDS", "modelGenderGoogle.RDS", "modelGenderGlove.RDS",
+  #"modelWeat1Reddit.RDS", #, "modelWeat1Glove.RDS
+  # "modelWeat7Reddit.RDS", "modelWeat7Google.RDS", "modelWeat7Glove.RDS"
+)
+
+
+simulated <- list()
+
+
+model <- readRDS(file = "models/modelReligionGoogle.RDS")
+simulated$religionGoogle <- sim(model)
+
+predictions$religionGoogle <- modelToPredictions(model, religionGoogle, "religionGoogle")
+
+model <- readRDS(file = "models/modelReligionGlove.RDS")
+simulated$religionGlove <- sim(model)
+
[…]

### 2023-05-17 functions/newModelBuilding.R

@@ -6,13 +6,6 @@ library(gridExtra)
 library(grid)
 
 
-religionReddit <- read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1]
-
-
-
-dataset <- religionReddit
-
-
 cleanDataset <- function (dataset) {
 
 colnames(dataset) <- c("pw","word","stereotype", "distance", "similarity", "connection")
@@ -20,7 +13,6 @@ dataset$pw <- as.factor(dataset$pw)
 dataset$word <- as.factor(dataset$word)
 dataset$stereotype <- as.factor(dataset$stereotype)
 dataset$connection <- as.factor(dataset$connection)
-
 
 
 ifelse(sum( dataset$similarity < -1 | dataset$distance < -1) != 0, 
@@ -45,78 +37,8 @@ return(dataset)
 }
 
 
-dataset <- cleanDataset(religionReddit)
-
-
-
-#full hierarhical
-modelResult <- ulam(
-  alist(
-    distance ~ dnorm(mu,sigma),
-    mu <- d[pwi] * different + a[pwi] * associated + h[pwi] * human + n[pw] * none,
-    d[pwi] ~ dnorm(dbar, dsigmabar),
-    a[pwi] ~ dnorm(abar, asigmabar),
-    h[pwi] ~ dnorm(hbar, hsigmabar),
-    n[pwi] ~ dnorm(nbar, nsigmabar),
-    dbar ~ dnorm(1,.3),
-    abar ~ dnorm(1,.3),
-    hbar ~ dnorm(1,.3),
-    nbar ~ dnorm(1,.3),
-    dsigmabar ~ dexp(2),
-    asigmabar ~ dexp(2),
-    hsigmabar ~ dexp(2),
-    nsigmabar ~ dexp(2),
-    sigma ~ dexp(2)
-  ),   data = dataset, chains=2 , iter=8000 , warmup=1000,  log_lik = TRUE, cores = 4
-)
-
-head(dataset)
-
-#relativized sigma
-modelSigma <- ulam(
-  alist(
-    distance ~ dnorm(mu,sigma),
-    mu <- d[pwi] * different + a[pwi] * associated + h[pwi] * human + n[pw] * none,
-    d[pwi] ~ dnorm(dbar, dsigmabar),
-    a[pwi] ~ dnorm(abar, asigmabar),
[…]

### 2023-05-17 paper/paperDraft.Rmd
revised first two figures

@@
 
-In what follows, we focus on  two popular  measures of bias applicable to many existing word embeddings ( \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.} \emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.} and \emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}): \emph{Word Embedding Association Test}  (WEAT) [@Caliskan2017semanticsBiases],  and \emph{Mean Average Cosine Distance} (MAC)  [@Manzini2019blackToCriminal].
+In what follows, we focus on  two popular  measures of bias applicable to many existing word embeddings ( \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.} \emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.} and \emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}): \emph{Word Embedding Association Test}  (\textsf{WEAT}) [@Caliskan2017semanticsBiases],  and \emph{Mean Average Cosine Distance} (\textsf{MAC})  [@Manzini2019blackToCriminal].
 We first explain how these measures are supposed to work.
@@
 We  propose to replace them with a Bayesian data analysis (hierarchical models), which not only provides more modest and realistic assessment of the uncertainty involved, but also allows for inspection at various levels of granularity.
-Once we introduce the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of WEAT or MAC (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
+Once we introduce the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of \textsf{WEAT} or \textsf{MAC} (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,^[Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25.]  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
 Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable.
@@
 
-A measure of bias in word embeddings which does not employ similarities to  pre-calculated protected-class-level directions, the Word Embedding Association Test (WEAT), has been proposed in  [@Caliskan2017semanticsBiases].
+A measure of bias in word embeddings which does not employ similarities to  pre-calculated protected-class-level directions, the Word Embedding Association Test (\textsf{WEAT}), has been proposed in  [@Caliskan2017semanticsBiases].
 The idea here is that the   bias between two sets of target words, $X$ and $Y$ (we call them protected words), should be quantified in terms of the cosine similarity  between the protected words and attribute words coming from  two sets of stereotype attribute words, $A$ and $B$ (we will call them attributes).
@@
 \noindent The authors use it as a test statistic in some tests, and the assumption that $X$ and $Y$ are of the same size is necessary for the statistic to make sense.
-In the final measure (of effect size), WEAT, means are taken, which makes it insensitive to size differences between $X$ and $Y$:
+In the final measure (of effect size), \textsf{WEAT}, means are taken, which makes it insensitive to size differences between $X$ and $Y$:
 \begin{align}
@@
 
- WEAT is inspired by  the Implicit Association Test (IAT) [@Nosek2002harvesting] used in psychology, and it uses almost the same word sets, allowing for a \emph{prima facie} sensible comparison with bias in humans.
+ \textsf{WEAT} is inspired by  the Implicit Association Test (IAT) [@Nosek2002harvesting] used in psychology, and it uses almost the same word sets, allowing for a \emph{prima facie} sensible comparison with bias in humans.
 In  [@Caliskan2017semanticsBiases] the authors argue that significant biases---thus measured--- similar to the ones discovered by IAT can be discovered in word embeddings.
@@
 \begin{figure}[H]
-```{r WEATexample,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "55%", warning = FALSE, message = FALSE}
+```{r WEATexample5,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "55%", warning = FALSE, message = FALSE}
 words <- c("man", "he", "boss", "architect","nurse","receptionist","woman","she")
@@
 ggplot(weat, aes(x=x,y=y, label = words))+geom_point()+geom_text(check_overlap = TRUE,
-                                                                 nudge_y = 0.018, size = 7 )+
+                                                                 nudge_y = 0.021, size = 7 )+
     geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.5, yend=.2),  linetype =2, arrow=arrow(), size=.05, color="blue") +
@@
 ```
-\caption{Example WEAT calculations: $s_1 = s(he,A,B) =  \frac{.6+.7}{2}  - \frac{.2+.1}{2} = .65-.15= .5$,  $s_2 = s(man,A,B) = .3$,  $s_3 = s(woman,A,B) = -.6, s_4 = s(she, A, B) = -.3$, $\mathsf{WEAT}(A,B)  = \frac{\nicefrac{(s_1+s_2)}{2} - \nicefrac{(s_3+s_4)}{2}}{sd(\{s_1,s_2,s_3,s_4\})} \approx 1.93$.}
+\caption{A simple example of a two-class set-up. Two groups of protected words, $X$ and $Y$ with two stereotypical attribute sets. An example of \textsf{WEAT} calculations follows.}
 \label{eq:WEATexample}
@@
 
-\footnotesize
-
-- $s_1 = s(he,A,B) =  \frac{.6+.7}{2}  - \frac{.2+.1}{2} = .65-.15= .5$
-
-- $s_2 = s(man,A,B) = .3$, \linebreak  $s_3 = s(woman,A,B) = -.6, s_4 = s(she, A, B) = -.3$
-
-\vspace{-4mm}
-
-\normalsize 
-\begin{align*}
+\noindent Here is an example of \textsf{WEAT} calculations for Figure \ref{eq:WEATexample}:
+\begin{align*} s_1 & = s(he,A,B)  =  \nicefrac{(.6+.7)}{2}  - \nicefrac{(.2+.1)}{2} = .65-.15= .5 \\
+s_2  & = s(man,A,B) = .3 \\
+s_3  & = s(woman,A,B) = -.6\\
+s_4 & = s(she, A, B) = -.3\\
 \mathsf{WEAT}(A,B)  & = \frac{\nicefrac{(s_1+s_2)}{2} - \nicefrac{(s_3+s_4)}{2}}{sd(\{s_1,s_2,s_3,s_4\})} \approx 1.93
@@
 
-WEAT, however, has been developed to investigate biases corresponding to a pair of supposedly opposing stereotypes, and so the question arises as to how to generalize the measure to contexts in which biases with respect to more than two stereotypical groups are to be measured.
+\textsf{WEAT}, however, has been developed to investigate biases corresponding to a pair of supposedly opposing stereotypes, and so the question arises as to how to generalize the measure to contexts in which biases with respect to more than two stereotypical groups are to be measured.
 Such a generalization can be found in [@Manzini2019blackToCriminal].
-The authors introduce Mean Average Cosine distance (MAC) as a measure of bias.
+The authors introduce Mean Average Cosine distance (\textsf{MAC}) as a measure of bias.
 Let $T = \{t_1, \dots, t_k\}$ be a set of protected words, and let each $A_j\in \mathcal{A}$ be a set of attributes stereotypically associated with a protected word where $\mathcal{A}$.
@@
 ```
-\caption{Example of MAC calculations.  $s_1  = s(muslim,A_1) = \frac{cos(muslim,dirty)+cos(muslim,terrorist)}{2}$, 
-$s_2  = s(muslim,A_2) = \frac{cos(muslim,familiar)+cos(muslim,conservative)}{2}$, \dots. Then  $\mathsf{MAC}(T,A)  = \mathsf{mean}(\{s_i \vert i \in 1, \dots, k\})$.}
+\caption{A small subset of the religion dataset.
+To each protected word in $T$  there corresponds one class of stereotypical attributes typically associated with it (and other classes of stereotypical attributes associated with different protected words).}
 \label{fig:MACexample}
 \end{figure}
+
+\noindent An example of \textsf{MAC} calculations for the situation depicted in Figure \ref{fig:MACexample} is as follows:
+\begin{align*}
+s_1   = s(muslim,A_1)  & = \frac{\mathsf{cos}(muslim,dirty)+\mathsf{cos}(muslim,terrorist)}{2} \\ 
+s_2   = s(muslim,A_2)  & = \frac{\mathsf{cos}(muslim,familiar)+\mathsf{cos}(muslim,conservative)}{2}\\ & \vdots \\ 
+\mathsf{MAC}(T,A)  & = \mathsf{mean}(\{s_i \vert i \in 1, \dots, k\})
+\end{align*}
+
+Notably, the inutitive distinction between different attribute sets plays no real role in the \textsf{MAC} calculations.
+Equally well one could calculate the mean distance of \emph{muslim} to all the predicates, mean distance of \emph{christian} to all the predicates, mean distance of \emph{jew} to all the predicates, and then to take the mean of these three means.
 
@@
 It presents the results of applying the procedure with respect to the Religion dataset.
-MAC scores are given for the original datasets, and two resulting from two debiasing methods applied to the Reddit embedding.
-The scores for debiasing come with p-values for the comparison to the original MAC.
+\textsf{MAC} scores are given for the original datasets, and two resulting from two debiasing methods applied to the Reddit embedding.
+The scores for debiasing come with p-values for the comparison to the original \textsf{MAC}.
 
@@
 \toprule
-Religion Debiasing & MAC & a p-value \\
+Religion Debiasing & \textsf{MAC} & a p-value \\
 \midrule
@@
 \caption{The associated mean average cosine similarity
-(MAC) and p-values for debiasing methods for religious bias.}
+(\textsf{MAC}) and p-values for debiasing methods for religious bias.}
 \label{tab:religionOriginal2}
@@
 
-The first conceptual question we should ask is whether   the initial MAC values lower than 1 indeed are  indicative of the presence of bias?
+The first conceptual question we should ask is whether   the initial \textsf{MAC} values lower than 1 indeed are  indicative of the presence of bias?
 Of course, thinking abstractly, 1 is the ideal distance for unrelated words.
-But clearly, there is variation in distances, which might lead to non-biased lists also having MAC smaller than 1.
+But clearly, there is variation in distances, which might lead to non-biased lists also having \textsf{MAC} smaller than 1.
 How much smaller?
@@
 
-The problem is that the original paper did not employ any control group of neutral attributes for comparison to obtain a more realistic gauge on how to understand MAC values.
+The problem is that the original paper did not employ any control group of neutral attributes for comparison to obtain a more realistic gauge on how to understand \textsf{MAC} values.
 Later on, in our approach, we introduce such control word lists.
@@
 Moreover, it might be the case that words that have to do with human activities in general, even if unbiased, are systematically closer to the protected words than merely neutral words.
-This, again, casts doubt on whether comparing MAC to the abstractly ideal value of 1 is a methodologically sound idea.
+This, again, casts doubt on whether comparing \textsf{MAC} to the abstractly ideal value of 1 is a methodologically sound idea.
 For this reason we also use a second list with intuitively non-stereotypical human attributes (again, see the Appendix for a word list).
 
-Another important observation is that  MAC calculations  do not distinguish whether a given  attribute is associated with a given protected word, simply averaging across all such groups.
+Another important observation is that  \textsf{MAC} calculations  do not distinguish whether a given  attribute is associated with a given protected word, simply averaging across all such groups.
 Let us use the  case of religion-related stereotypes to illustrate.
@@
 The horizontal lines represent the values of 1-\textsf{MAC} (that is, we now talk in terms of cosine similarity rather than cosine distance; this is a trivial linear transformation) that the authors considered indicative of bias for stereotypes corresponding to  given  word lists.
-For instance, in religion, MAC was .859 and considered a sign of bias, so we plot $0\pm (1-.859)\approx .14$ lines around similarity = 0 (that is, distance = 1).
+For instance, in religion, \textsf{MAC} was .859 and considered a sign of bias, so we plot $0\pm (1-.859)\approx .14$ lines around similarity = 0 (that is, distance = 1).
 Notice that most distributions are quite wide, and the proportions of even neutral or human neutral words with similarities higher than the value of 1-\textsf{MAC} deserving debiasing according to the authors is quite high.
@@
 
-The t-tests they employ are run on average cosines used to calculate MAC.
+The t-tests they employ are run on average cosines used to calculate \textsf{MAC}.
 
@@
 
-In the calculations of WEAT means are taken twice.
+In the calculations of \textsf{WEAT} means are taken twice.
 The $s$-values themselves are means, and then means of $s$-values are compared between groups.
@@
 The uncertainty resulting from including the raw data point variance into considerations is more extensive than the one suggested by the low p-values obtained from taking means or means of means as data points.
-In this section, we discussed the performance of the WEAT measure, but since the [@Manzini2019blackToCriminal] one is a generalization thereof, including the method of running statistical tests on pre-averaged data, our remarks, \emph{mutatis nutandis}, apply.
+In this section, we discussed the performance of the \textsf{WEAT} measure, but since the [@Manzini2019blackToCriminal] one is a generalization thereof, including the method of running statistical tests on pre-averaged data, our remarks, \emph{mutatis nutandis}, apply.
 Moreover, a general problem is that studies such as [@Caliskan2017semanticsBiases] or [@Manzini2019blackToCriminal]  do not use  control groups, and this makes the interpretation of the analysis additionally difficult.
@@
 
-After introducing the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of WEAT or MAC (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,^[Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25.]  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
+After introducing the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of \textsf{WEAT} or \textsf{MAC} (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small.
+posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
 Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable.

### 2023-05-17 paper/paperDraft.tex
revised first two figures

@@
 \documentclass[
-  10pt,
+  12pt,
   dvipsnames,enabledeprecatedfontcommands]{scrartcl}
@@
 \usepackage[textsize=footnotesize]{todonotes}
-\newcommand{\ali}[1]{\todo[color=gray!40]{\textbf{Alicja:} #1}}
-\newcommand{\mar}[1]{\todo[color=blue!40]{#1}}
-\newcommand{\raf}[1]{\todo[color=olive!40]{#1}}
 
@@
 assistants, search results, translations, text analysis and so on. Such
-tools require input words to be represented by numbers. This can be done
-with word embeddings, which represent words as vectors of real numbers.
-A common way to construct an embedding is to use a large natural
-language corpus to train a neural network to assign vectors to words
-that optimize for co-occurrence prediction accuracy. The vectors can
-then be compared in terms of their similarity -- the usual measure is
-cosine similarity -- and the results of such comparisons can be used in
-downstream tasks. Roughly speaking, cosine similarity is an imperfect
-mathematical proxy for semantic similarity {[}14{]}.
+tools require input words to be represented by vectors of real numbers.
+This can usually be done with word embeddings. A common way to construct
+an embedding is to use a large natural language corpus to train a neural
+network to assign vectors to words that optimize for co-occurrence
+prediction accuracy. The vectors can then be compared in terms of their
+similarity -- the usual measure is cosine similarity -- and the results
+of such comparisons can be used in downstream tasks. Roughly speaking,
+cosine similarity is an imperfect mathematical proxy for semantic
+similarity {[}14{]}.
 
@@
 One example of this approach is {[}3{]}. However, if our goal is to
-develop downstream tools that perform tasks that we care about well
-without further perpetuating or exacerbating harmful stereotypes, we
-still have good reasons to try to minimize the negative impact, at least
-as long as this does not damage the overall performance of the tool.
-Moreover, it is often not the case that the corpora mirror reality---to
-give a trivial example, heads are spoken of more often than kidneys, but
-this does not mean that kidneys occur much less often in reality than
-heads. To give a more relevant example, the disproportionate association
-of female words with female occupations in a corpus actually greatly
+develop downstream tools that perform tasks that we care about without
+further perpetuating or exacerbating harmful stereotypes, we still have
+good reasons to try to minimize the negative impact, at least as long as
+this does not damage the overall performance of the tool. Moreover, it
+is often not the case that the corpora mirror reality---to give a
+trivial example, heads are spoken of more often than kidneys, but this
+does not mean that kidneys occur much less often in reality than heads.
+To give a more relevant example, the disproportionate association of
+female words with female occupations in a corpus actually greatly
 exaggerates the actual lower disproportion in the real distribution of
@@
 \emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}):
-\emph{Word Embedding Association Test} (WEAT) {[}8{]}, and
-\emph{Mean Average Cosine Distance} (MAC) {[}12{]}. We first explain how
-these measures are supposed to work. Then we argue that they are
-problematic for various reasons, a key one being that by pre-averaging
-data they manufacture false confidence, which we illustrate in terms of
-simulations showing that the measures often suggest the existence of
-bias even if by design it is non-existent in a simulated dataset.
+\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}8{]}, and
+\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}12{]}. We first
+explain how these measures are supposed to work. Then we argue that they
+are problematic for various reasons, a key one being that by
+pre-averaging data they manufacture false confidence, which we
+illustrate in terms of simulations showing that the measures often
+suggest the existence of bias even if by design it is non-existent in a
+simulated dataset.
 
@@
 forward some general observations that are not exactly in line with the
-usual picture painted in terms of WEAT or MAC (and the problem
-generalizes to any approach that focuses on chasing a single numeric
-metric): the word list sizes and sample sizes used in the studies are
-usually
-small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}
-posterior density intervals are fairly wide, often the differences
-between associated, different and neutral human predicates, are not very
+usual picture painted in terms of \textsf{WEAT} or \textsf{MAC} (and the
+problem generalizes to any approach that focuses on chasing a single
+numeric metric): the word list sizes and sample sizes used in the
+studies are usually small,\footnote{Depending on a list for {[}8{]} the
+  range for protected words is between 13 and 100, and for attributes
+  between 16 and 25; for {[}12{]} the range for protected words is
+  between 14 and 18, and for attributes between 11 and 25.} posterior
+density intervals are fairly wide, often the differences between
+associated, different and neutral human predicates, are not very
 impressive. Also, a preliminary inspection suggests that the
@@
 similarity to the gender direction might be an indicator of bias with
-respect to a given class of words, but it is only one possible
-manifestation of it, and reducing the cosine similarity to such a
-projection may not be sufficient to eliminate bias. For instance,
-``math'' and ``delicate'' might be equally similar to a pair of opposed
-explicitly gendered words (\emph{she}, \emph{he}), while being closer to
-quite different stereotypical attribute words (such as \emph{scientific}
-or \emph{caring}). Further, it is observed in {[}5{]} that most word
-pairs retain similarity under debiasing meant to minimize
-projection-based bias.\footnote{In {[}1{]} another method which involves
-  analogies and their evaluations by human users on Mechanical Turk is
-  also used. We do not discuss this method in this paper, see its
-  criticism in {[}16{]}.}
+respect to a given class of words, it is only one possible manifestation
+of it, and reducing the cosine similarity to such a projection may not
+be sufficient to eliminate bias. For instance, ``math'' and ``delicate''
+might be equally similar to a pair of opposed explicitly gendered words
+(\emph{she}, \emph{he}), while being closer to quite different
+stereotypical attribute words (such as \emph{scientific} or
+\emph{caring}). Further, it is observed in {[}5{]} that most word pairs
+retain similarity under debiasing meant to minimize projection-based
+bias.\footnote{In {[}1{]} another method which involves analogies and
+  their evaluations by human users on Mechanical Turk is also used. We
+  do not discuss this method in this paper, see its criticism in
+  {[}16{]}.}
 
@@
 to pre-calculated protected-class-level directions, the Word Embedding
-Association Test (WEAT), has been proposed in {[}8{]}. The idea here is
-that the bias between two sets of target words, \(X\) and \(Y\) (we call
-them protected words), should be quantified in terms of the cosine
-similarity between the protected words and attribute words coming from
-two sets of stereotype attribute words, \(A\) and \(B\) (we will call
-them attributes). For instance, \(X\) might be a set of male names,
+Association Test (\textsf{WEAT}), has been proposed in {[}8{]}. The idea
+here is that the bias between two sets of target words, \(X\) and \(Y\)
+(we call them protected words), should be quantified in terms of the
+cosine similarity between the protected words and attribute words coming
+from two sets of stereotype attribute words, \(A\) and \(B\) (we will
+call them attributes). For instance, \(X\) might be a set of male names,
 \(Y\) a set of female names, \(A\) might contain stereotypically
@@
 
-\begin{align*}
+\begin{align}
+\label{eq:stAB}
 \mathsf{s}(t,A,B) & = \frac{\sum_{a\in A}\mathsf{cos}(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}\mathsf{cos}(t,b)}{\vert B\vert}
-\end{align*} \normalsize \noindent then, the association difference
+\end{align} \normalsize \noindent then, the association difference
 between \(A\) a \(B\) is:
 
-\begin{align*}
+\begin{align}
+\label{eq:sXYAB}
 \mathsf{s}(X,Y,A,B) & = \sum_{x\in X} \mathsf{s}(x,A,B) -  \sum_{y\in Y} \mathsf{s}(y,A,B)
-\end{align*}
+\end{align}
 
@@
 the statistic to make sense. In the final measure (of effect size),
-WEAT, means are taken, which makes it insensitive to size differences
-between \(X\) and \(Y\): \begin{align}
+\textsf{WEAT}, means are taken, which makes it insensitive to size
+differences between \(X\) and \(Y\): \begin{align}
 \mathsf{weat}(A,B) & = \frac{
@@
 
-WEAT is inspired by the Implicit Association Test (IAT) {[}17{]} used in
-psychology, and it uses almost the same word sets, allowing for a
-\emph{prima facie} sensible comparison with bias in humans. In {[}8{]}
-the authors argue that significant biases---thus measured--- similar to
-the ones discovered by IAT can be discovered in word embeddings. In
-{[}11{]} the methodology is extended to a multilingual and cross-lingual
-setting, arguing that using Euclidean distance instead of cosine
-similarity does not make much difference, while the bias effects vary
-greatly across embedding models (interestingly, with social media-text
-trained embeddings being less biased than those based on Wikipedia). A
-similar methodology is employed in {[}4{]}. The authors employ word
-embeddings trained on corpora from different decades to study the shifts
-in various biases.\footnote{Strictly speaking, these authors use
-  Euclidean distances and their differences, but the way they take
-  averages and averages thereof is analogous, and so what we will have
-  to say about pre-averaging leading to false confidence applies to this
-  methodology as well.}
-
-WEAT, however, has been developed to investigate biases corresponding to
-a pair of supposedly opposing stereotypes, and so the question arises as
-to how generalize the measure to contexts in which biases with respect
-to more than two stereotypical groups are to be measured. Such a
-generalization can be found in {[}12{]}. The authors introduce Mean
-Average Cosine distance (MAC) as a measure of bias. Let
-\(T = \{t_1, \dots, t_k\}\) be a set of protected words, and let each
-\(A_j\in \mathcal{A}\) be a set of attributes stereotypically associated
-with a protected word where \(\mathcal{A}\). For instance, when biases
-related to religion are to be investigated, they use a dataset of the
-format illustrated in Table \ref{tab:religionOriginal}. The measure is
-defined as follows:
+\textsf{WEAT} is inspired by the Implicit Association Test (IAT)
+{[}17{]} used in psychology, and it uses almost the same word sets,
+allowing for a \emph{prima facie} sensible comparison with bias in
+humans. In {[}8{]} the authors argue that significant biases---thus
+measured--- similar to the ones discovered by IAT can be discovered in
+word embeddings. In {[}11{]} the methodology is extended to a
+multilingual and cross-lingual setting, arguing that using Euclidean
+distance instead of cosine similarity does not make much difference,
+while the bias effects vary greatly across embedding models
+(interestingly, with social media-text trained embeddings being less
+biased than those based on Wikipedia). A similar methodology is employed
+in {[}4{]}. The authors employ word embeddings trained on corpora from
+different decades to study the shifts in various biases.\footnote{Strictly
+  speaking, these authors use Euclidean distances and their differences,
+  but the way they take averages and averages thereof is analogous, and
+  so what we will have to say about pre-averaging leading to false
+  confidence applies to this methodology as well.}
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.55\linewidth]{paperDraft_files/figure-latex/WEATexample5-1} \end{center}
+\caption{A simple example of a two-class set-up. Two groups of protected words, $X$ and $Y$ with two stereotypical attribute sets. An example of \textsf{WEAT} calculations follows.}
+\label{eq:WEATexample}
+\end{figure}
+
+\noindent Here is an example of \textsf{WEAT} calculations for Figure
+\ref{eq:WEATexample}:
+\begin{align*} s_1 & = s(he,A,B)  =  \nicefrac{(.6+.7)}{2}  - \nicefrac{(.2+.1)}{2} = .65-.15= .5 \\
+s_2  & = s(man,A,B) = .3 \\
+s_3  & = s(woman,A,B) = -.6\\
+s_4 & = s(she, A, B) = -.3\\
+\mathsf{WEAT}(A,B)  & = \frac{\nicefrac{(s_1+s_2)}{2} - \nicefrac{(s_3+s_4)}{2}}{sd(\{s_1,s_2,s_3,s_4\})} \approx 1.93
+\end{align*}
+
+\textsf{WEAT}, however, has been developed to investigate biases
+corresponding to a pair of supposedly opposing stereotypes, and so the
+question arises as to how to generalize the measure to contexts in which
+biases with respect to more than two stereotypical groups are to be
+measured. Such a generalization can be found in {[}12{]}. The authors
+introduce Mean Average Cosine distance (\textsf{MAC}) as a measure of
+bias. Let \(T = \{t_1, \dots, t_k\}\) be a set of protected words, and
+let each \(A_j\in \mathcal{A}\) be a set of attributes stereotypically
+associated with a protected word where \(\mathcal{A}\). For instance,
+when biases related to religion are to be investigated, they use a
+dataset of the format illustrated in Table \ref{tab:religionOriginal}.
+The measure is defined as follows:
 
@@
 
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.5\linewidth]{paperDraft_files/figure-latex/MACexample-1} \end{center}
+\caption{A small subset of the religion dataset.
+To each protected word in $T$  there corresponds one class of stereotypical attributes typically associated with it (and other classes of stereotypical attributes associated with different protected words).}
+\label{fig:MACexample}
+\end{figure}
+
+\noindent An example of \textsf{MAC} calculations for the situation
+depicted in Figure \ref{fig:MACexample} is as follows: \begin{align*}
+s_1   = s(muslim,A_1)  & = \frac{\mathsf{cos}(muslim,dirty)+\mathsf{cos}(muslim,terrorist)}{2} \\ 
+s_2   = s(muslim,A_2)  & = \frac{\mathsf{cos}(muslim,familiar)+\mathsf{cos}(muslim,conservative)}{2}\\ & \vdots \\ 
+\mathsf{MAC}(T,A)  & = \mathsf{mean}(\{s_i \vert i \in 1, \dots, k\})
+\end{align*}
+
+Notably, the inutitive distinction between different attribute sets
+plays no real role in the \textsf{MAC} calculations. Equally well one
+could calculate the mean distance of \emph{muslim} to all the
+predicates, mean distance of \emph{christian} to all the predicates,
+mean distance of \emph{jew} to all the predicates, and then to take the
+mean of these three means.
+
 Having introduced the measures, first, we will introduce a selection of
@@
 
-\hypertarget{general-methodological-problems-with-cosine-based-measures-of-bias}{%
-\section{General methodological problems with cosine-based measures of
-bias}\label{general-methodological-problems-with-cosine-based-measures-of-bias}}
-
-Consider the MAC results from Table REF. Are the initial MAC values
-lower than 1 indicative of the presence of bias? Of course, thinking
-abstractly, 1 is the ideal distance for unrelated words. But clearly,
-there is variation in distances, which might lead to non-biased lists
-also having MAC smaller than 1. How much smaller? While the original
-paper did not employ control lists for comparison in the paper. To
-obtain a more realistic gauge on how to understand MAC values (and in
-further development of our approach), we introduced control word lists.
-One of them is a list of words we intuitively considered neutral (see
-REF in the APPENDIX). Moreover, it might be the case that words that
-have to do with human activities are systematically closer to the
-protected words than merely neutral words, for which reason we also use
-a second list with intuitively non-stereotypical human attributes (see
-REF in the APPENDIX). Another important observation is that in their
-calculations the authors do not distinguish whether a given attribute is
-associated with a given protected word (e.g.~protected word \emph{jew}
-and the attribute \emph{greedy}) or whether it is associated with a
-different stereotype from the same list (e.g.~protected word
-\emph{quran} and the attribute \emph{greedy}), simply averaging across
-all such groups.
-
-In Figure \ref{fig:empirical} We look at the empirical distributions,
-while paying attention to such divisions. Horizontal lines represent
-1-MAC values that the authors considered indicative of bias for
-stereotypes corresponding to given word lists. For instance, in
-religion, MAC was .859, so we plot \(0\pm (1-.859)\approx .14\) lines
-around similarity = 0 (that is, distance = 1). Notice that most
-distributions are quite wide, and the proportions of even neutral or
-human neutral words with similarities higher than the value of 1-MAC
-deserving debiasing according to the authors is quite high.
-
-\pagebreak
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8b-1} \end{center}
-
-\caption{Empirical distributions of cosine similarities (1-distances) for three word lists used in  the original paper.  }
-
-\label{fig:empirical}
-\end{figure}
-
-One issue to consider is the selection of attributes for bias
-measurement. The word lists used in the literature are often fairly
-small (5-50).
-
-While the papers do employ statistical tests to measure the uncertainty
-involved, we will later on argue that these method are not proper for
-the goal at hand and show that a more appropriate use of statistical
-methods leads to estimates of uncertainty that are rather
-epistemologically pessimistic. \todo{sensitivity to choice of words}
-
-Let's think about MAC, using the case of religion-related stereotypes.
-The full lists from {[}12{]} can be found in
-\ref{appendix:manzini_word_lists}. In the original paper, words from all
-three religions were compared against all of the stereotypes. One reason
-this is problematic is that no distinction between cases in which the
-stereotype is associated with a given religion, as opposed to the
-situation in which it is associated with another one, is made. This is
-problematic, as not all of the stereotypical words have to be considered
-as harmful for all of the religions. One should investigate the
-religions separately as some of them may have stronger harmful
-associations that others.
-
-The interpretation of the results is also a challenge. In {[}12{]} we
-can find summaries of average cosine distances per group (such as
-gender, race, or religion). For instance, for religion, here is the
-relevant fragment of table:
-
-\begin{table}
+\hypertarget{methodological-problems-with-cosine-based-bias-metrics}{%
+\section{Methodological problems with cosine-based bias
+metrics}\label{methodological-problems-with-cosine-based-bias-metrics}}
+
+Here is an example of results of debiasing that can be found in {[}12{]}
+(Table \ref{tab:religionOriginal2}). It presents the results of applying
+the procedure with respect to the Religion dataset. \textsf{MAC} scores
+are given for the original datasets, and two resulting from two
+debiasing methods applied to the Reddit embedding. The scores for
+debiasing come with p-values for the comparison to the original
+\textsf{MAC}.
+
+\begin{table}[H]
 \footnotesize
@@
 \caption{The associated mean average cosine similarity
-(MAC) and p-values for debiasing methods for religious bias}
+(\textsf{MAC}) and p-values for debiasing methods for religious bias.}
 \label{tab:religionOriginal2}
@@
 
-\noindent (MAC stand for mean average cosine similarity, although in
-reality the the table contains mean cosine distances). What may attract
-attention is the fact that the value of cosine distance in ``Biased''
-category is already quite high (i.e.~close to 1) even before debiasing.
-High cosine distance indicates low cosine similarity between values. One
-could think that average cosine similarity equal to approximately 0.141
-is not significant enough to consider it as bias. However, the authors
-still aim to mitigate such ``bias'' to make the distance even larger.
-Methodologically the question is, on what basis is this small similarity
-still considered as a proof of the presence of bias, and whether these
-small changes are meaningful.
-
-The underlying problem here is that in the paper there is no control
-group. One should also include control groups to have a way of comparing
-the results for a supposedly stereotyped group with the results for sets
-of neutral or human-related neutral words. In our approach later on, we
-distinguish between stereotypes associated with a given group,
-stereotypes associated with different groups, and introduce control
-groups: neutral words and stereotype-free human predicates.
-
-\hypertarget{metrics-that-pre-average-are-a-bad-guide}{%
-\subsection{Metrics that pre-average are a bad
-guide}\label{metrics-that-pre-average-are-a-bad-guide}}
-
-In contrast, statistical intervals will help us decide whether a given
-cosine similarity is high enough to consider the words to be more
-similar than if we chose them at random. We will use highest posterior
-density intervals, in line with Bayesian methodology.
-
-Crucially, these approaches use means of mean average cosine
-similarities to measure similarity between protected word and harmful
-stereotypes. If one takes a closer look at the individual values that
-are taken for the calculations, it turns out that there are quite a few
-outliers and surprisingly dissimilar words. This problem becomes
-transparent when we examine the visualizations of the individual cosine
-distances, following the idea that one of the first steps in
-understanding data is to look at it.
+The first conceptual question we should ask is whether the initial
+\textsf{MAC} values lower than 1 indeed are indicative of the presence
+of bias? Of course, thinking abstractly, 1 is the ideal distance for
+unrelated words. But clearly, there is variation in distances, which
+might lead to non-biased lists also having \textsf{MAC} smaller than 1.
+How much smaller? What may attract attention is the fact that the value
+of cosine distance in ``Biased'' category is already quite high
+(i.e.~close to 1) even before debiasing. High cosine distance indicates
+low cosine similarity between values. One could think that average
+cosine similarity equal to approximately 0.141 is not large enough to
+consider it as bias. The authors, though, still aim to mitigate such
+``bias'' to make the distance even larger. The question is, on what
+basis is this small similarity still considered as a proof of the
+presence of bias, and whether these small changes are meaningful.
+
+The problem is that the original paper did not employ any control group
+of neutral attributes for comparison to obtain a more realistic gauge on
+how to understand \textsf{MAC} values. Later on, in our approach, we
+introduce such control word lists. One of them is a list of words we
+intuitively considered neutral (see the Appendix for word lists).
+
+Moreover, it might be the case that words that have to do with human
+activities in general, even if unbiased, are systematically closer to
+the protected words than merely neutral words. This, again, casts doubt
+on whether comparing \textsf{MAC} to the abstractly ideal value of 1 is
+a methodologically sound idea. For this reason we also use a second list
+with intuitively non-stereotypical human attributes (again, see the
+Appendix for a word list).
+
+Another important observation is that \textsf{MAC} calculations do not
+distinguish whether a given attribute is associated with a given
+protected word, simply averaging across all such groups. Let us use the
+case of religion-related stereotypes to illustrate. The full lists from
+{[}12{]} can be found in Appendix \ref{appendix:manzini_word_lists}. In
+the original paper, words from all three religions were compared against
+all of the stereotypes. No distinction between cases in which the
+stereotype is associated with a given religion, as opposed to the
+situation in which it is associated with another one, is made. For
+example, the protected word \emph{jew} is supposed to be stereotypically
+connected with the attribute \emph{greedy}, while from the protected
+word \emph{quran} the attribute \emph{greedy} comes from a different
+stereotype, and yet the distances between these pairs contribute equally
+to the final \textsf{MAC} score. This is problematic, as not all of the
+stereotypical words have to be considered as harmful for all of the
+religions. To avoid the masking effect, one should pay attention to how
+protected words and attributes are paired by stereotypes.
+
+In Figure \ref{fig:empirical} we look at the empirical distributions,
+while paying attention to such divisions. The horizontal lines represent
+the values of 1-\textsf{MAC} (that is, we now talk in terms of cosine
+similarity rather than cosine distance; this is a trivial linear
+transformation) that the authors considered indicative of bias for
+stereotypes corresponding to given word lists. For instance, in
+religion, \textsf{MAC} was .859 and considered a sign of bias, so we
+plot \(0\pm (1-.859)\approx .14\) lines around similarity = 0 (that is,
+distance = 1). Notice that most distributions are quite wide, and the
+proportions of even neutral or human neutral words with similarities
+higher than the value of 1-\textsf{MAC} deserving debiasing according to
+the authors is quite high.
+
+\pagebreak
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8b-1} \end{center}
+
+\caption{Empirical distributions of cosine similarities (1-distances) for three word lists used in  the original paper.  }
+
+\label{fig:empirical}
+\end{figure}
+
+Another issue to consider is the selection of attributes for bias
+measurement. The word lists used in the literature are often fairly
+small (5-50). The papers in the field do employ statistical tests to
+measure the uncertainty involved and do make claims of statistical
+significance. Yet, we will later on argue that these method are not
+proper for the goal at hand. By using Bayesian methods we will show that
+a more appropriate use of statistical methods leads to estimates of
+uncertainty which suggest that larger word lists would be advisable.
+
+To avoid such problems, later on, we employ control groups and in line
+with Bayesian methodology, use posterior distributions and highest
+posterior density intervals instead of chasing single-point metrics
+based on pre-averaged data. But before we do so, let us spend some time
+explaining why pre-averaging is a sub-optimal strategy.
+
+\hypertarget{pre-averaging-bootstrapping-and-testing-with-the-null-model}{%
+\section{Pre-averaging, bootstrapping and testing with the null
+model}\label{pre-averaging-bootstrapping-and-testing-with-the-null-model}}
+
+The approaches we have been describing use means of mean average cosine
+similarities to measure similarity between protected words and
+attributes coming from harmful stereotypes. However, a closer inspection
+look at the individual values that are taken for the calculations, it
+turns out that there are quite a few outliers and surprisingly
+dissimilar words. This problem becomes transparent when we examine the
+visualizations of the individual cosine distances, following the idea
+that one of the first steps in understanding data is to look at it.
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visMuslim-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visPriest-1} \end{center}
+
+\caption{Actual distances for two protected words.
+For the protected word "muslim", the most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates.
+For the word "priest", the situation is even less as expected: the nearest attributes are human attributes, and all there seems to be no clear pattern to the distances to other attributes.}
+\end{figure}
 
@@
 The t-tests they employ are run on average cosines used to calculate
-MAC.
-
-\hypertarget{pre-averaging-bootstrapping-and-testing-with-the-null-model}{%
-\section{Pre-averaging, bootstrapping and testing with the null
-model}\label{pre-averaging-bootstrapping-and-testing-with-the-null-model}}
+\textsf{MAC}.
 
@@
 
-In the calculations of WEAT means are taken twice. The \(s\)-values
-themselves are means, and then means of \(s\)-values are compared
-between groups. The test statistic used itself is of this sort as
-well---while it compares sums, these are sums of sets of the same size,
-so they are really linear transformations of means of means. Statistical
-troubles start when we run statistical tests on sets of means, for at
-least two reasons One, by pre-averaging data we throw away information
-about sample sizes. For the former point, think about proportions: 10
-out of 20 and 2 out of 4 give the same mean, but you would obtain more
-information by making the former observation rather than by making the
-latter. And especially in this context, in which the word lists are not
-huge, sample sizes should matter. Two, when we pre-average, we remove
-variation, and so pre-averaging tend to manufacture false confidence.
-Group means display less variation than the raw data points, and the
-standard deviation of a set of means of sets of means is bound to be
-lower than the original standard deviation in the row data. Now, if you
-calculate your effect size by dividing by the pre-averaged standard
-deviation, your quite likely to get something that looks like a strong
-effect size, but the results of your calculations might not track
-anything interesting.
+In the calculations of \textsf{WEAT} means are taken twice. The
+\(s\)-values themselves are means, and then means of \(s\)-values are
+compared between groups. The test statistic used itself is of this sort
+as well---while it compares sums, these are sums of sets of the same
+size, so they are really linear transformations of means of means.
+Statistical troubles start when we run statistical tests on sets of
+means, for at least two reasons One, by pre-averaging data we throw away
+information about sample sizes. For the former point, think about
+proportions: 10 out of 20 and 2 out of 4 give the same mean, but you
+would obtain more information by making the former observation rather
+than by making the latter. And especially in this context, in which the
+word lists are not huge, sample sizes should matter. Two, when we
+pre-average, we remove variation, and so pre-averaging tend to
+manufacture false confidence. Group means display less variation than
+the raw data points, and the standard deviation of a set of means of
+sets of means is bound to be lower than the original standard deviation
+in the row data. Now, if you calculate your effect size by dividing by
+the pre-averaged standard deviation, your quite likely to get something
+that looks like a strong effect size, but the results of your
+calculations might not track anything interesting.
 
@@
 the low p-values obtained from taking means or means of means as data
-points. In this section, we discussed the performance of the WEAT
-measure, but since the {[}12{]} one is a generalization thereof,
-including the method of running statistical tests on pre-averaged data,
-our remarks, \emph{mutatis nutandis}, apply. Moreover, a general problem
-is that studies such as {[}8{]} or {[}12{]} do not use control groups,
-and this makes the interpretation of the analysis additionally
-difficult.
+points. In this section, we discussed the performance of the
+\textsf{WEAT} measure, but since the {[}12{]} one is a generalization
+thereof, including the method of running statistical tests on
+pre-averaged data, our remarks, \emph{mutatis nutandis}, apply.
+Moreover, a general problem is that studies such as {[}8{]} or {[}12{]}
+do not use control groups, and this makes the interpretation of the
+analysis additionally difficult.
 
@@
 Furthermore, conceptually there are two levels of analysis in our
-approach. On the one hand, we are interested in the general question of
-whether related attributes are systematically closer across the dataset.
-On the other hand, we are interested in a more fine-grained picture of
-the role of the predictor for particular protected words. Learning in
-hierarchical Bayesian models involves using Bayesian inference to update
-the parameters of the model. This update is based on the observed data,
-and estimates are made at different levels of the data hierarchy. We use
-hierarchical Bayesian models in which we simultaneously estimate
-parameters at the protected word level and at the global level, assuming
-that all lower-level parameters are drawn from global distributions.
-Such models can be thought of as incorporating adaptive regularization,
-which avoids overfitting and leads to improved estimates for unbalanced
-datasets (and the datasets we need to use are unbalanced).
+approach (see Figure \ref{fig:visDag}). On the one hand, we are
+interested in the general question of whether related attributes are
+systematically closer across the dataset. On the other hand, we are
+interested in a more fine-grained picture of the role of the predictor
+for particular protected words. Learning in hierarchical Bayesian models
+involves using Bayesian inference to update the parameters of the model.
+This update is based on the observed data, and estimates are made at
+different levels of the data hierarchy. We use hierarchical Bayesian
+models in which we simultaneously estimate parameters at the protected
+word level and at the global level, assuming that all lower-level
+parameters are drawn from global distributions. Such models can be
+thought of as incorporating adaptive regularization, which avoids
+overfitting and leads to improved estimates for unbalanced datasets (and
+the datasets we need to use are unbalanced).
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/dagVis-1} \end{center}
+
+\caption{At a general level, we will be estimating the coefficients for distances as grouped by whether they are between protected words and attributes coming from their respective associated/different/human/neutral attribute groups.
+At a more fine-grained level, for each protected word we will be estimating the proximity of that word to attributes that are associated with its respective stereotype, come from a different stereotype, or come from the human/neutral attribute lists.}
+\label{fig:visDag}
+\end{figure}
 
@@
 observations that are not exactly in line with the usual picture painted
-in terms of WEAT or MAC (and the problem generalizes to any approach
-that focuses on chasing a single numeric metric): the word list sizes
-and sample sizes used in the studies are usually
-small,\footnote{Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25}
+in terms of \textsf{WEAT} or \textsf{MAC} (and the problem generalizes
+to any approach that focuses on chasing a single numeric metric): the
+word list sizes and sample sizes used in the studies are usually small.
 posterior density intervals are fairly wide, often the differences

### 2023-05-17 paper/paperDraft.Rmd
fixed figures

@@
 \begin{figure}[H]
-```{r visMuslim,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%",message =FALSE, warning = FALSE}
+```{r visMuslimd,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%",message =FALSE, warning = FALSE}
 religion <- read.csv("datasets/cleanedDatasets/religionRedditCleaned.csv")
@@
 
-visualiseProtected(muslimVis,"muslim")+theme_void(base_size = 20)
-```
-
-```{r visPriest,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
+visualiseProtected(muslimVis,"muslim")+theme_void(base_size = 11)
+```
+
+```{r visPriestd,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
 religion <- read.csv("datasets/cleanedDatasets/religionRedditCleaned.csv")
@@
 
-visualiseProtected(priestVis,"priest")+theme_void(base_size = 20)
+visualiseProtected(priestVis,"priest")+theme_void(base_size = 11)
 ```
@@
 \footnotesize
-```{r generatePlots2,echo=FALSE, eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
+```{r generatePlots2h,echo=FALSE, eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
 source("functions/cleanDataset.R")
@@
                                      embedding = "Google",
-                                     ylims = c(.75,1.1))
+                                     ylims = c(.8,1.05))
 
@@
 \begin{figure}[H]
-```{r fig:debiasedCosinePair2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, out.width="85%", warning = FALSE, message = FALSE}
+```{r fig:debiasedCosinePair2b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, out.width="85%", warning = FALSE, message = FALSE}
 
@@
 
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \begin{figure}[H] -->
+
+```{r weatjoint2r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat7Google$plotJoint
 ```
-\caption{dsds}
-\label{fig:weat7google}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:weat7google} -->
+<!-- \end{figure} -->
+
+\pagebreak
+
+<!-- \begin{figure}[H] -->
+
+```{r weat7glovf, echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat7Glove$plotJoint
 ```
-\caption{dsds}
-\label{fig:weat7glove}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:weat7glove} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsWeat7Reddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat7Reddit$plotJoint
 ```
-\caption{dsds}
-\label{fig:weat7reddit}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:weat7reddit} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsWeat1Google,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat1Google$plotJoint
 ```
-\caption{dsds}
-\label{fig:weat1google}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:weat1google} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsWeat1Glove,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat1Glove$plotJoint
 ```
-\caption{dsds}
-\label{fig:weat1glove}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:weat1glove} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsWeat1Reddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat1Reddit$plotJoint
 ```
-\caption{dsds}
-\label{fig:weat1reddit}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:weat1reddit} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsReligionGoogle,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsReligionGoogle$plotJoint
 ```
-\caption{dsds}
-\label{fig:religionGoogle}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:religionGoogle} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsReligionGlove,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsReligionGlove$plotJoint
 ```
-\caption{dsds}
-\label{fig:religionGlove}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:religionGlove} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsReligionReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsReligionReddit$plotJoint
 ```
-\caption{dsds}
-\label{fig:religionReddit}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:religionReddit} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsRaceGoogle,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsRaceGoogle$plotJoint
 ```
-\caption{dsds}
-\label{fig:raceGoogle}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:raceGoogle} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsRaceGlove,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsRaceGlove$plotJoint
 ```
-\caption{dsds}
-\label{fig:raceGlove}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:raceGlove} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsRaceReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsRaceReddit$plotJoint
 ```
-\caption{dsds}
-\label{fig:raceReddit}
-\end{figure}
-
-\begin{figure}[H]
-
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:raceReddit} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+
+```{r resultsGenderGoogle,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsGenderGoogle$plotJoint
 ```
-\caption{dsds}
-\label{fig:genderGoogle}
-\end{figure}
-
-\begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:genderGoogle} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+```{r resultsGenderGlove,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsGenderGlove$plotJoint
 ```
-\caption{dsds}
-\label{fig:genderGlove}
-\end{figure}
-
-\begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:genderGlove} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+```{r resultsGenderReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsGenderReddit$plotJoint
 ```
-\caption{dsds}
-\label{fig:genderReddit}
-\end{figure}
-
-\begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:genderReddit} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+```{r resultsDebiasedReligionReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsDebiasedReligionReddit$plotJoint
 ```
-\caption{dsds}
-\label{fig:debiasedReligion}
-\end{figure}
-
-\begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:debiasedReligion} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+```{r resultsDebiasedRaceReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsDebiasedRaceReddit$plotJoint
 ```
-\caption{dsds}
-\label{fig:debiasedRace}
-\end{figure}
-
-\begin{figure}[H]
-```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE}
+<!-- \caption{dsds} -->
+<!-- \label{fig:debiasedRace} -->
+<!-- \end{figure} -->
+
+<!-- \begin{figure}[H] -->
+```{r resultsDebiasedGenderReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsDebiasedGenderReddit$plotJoint
 ```
-\caption{dsds}
-\label{fig:debiasedGender}
-\end{figure}
+<!-- \caption{dsds} -->
+<!-- \label{fig:debiasedGender} -->
+<!-- \end{figure} -->
 

### 2023-05-17 paper/paperDraft.tex
fixed figures

@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visMuslim-1} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visPriest-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visMuslimd-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visPriestd-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2-1} \includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2-2} \end{center}
+\begin{center}\includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2b-1} \includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2b-2} \end{center}
 \caption{Means with 89\% highest posterior density intervals for gender before and after debiasing.}
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-1-1} \end{center}
-\caption{dsds}
-\label{fig:weat7google}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-2-1} \end{center}
-\caption{dsds}
-\label{fig:weat7glove}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-3-1} \end{center}
-\caption{dsds}
-\label{fig:weat7reddit}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-4-1} \end{center}
-\caption{dsds}
-\label{fig:weat1google}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-5-1} \end{center}
-\caption{dsds}
-\label{fig:weat1glove}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-6-1} \end{center}
-\caption{dsds}
-\label{fig:weat1reddit}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-7-1} \end{center}
-\caption{dsds}
-\label{fig:religionGoogle}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-8-1} \end{center}
-\caption{dsds}
-\label{fig:religionGlove}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-9-1} \end{center}
-\caption{dsds}
-\label{fig:religionReddit}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-10-1} \end{center}
-\caption{dsds}
-\label{fig:raceGoogle}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-11-1} \end{center}
-\caption{dsds}
-\label{fig:raceGlove}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-12-1} \end{center}
-\caption{dsds}
-\label{fig:raceReddit}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-13-1} \end{center}
-\caption{dsds}
-\label{fig:genderGoogle}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-14-1} \end{center}
-\caption{dsds}
-\label{fig:genderGlove}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-15-1} \end{center}
-\caption{dsds}
-\label{fig:genderReddit}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-16-1} \end{center}
-\caption{dsds}
-\label{fig:debiasedReligion}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-17-1} \end{center}
-\caption{dsds}
-\label{fig:debiasedRace}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-18-1} \end{center}
-\caption{dsds}
-\label{fig:debiasedGender}
-\end{figure}
+\begin{center}\includegraphics{paperDraft_files/figure-latex/weatjoint2r-1} \end{center}
+
+\pagebreak
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/weat7glovf-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat7Reddit-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Google-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Glove-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Reddit-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGoogle-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGlove-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionReddit-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGoogle-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGlove-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceReddit-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGoogle-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGlove-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderReddit-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedReligionReddit-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedRaceReddit-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedGenderReddit-1} \end{center}
 

### 2023-05-17 functions/plotFromPrecis.R

@@ -53,7 +53,7 @@ plotFromPrecis <- function (precis, dataset, ylims = c(.5,1.2), list, embedding)
     geom_segment(data = DFindividual,aes(x=reorder(pw, desc(pw)),xend=reorder(pw, desc(pw)), y  = `5.5%`, yend= `94.5%`,
                                          color = type),
                  position = position_nudge((as.integer(DFindividual$type)-1)*0.15), size = .4)+
-    geom_hline(yintercept = 1, lty = 2, size = .2, alpha = .3) +labs(title = paste(list, ", ", embedding, ", cosine distances, by protected-word", sep = ""), subtitle = "Means with 89% highest posterior density intervals")+ 
+    geom_hline(yintercept = 1, lty = 2, size = .2, alpha = .3) +labs(title = paste(list, ", ", embedding, ", cosine distances, by protected word", sep = ""), subtitle = "Means with 89% highest posterior density intervals")+ 
     theme(legend.position='none', plot.title.position = "plot") + ylim(ylims)
   
   
@@ -74,14 +74,15 @@ plotFromPrecis <- function (precis, dataset, ylims = c(.5,1.2), list, embedding)
   
   
   grobIndividual <- ggplotGrob(results$plotIndividual+theme_tufte(base_size = 6) + theme(legend.position='none'))
-  grobOverall <- ggplotGrob(results$plotOverall+theme_tufte(base_size = 6)+ theme(legend.position = c(0.9, 0.7)))
+  grobOverall <- ggplotGrob(results$plotOverall+theme_tufte(base_size = 6)+ theme(legend.position = "top"))
   
   
-  results$plotJoint <- ggplot(data.frame(a=1)) + xlim(1, 20) + ylim(10, 60)+theme_void()+
-    annotation_custom(grobIndividual, xmin = 1, xmax = 20, ymin = 10, ymax = 50)+
-    annotation_custom(grobOverall, xmin = 1, xmax = 20, ymin = 50, ymax = 60)  
+  results$plotJoint <- ggplot(data.frame(a=1)) + xlim(1, 20) + ylim(10, 100)+theme_void()+
+    annotation_custom(grobIndividual, xmin = 1, xmax = 20, ymin = 10, ymax = 80)+
+    annotation_custom(grobOverall, xmin = 1, xmax = 20, ymin = 80, ymax = 100)  
   
   return(results)
 }
+
 
 

### 2023-05-17 functions/visualisationTools.R

@@ -37,8 +37,8 @@ visualiseProtected <- function(dataset, protected){
     geom_point(aes(x=xaxis,y=yaxis, color = connection, size = cosineSimilarity), alpha = 0.3)+ 
     geom_segment(aes(x= 0, y = 0, xend = closestx,yend=closesty),alpha= 0.1, size = 0.2)+
     geom_segment(aes(x= 0, y = 0, xend = furthestx,yend=furthesty),alpha=0.1,size = 0.2)+
-    annotate("text", x = 0.5*closestx, y = 0.5*closesty-0.05, label = round(closestDistance,3), size = 3)+
-    annotate("text", x = 1.5*furthestx, y = 0.5*furthesty, label = round(furthestDistance,3), size = 3)+
+    annotate("text", x = 0.5*closestx, y = 0.5*closesty-0.05, label = round(closestDistance,2), size = 3)+
+    annotate("text", x = 1.9*furthestx, y = 0.5*furthesty, label = round(furthestDistance,2), size = 3)+
     annotate("text", x = -0.02, y = -0.13, label = protected, size = 4)+
     geom_text(aes(label = wordToCompare,x=xaxis+.07,y=yaxis+0.03,color = connection), size = 3)+
     theme_void()+

### 2023-05-25 paper/paperDraft.Rmd
revised visualizations, fixed intro

@@
 
-Natural language processing (NLP) models are used to perform various language-related tasks such as providing email filters, intelligent assistants, search results,  translations, text analysis and so on.
-Such tools require input words to be represented by  vectors of real numbers.
-This can usually be done with word embeddings.
-A common way to construct an embedding is to use  a large natural language corpus to train  a  neural network to assign vectors to words that optimize for co-occurrence prediction accuracy.
-The vectors can then be compared in terms of their similarity -- the usual measure is cosine similarity -- and the results of such comparisons can be used in downstream tasks.
-Roughly speaking, cosine similarity is an imperfect mathematical proxy for semantic similarity [@Mikolov2013efficient].
-
-It has been suggested [@Bolukbasi2016man; @Caliskan2017semanticsBiases; @Gonen2019lipstick; @Lauscher2019multidimensional; @Garg2018years; @Manzini2019blackToCriminal]  that  such models can learn implicit biases that reflect harmful stereotypical thinking---for example,  the (vector corresponding to the) word \textit{she} might be  much closer in the vector space to  the word \textit{cooking} than  the word \textit{he}.
-Such phenomena may be undesirable at least  in some  downstream tasks, such as web search, recommendations, and so on.
+It has been suggested^[See for instance [@Bolukbasi2016man; @Caliskan2017semanticsBiases; @Gonen2019lipstick; @Lauscher2019multidimensional; @Garg2018years; @Manzini2019blackToCriminal].]  that  language models can learn implicit biases that reflect harmful stereotypical thinking---for example,  the (vector corresponding to the) word \textit{she} might be  much closer in the vector space to  the word \textit{cooking} than  the word \textit{he}.
+Such phenomena are undesirable at least  in some  downstream tasks, such as web search, recommendations, and so on.
 To investigate such issues, several measures of bias in word embeddings have been formulated and applied.
 Our goal is to use two prominent examples of such measures to argue that this approach oversimplifies the situation and to develop a Bayesian alternative.
+
+A common approach in natural language processing is to  represent words  by  vectors of real numbers---such representations are called \emph{embeddings}.
+One way to construct an embedding---we will focus our attention on non-contextual language models^[One example of a contextualized representation is BERT.
+Another is GPT.]---is to use  a large corpus to train  a  neural network to assign vectors to words in a way that optimizes for co-occurrence prediction accuracy.
+Such vectors can then be compared in terms of their similarity---the usual measure is cosine similarity---and the results of such comparisons can be used in downstream tasks.
+Roughly speaking, cosine similarity is an imperfect mathematical proxy for semantic similarity [@Mikolov2013efficient].
 
@@
 Another response might be to insist that there is no bias introduced by the use of machine learning methods here, since the algorithm is simply learning to correctly predict co-occurrences based on what "reality" looks like.
-However, this objection overlooks the fact that we, humans, are the ones who construct this linguistic reality, is shaped in part by the natural language processing tools we use on a massive scale.
-Sure, if there is unfairness and our goal is to diagnose it, we should do complete justice to learning it in the model used in a study thereof.
-One example of this approach is [@Garg2017hundredYears].
-However, if our goal is to develop downstream tools that perform tasks that we care about without further perpetuating  or exacerbating harmful stereotypes, we still have good reasons to try to minimize the negative impact, at least as long as this does not damage the overall performance of the tool.
+However, this objection overlooks the fact that we, humans, are the ones who construct this linguistic reality, which is shaped in part by the natural language processing tools we use on a massive scale.
+Sure, if there is unfairness and our goal is to diagnose it, we should do complete justice to learning it in the model used to study it.
+One example of this approach is [@Garg2017hundredYears], where the authors use language  models to study the shape of certain biases across a century.
+
+However, if our goal is to develop downstream tools that perform tasks that we care about without further perpetuating  or exacerbating harmful stereotypes, we still have good reasons to try to minimize the negative impact.
 Moreover, it is often not the case that the corpora mirror reality---to give a trivial example, heads are spoken of more often than kidneys, but this does not mean that kidneys occur much less often in reality than heads.
-To give a more relevant example, the disproportionate association of female words with female occupations in a corpus actually greatly exaggerates the actual lower disproportion in the real distribution of occupations.
-[@gordon2012reporting].
-
-In what follows, we focus on  two popular  measures of bias applicable to many existing word embeddings ( \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.} \emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.} and \emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}): \emph{Word Embedding Association Test}  (\textsf{WEAT}) [@Caliskan2017semanticsBiases],  and \emph{Mean Average Cosine Distance} (\textsf{MAC})  [@Manzini2019blackToCriminal].
+To give a more relevant example, the disproportionate association of female words with female occupations in a corpus actually greatly exaggerates the actual lower disproportion in the real distribution of occupations [@gordon2012reporting].
+
+In what follows, we focus on  two popular  measures of bias applicable to many existing word embeddings, such as \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.} \emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.} and \emph{Reddit Corpus}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}: \emph{Word Embedding Association Test}  (\textsf{WEAT}) [@Caliskan2017semanticsBiases],  and \emph{Mean Average Cosine Distance} (\textsf{MAC})  [@Manzini2019blackToCriminal].
 We first explain how these measures are supposed to work.
-Then we argue that they are problematic for various reasons, a key one being that by pre-averaging data they manufacture false confidence, which we illustrate in terms of simulations showing that the measures often suggest the existence of bias even if by design it is non-existent in a simulated dataset.
-
-We  propose to replace them with a Bayesian data analysis (hierarchical models), which not only provides more modest and realistic assessment of the uncertainty involved, but also allows for inspection at various levels of granularity.
-Once we introduce the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of \textsf{WEAT} or \textsf{MAC} (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small,^[Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25.]  posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
-Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable.
-The tools that we have proposed, however, allow for a more fine-grained and multi-level evaluation of bias and debiasing in language models without abandoning modesty about the uncertainties involved.
+Then we argue that they are problematic for various reasons---the key one being that by pre-averaging data they manufacture false confidence, which we illustrate in terms of simulations showing that the measures often suggest the existence of bias even if by design it is non-existent in a simulated dataset.
+
+We  propose to replace them with a Bayesian data analysis, which not only provides more modest and realistic assessment of the uncertainty involved, but in which hierarchical models allow for inspection at various levels of granularity.
+Once we introduce the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of \textsf{WEAT} or \textsf{MAC}.
+
+Most of the  problems that we point out generalize to any existing approach that focuses on chasing a single numeric metric of bias.
+(1) They treat the results of pre-averaging as raw data in statistical significance tests, which in this context is bound to overestimate significance.
+We show similar results can easily be obtained when sampling from null models with no bias.
+(2) The word list sizes and sample sizes used in the studies are usually small,^[Depending on a list for [@Caliskan2017semanticsBiases] the range for protected words is between 13 and 100, and for attributes between 16 and 25; for [@Manzini2019blackToCriminal] the range for protected words is between 14 and 18, and for attributes between 11 and 25.]
+(3) Many studies do not use any control predicates, such as random neutral words or neutral human predicates for comparison.
+
+On the constructive side, we develop and deploy our method, and the results are, roughly, as follows.
+(A) Posterior density intervals are fairly wide and    the average differences between associated, different and neutral human predicates, are not very large.
+(B) A preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable.
+
+In Section \ref{sec:two} we describe the two key measures discussed in this paper, \textsf{WEAT} and \textsf{MAC}, explaining how they are calculated and how they are supposed to work.
+In Section \ref{sec:challenges} we first argue in Subsection \ref{subsec:interpretability}, that it is far from clear how results given in terms of \textsf{WEAT} or \textsf{MAC} are to be interpreted.
+Second, in Subsection \ref{subsec:problems} we explain the statistical problems that arise when one uses pre-averaged data in such contexts, as these measures do.
+In Section \ref{sec:bayesian} we explain the alternative Bayesian approach that we propose.
+In Section \ref{sec:results} we elaborate on the results that it leads to, including a somewhat skeptical view of the efficiency of debiasing methods, discussed in Subsection \ref{subsec:rethinking}.
+Finally, in Section \ref{sec:related} we spend some time placing our results in the ongoing discussions.^[\textbf{Disclaimer:} throughout the paper we will be mentioning and using word lists and stereotypes we did not formulate, which does not mean we condone any judgment made therein or underlying a given word selection.
+For instance, the Gender dataset does not recognize non-binary categories, and yet we use it without claiming that such categories should be ignored.]
 
 # Two measures of bias: WEAT and MAC
-
-Conceptually the idea is that if a particular harmful stereotype is learned in a particular embedding, then certain groups of words will be systematically closer to (or further from) each other.
-This gives rise to the idea of protected groups---for example, in guiding online search completion or recommendation, female words might require protection in that they should not be systematically closer to stereotypically female job names, and male words require protection in that they should not be systematically closer to toxic masculinity stereotypes, such as "tough", "never complaining" or "macho".^[However, for some research-related purposes, such as the study of stereotypes across history [@Garg2017hundredYears], embeddings which do not protect certain classes may also be useful.]
+\label{sec:two}
+
+Conceptually, the idea is that if a particular harmful stereotype is learned in a particular embedding, then certain groups of words will be systematically closer to (or further from) each other.
+This gives rise to the idea of protected groups---for example, in guiding online search completion or recommendation, female words might require protection in that they should not be systematically closer to stereotypically female job names, such as "nurse", "librarian", "waitress", and male words require protection in that they should not be systematically closer to toxic masculinity stereotypes, such as "tough", "never complaining" or "macho".^[However, for some research-related purposes, such as the study of stereotypes across history [@Garg2017hundredYears], embeddings which do not protect certain classes may also be useful.]
 
 The key role in the measures to be discussed is played by the notion of cosine distance (or, symmetrically, by cosine similarity).
-These are defined as follows:^[Here, "$-$" stands for point-wise difference, "$\cdot$" stands for the dot product operation, and $\lVert a\rVert  = \sqrt{(a \cdot a)}$.]
+These are defined as follows:^[Here, "$-$" stands for point-wise difference, "$\cdot$" stands for the dot product operation, and $\lVert a\rVert  = \sqrt{(a \cdot a)}$.] $^{\!\!\!
+, \,}$^[Note that this terminology is slightly misleading, as mathematically cosine distance is not a distance measure, because it does not satisfy the triangle inequality, as generally $\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)$.
+We will keep using this mainstream terminology.]
 \begin{align} \tag{Sim}
@@
 \end{align}
-Note  that this terminology is slightly misleading, as mathematically cosine distance is not a distance measure, because it does not satisfy the triangle inequality, as generally $\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)$.
-We will keep using this mainstream terminology.
 
@@
 \midrule
-\cellcolor{gray!15}{rabbi} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{1.0306175}\\
-church & familial & christianStereotype & 0.7087424\\
-\cellcolor{gray!15}{synagogue} & \cellcolor{gray!15}{liberal} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{0.7922607}\\
-jew & familial & christianStereotype & 0.9783245\\
-\cellcolor{gray!15}{quran} & \cellcolor{gray!15}{dirty} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{1.1207093}\\
-muslim & uneducated & muslimStereotype & 0.5160429\\
-\cellcolor{gray!15}{torah} & \cellcolor{gray!15}{terrorist} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{0.9341137}\\
-quran & hairy & jewStereotype & 1.1764642\\
-\cellcolor{gray!15}{synagogue} & \cellcolor{gray!15}{violent} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{0.9549743}\\
-bible & cheap & jewStereotype & 1.2234364\\
-\cellcolor{gray!15}{christianity} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{0.9728545}\\
-muslim & hairy & jewStereotype & 0.8788219\\
-\cellcolor{gray!15}{islam} & \cellcolor{gray!15}{critical} & \cellcolor{gray!15}{christianStereotype} & \cellcolor{gray!15}{0.7880706}\\
-muslim & conservative & christianStereotype & 0.4453191\\
-\cellcolor{gray!15}{mosque} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{1.1541524}\\
+\cellcolor{gray!15}{rabbi} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{1.03}\\
+church & familial & christianStereotype & 0.70\\
+\cellcolor{gray!15}{synagogue} & \cellcolor{gray!15}{liberal} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{0.79}\\
+jew & familial & christianStereotype & 0.98\\
+\cellcolor{gray!15}{quran} & \cellcolor{gray!15}{dirty} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{1.12}\\
+muslim & uneducated & muslimStereotype & 0.52\\
+\cellcolor{gray!15}{torah} & \cellcolor{gray!15}{terrorist} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{0.93}\\
+quran & hairy & jewStereotype & 1.18\\
+\cellcolor{gray!15}{synagogue} & \cellcolor{gray!15}{violent} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{0.95}\\
+bible & cheap & jewStereotype & 1.22\\
+\cellcolor{gray!15}{christianity} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{0.97}\\
+muslim & hairy & jewStereotype & 0.88\\
+\cellcolor{gray!15}{islam} & \cellcolor{gray!15}{critical} & \cellcolor{gray!15}{christianStereotype} & \cellcolor{gray!15}{0.79}\\
+muslim & conservative & christianStereotype & 0.45\\
+\cellcolor{gray!15}{mosque} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{1.15}\\
 \bottomrule
@@
 
-# Methodological problems with cosine-based bias metrics
+# Challenges to cosine-based bias metrics
+\label{sec:challenges}
+
+## Interpretability issues
+\label{subsec:interpretability}
 
@@
 
-In Figure  \ref{fig:empirical} we look at the empirical distributions, while paying attention to such divisions.
+In Figures  (\ref{fig:empirical0}-\ref{fig:empirical3}) we look at the empirical distributions, while paying attention to such divisions.
 The horizontal lines represent the values of 1-\textsf{MAC} (that is, we now talk in terms of cosine similarity rather than cosine distance; this is a trivial linear transformation) that the authors considered indicative of bias for stereotypes corresponding to  given  word lists.
@@
 
-```{r cosineDistributions6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
+```{r cosineDistributions6d,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
 
@@
   geom_vline(xintercept = .14) + geom_hline(yintercept = -.14, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
-  labs(title =  "Religion (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
+  ylab("similarity")+theme_tufte(base_size = 13)+ geom_hline(yintercept = .14, color = "grey")+
+  labs(title =  "Religion (Reddit)")+
   geom_jitter(size = 0.5, alpha = .2, width=0.08)+
@@
   geom_vline(xintercept = .38) + geom_hline(yintercept = -.38, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .38, color = "grey")+
-  ggtitle("Gender: empirical distribution of cosine similarity")+
+  ylab("similarity")+theme_tufte(base_size = 13)+ geom_hline(yintercept = .38, color = "grey")+
   stat_summary(fun.data = "mean_cl_boot", fun.args = list(
@@
   annotate(geom = "label", x = 1.2, y = 0, label = "47%")+
-  labs(title =  "Gender (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
+  labs(title =  "Gender (Reddit)")+
   theme(plot.title.position = "plot",legend.position='none')
@@
   geom_vline(xintercept = .11) + geom_hline(yintercept = -.11, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .11, color = "grey")+
-   labs(title =  "Race (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
+  ylab("similarity")+theme_tufte(base_size = 13)+ geom_hline(yintercept = .11, color = "grey")+
+   labs(title =  "Race (Reddit)")+
   stat_summary(fun.data = "mean_cl_boot", fun.args = list(
@@
 ```
-\pagebreak
 
 \begin{figure}[H]
-```{r fig:cosineDistributions8b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", fig.height = 16, warning = FALSE, message = FALSE}
-
-grid.arrange(violinReligionReddit, violinGenderReddit, violinRaceReddit, ncol = 1)
-
-```
-
-\caption{Empirical distributions of cosine similarities (1-distances) for three word lists used in  the original paper.  }
-
-\label{fig:empirical}
+```{r fig:cosineDistributions8d,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", warning = FALSE, message = FALSE}
+
+violinReligionReddit
+
+```
+
+\caption{Empirical distributions of cosine similarities  for the Religion word list  used in  the original paper. }
+
+\label{fig:empirical0}
+\end{figure}
+
+\begin{figure}[H]
+```{r fig:cosineDistributions8df,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%",  warning = FALSE, message = FALSE}
+
+violinGenderReddit
+```
+
+\caption{Empirical distributions of cosine similarities for the Gender word list used in  the original paper.  }
+
+\label{fig:empirical2}
+\end{figure}
+
+\begin{figure}[H]
+```{r fig:cosineDistributions8dg,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", warning = FALSE, message = FALSE}
+
+violinRaceReddit
+```
+
+\caption{Empirical distributions of cosine similarities  for the Race word list used in  the original paper.  }
+
+\label{fig:empirical3}
 \end{figure}
@@
 
-#  Pre-averaging, bootstrapping and testing with the null model
+##  Problems with pre-averaging
+\label{subsec:problems}
 
@@
 This problem becomes transparent when we examine the visualizations of the individual cosine distances, following the idea that one of the first steps in understanding data is to look at it.
+Let's start with inspecting two examples in Figures \ref{fig:muslim} and \ref{fig:priest}.
 
@@
 ```
-
+\caption{Actual distances for the protected word "muslim".
+The most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates.
+}
+\label{fig:muslim}
+\end{figure}
+
+\begin{figure}[H]
 ```{r visPriestd,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
@@
 
-\caption{Actual distances for two protected words.
-For the protected word "muslim", the most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates.
-For the word "priest", the situation is even less as expected: the nearest attributes are human attributes, and all there seems to be no clear pattern to the distances to other attributes.}
+\caption{Actual distances for the protected word "priest", the situation is even less as expected: the nearest attributes are human attributes, and all there seems to be no clear pattern to the distances to other attributes.}
+\label{fig:priest}
 \end{figure}
@@
 ourESplot <- ggplot(dataPLOT )+geom_histogram(aes(x=biasesD0, y = ..density..), bins = 50, alpha = .6)+
-  theme_tufte(base_size = 10)+
-  labs(title="Distribution of effect sizes given a null model",
-       subtitle = "Simulation with Norm(0,0.8), n=10k")+ xlab("effect size")+
-    theme(plot.title.position = "plot", axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank()) +
+  labs(title="Distribution of effect sizes given a null model")+ xlab("effect size") +
   geom_vline(xintercept = ourEffectSize, linewidth = .5,
-             lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
+             lty = 2, alpha = .5)+theme_tufte(base_size = 13)+
   annotate("label", x = ourEffectSize, y = .06,
            label =  paste(round(ourEffectSize,2),
-                          ", p=", 0.38, sep = ""), size = 3)
+                          ", p=", 0.38, sep = ""), size = 4)+
+    theme(plot.title.position = "plot", axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank())
 
 ourStatPlot <- ggplot(dataPLOT )+ geom_histogram(aes(x=svaluesD0, y = ..density..), alpha = .6, bins = 50) +
-  theme_tufte(base_size = 10)+
-  labs(title="Distribution of test statistics given a null model",
-       subtitle = "Simulation with Norm(0,0.8), n=10k")+ xlab("test statistic")+
-    theme(plot.title.position = "plot", axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank()) +
+  labs(title="Distribution of test statistics given a null model")+ xlab("test statistic") +
   geom_vline(xintercept = ourStatistic, linewidth = .5,
-             lty = 2, alpha = .5)+theme_tufte(base_size = 10)+
+             lty = 2, alpha = .5)+theme_tufte(base_size = 13)+
   annotate("label", x = ourStatistic, y = 1.2,
            label =  paste(round(ourStatistic,2),
-                          ", p=", 0.01, sep = ""), size = 3)
+                          ", p=", 0.01, sep = ""), size = 4)+
+    theme(plot.title.position = "plot", axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank())
 
@@
 \begin{figure}[H]
-```{r fig:ourDistancesPlot2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
+```{r fig:ourDistancesPlot2e,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
 
@@
 
-\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08$.}
+\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08, n=10k$.}
 \label{fig:ourDistances}
@@
 # Bayesian method
-
-## Introduction and model definition
+\label{sec:bayesian}
+
+## Model construction
+\label{subsec:model}
 
@@
 ## Posterior predictive check
+\label{subsec:posterior}
 
@@
 \begin{figure}[H]
-```{r figposteriorPrCheckA6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", dpi = 300, warning = FALSE, message = FALSE}
-grid.arrange(posteriorPlotsA[[3]], posteriorPlotsB[[3]], ncol = 2)
+```{r figposteriorPrCheckA6b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", fig.height= 6, dpi = 300, warning = FALSE, message = FALSE}
+grid.arrange(posteriorPlotsA[[3]]+theme_tufte(base_size = 15),
+             posteriorPlotsB[[3]] +theme_tufte(base_size = 15), ncol = 1)
 ```
@@
 # Results and discussion
+\label{sec:results}
 
 \vspace{1mm}
-\footnotesize
-```{r generatePlots2h,echo=FALSE, eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
+
+```{r generatePlots2p,echo=FALSE, eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
 source("functions/cleanDataset.R")
@@
 
+\normalsize
+
 ## Observations
+\label{subsec:observations}
 
@@
 ## Rethinking debiasing
+\label{subsec:rethinking}
 
 These visualizations can be also handy when it comes to the investigation of the effect that debiasing has on the embedding space.
-Below one may see chosen pair depicting the difference in means with 89% highest posterior density intervals before and after applying debiasing.
+In Figures  \ref{fig:empiricalPriorToDebiasing} and  \ref{fig:empiricalDebiased}
+ we  see chosen pair depicting the difference in means with 89% highest posterior density intervals before and after applying debiasing.
 
 \begin{figure}[H]
-```{r fig:debiasedCosinePair2b,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, out.width="85%", warning = FALSE, message = FALSE}
+```{r fig:debiasedCosinePair2e,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, fig.height= 9, fig.width= 6, warning = FALSE, message = FALSE}
 
@@
 
+```
+\caption{Mean cosine distances with 89\% highest posterior density intervals for the  gender dataset before debiasing.}
+\label{fig:empiricalPriorToDebiasing}
+\end{figure}
+
+\begin{figure}[H]
+```{r fig:debiasedCosinePair2_e,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300,  fig.height= 9, fig.width= 6, warning = FALSE, message = FALSE}
+
 resultsDebiasedGenderReddit$plotJoint
@@
 ```
-\caption{Means with 89\% highest posterior density intervals for gender before and after debiasing.}
-\label{fig:empiricalDebiasedPair}
+\caption{Mean with 89\% highest posterior density intervals for gender after debiasing.}
+\label{fig:empiricalDebiased}
 \end{figure}
 
-In Figure \ref{fig:empiricalDebiased} we look at the empirical distributions for the debiased embeddings.
+In Figures  \ref{fig:empiricalDebiased}-\ref{fig:empiricalDebiased3} we inspect the empirical distributions for the debiased embeddings.
 Comparing the results to the original embedding, one may notice that for the Religion group the neutral and human distribution has changed slightly.
@@
 
-```{r debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
+```{r debiasedCosineb,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
 
@@
   ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
-  labs(title =  "Debiased Religion (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
+  labs(title =  "Debiased Religion (Reddit)")+
     geom_jitter(size = 0.5, alpha = .2, width=0.08)+
@@
   annotate(geom = "label", x = 1.15, y = 0, label = "63%")+
-  theme(plot.title.position = "plot",legend.position="top")
+  theme_tufte(base_size = 12)+  
+  theme(plot.title.position = "plot", legend.position='none')
 
@@
   geom_vline(xintercept = .38) + geom_hline(yintercept = -.38, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .38, color = "grey")+
-  labs(title =  "Debiased Gender (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
+  ylab("similarity")+ geom_hline(yintercept = .38, color = "grey")+
+  labs(title =  "Debiased Gender (Reddit)")+
     stat_summary(fun.data = "mean_cl_boot", fun.args = list(
@@
   annotate(geom = "label", x = 2.15, y = 0, label = "51%")+
-  annotate(geom = "label", x = 1.15, y = 0, label = "51%")+
+  annotate(geom = "label", x = 1.15, y = 0, label = "51%")+   theme_tufte(base_size = 12)+
   theme(plot.title.position = "plot",legend.position='none')
@@
   geom_vline(xintercept = .14) + geom_hline(yintercept = -.14, color = "grey") +
-  ylab("similarity")+theme_tufte(base_size = 10)+ geom_hline(yintercept = .14, color = "grey")+
-  labs(title =  "Debiased Race (Reddit)", subtitle = "Empirical distribution of cosine similarities")+
+  ylab("similarity")+ geom_hline(yintercept = .14, color = "grey")+
+  labs(title =  "Debiased Race (Reddit)")+
     stat_summary(fun.data = "mean_cl_boot", fun.args = list(
@@
   annotate(geom = "label", x = 2.15, y = 0, label = "36%")+
-  annotate(geom = "label", x = 1.15, y = 0, label = "38%")+
-  theme(plot.title.position = "plot")+geom_point(size = 1.2, alpha = .3)+
+  annotate(geom = "label", x = 1.15, y = 0, label = "38%")+geom_point(size = 1.2, alpha = .3)+
+  theme_tufte(base_size = 12)+
   theme(plot.title.position = "plot",legend.position='none')
 ```
-\pagebreak
 
 \begin{figure}[H]
-```{r fig:debiasedCosine,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "70%", fig.height = 16, warning = FALSE, message = FALSE}
-
-grid.arrange(violinReligionReddit, debiasedViolinReligionReddit, violinGenderReddit, debiasedViolinGenderReddit, violinRaceReddit, debiasedViolinRaceReddit, ncol = 2)
-
-```
-\caption{Empirical distributions of debiased cosine similarities (1-distances) for three word lists used in  the original paper.}
+```{r fig:debiasedCosine21c,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "95%", fig.height = 6, warning = FALSE, message = FALSE}
+
+grid.arrange(violinReligionReddit+theme_tufte(base_size = 12)+
+               theme(legend.position = 'top'), debiasedViolinReligionReddit, ncol = 2)
+
+```
+\caption{Empirical distributions of cosine similarities before and after  debiasing  for  the Religion word list  used in  the original paper.}
 \label{fig:empiricalDebiased}
@@
 
+\begin{figure}[H]
+```{r fig:debiasedCosine22c,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "95%", fig.height = 6, warning = FALSE, message = FALSE}
+
+grid.arrange(violinGenderReddit+theme_tufte(base_size = 12)+
+               theme(legend.position = 'top'),
+             debiasedViolinGenderReddit,
+             ncol = 2)
+
+```
+\caption{Empirical distributions of cosine similarities before and after  debiasing for  the Gender word list used in  the original paper.}
+\label{fig:empiricalDebiased2}
+\end{figure}
+
+\begin{figure}[H]
+```{r fig:debiasedCosine23c,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "95%", fig.height = 6, warning = FALSE, message = FALSE}
+
+grid.arrange( violinRaceReddit+theme_tufte(base_size = 12)+
+               theme(legend.position = 'top'),
+              debiasedViolinRaceReddit, ncol = 2)
+
+```
+\caption{Empirical distributions of cosine similarities before and after  debiasing for the Race word list used in  the original paper.}
+\label{fig:empiricalDebiased3}
+\end{figure}
+
 # Related work and conclusions
+\label{sec:related}
 
@@
 \newpage
+
+# References {-}
+
+<div id="refs"></div>
+
+\newpage
 \appendix
 # Appendix 
+\label{sec:appendix}
 
@@
 
-```{r weatjoint2r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r weatjoint2z,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat7Google$plotJoint
@@
 
-```{r weat7glovf, echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r weat7glovh, echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat7Glove$plotJoint
@@
 
-```{r resultsWeat7Reddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsWeat7Reddita,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat7Reddit$plotJoint
@@
 
-```{r resultsWeat1Google,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsWeat1Googlea,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat1Google$plotJoint
@@
 
-```{r resultsWeat1Glove,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsWeat1Glovea,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat1Glove$plotJoint
@@
 
-```{r resultsWeat1Reddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsWeat1Reddita,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsWeat1Reddit$plotJoint
@@
 
-```{r resultsReligionGoogle,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsReligionGooglea,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsReligionGoogle$plotJoint
@@
 
-```{r resultsReligionGlove,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsReligionGlovea,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsReligionGlove$plotJoint
@@
 
-```{r resultsReligionReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsReligionReddita,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsReligionReddit$plotJoint
@@
 
-```{r resultsRaceGoogle,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsRaceGooglea,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsRaceGoogle$plotJoint
@@
 
-```{r resultsRaceGlove,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsRaceGlovea,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsRaceGlove$plotJoint
@@
 
-```{r resultsRaceReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsRaceReddita,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsRaceReddit$plotJoint
@@
 
-```{r resultsGenderGoogle,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsGenderGooglea,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsGenderGoogle$plotJoint
@@
 <!-- \begin{figure}[H] -->
-```{r resultsGenderGlove,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsGenderGlovea,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsGenderGlove$plotJoint
@@
 <!-- \begin{figure}[H] -->
-```{r resultsGenderReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsGenderReddita,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsGenderReddit$plotJoint
@@
 <!-- \begin{figure}[H] -->
-```{r resultsDebiasedReligionReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsDebiasedReligionReddita,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsDebiasedReligionReddit$plotJoint
@@
 <!-- \begin{figure}[H] -->
-```{r resultsDebiasedRaceReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsDebiasedRaceReddita,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsDebiasedRaceReddit$plotJoint
@@
 <!-- \begin{figure}[H] -->
-```{r resultsDebiasedGenderReddit,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
+```{r resultsDebiasedGenderReddita,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "110%", dpi = 300, warning = FALSE, message = FALSE, fig.height= 9, fig.width= 6}
 resultsDebiasedGenderReddit$plotJoint
@@
 ## Posterior predictive checks
+\label{appendix:posterior}
 
@@
 ## Word lists
-
-### Lists for [@Manzini2019blackToCriminal]:
+\label{appendix:word}
+
+### Lists used in previous research
+
 \label{appendix:manzini_word_lists}
-The lists are available here:
+The lists from [@Manzini2019blackToCriminal] are available here:
 
@@
 
-### Our custom word lists
+### Custom lists used in this paper
 

### 2023-05-25 paper/paperDraft.tex
revised visualizations, fixed intro

@@
 
-\usepackage{sectsty}
-\sectionfont{\fontsize{11}{11}\selectfont}
+% \usepackage{sectsty}
+% \sectionfont{\fontsize{12}{11}\selectfont}
+% \subsectionfont{\fontsize{11}{11}\selectfont}
 
@@
 
-Natural language processing (NLP) models are used to perform various
-language-related tasks such as providing email filters, intelligent
-assistants, search results, translations, text analysis and so on. Such
-tools require input words to be represented by vectors of real numbers.
-This can usually be done with word embeddings. A common way to construct
-an embedding is to use a large natural language corpus to train a neural
-network to assign vectors to words that optimize for co-occurrence
-prediction accuracy. The vectors can then be compared in terms of their
-similarity -- the usual measure is cosine similarity -- and the results
-of such comparisons can be used in downstream tasks. Roughly speaking,
-cosine similarity is an imperfect mathematical proxy for semantic
-similarity {[}14{]}.
-
-It has been suggested {[}1,4,5,8,11,12{]} that such models can learn
-implicit biases that reflect harmful stereotypical thinking---for
-example, the (vector corresponding to the) word \textit{she} might be
-much closer in the vector space to the word \textit{cooking} than the
-word \textit{he}. Such phenomena may be undesirable at least in some
-downstream tasks, such as web search, recommendations, and so on. To
-investigate such issues, several measures of bias in word embeddings
-have been formulated and applied. Our goal is to use two prominent
-examples of such measures to argue that this approach oversimplifies the
-situation and to develop a Bayesian alternative.
+It has been suggested\footnote{See for instance {[}1,4,5,8,11,12{]}.}
+that language models can learn implicit biases that reflect harmful
+stereotypical thinking---for example, the (vector corresponding to the)
+word \textit{she} might be much closer in the vector space to the word
+\textit{cooking} than the word \textit{he}. Such phenomena are
+undesirable at least in some downstream tasks, such as web search,
+recommendations, and so on. To investigate such issues, several measures
+of bias in word embeddings have been formulated and applied. Our goal is
+to use two prominent examples of such measures to argue that this
+approach oversimplifies the situation and to develop a Bayesian
+alternative.
+
+A common approach in natural language processing is to represent words
+by vectors of real numbers---such representations are called
+\emph{embeddings}. One way to construct an embedding---we will focus our
+attention on non-contextual language models\footnote{One example of a
+  contextualized representation is BERT. Another is GPT.}---is to use a
+large corpus to train a neural network to assign vectors to words in a
+way that optimizes for co-occurrence prediction accuracy. Such vectors
+can then be compared in terms of their similarity---the usual measure is
+cosine similarity---and the results of such comparisons can be used in
+downstream tasks. Roughly speaking, cosine similarity is an imperfect
+mathematical proxy for semantic similarity {[}14{]}.
 
@@
 looks like. However, this objection overlooks the fact that we, humans,
-are the ones who construct this linguistic reality, is shaped in part by
-the natural language processing tools we use on a massive scale. Sure,
-if there is unfairness and our goal is to diagnose it, we should do
-complete justice to learning it in the model used in a study thereof.
-One example of this approach is {[}3{]}. However, if our goal is to
-develop downstream tools that perform tasks that we care about without
-further perpetuating or exacerbating harmful stereotypes, we still have
-good reasons to try to minimize the negative impact, at least as long as
-this does not damage the overall performance of the tool. Moreover, it
-is often not the case that the corpora mirror reality---to give a
-trivial example, heads are spoken of more often than kidneys, but this
-does not mean that kidneys occur much less often in reality than heads.
-To give a more relevant example, the disproportionate association of
-female words with female occupations in a corpus actually greatly
-exaggerates the actual lower disproportion in the real distribution of
-occupations. {[}6{]}.
+are the ones who construct this linguistic reality, which is shaped in
+part by the natural language processing tools we use on a massive scale.
+Sure, if there is unfairness and our goal is to diagnose it, we should
+do complete justice to learning it in the model used to study it. One
+example of this approach is {[}3{]}, where the authors use language
+models to study the shape of certain biases across a century.
+
+However, if our goal is to develop downstream tools that perform tasks
+that we care about without further perpetuating or exacerbating harmful
+stereotypes, we still have good reasons to try to minimize the negative
+impact. Moreover, it is often not the case that the corpora mirror
+reality---to give a trivial example, heads are spoken of more often than
+kidneys, but this does not mean that kidneys occur much less often in
+reality than heads. To give a more relevant example, the
+disproportionate association of female words with female occupations in
+a corpus actually greatly exaggerates the actual lower disproportion in
+the real distribution of occupations {[}6{]}.
 
 In what follows, we focus on two popular measures of bias applicable to
-many existing word embeddings (
+many existing word embeddings, such as
 \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.}
@@
 and
-\emph{Reddit Corpus}\footnote{The Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}):
+\emph{Reddit Corpus}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
 \emph{Word Embedding Association Test} (\textsf{WEAT}) {[}8{]}, and
@@
 explain how these measures are supposed to work. Then we argue that they
-are problematic for various reasons, a key one being that by
+are problematic for various reasons---the key one being that by
 pre-averaging data they manufacture false confidence, which we
@@
 
-We propose to replace them with a Bayesian data analysis (hierarchical
-models), which not only provides more modest and realistic assessment of
-the uncertainty involved, but also allows for inspection at various
-levels of granularity. Once we introduce the method, we apply it to
-multiple word embeddings and results of supposed debiasing, putting
+We propose to replace them with a Bayesian data analysis, which not only
+provides more modest and realistic assessment of the uncertainty
+involved, but in which hierarchical models allow for inspection at
+various levels of granularity. Once we introduce the method, we apply it
+to multiple word embeddings and results of supposed debiasing, putting
 forward some general observations that are not exactly in line with the
-usual picture painted in terms of \textsf{WEAT} or \textsf{MAC} (and the
-problem generalizes to any approach that focuses on chasing a single
-numeric metric): the word list sizes and sample sizes used in the
-studies are usually small,\footnote{Depending on a list for {[}8{]} the
-  range for protected words is between 13 and 100, and for attributes
-  between 16 and 25; for {[}12{]} the range for protected words is
-  between 14 and 18, and for attributes between 11 and 25.} posterior
-density intervals are fairly wide, often the differences between
-associated, different and neutral human predicates, are not very
-impressive. Also, a preliminary inspection suggests that the
-desirability of changes obtained by the usual debiasing methods is
-debatable. The tools that we have proposed, however, allow for a more
-fine-grained and multi-level evaluation of bias and debiasing in
-language models without abandoning modesty about the uncertainties
-involved.
+usual picture painted in terms of \textsf{WEAT} or \textsf{MAC}.
+
+Most of the problems that we point out generalize to any existing
+approach that focuses on chasing a single numeric metric of bias. (1)
+They treat the results of pre-averaging as raw data in statistical
+significance tests, which in this context is bound to overestimate
+significance. We show similar results can easily be obtained when
+sampling from null models with no bias. (2) The word list sizes and
+sample sizes used in the studies are usually small,\footnote{Depending
+  on a list for {[}8{]} the range for protected words is between 13 and
+  100, and for attributes between 16 and 25; for {[}12{]} the range for
+  protected words is between 14 and 18, and for attributes between 11
+  and 25.} (3) Many studies do not use any control predicates, such as
+random neutral words or neutral human predicates for comparison.
+
+On the constructive side, we develop and deploy our method, and the
+results are, roughly, as follows. (A) Posterior density intervals are
+fairly wide and the average differences between associated, different
+and neutral human predicates, are not very large. (B) A preliminary
+inspection suggests that the desirability of changes obtained by the
+usual debiasing methods is debatable.
+
+In Section \ref{sec:two} we describe the two key measures discussed in
+this paper, \textsf{WEAT} and \textsf{MAC}, explaining how they are
+calculated and how they are supposed to work. In Section
+\ref{sec:challenges} we first argue in Subsection
+\ref{subsec:interpretability}, that it is far from clear how results
+given in terms of \textsf{WEAT} or \textsf{MAC} are to be interpreted.
+Second, in Subsection \ref{subsec:problems} we explain the statistical
+problems that arise when one uses pre-averaged data in such contexts, as
+these measures do. In Section \ref{sec:bayesian} we explain the
+alternative Bayesian approach that we propose. In Section
+\ref{sec:results} we elaborate on the results that it leads to,
+including a somewhat skeptical view of the efficiency of debiasing
+methods, discussed in Subsection \ref{subsec:rethinking}. Finally, in
+Section \ref{sec:related} we spend some time placing our results in the
+ongoing discussions.\footnote{\textbf{Disclaimer:} throughout the paper
+  we will be mentioning and using word lists and stereotypes we did not
+  formulate, which does not mean we condone any judgment made therein or
+  underlying a given word selection. For instance, the Gender dataset
+  does not recognize non-binary categories, and yet we use it without
+  claiming that such categories should be ignored.}
 
@@
 
-Conceptually the idea is that if a particular harmful stereotype is
+\label{sec:two}
+
+Conceptually, the idea is that if a particular harmful stereotype is
 learned in a particular embedding, then certain groups of words will be
@@
 that they should not be systematically closer to stereotypically female
-job names, and male words require protection in that they should not be
-systematically closer to toxic masculinity stereotypes, such as
-``tough'', ``never complaining'' or ``macho''.\footnote{However, for
-  some research-related purposes, such as the study of stereotypes
-  across history {[}3{]}, embeddings which do not protect certain
-  classes may also be useful.}
+job names, such as ``nurse'', ``librarian'', ``waitress'', and male
+words require protection in that they should not be systematically
+closer to toxic masculinity stereotypes, such as ``tough'', ``never
+complaining'' or ``macho''.\footnote{However, for some research-related
+  purposes, such as the study of stereotypes across history {[}3{]},
+  embeddings which do not protect certain classes may also be useful.}
 
@@
   difference, ``\(\cdot\)'' stands for the dot product operation, and
-  \(\lVert a\rVert = \sqrt{(a \cdot a)}\).} \begin{align} \tag{Sim}
+  \(\lVert a\rVert = \sqrt{(a \cdot a)}\).}
+\(^{\!\!\! , \,}\)\footnote{Note that this terminology is slightly
+  misleading, as mathematically cosine distance is not a distance
+  measure, because it does not satisfy the triangle inequality, as
+  generally
+  \(\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)\).
+  We will keep using this mainstream terminology.}
+\begin{align} \tag{Sim}
 \mathsf{cosineSimilarity}(A,B) & = \frac{A \cdot B}{\lVert  A \rVert \,\lVert B \rVert}
@@
 \mathsf{cosineDistance}(A,B) &  = 1 - \mathsf{cosineSimilarity}(A,B).
-\end{align} Note that this terminology is slightly misleading, as
-mathematically cosine distance is not a distance measure, because it
-does not satisfy the triangle inequality, as generally
-\(\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)\).
-We will keep using this mainstream terminology.
+\end{align}
 
@@
 
-\hypertarget{methodological-problems-with-cosine-based-bias-metrics}{%
-\section{Methodological problems with cosine-based bias
-metrics}\label{methodological-problems-with-cosine-based-bias-metrics}}
+\hypertarget{challenges-to-cosine-based-bias-metrics}{%
+\section{Challenges to cosine-based bias
+metrics}\label{challenges-to-cosine-based-bias-metrics}}
+
+\label{sec:challenges}
+
+\hypertarget{interpretability-issues}{%
+\subsection{Interpretability issues}\label{interpretability-issues}}
+
+\label{subsec:interpretability}
 
@@
 
-In Figure \ref{fig:empirical} we look at the empirical distributions,
-while paying attention to such divisions. The horizontal lines represent
-the values of 1-\textsf{MAC} (that is, we now talk in terms of cosine
-similarity rather than cosine distance; this is a trivial linear
-transformation) that the authors considered indicative of bias for
-stereotypes corresponding to given word lists. For instance, in
-religion, \textsf{MAC} was .859 and considered a sign of bias, so we
+In Figures (\ref{fig:empirical0}-\ref{fig:empirical3}) we look at the
+empirical distributions, while paying attention to such divisions. The
+horizontal lines represent the values of 1-\textsf{MAC} (that is, we now
+talk in terms of cosine similarity rather than cosine distance; this is
+a trivial linear transformation) that the authors considered indicative
+of bias for stereotypes corresponding to given word lists. For instance,
+in religion, \textsf{MAC} was .859 and considered a sign of bias, so we
 plot \(0\pm (1-.859)\approx .14\) lines around similarity = 0 (that is,
@@
 
-\pagebreak
-
 \begin{figure}[H]
 
-\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8b-1} \end{center}
-
-\caption{Empirical distributions of cosine similarities (1-distances) for three word lists used in  the original paper.  }
-
-\label{fig:empirical}
+\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8d-1} \end{center}
+
+\caption{Empirical distributions of cosine similarities  for the Religion word list  used in  the original paper. }
+
+\label{fig:empirical0}
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8df-1} \end{center}
+
+\caption{Empirical distributions of cosine similarities for the Gender word list used in  the original paper.  }
+
+\label{fig:empirical2}
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8dg-1} \end{center}
+
+\caption{Empirical distributions of cosine similarities  for the Race word list used in  the original paper.  }
+
+\label{fig:empirical3}
 \end{figure}
@@
 
-\hypertarget{pre-averaging-bootstrapping-and-testing-with-the-null-model}{%
-\section{Pre-averaging, bootstrapping and testing with the null
-model}\label{pre-averaging-bootstrapping-and-testing-with-the-null-model}}
+\hypertarget{problems-with-pre-averaging}{%
+\subsection{Problems with
+pre-averaging}\label{problems-with-pre-averaging}}
+
+\label{subsec:problems}
 
@@
 that one of the first steps in understanding data is to look at it.
+Let's start with inspecting two examples in Figures \ref{fig:muslim} and
+\ref{fig:priest}.
 
@@
 \begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visMuslimd-1} \end{center}
+\caption{Actual distances for the protected word "muslim".
+The most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates.
+}
+\label{fig:muslim}
+\end{figure}
+
+\begin{figure}[H]
 
@@
 
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2-1} \end{center}
-
-\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08$.}
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2e-1} \end{center}
+
+\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08, n=10k$.}
 \label{fig:ourDistances}
@@
 
-\hypertarget{introduction-and-model-definition}{%
-\subsection{Introduction and model
-definition}\label{introduction-and-model-definition}}
+\label{sec:bayesian}
+
+\hypertarget{model-construction}{%
+\subsection{Model construction}\label{model-construction}}
+
+\label{subsec:model}
 
@@
 
+\label{subsec:posterior}
+
 A posterior predictive check is a technique used to evaluate the fit of
@@
 
+\label{sec:results}
+
 \vspace{1mm}
-\footnotesize
+
+\normalsize
 
@@
 \subsection{Observations}\label{observations}}
+
+\label{subsec:observations}
 
@@
 
+\label{subsec:rethinking}
+
 These visualizations can be also handy when it comes to the
 investigation of the effect that debiasing has on the embedding space.
-Below one may see chosen pair depicting the difference in means with
-89\% highest posterior density intervals before and after applying
-debiasing.
+In Figures \ref{fig:empiricalPriorToDebiasing} and
+\ref{fig:empiricalDebiased} we see chosen pair depicting the difference
+in means with 89\% highest posterior density intervals before and after
+applying debiasing.
 
@@
 
-\begin{center}\includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2b-1} \includegraphics[width=0.85\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosinePair2b-2} \end{center}
-\caption{Means with 89\% highest posterior density intervals for gender before and after debiasing.}
-\label{fig:empiricalDebiasedPair}
+\begin{center}\includegraphics{paperDraft_files/figure-latex/fig:debiasedCosinePair2e-1} \end{center}
+\caption{Mean cosine distances with 89\% highest posterior density intervals for the  gender dataset before debiasing.}
+\label{fig:empiricalPriorToDebiasing}
 \end{figure}
 
-In Figure \ref{fig:empiricalDebiased} we look at the empirical
-distributions for the debiased embeddings. Comparing the results to the
-original embedding, one may notice that for the Religion group the
-neutral and human distribution has changed slightly. Before within the
-``correct'' cosine similarity boundaries there were 56\% of neutral and
-55\% of human word lists. After the debiasing the values changed to 59\%
-(for neutral) and 59\% (for human). The different and associated word
-lists were more influenced. The general shape of both distributions is
-less stretched. Before debiasing 43\% of the different word lists and
-35\% of the associated word lists were within the accepted boundaries.
-After the embedding manipulation the percentage has increased for both
-lists to 63\%. Visualization for Gender group illustrates almost no
-change for the neutral and human word lists before and after debiasing.
-The values for different and associated word lists are also barely
-impacted by the embedding modification. In the Race group, the
-percentage within the boundaries for neutral and associated word lists
-have increased. The opposite happened for human and different word
-lists, where the percentage of ``correct'' cosine similarity dropped
-from 67\% to 55\% (human) and from 39\% to 36\% (different).
-
-\pagebreak
-
 \begin{figure}[H]
 
-\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine-1} \end{center}
-\caption{Empirical distributions of debiased cosine similarities (1-distances) for three word lists used in  the original paper.}
+\begin{center}\includegraphics{paperDraft_files/figure-latex/fig:debiasedCosinePair2_e-1} \end{center}
+\caption{Mean with 89\% highest posterior density intervals for gender after debiasing.}
 \label{fig:empiricalDebiased}
+\end{figure}
+
+In Figures \ref{fig:empiricalDebiased}-\ref{fig:empiricalDebiased3} we
+inspect the empirical distributions for the debiased embeddings.
+Comparing the results to the original embedding, one may notice that for
+the Religion group the neutral and human distribution has changed
+slightly. Before within the ``correct'' cosine similarity boundaries
+there were 56\% of neutral and 55\% of human word lists. After the
+debiasing the values changed to 59\% (for neutral) and 59\% (for human).
+The different and associated word lists were more influenced. The
+general shape of both distributions is less stretched. Before debiasing
+43\% of the different word lists and 35\% of the associated word lists
+were within the accepted boundaries. After the embedding manipulation
+the percentage has increased for both lists to 63\%. Visualization for
+Gender group illustrates almost no change for the neutral and human word
+lists before and after debiasing. The values for different and
+associated word lists are also barely impacted by the embedding
+modification. In the Race group, the percentage within the boundaries
+for neutral and associated word lists have increased. The opposite
+happened for human and different word lists, where the percentage of
+``correct'' cosine similarity dropped from 67\% to 55\% (human) and from
+39\% to 36\% (different).
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine21c-1} \end{center}
+\caption{Empirical distributions of cosine similarities before and after  debiasing  for  the Religion word list  used in  the original paper.}
+\label{fig:empiricalDebiased}
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine22c-1} \end{center}
+\caption{Empirical distributions of cosine similarities before and after  debiasing for  the Gender word list used in  the original paper.}
+\label{fig:empiricalDebiased2}
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine23c-1} \end{center}
+\caption{Empirical distributions of cosine similarities before and after  debiasing for the Race word list used in  the original paper.}
+\label{fig:empiricalDebiased3}
 \end{figure}
@@
 conclusions}\label{related-work-and-conclusions}}
+
+\label{sec:related}
 
@@
 \newpage
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Bolukbasi2016man}{}}%
+\CSLLeftMargin{{[}1{]} }%
+\CSLRightInline{Tolga Bolukbasi, Kai-Wei Chang, James Y. Zou, Venkatesh
+Saligrama, and Adam Kalai. 2016. Man is to computer programmer as woman
+is to homemaker? Debiasing word embeddings. \emph{CoRR} abs/1607.06520,
+(2016). Retrieved from \url{http://arxiv.org/abs/1607.06520}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ethayarajh2020Bernstein}{}}%
+\CSLLeftMargin{{[}2{]} }%
+\CSLRightInline{Kawin Ethayarajh. 2020. Is your classifier actually
+biased? Measuring fairness under uncertainty with bernstein bounds.
+\emph{CoRR} abs/2004.12332, (2020). Retrieved from
+\url{https://arxiv.org/abs/2004.12332}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Garg2017hundredYears}{}}%
+\CSLLeftMargin{{[}3{]} }%
+\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
+Zou. 2017. Word embeddings quantify 100 years of gender and ethnic
+stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
+(November 2017).
+DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Garg2018years}{}}%
+\CSLLeftMargin{{[}4{]} }%
+\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
+Zou. 2018. Word embeddings quantify 100 years of gender and ethnic
+stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
+16 (April 2018), E3635--E3644.
+DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gonen2019lipstick}{}}%
+\CSLLeftMargin{{[}5{]} }%
+\CSLRightInline{Hila Gonen and Yoav Goldberg. 2019. Lipstick on a pig:
+{D}ebiasing methods cover up systematic gender biases in word embeddings
+but do not remove them. In \emph{Proceedings of the 2019 conference of
+the north {A}merican chapter of the association for computational
+linguistics: Human language technologies, volume 1 (long and short
+papers)}, Association for Computational Linguistics, Minneapolis,
+Minnesota, 609--614.
+DOI:https://doi.org/\href{https://doi.org/10.18653/v1/N19-1061}{10.18653/v1/N19-1061}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gordon2012reporting}{}}%
+\CSLLeftMargin{{[}6{]} }%
+\CSLRightInline{Jonathan Gordon and Benjamin Durme. 2013. Reporting bias
+and knowledge acquisition. In \emph{AKBC 2013 - Proceedings of the 2013
+Workshop on Automated Knowledge Base Construction, Co-located with CIKM
+2013}, 25--30.
+DOI:https://doi.org/\href{https://doi.org/10.1145/2509558.2509563}{10.1145/2509558.2509563}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hoekstra2014Misinterpretation}{}}%
+\CSLLeftMargin{{[}7{]} }%
+\CSLRightInline{Rink Hoekstra, Richard D. Morey, Jeffrey N. Rouder, and
+Eric-Jan Wagenmakers. 2014. Robust misinterpretation of confidence
+intervals. \emph{Psychonomic Bulletin \& Review} 21, 5 (October 2014),
+1157--1164.
+DOI:https://doi.org/\href{https://doi.org/10.3758/s13423-013-0572-3}{10.3758/s13423-013-0572-3}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Caliskan2017semanticsBiases}{}}%
+\CSLLeftMargin{{[}8{]} }%
+\CSLRightInline{Aylin Caliskan Islam, Joanna J. Bryson, and Arvind
+Narayanan. 2016. Semantics derived automatically from language corpora
+necessarily contain human biases. \emph{CoRR} abs/1608.07187, (2016).
+Retrieved from \url{http://arxiv.org/abs/1608.07187}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-JohnsonValueFree}{}}%
+\CSLLeftMargin{{[}9{]} }%
+\CSLRightInline{Gabbrielle Johnson. forthcoming. Are algorithms
+value-free? Feminist theoretical virtues in machine learning.
+\emph{Journal Moral Philosophy} (forthcoming).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kruschke2015bayesian}{}}%
+\CSLLeftMargin{{[}10{]} }%
+\CSLRightInline{John Kruschke. 2015. \emph{Doing bayesian data analysis
+(second edition)}. Academic Press, Boston.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Lauscher2019multidimensional}{}}%
+\CSLLeftMargin{{[}11{]} }%
+\CSLRightInline{Anne Lauscher and Goran Glavas. 2019. Are we
+consistently biased? Multidimensional analysis of biases in
+distributional word vectors. \emph{CoRR} abs/1904.11783, (2019).
+Retrieved from \url{http://arxiv.org/abs/1904.11783}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Manzini2019blackToCriminal}{}}%
+\CSLLeftMargin{{[}12{]} }%
+\CSLRightInline{Thomas Manzini, Yao Chong Lim, Yulia Tsvetkov, and Alan
+W Black. 2019. Black is to criminal as caucasian is to police: Detecting
+and removing multiclass bias in word embeddings. Retrieved from
+\url{https://arxiv.org/abs/1904.04047}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-statrethinkingbook2020}{}}%
+\CSLLeftMargin{{[}13{]} }%
+\CSLRightInline{Richard McElreath. 2020. \emph{Statistical rethinking: A
+bayesian course with examples in r and stan, 2nd edition} (2nd ed.). CRC
+Press. Retrieved from
+\url{http://xcelab.net/rm/statistical-rethinking/}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Mikolov2013efficient}{}}%
+\CSLLeftMargin{{[}14{]} }%
+\CSLRightInline{Tomas Mikolov, Kai Chen, Greg Corrado, and Jeffrey Dean.
+2013. Efficient estimation of word representations in vector space.
+DOI:https://doi.org/\href{https://doi.org/10.48550/ARXIV.1301.3781}{10.48550/ARXIV.1301.3781}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Morey2015confidenceFallacy}{}}%
+\CSLLeftMargin{{[}15{]} }%
+\CSLRightInline{Richard Morey, Rink Hoekstra, Jeffrey Rouder, Michael
+Lee, and EJ Wagenmakers. 2015. The fallacy of placing confidence in
+confidence intervals. \emph{Psychonomic Bulletin \& Review} (September
+2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nissim2020fair}{}}%
+\CSLLeftMargin{{[}16{]} }%
+\CSLRightInline{Malvina Nissim, Rik van Noord, and Rob van der Goot.
+2020. Fair is better than sensational: Man is to doctor as woman is to
+doctor. \emph{Computational Linguistics} 46, 2 (June 2020), 487--497.
+DOI:https://doi.org/\href{https://doi.org/10.1162/coli_a_00379}{10.1162/coli\_a\_00379}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nosek2002harvesting}{}}%
+\CSLLeftMargin{{[}17{]} }%
+\CSLRightInline{Brian A. Nosek, Mahzarin R. Banaji, and Anthony G.
+Greenwald. 2002. Harvesting implicit group attitudes and beliefs from a
+demonstration web site. \emph{Group Dynamics: Theory, Research, and
+Practice} 6, 1 (2002), 101--115.
+DOI:https://doi.org/\href{https://doi.org/10.1037/1089-2699.6.1.101}{10.1037/1089-2699.6.1.101}}
+
+\end{CSLReferences}
+
+\newpage
 \appendix
@@
 
+\label{sec:appendix}
+
 \hypertarget{visualizations}{%
@@
 
-\begin{center}\includegraphics{paperDraft_files/figure-latex/weatjoint2r-1} \end{center}
+\begin{center}\includegraphics{paperDraft_files/figure-latex/weatjoint2z-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/weat7glovf-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat7Reddit-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Google-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Glove-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Reddit-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGoogle-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGlove-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionReddit-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGoogle-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGlove-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceReddit-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGoogle-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGlove-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderReddit-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedReligionReddit-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedRaceReddit-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedGenderReddit-1} \end{center}
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/weat7glovh-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat7Reddita-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Googlea-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Glovea-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Reddita-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGooglea-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGlovea-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionReddita-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGooglea-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGlovea-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceReddita-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGooglea-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGlovea-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderReddita-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedReligionReddita-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedRaceReddita-1} \end{center}
+
+\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedGenderReddita-1} \end{center}
 
@@
 
+\label{appendix:posterior}
+
 \begin{center}\includegraphics[width=0.85\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixa-1} \end{center}
@@
 
-\hypertarget{lists-for-manzini2019blacktocriminal}{%
-\subsubsection{Lists for
-{[}12{]}:}\label{lists-for-manzini2019blacktocriminal}}
-
-\label{appendix:manzini_word_lists} The lists are available here:
+\label{appendix:word}
+
+\hypertarget{lists-used-in-previous-research}{%
+\subsubsection{Lists used in previous
+research}\label{lists-used-in-previous-research}}
+
+\label{appendix:manzini_word_lists} The lists from {[}12{]} are
+available here:
 
@@
 
-\hypertarget{our-custom-word-lists}{%
-\subsubsection{Our custom word lists}\label{our-custom-word-lists}}
+\hypertarget{custom-lists-used-in-this-paper}{%
+\subsubsection{Custom lists used in this
+paper}\label{custom-lists-used-in-this-paper}}
 
@@
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Bolukbasi2016man}{}}%
-\CSLLeftMargin{{[}1{]} }%
-\CSLRightInline{Tolga Bolukbasi, Kai-Wei Chang, James Y. Zou, Venkatesh
-Saligrama, and Adam Kalai. 2016. Man is to computer programmer as woman
-is to homemaker? Debiasing word embeddings. \emph{CoRR} abs/1607.06520,
-(2016). Retrieved from \url{http://arxiv.org/abs/1607.06520}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ethayarajh2020Bernstein}{}}%
-\CSLLeftMargin{{[}2{]} }%
-\CSLRightInline{Kawin Ethayarajh. 2020. Is your classifier actually
-biased? Measuring fairness under uncertainty with bernstein bounds.
-\emph{CoRR} abs/2004.12332, (2020). Retrieved from
-\url{https://arxiv.org/abs/2004.12332}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Garg2017hundredYears}{}}%
-\CSLLeftMargin{{[}3{]} }%
-\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
-Zou. 2017. Word embeddings quantify 100 years of gender and ethnic
-stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
-(November 2017).
-DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Garg2018years}{}}%
-\CSLLeftMargin{{[}4{]} }%
-\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
-Zou. 2018. Word embeddings quantify 100 years of gender and ethnic
-stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
-16 (April 2018), E3635--E3644.
-DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gonen2019lipstick}{}}%
-\CSLLeftMargin{{[}5{]} }%
-\CSLRightInline{Hila Gonen and Yoav Goldberg. 2019. Lipstick on a pig:
-{D}ebiasing methods cover up systematic gender biases in word embeddings
-but do not remove them. In \emph{Proceedings of the 2019 conference of
-the north {A}merican chapter of the association for computational
-linguistics: Human language technologies, volume 1 (long and short
-papers)}, Association for Computational Linguistics, Minneapolis,
-Minnesota, 609--614.
-DOI:https://doi.org/\href{https://doi.org/10.18653/v1/N19-1061}{10.18653/v1/N19-1061}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gordon2012reporting}{}}%
-\CSLLeftMargin{{[}6{]} }%
-\CSLRightInline{Jonathan Gordon and Benjamin Durme. 2013. Reporting bias
-and knowledge acquisition. In \emph{AKBC 2013 - Proceedings of the 2013
-Workshop on Automated Knowledge Base Construction, Co-located with CIKM
-2013}, 25--30.
-DOI:https://doi.org/\href{https://doi.org/10.1145/2509558.2509563}{10.1145/2509558.2509563}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hoekstra2014Misinterpretation}{}}%
-\CSLLeftMargin{{[}7{]} }%
-\CSLRightInline{Rink Hoekstra, Richard D. Morey, Jeffrey N. Rouder, and
-Eric-Jan Wagenmakers. 2014. Robust misinterpretation of confidence
-intervals. \emph{Psychonomic Bulletin \& Review} 21, 5 (October 2014),
-1157--1164.
-DOI:https://doi.org/\href{https://doi.org/10.3758/s13423-013-0572-3}{10.3758/s13423-013-0572-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Caliskan2017semanticsBiases}{}}%
-\CSLLeftMargin{{[}8{]} }%
-\CSLRightInline{Aylin Caliskan Islam, Joanna J. Bryson, and Arvind
-Narayanan. 2016. Semantics derived automatically from language corpora
-necessarily contain human biases. \emph{CoRR} abs/1608.07187, (2016).
-Retrieved from \url{http://arxiv.org/abs/1608.07187}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-JohnsonValueFree}{}}%
-\CSLLeftMargin{{[}9{]} }%
-\CSLRightInline{Gabbrielle Johnson. forthcoming. Are algorithms
-value-free? Feminist theoretical virtues in machine learning.
-\emph{Journal Moral Philosophy} (forthcoming).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruschke2015bayesian}{}}%
-\CSLLeftMargin{{[}10{]} }%
-\CSLRightInline{John Kruschke. 2015. \emph{Doing bayesian data analysis
-(second edition)}. Academic Press, Boston.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lauscher2019multidimensional}{}}%
-\CSLLeftMargin{{[}11{]} }%
-\CSLRightInline{Anne Lauscher and Goran Glavas. 2019. Are we
-consistently biased? Multidimensional analysis of biases in
-distributional word vectors. \emph{CoRR} abs/1904.11783, (2019).
-Retrieved from \url{http://arxiv.org/abs/1904.11783}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Manzini2019blackToCriminal}{}}%
-\CSLLeftMargin{{[}12{]} }%
-\CSLRightInline{Thomas Manzini, Yao Chong Lim, Yulia Tsvetkov, and Alan
-W Black. 2019. Black is to criminal as caucasian is to police: Detecting
-and removing multiclass bias in word embeddings. Retrieved from
-\url{https://arxiv.org/abs/1904.04047}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-statrethinkingbook2020}{}}%
-\CSLLeftMargin{{[}13{]} }%
-\CSLRightInline{Richard McElreath. 2020. \emph{Statistical rethinking: A
-bayesian course with examples in r and stan, 2nd edition} (2nd ed.). CRC
-Press. Retrieved from
-\url{http://xcelab.net/rm/statistical-rethinking/}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Mikolov2013efficient}{}}%
-\CSLLeftMargin{{[}14{]} }%
-\CSLRightInline{Tomas Mikolov, Kai Chen, Greg Corrado, and Jeffrey Dean.
-2013. Efficient estimation of word representations in vector space.
-DOI:https://doi.org/\href{https://doi.org/10.48550/ARXIV.1301.3781}{10.48550/ARXIV.1301.3781}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Morey2015confidenceFallacy}{}}%
-\CSLLeftMargin{{[}15{]} }%
-\CSLRightInline{Richard Morey, Rink Hoekstra, Jeffrey Rouder, Michael
-Lee, and EJ Wagenmakers. 2015. The fallacy of placing confidence in
-confidence intervals. \emph{Psychonomic Bulletin \& Review} (September
-2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nissim2020fair}{}}%
-\CSLLeftMargin{{[}16{]} }%
-\CSLRightInline{Malvina Nissim, Rik van Noord, and Rob van der Goot.
-2020. Fair is better than sensational: Man is to doctor as woman is to
-doctor. \emph{Computational Linguistics} 46, 2 (June 2020), 487--497.
-DOI:https://doi.org/\href{https://doi.org/10.1162/coli_a_00379}{10.1162/coli\_a\_00379}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nosek2002harvesting}{}}%
-\CSLLeftMargin{{[}17{]} }%
-\CSLRightInline{Brian A. Nosek, Mahzarin R. Banaji, and Anthony G.
-Greenwald. 2002. Harvesting implicit group attitudes and beliefs from a
-demonstration web site. \emph{Group Dynamics: Theory, Research, and
-Practice} 6, 1 (2002), 101--115.
-DOI:https://doi.org/\href{https://doi.org/10.1037/1089-2699.6.1.101}{10.1037/1089-2699.6.1.101}}
-
-\end{CSLReferences}
-
 \end{document}

### 2023-05-25 functions/plotFromPrecis.R

@@ -53,7 +53,8 @@ plotFromPrecis <- function (precis, dataset, ylims = c(.5,1.2), list, embedding)
     geom_segment(data = DFindividual,aes(x=reorder(pw, desc(pw)),xend=reorder(pw, desc(pw)), y  = `5.5%`, yend= `94.5%`,
                                          color = type),
                  position = position_nudge((as.integer(DFindividual$type)-1)*0.15), size = .4)+
-    geom_hline(yintercept = 1, lty = 2, size = .2, alpha = .3) +labs(title = paste(list, ", ", embedding, ", cosine distances, by protected word", sep = ""), subtitle = "Means with 89% highest posterior density intervals")+ 
+    geom_hline(yintercept = 1, lty = 2, size = .2, alpha = .3) + 
+    labs(title = paste(list, ", ", embedding, ", cosine distances, by protected word", sep = ""))+ 
     theme(legend.position='none', plot.title.position = "plot") + ylim(ylims)
   
   
@@ -66,15 +67,24 @@ plotFromPrecis <- function (precis, dataset, ylims = c(.5,1.2), list, embedding)
                                   "different" = "chartreuse4",
                                   "human" = "skyblue",
                                   "none" = "grey"))+ theme(legend.position = c(0.9, 0.8), plot.title.position = "plot")+
-    labs(title = paste(list, ", ", embedding, ",  cosine distances, by connection type", sep = ""), 
-       subtitle = "Means with 89% highest posterior density intervals")+
+    labs(title = paste(list, ", ", embedding, ",  cosine distances, by connection type", sep = ""))+
     ylab("cosine distance")+
     geom_hline(yintercept = 1, lty = 2, size = .2, alpha = .3)
   
   
   
-  grobIndividual <- ggplotGrob(results$plotIndividual+theme_tufte(base_size = 6) + theme(legend.position='none'))
-  grobOverall <- ggplotGrob(results$plotOverall+theme_tufte(base_size = 6)+ theme(legend.position = "top"))
+  grobIndividual <- ggplotGrob(results$plotIndividual+theme_tufte(base_size = 10) +
+                                 theme(legend.position='none',
+                                 plot.title.position = "plot",
+                                 axis.title.x = element_text(size = 8),
+                                 axis.title.y = element_text(size = 8),
+                                 plot.title=element_text(size=11)))
+  grobOverall <- ggplotGrob(results$plotOverall+theme_tufte(base_size = 10)+
+                              theme(legend.position = "top",
[…]

### 2023-06-01 paper/paperDraft.Rmd
literature remains

@@
 knitr::opts_knit$set(root.dir = '../')
-
+library("viridis") 
 library(ggplot2)
@@
 
-Conceptually, the idea is that if a particular harmful stereotype is learned in a particular embedding, then certain groups of words will be systematically closer to (or further from) each other.
+The underlying intuition is that if a particular harmful stereotype is learned in a particular embedding, then certain groups of words will be systematically closer to (or further from) each other.
 This gives rise to the idea of protected groups---for example, in guiding online search completion or recommendation, female words might require protection in that they should not be systematically closer to stereotypically female job names, such as "nurse", "librarian", "waitress", and male words require protection in that they should not be systematically closer to toxic masculinity stereotypes, such as "tough", "never complaining" or "macho".^[However, for some research-related purposes, such as the study of stereotypes across history [@Garg2017hundredYears], embeddings which do not protect certain classes may also be useful.]
@@
 Given a list  $N$ of  supposedly gender neutral words,^[We follow the methodology used in the debate in assuming that there is a class of words  identified as more or less neutral, such as \emph{ballpark, eat, walk, sleep, table}, whose average similarity to the gender direction (or other protected words) is around 0.
-We will have more to say about this assumption when we describe our dataset construction.]  and the gender direction $\mathsf{gd}$,    the direct gender bias is defined as the average cosine similarity of the words in $N$ from $\mathsf{gd}$ ($c$ is a parameter determining how strict we want to be):
+See our list in Appendix \ref{app:custom} and a brief discussion in Subsection \ref{subsec:interpretability}.]  and the gender direction $\mathsf{gd}$,    the direct gender bias is defined as the average cosine similarity of the words in $N$ from $\mathsf{gd}$ ($c$ is a parameter determining how strict we want to be):
 \begin{align*}
@@
 
-A measure of bias in word embeddings which does not employ similarities to  pre-calculated protected-class-level directions, the Word Embedding Association Test (\textsf{WEAT}), has been proposed in  [@Caliskan2017semanticsBiases].
+A measure of bias in word embeddings which does not proceed by identifying bias directions (such as a gender vector), the Word Embedding Association Test (\textsf{WEAT}), has been proposed in  [@Caliskan2017semanticsBiases].
 The idea here is that the   bias between two sets of target words, $X$ and $Y$ (we call them protected words), should be quantified in terms of the cosine similarity  between the protected words and attribute words coming from  two sets of stereotype attribute words, $A$ and $B$ (we will call them attributes).
@@
 
-\noindent The authors use it as a test statistic in some tests, and the assumption that $X$ and $Y$ are of the same size is necessary for the statistic to make sense.
-In the final measure (of effect size), \textsf{WEAT}, means are taken, which makes it insensitive to size differences between $X$ and $Y$:
-\begin{align}
-\mathsf{weat}(A,B) & = \frac{
+\noindent The intention is that large values of $s$ scores suggest systematic differences between how $X$ and $Y$ are related to $A$ and $B$, and therefore are indicative of the presence of bias.
+The authors use it as a test statistic in some tests,^[Note their method assumes $X$ and $Y$ are of the same size.] and the final measure of effect size, \textsf{WEAT}, is constructed by taking means of these values and standardizing:
+\begin{align} \label{eq:weat}
+\mathsf{WEAT}(A,B) & = \frac{
 \mu(\{\mathsf{s}(x,A,B)\}_{x\in X}) -\mu(\{\mathsf{s}(y,A,B)\}_{y\in Y}) 
@@
 
- \textsf{WEAT} is inspired by  the Implicit Association Test (IAT) [@Nosek2002harvesting] used in psychology, and it uses almost the same word sets, allowing for a \emph{prima facie} sensible comparison with bias in humans.
+ \textsf{WEAT} is inspired by  the Implicit Association Test (IAT) [@Nosek2002harvesting] used in psychology, and in some applications it uses almost the same word sets, allowing for a \emph{prima facie} sensible comparison with bias in humans.
 In  [@Caliskan2017semanticsBiases] the authors argue that significant biases---thus measured--- similar to the ones discovered by IAT can be discovered in word embeddings.
-In [@Lauscher2019multidimensional] the methodology is  extended to a multilingual and cross-lingual setting, arguing that using Euclidean distance instead of cosine similarity does not make much difference, while the bias effects vary greatly across embedding models (interestingly, with social media-text trained embeddings being less biased than those based on Wikipedia).
+In [@Lauscher2019multidimensional] the methodology is  extended to a multilingual and cross-lingual setting, arguing that using Euclidean distance instead of cosine similarity does not make much difference, while the bias effects vary greatly across embedding models.^[Interestingly, with social media-text trained embeddings being less biased than those based on Wikipedia.]
  A similar methodology is employed in [@Garg2018years].
-The authors employ word embeddings trained on corpora from different decades to study the shifts in various biases.^[Strictly speaking, these authors use Euclidean distances and their differences, but the way they take averages and averages thereof is analogous, and so what we will have to say about pre-averaging leading to false confidence applies to this methodology as well.]
+The authors employ word embeddings trained on corpora from different decades to study the shifts in various biases through the century.^[Strictly speaking, these authors use Euclidean distances and their differences, but the way they take averages and averages thereof is analogous, and so what we will have to say about pre-averaging leading to false confidence applies to this methodology as well.]
 
@@
 
-\textsf{WEAT}, however, has been developed to investigate biases corresponding to a pair of supposedly opposing stereotypes, and so the question arises as to how to generalize the measure to contexts in which biases with respect to more than two stereotypical groups are to be measured.
+\textsf{WEAT} has been developed to investigate biases corresponding to a pair of supposedly opposing stereotypes,  so the question arises as to how to generalize the measure to contexts in which biases with respect to more than two stereotypical groups are to be measured.
 Such a generalization can be found in [@Manzini2019blackToCriminal].
@@
 \mathsf{s}(t, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cosineDistance}(t,a) \\
-\mathsf{mac}(T,\mathcal{A}) & = \frac{1}{\vert T \vert \,\vert \mathcal{A}\vert}\sum_{t \in T }\sum_{A_j \in \mathcal{A}}    \mathsf{s}(t,A_j)
+\mathsf{MAC}(T,\mathcal{A}) & = \frac{1}{\vert T \vert \,\vert \mathcal{A}\vert}\sum_{t \in T }\sum_{A_j \in \mathcal{A}}    \mathsf{s}(t,A_j)
 \end{align*}
 
-\noindent That is, for each  protected word $t\in T$, and each attribute set $A_j$, they first take the mean of distances for this protected word and all attributes in a given attribute class, and then take the mean of thus obtained means for all the protected words and all the protected classes.^[The authors'  code are in their github repository, [https://github.com/TManzini/DebiasMulticlassWordEmbedding](https://github.com/TManzini/DebiasMulticlassWordEmbedding).]
+\noindent That is, for each  protected word $t\in T$, and each attribute set $A_j$, they first take the mean of distances for this protected word and all attributes in a given attribute class, and then take the mean of thus obtained means for all the protected words and all the protected classes.^[The authors'  code is available through their github repository at  [https://github.com/TManzini/DebiasMulticlassWordEmbedding](https://github.com/TManzini/DebiasMulticlassWordEmbedding).]
 
@@
 
-ggplot(mac, aes(x=x,y=y, label = words))+geom_point()+geom_text(check_overlap = TRUE,
-                                                                 nudge_y = 0.02, size = 9 )+
-  geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.4, yend=.1),  linetype =2, arrow=arrow(), size=.03, color="blue")+geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.4, yend=.2), linetype =4, arrow=arrow(), size=.0, color="blue") + geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.7, yend=.3), linetype =2, arrow=arrow(), size=.05, color="orangered")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.7, yend=.4), linetype =2, arrow=arrow(), size=.05, color="orangered")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=1.1, yend=.1), linetype =3, arrow=arrow(), size=.05, color="chartreuse4")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=1.1, yend=.2), linetype =2, arrow=arrow(), size=.05, color="chartreuse4")+ annotate("text", x = c(.1,.4,.7,1.2), y = c(.45,.45,.45,.45),  label = c("T", "A[1]","A[2]", "A[3]") ,parse = TRUE, color=c("black", "blue","orangered","chartreuse4"),            size=10,  fontface="bold")+xlim(0,1.3)+theme_void(base_size = 25)
+ggplot(mac, aes(x=x,y=y, label = words))+geom_point()+geom_text(check_overlap = TRUE,                               nudge_y = 0.02, size = 9 )+
+  geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.4, yend=.1),  linetype =2, arrow=arrow(), size=.03, color="blue")+
+  geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.4, yend=.2),
+               linetype =4, arrow=arrow(), size=.0, color="blue") + geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.7, yend=.3), linetype =2, arrow=arrow(), size=.05, color="orangered")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.7, yend=.4), linetype =2, arrow=arrow(), size=.05, color="orangered")+
+  geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=1.1, yend=.1), linetype =3, arrow=arrow(), size=.05, color="chartreuse4")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=1.1, yend=.2), linetype =2, arrow=arrow(), size=.05, color="chartreuse4")+
+  annotate("text", x = c(.1,.4,.7,1.1), y = c(.45,.45,.45,.45),  label = c("T", "A[1]","A[2]", "A[3]") ,parse = TRUE, color=c("black", "blue","orangered","chartreuse4"),            size=10,  fontface="bold")+xlim(0,1.3)+theme_void(base_size = 25)
 
@@
 
-Here is an example of results of debiasing that can be found in [@Manzini2019blackToCriminal] (Table \ref{tab:religionOriginal2}).
-It presents the results of applying the procedure with respect to the Religion dataset.
-\textsf{MAC} scores are given for the original datasets, and two resulting from two debiasing methods applied to the Reddit embedding.
-The scores for debiasing come with p-values for the comparison to the original \textsf{MAC}.
+Table \ref{tab:religionOriginal2} contains  an example of \textsf{MAC} scores (and $p$ values, we explain how these are obtained in Subsection \ref{subsec:problems}) before and after  deploying two debiasing methods to the Reddit embedding, where the score is calculated using the Religion word lists from  [@Manzini2019blackToCriminal].
+For our purpose the details of the debiasing method are not important: what matters is that \textsf{MAC} is used in the evaluation of these methods.
 
@@
 
-The first conceptual question we should ask is whether   the initial \textsf{MAC} values lower than 1 indeed are  indicative of the presence of bias?
-Of course, thinking abstractly, 1 is the ideal distance for unrelated words.
-But clearly, there is variation in distances, which might lead to non-biased lists also having \textsf{MAC} smaller than 1.
+The first question we should ask is whether   the initial \textsf{MAC} values lower than 1 indeed are  indicative of the presence of bias?
+Thinking abstractly, 1 is the ideal distance for unrelated words.
+But in fact there is some variation in distances, which might lead to non-biased lists also having \textsf{MAC} scores smaller than 1.
 How much smaller?
@@
 High cosine distance indicates low cosine similarity between values.
-One could think that average cosine similarity equal to approximately 0.141 is not large enough to consider it as bias.
-The authors, though, still aim to mitigate such "bias" to make the distance even larger.
+One could think that average cosine similarity equal to approximately 0.141 is not large enough to claim the presence of a bias to start with.
+The authors, though, still aim to mitigate it by making the distances involved in the \textsf{MAC} calculations even larger.
 The question is, on what basis is this small similarity still considered as a proof of the presence of bias, and whether these small changes are meaningful.
@@
 Later on, in our approach, we introduce such control word lists.
-One of them is a list of words we intuitively considered neutral (see the Appendix for word lists).
-
+One of them is a list of words we intuitively considered neutral.
 Moreover, it might be the case that words that have to do with human activities in general, even if unbiased, are systematically closer to the protected words than merely neutral words.
 This, again, casts doubt on whether comparing \textsf{MAC} to the abstractly ideal value of 1 is a methodologically sound idea.
-For this reason we also use a second list with intuitively non-stereotypical human attributes (again, see the Appendix for a word list).
+For this reason we also use a second list with intuitively non-stereotypical human attributes.^[See  Appendix \ref{app:custom} for the word lists.]
 
@@
 In Figures  (\ref{fig:empirical0}-\ref{fig:empirical3}) we look at the empirical distributions, while paying attention to such divisions.
-The horizontal lines represent the values of 1-\textsf{MAC} (that is, we now talk in terms of cosine similarity rather than cosine distance; this is a trivial linear transformation) that the authors considered indicative of bias for stereotypes corresponding to  given  word lists.
-For instance, in religion, \textsf{MAC} was .859 and considered a sign of bias, so we plot $0\pm (1-.859)\approx .14$ lines around similarity = 0 (that is, distance = 1).
-Notice that most distributions are quite wide, and the proportions of even neutral or human neutral words with similarities higher than the value of 1-\textsf{MAC} deserving debiasing according to the authors is quite high.
+The horizontal lines represent the values of $1 - \mathsf{MAC}$ (that is, we now talk in terms of cosine similarity rather than cosine distance) that the authors considered indicative of bias for stereotypes corresponding to  given  word lists.
+For instance, in religion, \textsf{MAC} was .859, which  was considered a sign of bias, so we plot $0\pm (1-.859)\approx .14$ lines around similarity = 0 (that is, distance = 1).
+Notice that most distributions are quite wide, and the proportions of even neutral or human neutral words with similarities higher than the value of $1 - \mathsf{MAC}$ deserving debiasing according to the authors are quite high.
 
@@
 
-To avoid such problems, later on, we employ control groups and in line with Bayesian methodology, use posterior distributions and  highest posterior density intervals instead of chasing single-point metrics based on pre-averaged data.
-But before we do so, let us spend some time explaining why pre-averaging is a sub-optimal strategy.
+To avoid the problem brought up in this subsection, we employ control groups and in line with Bayesian methodology, use posterior distributions and  highest posterior density intervals instead of chasing single-point metrics based on pre-averaged data.
+Before we do so, we first explain why pre-averaging and chasing single-number metrics is a sub-optimal strategy.
 
@@
 The approaches we have been describing use  means of mean average cosine similarities to measure  similarity between protected words and attributes coming from harmful stereotypes.
-However, a closer inspection look at the individual values that are taken for the calculations, it turns out that there are quite a few  outliers and surprisingly dissimilar words.
+But once we take a look at the individual values, it turns out that the raw data variance is rather high, and there are quite a few  outliers and surprisingly dissimilar words.
 This problem becomes transparent when we examine the visualizations of the individual cosine distances, following the idea that one of the first steps in understanding data is to look at it.
-Let's start with inspecting two examples in Figures \ref{fig:muslim} and \ref{fig:priest}.
+Let's start with inspecting two examples of such visualizations in Figures \ref{fig:muslim} and \ref{fig:priest} (we also include neutral and human predicates to make our point more transparent).
+Again, we emphasize that **we do not condone the associations which we are about to illustrate.**
 
 \begin{figure}[H]
-```{r visMuslimd,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%",message =FALSE, warning = FALSE}
+```{r visMuslimd,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%",message =FALSE, warning = FALSE}
 religion <- read.csv("datasets/cleanedDatasets/religionRedditCleaned.csv")
@@
 ```
-\caption{Actual distances for the protected word "muslim".
-The most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates.
-}
+\caption{Actual distances for the protected word \textsf{muslim}. }
 \label{fig:muslim}
@@
 \begin{figure}[H]
-```{r visPriestd,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
+```{r visPriestd,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%", warning = FALSE, message = FALSE}
 religion <- read.csv("datasets/cleanedDatasets/religionRedditCleaned.csv")
@@
 
-\caption{Actual distances for the protected word "priest", the situation is even less as expected: the nearest attributes are human attributes, and all there seems to be no clear pattern to the distances to other attributes.}
+\caption{Actual distances for the protected word \textsf{priest}.}
 \label{fig:priest}
@@
 
-With such a method the uncertainty involved is not really considered which makes it even more difficult to give reasonable interpretations of the results.
-We propose the use of Bayesian method to obtain some understanding of the influence the uncertainty has on the interpretation of final results.
-
-\noindent
-$s(X,Y,A,B)$ is the statistic used in the significance test, and the $p$-value is obtained by bootstrapping: it is the frequency of $s(X_i,Y_i,A,B)>s(X,Y,A,B)$ for all equally sized partitions $X_i, Y_i$ of $X\cup Y$.
-The effect size is computed by normalizing the difference in means as follows:
-
-\vspace{-2mm}
-
-\footnotesize
-
-\begin{align}
-bias(A,B) & = \frac{
-\mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y}) 
-}{
-\sigma(\{s(w,A,B)\}_{w\in X\cup Y})
-}
-\end{align}
+As transparent in Figures \ref{fig:muslim} and \ref{fig:priest}, for the protected word \textsf{muslim}, the most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates.
+For the protected word \textsf{priest}, the situation is even less as expected: the nearest attributes are human attributes, and all there seems to be no clear pattern to the distances to other attributes.
+
+The general phenomenon that makes us skeptical about running statistical tests on pre-averaged data is that raw datasets of different variance can result in the same pre-averaged data and consequently the same single-number metric.
+In other words, a method that proceeds this way is not very sensitive to the real sample variance.
+
+Let us illustrate how this problem arises in the context of \textsf{WEAT}.
+Once a particular  $s(X,Y,A,B)$ is calculated, the question arises as to whether a value that high could have arisen by chance.
+To address the question, each $s(X,Y,A,B)$ is used in the original paper to generate a  $p$-value by bootstrapping.
+The $p$-value  is the frequency of how often it is the case that  $s(X_i,Y_i,A,B)>s(X,Y,A,B)$ for sampled equally sized partitions $X_i, Y_i$ of $X\cup Y$.
+The WEAT score is then computed by standardizing the difference in means of means by dividing by the standard deviation of means, see equation \eqref{eq:weat}.
+
+<!-- \vspace{-2mm} -->
+
+<!-- \footnotesize  -->
+
+<!-- \begin{align} -->
+<!-- \mathsf{bias}(A,B) & = \frac{ -->
+<!-- \mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y})  -->
+<!-- }{ -->
+<!-- \sigma(\{s(w,A,B)\}_{w\in X\cup Y}) -->
+<!-- } -->
+<!-- \end{align} -->
 
@@
 
-The t-tests they employ are run on average cosines used to calculate \textsf{MAC}.
-
-The effect sizes reported by [@Caliskan2017semanticsBiases] for lists of words for which the embeddings are supposedly biased range from 2.06 to 1.81, and the reported $p$-values are in the range of $10^{-7}-10^{-2}$ with one exception for Math vs arts, where it is $.018$.
+\noindent The \textsf{WEAT} scores reported by [@Caliskan2017semanticsBiases] for lists of words for which the embeddings are supposedly biased range from 2.06 to 1.81, and the reported $p$-values are in the range of $10^{-7}-10^{-2}$ with one exception for *Math vs Arts*, where it is $.018$.
+
 The question is, are those results meaningful?
-
-One way to answer this question is to think in terms of null models.
-If the words actually are samples from two populations with equal mean, how often would we see effect sizes in this range?
+One way to answer this question is to think in terms of null generative models.
+If the words actually are samples from two populations with equal means, how often would we see \textsf{WEAT} scores in this range?
 How often would we reach the $p$-values that the authors reported?
 
-We start our exploration with considering a list of 16 protected words, each class containing 8, with 16 attributes, 8 in each attribute class.
-In each particular sample we will investigate the consequence of the model being in fact null: we draw the cosine distances, every time, from the $\mathsf{Normal}(0,.08)$ distribution.
-$0.08$ is approximately the empirical standard deviation observed in fairly large samples of neutral words, and 16 is the sample size used in the WEAT7 word list, which is not much different from the other sample sizes in word lists used by [@Caliskan2017semanticsBiases]
+Imagine there are two groups of protected words, each of size 8, and two groups of stereotypical attributes, of the same size.^[16 is the sample size used in the WEAT7 word list, which is not much different from the other sample sizes in word lists used by [@Caliskan2017semanticsBiases]).] Each such a collection of samples, as far as our question is involved, is equivalent to a sample of $16^2$ cosine distances.
+Further, imagine there really is no difference between these groups of word and the  model is in fact null.
+That is,  we draw the cosine distances from the $\mathsf{Normal}(0,.08)$ distribution.^[$.08$ is approximately the empirical standard deviation observed in fairly large samples of neutral words.]
 
@@
         axis.ticks.y = element_blank()) +
-  labs(title = "Boostrapped distribution of test statistics",
-       subtitle = "Random null model with N(0,.008) and n = 16")+
+  labs(title = "Boostrapped distribution of test statistics: s values")+
    annotate("label", x = ourStatistic, y = 2,
@@
                            ", p=", 0.011, sep = ""), size = 3)+
-  xlim(-2,2)+xlab("test statistic")+ylab("density")
+  xlim(-2,2)+xlab("s values")+ylab("density")
 
 plotEff <- ggplot(combinations) + geom_density(aes(x = es))+
-  xlab("effect size")+ylab("density")+
+  xlab("WEAT")+ylab("density")+
   geom_vline(xintercept = ourEffectSize, linewidth = .5, 
@@
         axis.ticks.y = element_blank()) +
-  labs(title = "Boostrapped distribution of effect sizes",
-       subtitle = "Random null model with N(0,.008) and n = 16")+
+  labs(title = "Boostrapped distribution of effect sizes: WEAT scores")+
   annotate("label", x = ourEffectSize, y = .5,
@@
         legend.position = c(0.08, 0.89),
-        legend.background = element_rect(fill="lightblue", 
-         size=0.01, linetype="solid"))
-
-```
-
-So let's draw one sample of this type.
-We calculate the s-values, the test statistics and the effect sizes.
-Then, following the original methodology, we obtain bootstrapped distributions of the effect sizes and the test statistics by calculating the same for each possible equal split of the initial random data set.
-The observed test statistic is `r ourStatistic` and `r ourEffectSize`.
-The bootstrapped distributions of the test statistics and effect sizes are illustrated in Figure \ref{fig:caliskanCalc}.
-Quite notably both (two-sided) $p$ values are uneven and rather low.
+        legend.background = element_rect(fill="gray89", 
+         size=0.01, linetype="solid"))+
+  scale_colour_manual(values = c("darkorange3", "grey40"))+
+  xlab("distance to A")+ylab("distance to B")
+
+```
+
+In Figure \ref{fig:caliskanCalc} we illustrate one iteration of the procedure.
+We draw one such sample of size $16^2$.
+Then we actually list all possible ways to split the 16 words in two equal sets (each such a split is one bootstrapped sample) and for each of them we calculate the \textsf{s} values and \textsf{WEAT}.
+What are the resulting distributions of  $\textsf{s}$ scores and what $p$-values do they lead to?
+What are the resulting effect sizes for each bootstrapped sample, and how often can we get an effect size as large as the ones reported in the original paper?
+
+In the bootstrapped samples we would rather expect low \textsf{s} values and low \textsf{WEAT}: after all, these are just random permutations of random distances all of which are drawn from the same null distribution.
+Let's take a look at one such a bootstrapped sample.
+
+On purpose we picked a rather unusual one: the observed test statistic is `r round(ourStatistic,2)` and `r round(ourEffectSize,2)`.
+The bootstrapped distributions of the test statistics and effect sizes are illustrated in Figure \ref{fig:caliskanCalc}, together with this particular example.
+Quite notably both (two-sided) $p$ values for our example are rather low (Figure \ref{fig:caliskanCalc}).
+These facts might suggested that we ended up with a situation where "bias" is present (albeit, due to random noise).
+The reason why we picked it is that it is an example of a word list that ends up with  relatively low $p$-value and a relatively unusual effect size, but nevertheless its closer inspection shows that even for a word list with such properties there is no clear reason to think that the bias is present.
 
 \begin{figure}[H]
-```{r fig:caliskanCalcB2B,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
+```{r fig:caliskanCalcB2D,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
 grid.arrange(plotStat,plotEff)
@@
 
-\caption{Bootstrapped distributions of test statistics and effect sizes in a random sample given the null hypothesis.}
+\caption{Bootstrapped distributions of test statistics and effect sizes in a random sample given the null hypothesis.
+We used a sample from the null  model with N(0,.08) and 16 protected words, and then boostrapped from it, following the original methodology.
+One particular boostrapped sample is highlighted, and discussed further in the text.
+}
 \label{fig:caliskanCalc}
@@
 
-At this point, we might think---right, so while drawing on the assumption of the null hypothesis we just stumbled into a data set that randomly happened to display strong bias.
+At this point, we might think that  we just stumbled into a bootstrapped sample that randomly happened to display strong bias.
 We decide to double-check this by visual inspection expecting exactly this: a strong, clearly visible bias (Figure \ref{fig:caliskanDistances}).
@@
 \begin{figure}[H]
-```{r fig:caliskanDistC2B,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
+```{r fig:caliskanDistC2K,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
 plotCaliskanDist 
@@
 In the calculations of \textsf{WEAT} means are taken twice.
-The $s$-values themselves are means, and then means of $s$-values are compared between groups.
-The test statistic used itself is of this sort as well---while it compares sums, these are sums of sets of the same size, so they are really linear transformations of means of means.
-Statistical troubles start when we run statistical tests on sets of means, for at least two reasons  One, by pre-averaging data we throw away information about sample sizes.
+The \textsf{s}-values themselves are means, and then means of \textsf{s}-values are compared between groups.
+Statistical troubles start when we run statistical tests on sets of means, for at least two reasons.
+
+\begin{enumerate}
+
+\item By pre-averaging data we throw away information about sample sizes.
 For the former point, think about proportions: 10 out of 20 and 2 out of 4 give the same mean, but you would obtain more information by making the former observation rather than by making the latter.
 And especially in this context, in which the word lists are not huge, sample sizes should matter.
-Two, when we pre-average, we remove variation, and so pre-averaging  tend to manufacture false confidence.
+
+\item When we pre-average, we disregard variation, and therefore pre-averaging  tend to manufacture false confidence.
 Group means display less variation than the raw data points, and the standard deviation of a set of means of sets of means is bound to be lower than the original standard deviation in the row data.
 Now, if you calculate your effect size by dividing by the pre-averaged standard deviation, your quite likely to get something that looks like a strong effect size, but the results of your calculations might not track anything interesting.
-
-So let us think again about the question that we are  ultimately interested in.
+\end{enumerate}
+
+Let us think again about the question that we are  ultimately interested in.
 Are the $X$ terms systematically closer to (further from) the $A$ attributes ($B$ attributes) than the $Y$ words?
 But now let's use the raw data points to try to answer these questions.
-As the first stab, let us run two quick $t$-tests to gauge what the raw data illustrated in Figure \ref{fig:caliskanDistances} tell us.
-
+
+To start with, let us run two quick $t$-tests to gauge what the raw data illustrated in Figure \ref{fig:caliskanDistances} tell us.
 First, distances to $A$ attributes for $X$ words and $Y$ words.
-Whoa, the result is significant!
-The $p$-value is $0.02$.
-So sure, the sample is in some sense unusual.
-But the 95\% confidence interval for the difference in means is $[.0052 .061]$, clearly nothing that a reader would expect given that the calculated effect size seemed quite large.
-How about the $B$ attributes?
+Well, the result is---strictly speaking---statistically significant.
+The $p$-value is $0.02$ (more than ten times higher than the $p$-valued obtained by the boostratpping procedure.
+So the sample is in some sense unusual.
+But the 95\% confidence interval for the difference in means is $[.0052, .061]$, clearly nothing that a reader would expect given that the calculated effect size seemed quite large.
+How about the distances to the  $B$ attributes?
 Here the $p$-value is  $.22$ and the 95\% confidence interval is $[-0.03, .009]$, even less of a reason to think a bias is present.
@@
 Calculate the test statistic and effect size the way [@Caliskan2017semanticsBiases] did.
-Do this 10000 times and look at what the distributions of these values are on the assumption of the null model with realistic empirically motivated raw data point standard deviation.
+Do this 10000 times, each time calculating \textsf{WEAT} and \textsf{s} values, and look at what the distributions of these values are on the assumption of the null model with realistic empirically motivated raw data point standard deviation.
 
@@
 ourESplot <- ggplot(dataPLOT )+geom_histogram(aes(x=biasesD0, y = ..density..), bins = 50, alpha = .6)+
-  labs(title="Distribution of effect sizes given a null model")+ xlab("effect size") +
+  labs(title="Distribution of effect sizes given a null model")+ xlab("WEAT ") +
   geom_vline(xintercept = ourEffectSize, linewidth = .5,
@@
 ourStatPlot <- ggplot(dataPLOT )+ geom_histogram(aes(x=svaluesD0, y = ..density..), alpha = .6, bins = 50) +
-  labs(title="Distribution of test statistics given a null model")+ xlab("test statistic") +
+  labs(title="Distribution of test statistics given a null model")+ xlab("s values") +
   geom_vline(xintercept = ourStatistic, linewidth = .5,
@@
 \begin{figure}[H]
-```{r fig:ourDistancesPlot2e,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
-
-grid.arrange(ourESplot, ourStatPlot, ncol = 1)
-
-```
-
-\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08, n=10k$.}
+```{r fig:ourDistancesPlot2f,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
+
+grid.arrange(ourStatPlot, ourESplot,  ncol = 1)
+
+```
+
+\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08, n=10k$.
+We also mark the sample we have been using as an example.}
 \label{fig:ourDistances}
@@
 The first observation is that the supposedly large effect size we obtained is not that unusual even assuming a null model.
-Around 38\% of samples result in effect size at least as extreme.
-This illustrates the point that the effect size used does not track anything interesting.
-Second, the distribution of test statistics is much more narrow, which means that if we use it to calculate $p$-values, it is not too difficult to obtain a supposedly significant test statistic which nevertheless does not correspond to anything interesting happening in the data set.
-
-Where do we stand now?
+Around 38\% of samples result in \textsf{WEAT} score at least as extreme.
+This illustrates the point that it  does not constitute a strong evidence of bias.
+Second, the distribution of $\textsf{s}$ values is much more narrow, which means that if we use it to calculate $p$-values, it is not too difficult to obtain a supposedly significant test statistic which nevertheless does not correspond to anything interesting happening in the data set.
+
  We have seen  that seemingly high effect sizes  might arise even if the underlying processes actually have the same mean.
-The uncertainty resulting from including the raw data point variance into considerations is more extensive than the one suggested by the low p-values obtained from taking means or means of means as data points.
-In this section, we discussed the performance of the \textsf{WEAT} measure, but since the [@Manzini2019blackToCriminal] one is a generalization thereof, including the method of running statistical tests on pre-averaged data, our remarks, \emph{mutatis nutandis}, apply.
-Moreover, a general problem is that studies such as [@Caliskan2017semanticsBiases] or [@Manzini2019blackToCriminal]  do not use  control groups, and this makes the interpretation of the analysis additionally difficult.
+The uncertainty resulting from including the raw data point variance into considerations is more extensive than the one suggested by the low $p$-values obtained from taking means or means of means as data points.
+In the section we discussed the performance of the \textsf{WEAT} measure, but since the [@Manzini2019blackToCriminal] one is a generalization thereof, including the method of running statistical tests on pre-averaged data, our remarks, \emph{mutatis nutandis}, apply.
 
 What is the alternative?
-As we already emphasized: focusing on what the real underlying question is, and trying to answer it using a statistical analysis of the raw data.
+As we already emphasized: focusing on what the real underlying question is, and trying to answer it using a statistical analysis of the raw data  using meaningful  control groups, to ensure interpretability.
 Moreover, since the data sets are not too large and since multiple evaluations are to be made, we will pursue this method from the Bayesian perspective.
-
-# Bayesian method
+Now we have to describe it.
+
+# A Bayesian approach to cosine-based bias
+
 \label{sec:bayesian}
@@
 In this study, 120 researchers and 442 students were asked to assess the truth value of six false statements involving different interpretations of a CI.
-Both researchers and students endorsed, on average, more than three of these statements.] [@Hoekstra2014Misinterpretation]
+Both researchers and students endorsed, on average, more than three of these statements.
+See also [@Hoekstra2014Misinterpretation] for a list of problems with the classical approach.]
 
@@
 
-In a standard Bayesian analysis, the first step is to understand the data and select potential predictors and the outcome variable.
-Once the data is understood, the next step is to formulate a mathematical description of the generative model of the relationships between the predictors and the outcome variable.
+In a standard Bayesian analysis, the first step is to understand the data, think hard about the underlying process  and select potential predictors and the outcome variable.
+The next step is to formulate a mathematical description of the generative model of the relationships between the predictors and the outcome variable.
 Prior distributions must then be chosen for the parameters used in the model.
-Next, Bayesian inference must be applied to find posterior probabilities over the possible parameter values.
-Finally, after finding the posterior probabilities, it is important to check how well the posterior predictions reflect the data with a posterior predictive check.
+Next, Bayesian inference must be applied to find posterior distributions over the possible parameter values.
+Finally, we need to  check how well the posterior predictions reflect the data with a posterior predictive check.
+
 In our analysis the outcome variable is  the cosine distances  between the protected words and attribute words.
 The predictor is a factor determining whether a given attribute word is a neutral word, a human predicate, is stereotypically associated with the protected word, or comes from a different stereotype connected with another protected word.
-The idea is, if bias is present in the embedding, distances to associated attribute words should be systematically lower than to other attribute words.
+The idea is really straightforward: if bias is present in the embedding, distances to associated attribute words should be systematically lower than to other attribute words.
 
@@
 \noindent where $\mathsf{different}, \mathsf{associated},\mathsf{human}$ and $\mathsf{neutral}$ are binary variables.
-
-This completes our description of the underlying mechanism.
+This completes our description of the simple underlying process that we would like to investigate.
+
 Now the priors and the hierarchy.
 We assume all the $a$ parameters come from one distribution, that is normal around a higher-level parameter $\bar{a}$ and so on for the other three groups of parameters.
-That is, $a_{\mathsf{pw[i]}}$ is the average distance of a given particular protected word to attributes stereotypically associated with it, while $\bar{a}$ is the overall average distance of protected words to attributes associated with them     [@kruschke2015bayesian; @statrethinkingbook2020].
+That is, $a_{\mathsf{pw[i]}}$ is the average distance of a given particular protected word to attributes stereotypically associated with it, while $\bar{a}$ is the overall average distance of protected words to attributes associated with them.^[For a thorough introduction to the concepts we're using, see  [@kruschke2015bayesian; @statrethinkingbook2020].]
 \begin{align*}
-d_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{d}, \overline{\sigma_d}) \\
-a_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{a}, \overline{\sigma_a}) \\
-h_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{h}, \overline{\sigma_h}) \\
-n_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{n}, \overline{\sigma_n}) 
+d_{\mathsf{pw[i]}} \sim \mathsf{Norm}(\bar{d}, \overline{\sigma_d}) &  & 
+a_{\mathsf{pw[i]}} \sim \mathsf{Norm}(\bar{a}, \overline{\sigma_a}) \\
+h_{\mathsf{pw[i]}} \sim \mathsf{Norm}(\bar{h}, \overline{\sigma_h}) & & 
+n_{\mathsf{pw[i]}} \sim \mathsf{Norm}(\bar{n}, \overline{\sigma_n}) 
 \end{align*}
@@
 According to our priors, the group means  $\bar{a}$, $\bar{d}$, $\bar{h}$ and $\bar{n}$  all come from one normal distribution with mean equal to $1$ and standard deviation equal to $.3$.
-The standard deviations $\bar{\sigma_a}$, $\bar{\sigma_d}$, $\bar{\sigma_h}$ and $\bar{sigma_n}$ to be estimated, according to our prior,  come from one  distribution, exponential with rate parameter equal to $2$.
+The standard deviations $\bar{\sigma_a}$, $\bar{\sigma_d}$, $\bar{\sigma_h}$ and $\bar{\sigma_n}$ to be estimated, according to our prior,  come from one  distribution, exponential with rate parameter equal to $2$.
 Our priors are slightly skeptical.
@@
 However, we use a rather wide standard deviation ($.3$) to easily account for cases where there is actually much higher similarity between two vectors (especially in cases where the embedding is supposed to be biased).
-Our priors for the standard deviations are also fairly weak.\footnote{To build a model one may use original \textit{Stan} which is a state-of-the-art platform for statistical modeling and high-performance statistical computation (https://mc-stan.org/).
-The simpler option is to use \textit{Rethinking} package from the \textit{Statistical rethinking} course.
-With one of the functions \textit{ulam} one may build RStan models from the formulas in an easy manner}
-
+Our priors for the standard deviations are also fairly weak.
 \begin{align*}
@@
 If the model is a good fit to the data, the simulated data should resemble the observed data.
-A posterior predictive check is crucial because it provides a way to assess the validity of a Bayesian model.
-This is important because a model that fits the observed data well may still fail to generalize to new data.
-In the figure below we can see that the model fits the the data well.
+In Figure \ref{fig:posteriorCheck1} we illustrate a posterior predictive check for one corpus (Reddit) and one word list.
+The remaining posterior predictive checks are in section \ref{appendix:posterior}.
 
@@
 \caption{Example of a posterior predictive check.
-(Left) Actual cosine distances are plotted against mean predictions with 89\% highest posterior density intervals.
+(Top) Actual cosine distances are plotted against mean predictions with 89\% highest posterior density intervals.
 Notice that 90\% of actual values fall within the 89\% HPDI and 55\% of actual values fall into 50\% HPDI, which indicates appropriate performance of the model.
 The left-right alignment of differrent colors coressponds to the fact that cosine differences between elements of different categories differ, to some extent systematically (this will be studied in the results section).
-(Right) Densities of predicted and observed distances.}
+(Bottom) Densities of predicted and observed distances.}
 \label{fig:posteriorCheck1}
@@
 - In Religion (Google, Glove, Reddit) and Race (Google, Glove), the associated attributes are not systematically closer than attributes belonging to different stereotypes, and the difference from neutral and human predicates is rather low, if noticeable.
-The situation is interestingly different in Race (Reddit) where both human and neutral predicates are systematically further than associated and different attributes - but even then, there is no clear difference between associated and different attributes.
+The situation is interestingly different in Race (Reddit) where both human and neutral predicates are systematically further than associated and different attributes---but even then, there is no clear difference between associated and different attributes.
 
@@
 We will just go over a few notable examples, leaving the visual inspection of particular results for other protected words to the reader.
-One general phenomenon is that--as we already pointed out--the word lists are quite short, which contributes to large uncertainty involved in some cases.
+One general phenomenon is that---as we already pointed out---the word lists are quite short, which contributes to large uncertainty involved in some cases.
 
@@
 
-These visualizations can be also handy when it comes to the investigation of the effect that debiasing has on the embedding space.
-In Figures  \ref{fig:empiricalPriorToDebiasing} and  \ref{fig:empiricalDebiased}
- we  see chosen pair depicting the difference in means with 89% highest posterior density intervals before and after applying debiasing.
+Bayesian analyses and  visualizations thereof can be also handy when it comes to the investigation of the effect that debiasing has on the embedding space.
+In Figures  \ref{fig:empiricalPriorToDebiasing} and  \ref{fig:empiricalDebiased}  we  see an example of two visualizations  depicting the difference in means with 89% highest posterior density intervals before and after applying debiasing (the remaining visualizations are in the Appendix).
+
+- In *Gender (Reddit)*, minor differences between different and associated predicates end up being smaller.
+However, this is not achieved by any major change in the relative positions of associated and different predicates with respect to protected words, but rather by shifting them jointly together.
+The only protected word for which a major difference is noticeable is *hers*.
+
+- In *Religion (Reddit)* debiasing aligns general coefficients for all groups together, all of them getting closer to where neutral words were prior to debiasing (this is true also for human predicates in general, which intuitively did not require debiasing).
+For some protected words such as *christian*, *jew*,  the proximity ordering between associated and different predicates have been reversed, and most of the distances shifted a bit towards 1 (sometimes even beyond, such as predicates associated with the word *quran*), but for most protected words, the relative differences between the coefficient did not change much (for instance, there is no change in the way the protected word *muslim* is mistreated).
+
+- For *Race (Reddit)*, general coefficients for different and associated  predicates became aligned.
+However, most of the changes roughly preserve the structure of bias for particular protected words with minor exceptions, such as making the proximities of different predicates for protected words *asian* and *asia*  much lower than associated predicates, which is the main factor responsible for the alignment of the general level coefficients.
+
+In general, debiasing might end up leading to lower differences between general level coefficients for associated and different attributes.
+But that usually happens without any major change to the structure of the coefficients for protected words, sporadic extreme and undesirable changes for some protected words, usually with the side-effect of changing what happens with neutral and human predicates.
+
+We wouldn't be even able to notice these phenomena had we restricted our attention to \textsf{MAC} or \textsf{WEAT} scores only.
+To be able to diagnose and remove biases at the right level of granularity, we need to go beyond single metric chasing.
 
@@
 
-# Related work and conclusions
+# Loose ends and conclusions
 \label{sec:related}
 
-It has been suggested [@Ethayarajh2020Bernstein] to use Bernstein bounds to state the uncertainty of the sample bias in a form of confidence interval.
-The interpretation of this interval is as follows, 95\% confidence interval contains such a range of values, that one can be 95% certain that it contains the population mean.
-As we will prove later in the text, confidence intervals are quite problematic for several reasons, among others the confusing interpretability.
-In [@Ethayarajh2020Bernstein] the current problem of bias estimation was correctly identified.
-Indeed, the bias estimate should not be expressed as a single number without taking into account that the estimate is made on a sample of data and therefore has intrinsic uncertainty.
-However, we argue that Bernstein bounds do not provide the best solution to this problem.
-Applying their method to a popular WinoBias dataset leads to the conclusion that more than 11903 samples are needed to claim a 95\% confidence interval for a bias estimate.
-This amount exceeds existing datasets for bias measurement.
-We propose a more realistic method that can be applied to commonly used datasets to estimate the uncertainty of the bias measurement.
+[@DBLP:journals/corr/abs-1811-07253] employ Bayesian methods to estimate uncertainty in NLP tasks, but they apply their Bayesian Neural Networks-based method to sentiment analysis or named entity recognition, not to bias.
+[@Ethayarajh2020Bernstein] correctly argues that bias estimate should not be expressed as a single number without taking into account that the estimate is made using a sample of data and therefore has intrinsic uncertainty.
+The authors   suggest to use Bernstein bounds to gauge  the uncertainty in terms of  confidence intervals.
+We do not discuss this approach extensively, as we think that confidence intervals are quite problematic for several reasons, among others the confusing interpretability.
+We do not think that  Bernstein bounds provide the best solution to the problem.
+Applying this method to a popular WinoBias dataset leads to the conclusion that more than 11903 samples are needed to claim a 95\% confidence interval for a bias estimate.
+This amount vastly exceeds the existing word lists for bias estimation.
+We propose a more realistic Bayesian method.
+Our conclusion is still that the word lists are sometimes too small, but at least allows for gauging uncertainty as we go on to improve our methodology and  extend the lists gradually.
+
+ [@schröder2021evaluating]  criticize some exisisting bias metrics such as \textsf{MAC} or \textsf{WEAT} on the grounds of them not satisfying some general formal principles, such as magnitude-comparability, and they propose a modification, called \textsf{SAME}.
+[@may-etal-2019-measuring] develop a generalization of \textsf{WEAT} meant to apply to sets of sentences, \textsf{SEAT}, which basically applies \textsf{WEAT} to vector representations of sentences.
+The authors, however, still pre-average and play the game of finding a single-number metric, so our remarks  apply.
 
@@
 The tools that we propose, however, allow for a more fine-grained and multi-level evaluation of bias and debiasing in language models without losing modesty about the uncertainties involved.
+
+The short general lesson here is this: it's complicated.
+Instead of pretending it isn't and chasing single-number metrics, we should rather devote attention to more nuanced and methodologically sound analysis.
 
@@
 
-### Custom lists used in this paper
+### Custom lists used in this paper \label{app:custom}
 

### 2023-06-01 paper/paperDraft.tex
literature remains

@@
 downstream tasks. Roughly speaking, cosine similarity is an imperfect
-mathematical proxy for semantic similarity {[}14{]}.
+mathematical proxy for semantic similarity {[}15{]}.
 
@@
 
-Conceptually, the idea is that if a particular harmful stereotype is
+The underlying intuition is that if a particular harmful stereotype is
 learned in a particular embedding, then certain groups of words will be
@@
   average similarity to the gender direction (or other protected words)
-  is around 0. We will have more to say about this assumption when we
-  describe our dataset construction.} and the gender direction
-\(\mathsf{gd}\), the direct gender bias is defined as the average cosine
-similarity of the words in \(N\) from \(\mathsf{gd}\) (\(c\) is a
-parameter determining how strict we want to be): \begin{align*}
+  is around 0. See our list in Appendix \ref{app:custom} and a brief
+  discussion in Subsection \ref{subsec:interpretability}.} and the
+gender direction \(\mathsf{gd}\), the direct gender bias is defined as
+the average cosine similarity of the words in \(N\) from \(\mathsf{gd}\)
+(\(c\) is a parameter determining how strict we want to be):
+\begin{align*}
 \mathsf{directBias_c(N,gd)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},\mathsf{gd})\vert^c}{\vert N \vert }
@@
   do not discuss this method in this paper, see its criticism in
-  {[}16{]}.}
-
-A measure of bias in word embeddings which does not employ similarities
-to pre-calculated protected-class-level directions, the Word Embedding
-Association Test (\textsf{WEAT}), has been proposed in {[}8{]}. The idea
-here is that the bias between two sets of target words, \(X\) and \(Y\)
-(we call them protected words), should be quantified in terms of the
-cosine similarity between the protected words and attribute words coming
-from two sets of stereotype attribute words, \(A\) and \(B\) (we will
-call them attributes). For instance, \(X\) might be a set of male names,
-\(Y\) a set of female names, \(A\) might contain stereotypically
-male-related, and \(B\) stereotypically female-related career words. The
-association difference for a particular word \(t\) (belonging to either
-\(X\) or \(Y\)) is:
+  {[}17{]}.}
+
+A measure of bias in word embeddings which does not proceed by
+identifying bias directions (such as a gender vector), the Word
+Embedding Association Test (\textsf{WEAT}), has been proposed in
+{[}8{]}. The idea here is that the bias between two sets of target
+words, \(X\) and \(Y\) (we call them protected words), should be
+quantified in terms of the cosine similarity between the protected words
+and attribute words coming from two sets of stereotype attribute words,
+\(A\) and \(B\) (we will call them attributes). For instance, \(X\)
+might be a set of male names, \(Y\) a set of female names, \(A\) might
+contain stereotypically male-related, and \(B\) stereotypically
+female-related career words. The association difference for a particular
+word \(t\) (belonging to either \(X\) or \(Y\)) is:
 
@@
 
-\noindent The authors use it as a test statistic in some tests, and the
-assumption that \(X\) and \(Y\) are of the same size is necessary for
-the statistic to make sense. In the final measure (of effect size),
-\textsf{WEAT}, means are taken, which makes it insensitive to size
-differences between \(X\) and \(Y\): \begin{align}
-\mathsf{weat}(A,B) & = \frac{
+\noindent The intention is that large values of \(s\) scores suggest
+systematic differences between how \(X\) and \(Y\) are related to \(A\)
+and \(B\), and therefore are indicative of the presence of bias. The
+authors use it as a test statistic in some tests,\footnote{Note their
+  method assumes \(X\) and \(Y\) are of the same size.} and the final
+measure of effect size, \textsf{WEAT}, is constructed by taking means of
+these values and standardizing: \begin{align} \label{eq:weat}
+\mathsf{WEAT}(A,B) & = \frac{
 \mu(\{\mathsf{s}(x,A,B)\}_{x\in X}) -\mu(\{\mathsf{s}(y,A,B)\}_{y\in Y}) 
@@
 \textsf{WEAT} is inspired by the Implicit Association Test (IAT)
-{[}17{]} used in psychology, and it uses almost the same word sets,
-allowing for a \emph{prima facie} sensible comparison with bias in
-humans. In {[}8{]} the authors argue that significant biases---thus
-measured--- similar to the ones discovered by IAT can be discovered in
-word embeddings. In {[}11{]} the methodology is extended to a
-multilingual and cross-lingual setting, arguing that using Euclidean
-distance instead of cosine similarity does not make much difference,
-while the bias effects vary greatly across embedding models
-(interestingly, with social media-text trained embeddings being less
-biased than those based on Wikipedia). A similar methodology is employed
-in {[}4{]}. The authors employ word embeddings trained on corpora from
-different decades to study the shifts in various biases.\footnote{Strictly
-  speaking, these authors use Euclidean distances and their differences,
-  but the way they take averages and averages thereof is analogous, and
-  so what we will have to say about pre-averaging leading to false
-  confidence applies to this methodology as well.}
+{[}18{]} used in psychology, and in some applications it uses almost the
+same word sets, allowing for a \emph{prima facie} sensible comparison
+with bias in humans. In {[}8{]} the authors argue that significant
+biases---thus measured--- similar to the ones discovered by IAT can be
+discovered in word embeddings. In {[}11{]} the methodology is extended
+to a multilingual and cross-lingual setting, arguing that using
+Euclidean distance instead of cosine similarity does not make much
+difference, while the bias effects vary greatly across embedding
+models.\footnote{Interestingly, with social media-text trained
+  embeddings being less biased than those based on Wikipedia.} A similar
+methodology is employed in {[}4{]}. The authors employ word embeddings
+trained on corpora from different decades to study the shifts in various
+biases through the century.\footnote{Strictly speaking, these authors
+  use Euclidean distances and their differences, but the way they take
+  averages and averages thereof is analogous, and so what we will have
+  to say about pre-averaging leading to false confidence applies to this
+  methodology as well.}
 
@@
 
-\textsf{WEAT}, however, has been developed to investigate biases
-corresponding to a pair of supposedly opposing stereotypes, and so the
-question arises as to how to generalize the measure to contexts in which
-biases with respect to more than two stereotypical groups are to be
-measured. Such a generalization can be found in {[}12{]}. The authors
-introduce Mean Average Cosine distance (\textsf{MAC}) as a measure of
-bias. Let \(T = \{t_1, \dots, t_k\}\) be a set of protected words, and
-let each \(A_j\in \mathcal{A}\) be a set of attributes stereotypically
-associated with a protected word where \(\mathcal{A}\). For instance,
-when biases related to religion are to be investigated, they use a
-dataset of the format illustrated in Table \ref{tab:religionOriginal}.
-The measure is defined as follows:
+\textsf{WEAT} has been developed to investigate biases corresponding to
+a pair of supposedly opposing stereotypes, so the question arises as to
+how to generalize the measure to contexts in which biases with respect
+to more than two stereotypical groups are to be measured. Such a
+generalization can be found in {[}12{]}. The authors introduce Mean
+Average Cosine distance (\textsf{MAC}) as a measure of bias. Let
+\(T = \{t_1, \dots, t_k\}\) be a set of protected words, and let each
+\(A_j\in \mathcal{A}\) be a set of attributes stereotypically associated
+with a protected word where \(\mathcal{A}\). For instance, when biases
+related to religion are to be investigated, they use a dataset of the
+format illustrated in Table \ref{tab:religionOriginal}. The measure is
+defined as follows:
 
@@
 \mathsf{s}(t, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cosineDistance}(t,a) \\
-\mathsf{mac}(T,\mathcal{A}) & = \frac{1}{\vert T \vert \,\vert \mathcal{A}\vert}\sum_{t \in T }\sum_{A_j \in \mathcal{A}}    \mathsf{s}(t,A_j)
+\mathsf{MAC}(T,\mathcal{A}) & = \frac{1}{\vert T \vert \,\vert \mathcal{A}\vert}\sum_{t \in T }\sum_{A_j \in \mathcal{A}}    \mathsf{s}(t,A_j)
 \end{align*}
@@
 take the mean of thus obtained means for all the protected words and all
-the protected classes.\footnote{The authors' code are in their github
-  repository,
+the protected classes.\footnote{The authors' code is available through
+  their github repository at
   \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding}.}
@@
 
-Here is an example of results of debiasing that can be found in {[}12{]}
-(Table \ref{tab:religionOriginal2}). It presents the results of applying
-the procedure with respect to the Religion dataset. \textsf{MAC} scores
-are given for the original datasets, and two resulting from two
-debiasing methods applied to the Reddit embedding. The scores for
-debiasing come with p-values for the comparison to the original
-\textsf{MAC}.
+Table \ref{tab:religionOriginal2} contains an example of \textsf{MAC}
+scores (and \(p\) values, we explain how these are obtained in
+Subsection \ref{subsec:problems}) before and after deploying two
+debiasing methods to the Reddit embedding, where the score is calculated
+using the Religion word lists from {[}12{]}. For our purpose the details
+of the debiasing method are not important: what matters is that
+\textsf{MAC} is used in the evaluation of these methods.
 
@@
 
-The first conceptual question we should ask is whether the initial
-\textsf{MAC} values lower than 1 indeed are indicative of the presence
-of bias? Of course, thinking abstractly, 1 is the ideal distance for
-unrelated words. But clearly, there is variation in distances, which
-might lead to non-biased lists also having \textsf{MAC} smaller than 1.
-How much smaller? What may attract attention is the fact that the value
-of cosine distance in ``Biased'' category is already quite high
-(i.e.~close to 1) even before debiasing. High cosine distance indicates
-low cosine similarity between values. One could think that average
-cosine similarity equal to approximately 0.141 is not large enough to
-consider it as bias. The authors, though, still aim to mitigate such
-``bias'' to make the distance even larger. The question is, on what
-basis is this small similarity still considered as a proof of the
-presence of bias, and whether these small changes are meaningful.
+The first question we should ask is whether the initial \textsf{MAC}
+values lower than 1 indeed are indicative of the presence of bias?
+Thinking abstractly, 1 is the ideal distance for unrelated words. But in
+fact there is some variation in distances, which might lead to
+non-biased lists also having \textsf{MAC} scores smaller than 1. How
+much smaller? What may attract attention is the fact that the value of
+cosine distance in ``Biased'' category is already quite high (i.e.~close
+to 1) even before debiasing. High cosine distance indicates low cosine
+similarity between values. One could think that average cosine
+similarity equal to approximately 0.141 is not large enough to claim the
+presence of a bias to start with. The authors, though, still aim to
+mitigate it by making the distances involved in the \textsf{MAC}
+calculations even larger. The question is, on what basis is this small
+similarity still considered as a proof of the presence of bias, and
+whether these small changes are meaningful.
 
@@
 introduce such control word lists. One of them is a list of words we
-intuitively considered neutral (see the Appendix for word lists).
-
-Moreover, it might be the case that words that have to do with human
-activities in general, even if unbiased, are systematically closer to
-the protected words than merely neutral words. This, again, casts doubt
-on whether comparing \textsf{MAC} to the abstractly ideal value of 1 is
-a methodologically sound idea. For this reason we also use a second list
-with intuitively non-stereotypical human attributes (again, see the
-Appendix for a word list).
+intuitively considered neutral. Moreover, it might be the case that
+words that have to do with human activities in general, even if
+unbiased, are systematically closer to the protected words than merely
+neutral words. This, again, casts doubt on whether comparing
+\textsf{MAC} to the abstractly ideal value of 1 is a methodologically
+sound idea. For this reason we also use a second list with intuitively
+non-stereotypical human attributes.\footnote{See Appendix
+  \ref{app:custom} for the word lists.}
 
@@
 empirical distributions, while paying attention to such divisions. The
-horizontal lines represent the values of 1-\textsf{MAC} (that is, we now
-talk in terms of cosine similarity rather than cosine distance; this is
-a trivial linear transformation) that the authors considered indicative
-of bias for stereotypes corresponding to given word lists. For instance,
-in religion, \textsf{MAC} was .859 and considered a sign of bias, so we
-plot \(0\pm (1-.859)\approx .14\) lines around similarity = 0 (that is,
+horizontal lines represent the values of \(1 - \mathsf{MAC}\) (that is,
+we now talk in terms of cosine similarity rather than cosine distance)
+that the authors considered indicative of bias for stereotypes
+corresponding to given word lists. For instance, in religion,
+\textsf{MAC} was .859, which was considered a sign of bias, so we plot
+\(0\pm (1-.859)\approx .14\) lines around similarity = 0 (that is,
 distance = 1). Notice that most distributions are quite wide, and the
 proportions of even neutral or human neutral words with similarities
-higher than the value of 1-\textsf{MAC} deserving debiasing according to
-the authors is quite high.
+higher than the value of \(1 - \mathsf{MAC}\) deserving debiasing
+according to the authors are quite high.
 
@@
 
-To avoid such problems, later on, we employ control groups and in line
-with Bayesian methodology, use posterior distributions and highest
-posterior density intervals instead of chasing single-point metrics
-based on pre-averaged data. But before we do so, let us spend some time
-explaining why pre-averaging is a sub-optimal strategy.
+To avoid the problem brought up in this subsection, we employ control
+groups and in line with Bayesian methodology, use posterior
+distributions and highest posterior density intervals instead of chasing
+single-point metrics based on pre-averaged data. Before we do so, we
+first explain why pre-averaging and chasing single-number metrics is a
+sub-optimal strategy.
 
@@
 similarities to measure similarity between protected words and
-attributes coming from harmful stereotypes. However, a closer inspection
-look at the individual values that are taken for the calculations, it
-turns out that there are quite a few outliers and surprisingly
-dissimilar words. This problem becomes transparent when we examine the
+attributes coming from harmful stereotypes. But once we take a look at
+the individual values, it turns out that the raw data variance is rather
+high, and there are quite a few outliers and surprisingly dissimilar
+words. This problem becomes transparent when we examine the
 visualizations of the individual cosine distances, following the idea
 that one of the first steps in understanding data is to look at it.
-Let's start with inspecting two examples in Figures \ref{fig:muslim} and
-\ref{fig:priest}.
+Let's start with inspecting two examples of such visualizations in
+Figures \ref{fig:muslim} and \ref{fig:priest} (we also include neutral
+and human predicates to make our point more transparent). Again, we
+emphasize that \textbf{we do not condone the associations which we are
+about to illustrate.}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visMuslimd-1} \end{center}
-\caption{Actual distances for the protected word "muslim".
-The most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates.
-}
+\begin{center}\includegraphics[width=0.9\linewidth]{paperDraft_files/figure-latex/visMuslimd-1} \end{center}
+\caption{Actual distances for the protected word \textsf{muslim}. }
 \label{fig:muslim}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visPriestd-1} \end{center}
-
-\caption{Actual distances for the protected word "priest", the situation is even less as expected: the nearest attributes are human attributes, and all there seems to be no clear pattern to the distances to other attributes.}
+\begin{center}\includegraphics[width=0.9\linewidth]{paperDraft_files/figure-latex/visPriestd-1} \end{center}
+
+\caption{Actual distances for the protected word \textsf{priest}.}
 \label{fig:priest}
@@
 
-With such a method the uncertainty involved is not really considered
-which makes it even more difficult to give reasonable interpretations of
-the results. We propose the use of Bayesian method to obtain some
-understanding of the influence the uncertainty has on the interpretation
-of final results.
-
-\noindent \(s(X,Y,A,B)\) is the statistic used in the significance test,
-and the \(p\)-value is obtained by bootstrapping: it is the frequency of
-\(s(X_i,Y_i,A,B)>s(X,Y,A,B)\) for all equally sized partitions
-\(X_i, Y_i\) of \(X\cup Y\). The effect size is computed by normalizing
-the difference in means as follows:
-
-\vspace{-2mm}
-
-\footnotesize
-
-\begin{align}
-bias(A,B) & = \frac{
-\mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y}) 
-}{
-\sigma(\{s(w,A,B)\}_{w\in X\cup Y})
+As transparent in Figures \ref{fig:muslim} and \ref{fig:priest}, for the
+protected word \textsf{muslim}, the most similar attributes tend to be
+the ones associated with it stereotypically, but then words associated
+with other stereotypes come closer than neutral or human predicates. For
+the protected word \textsf{priest}, the situation is even less as
+expected: the nearest attributes are human attributes, and all there
+seems to be no clear pattern to the distances to other attributes.
+
+The general phenomenon that makes us skeptical about running statistical
+tests on pre-averaged data is that raw datasets of different variance
+can result in the same pre-averaged data and consequently the same
+single-number metric. In other words, a method that proceeds this way is
+not very sensitive to the real sample variance.
+
+Let us illustrate how this problem arises in the context of
+\textsf{WEAT}. Once a particular \(s(X,Y,A,B)\) is calculated, the
+question arises as to whether a value that high could have arisen by
+chance. To address the question, each \(s(X,Y,A,B)\) is used in the
+original paper to generate a \(p\)-value by bootstrapping. The
+\(p\)-value is the frequency of how often it is the case that
+\(s(X_i,Y_i,A,B)>s(X,Y,A,B)\) for sampled equally sized partitions
+\(X_i, Y_i\) of \(X\cup Y\). The WEAT score is then computed by
+standardizing the difference in means of means by dividing by the
+standard deviation of means, see equation \eqref{eq:weat}.
+
+\normalsize
+
+\noindent The \textsf{WEAT} scores reported by {[}8{]} for lists of
+words for which the embeddings are supposedly biased range from 2.06 to
+1.81, and the reported \(p\)-values are in the range of
+\(10^{-7}-10^{-2}\) with one exception for \emph{Math vs Arts}, where it
+is \(.018\).
+
+The question is, are those results meaningful? One way to answer this
+question is to think in terms of null generative models. If the words
+actually are samples from two populations with equal means, how often
+would we see \textsf{WEAT} scores in this range? How often would we
+reach the \(p\)-values that the authors reported?
+
+Imagine there are two groups of protected words, each of size 8, and two
+groups of stereotypical attributes, of the same size.\footnote{16 is the
+  sample size used in the WEAT7 word list, which is not much different
+  from the other sample sizes in word lists used by {[}8{]}).} Each such
+a collection of samples, as far as our question is involved, is
+equivalent to a sample of \(16^2\) cosine distances. Further, imagine
+there really is no difference between these groups of word and the model
+is in fact null. That is, we draw the cosine distances from the
+\(\mathsf{Normal}(0,.08)\) distribution.\footnote{\(.08\) is
+  approximately the empirical standard deviation observed in fairly
+  large samples of neutral words.}
+
+In Figure \ref{fig:caliskanCalc} we illustrate one iteration of the
+procedure. We draw one such sample of size \(16^2\). Then we actually
+list all possible ways to split the 16 words in two equal sets (each
+such a split is one bootstrapped sample) and for each of them we
+calculate the \textsf{s} values and \textsf{WEAT}. What are the
+resulting distributions of \(\textsf{s}\) scores and what \(p\)-values
+do they lead to? What are the resulting effect sizes for each
+bootstrapped sample, and how often can we get an effect size as large as
+the ones reported in the original paper?
+
+In the bootstrapped samples we would rather expect low \textsf{s} values
+and low \textsf{WEAT}: after all, these are just random permutations of
+random distances all of which are drawn from the same null distribution.
+Let's take a look at one such a bootstrapped sample.
+
+On purpose we picked a rather unusual one: the observed test statistic
+is 0.39 and 1.27. The bootstrapped distributions of the test statistics
+and effect sizes are illustrated in Figure \ref{fig:caliskanCalc},
+together with this particular example. Quite notably both (two-sided)
+\(p\) values for our example are rather low (Figure
+\ref{fig:caliskanCalc}). These facts might suggested that we ended up
+with a situation where ``bias'' is present (albeit, due to random
+noise). The reason why we picked it is that it is an example of a word
+list that ends up with relatively low \(p\)-value and a relatively
+unusual effect size, but nevertheless its closer inspection shows that
+even for a word list with such properties there is no clear reason to
+think that the bias is present.
+
+\begin{figure}[H]
+
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanCalcB2D-1} \end{center}
+
+\caption{Bootstrapped distributions of test statistics and effect sizes in a random sample given the null hypothesis.
+We used a sample from the null  model with N(0,.08) and 16 protected words, and then boostrapped from it, following the original methodology.
+One particular boostrapped sample is highlighted, and discussed further in the text.
 }
-\end{align}
-
-\normalsize
-
-The t-tests they employ are run on average cosines used to calculate
-\textsf{MAC}.
-
-The effect sizes reported by {[}8{]} for lists of words for which the
-embeddings are supposedly biased range from 2.06 to 1.81, and the
-reported \(p\)-values are in the range of \(10^{-7}-10^{-2}\) with one
-exception for Math vs arts, where it is \(.018\). The question is, are
-those results meaningful?
-
-One way to answer this question is to think in terms of null models. If
-the words actually are samples from two populations with equal mean, how
-often would we see effect sizes in this range? How often would we reach
-the \(p\)-values that the authors reported?
-
-We start our exploration with considering a list of 16 protected words,
-each class containing 8, with 16 attributes, 8 in each attribute class.
-In each particular sample we will investigate the consequence of the
-model being in fact null: we draw the cosine distances, every time, from
-the \(\mathsf{Normal}(0,.08)\) distribution. \(0.08\) is approximately
-the empirical standard deviation observed in fairly large samples of
-neutral words, and 16 is the sample size used in the WEAT7 word list,
-which is not much different from the other sample sizes in word lists
-used by {[}8{]}
-
-So let's draw one sample of this type. We calculate the s-values, the
-test statistics and the effect sizes. Then, following the original
-methodology, we obtain bootstrapped distributions of the effect sizes
-and the test statistics by calculating the same for each possible equal
-split of the initial random data set. The observed test statistic is
-0.386222 and 1.2671948. The bootstrapped distributions of the test
-statistics and effect sizes are illustrated in Figure
-\ref{fig:caliskanCalc}. Quite notably both (two-sided) \(p\) values are
-uneven and rather low.
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanCalcB2B-1} \end{center}
-
-\caption{Bootstrapped distributions of test statistics and effect sizes in a random sample given the null hypothesis.}
 \label{fig:caliskanCalc}
@@
 
-At this point, we might think---right, so while drawing on the
-assumption of the null hypothesis we just stumbled into a data set that
-randomly happened to display strong bias. We decide to double-check this
-by visual inspection expecting exactly this: a strong, clearly visible
-bias (Figure \ref{fig:caliskanDistances}).
+At this point, we might think that we just stumbled into a bootstrapped
+sample that randomly happened to display strong bias. We decide to
+double-check this by visual inspection expecting exactly this: a strong,
+clearly visible bias (Figure \ref{fig:caliskanDistances}).
 
@@
 
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanDistC2B-1} \end{center}
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanDistC2K-1} \end{center}
 
@@
 In the calculations of \textsf{WEAT} means are taken twice. The
-\(s\)-values themselves are means, and then means of \(s\)-values are
-compared between groups. The test statistic used itself is of this sort
-as well---while it compares sums, these are sums of sets of the same
-size, so they are really linear transformations of means of means.
-Statistical troubles start when we run statistical tests on sets of
-means, for at least two reasons One, by pre-averaging data we throw away
-information about sample sizes. For the former point, think about
-proportions: 10 out of 20 and 2 out of 4 give the same mean, but you
-would obtain more information by making the former observation rather
-than by making the latter. And especially in this context, in which the
-word lists are not huge, sample sizes should matter. Two, when we
-pre-average, we remove variation, and so pre-averaging tend to
-manufacture false confidence. Group means display less variation than
-the raw data points, and the standard deviation of a set of means of
-sets of means is bound to be lower than the original standard deviation
-in the row data. Now, if you calculate your effect size by dividing by
-the pre-averaged standard deviation, your quite likely to get something
-that looks like a strong effect size, but the results of your
-calculations might not track anything interesting.
-
-So let us think again about the question that we are ultimately
-interested in. Are the \(X\) terms systematically closer to (further
-from) the \(A\) attributes (\(B\) attributes) than the \(Y\) words? But
-now let's use the raw data points to try to answer these questions. As
-the first stab, let us run two quick \(t\)-tests to gauge what the raw
-data illustrated in Figure \ref{fig:caliskanDistances} tell us.
-
-First, distances to \(A\) attributes for \(X\) words and \(Y\) words.
-Whoa, the result is significant! The \(p\)-value is \(0.02\). So sure,
-the sample is in some sense unusual. But the 95\% confidence interval
-for the difference in means is \([.0052 .061]\), clearly nothing that a
-reader would expect given that the calculated effect size seemed quite
-large. How about the \(B\) attributes? Here the \(p\)-value is \(.22\)
-and the 95\% confidence interval is \([-0.03, .009]\), even less of a
-reason to think a bias is present.
+\textsf{s}-values themselves are means, and then means of
+\textsf{s}-values are compared between groups. Statistical troubles
+start when we run statistical tests on sets of means, for at least two
+reasons.
+
+\begin{enumerate}
+
+\item By pre-averaging data we throw away information about sample sizes.
+For the former point, think about proportions: 10 out of 20 and 2 out of 4 give the same mean, but you would obtain more information by making the former observation rather than by making the latter.
+And especially in this context, in which the word lists are not huge, sample sizes should matter.
+
+\item When we pre-average, we disregard variation, and therefore pre-averaging  tend to manufacture false confidence.
+Group means display less variation than the raw data points, and the standard deviation of a set of means of sets of means is bound to be lower than the original standard deviation in the row data.
+Now, if you calculate your effect size by dividing by the pre-averaged standard deviation, your quite likely to get something that looks like a strong effect size, but the results of your calculations might not track anything interesting.
+\end{enumerate}
+
+Let us think again about the question that we are ultimately interested
+in. Are the \(X\) terms systematically closer to (further from) the
+\(A\) attributes (\(B\) attributes) than the \(Y\) words? But now let's
+use the raw data points to try to answer these questions.
+
+To start with, let us run two quick \(t\)-tests to gauge what the raw
+data illustrated in Figure \ref{fig:caliskanDistances} tell us. First,
+distances to \(A\) attributes for \(X\) words and \(Y\) words. Well, the
+result is---strictly speaking---statistically significant. The
+\(p\)-value is \(0.02\) (more than ten times higher than the
+\(p\)-valued obtained by the boostratpping procedure. So the sample is
+in some sense unusual. But the 95\% confidence interval for the
+difference in means is \([.0052, .061]\), clearly nothing that a reader
+would expect given that the calculated effect size seemed quite large.
+How about the distances to the \(B\) attributes? Here the \(p\)-value is
+\(.22\) and the 95\% confidence interval is \([-0.03, .009]\), even less
+of a reason to think a bias is present.
 
@@
 Calculate the test statistic and effect size the way {[}8{]} did. Do
-this 10000 times and look at what the distributions of these values are
-on the assumption of the null model with realistic empirically motivated
-raw data point standard deviation.
+this 10000 times, each time calculating \textsf{WEAT} and \textsf{s}
+values, and look at what the distributions of these values are on the
+assumption of the null model with realistic empirically motivated raw
+data point standard deviation.
 
@@
 
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2e-1} \end{center}
-
-\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08, n=10k$.}
+\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2f-1} \end{center}
+
+\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08, n=10k$.
+We also mark the sample we have been using as an example.}
 \label{fig:ourDistances}
@@
 obtained is not that unusual even assuming a null model. Around 38\% of
-samples result in effect size at least as extreme. This illustrates the
-point that the effect size used does not track anything interesting.
-Second, the distribution of test statistics is much more narrow, which
-means that if we use it to calculate \(p\)-values, it is not too
-difficult to obtain a supposedly significant test statistic which
-nevertheless does not correspond to anything interesting happening in
-the data set.
-
-Where do we stand now? We have seen that seemingly high effect sizes
-might arise even if the underlying processes actually have the same
-mean. The uncertainty resulting from including the raw data point
-variance into considerations is more extensive than the one suggested by
-the low p-values obtained from taking means or means of means as data
-points. In this section, we discussed the performance of the
-\textsf{WEAT} measure, but since the {[}12{]} one is a generalization
-thereof, including the method of running statistical tests on
-pre-averaged data, our remarks, \emph{mutatis nutandis}, apply.
-Moreover, a general problem is that studies such as {[}8{]} or {[}12{]}
-do not use control groups, and this makes the interpretation of the
-analysis additionally difficult.
+samples result in \textsf{WEAT} score at least as extreme. This
+illustrates the point that it does not constitute a strong evidence of
+bias. Second, the distribution of \(\textsf{s}\) values is much more
+narrow, which means that if we use it to calculate \(p\)-values, it is
+not too difficult to obtain a supposedly significant test statistic
+which nevertheless does not correspond to anything interesting happening
+in the data set.
+
+We have seen that seemingly high effect sizes might arise even if the
+underlying processes actually have the same mean. The uncertainty
+resulting from including the raw data point variance into considerations
+is more extensive than the one suggested by the low \(p\)-values
+obtained from taking means or means of means as data points. In the
+section we discussed the performance of the \textsf{WEAT} measure, but
+since the {[}12{]} one is a generalization thereof, including the method
+of running statistical tests on pre-averaged data, our remarks,
+\emph{mutatis nutandis}, apply.
 
@@
 real underlying question is, and trying to answer it using a statistical
-analysis of the raw data. Moreover, since the data sets are not too
-large and since multiple evaluations are to be made, we will pursue this
-method from the Bayesian perspective.
-
-\hypertarget{bayesian-method}{%
-\section{Bayesian method}\label{bayesian-method}}
+analysis of the raw data using meaningful control groups, to ensure
+interpretability. Moreover, since the data sets are not too large and
+since multiple evaluations are to be made, we will pursue this method
+from the Bayesian perspective. Now we have to describe it.
+
+\hypertarget{a-bayesian-approach-to-cosine-based-bias}{%
+\section{A Bayesian approach to cosine-based
+bias}\label{a-bayesian-approach-to-cosine-based-bias}}
 
@@
   procedure is performed a fixed amount of times. For a nice survey and
-  explanation of these misinterpretations, see {[}15{]}. For a
+  explanation of these misinterpretations, see {[}16{]}. For a
   psychological study of the occurrence of such misinterpretations, see
@@
   interpretations of a CI. Both researchers and students endorsed, on
-  average, more than three of these statements.} {[}7{]}
+  average, more than three of these statements. See also {[}7{]} for a
+  list of problems with the classical approach.}
 
@@
 In a standard Bayesian analysis, the first step is to understand the
-data and select potential predictors and the outcome variable. Once the
-data is understood, the next step is to formulate a mathematical
-description of the generative model of the relationships between the
-predictors and the outcome variable. Prior distributions must then be
-chosen for the parameters used in the model. Next, Bayesian inference
-must be applied to find posterior probabilities over the possible
-parameter values. Finally, after finding the posterior probabilities, it
-is important to check how well the posterior predictions reflect the
-data with a posterior predictive check. In our analysis the outcome
-variable is the cosine distances between the protected words and
-attribute words. The predictor is a factor determining whether a given
-attribute word is a neutral word, a human predicate, is stereotypically
-associated with the protected word, or comes from a different stereotype
-connected with another protected word. The idea is, if bias is present
-in the embedding, distances to associated attribute words should be
-systematically lower than to other attribute words.
+data, think hard about the underlying process and select potential
+predictors and the outcome variable. The next step is to formulate a
+mathematical description of the generative model of the relationships
+between the predictors and the outcome variable. Prior distributions
+must then be chosen for the parameters used in the model. Next, Bayesian
+inference must be applied to find posterior distributions over the
+possible parameter values. Finally, we need to check how well the
+posterior predictions reflect the data with a posterior predictive
+check.
+
+In our analysis the outcome variable is the cosine distances between the
+protected words and attribute words. The predictor is a factor
+determining whether a given attribute word is a neutral word, a human
+predicate, is stereotypically associated with the protected word, or
+comes from a different stereotype connected with another protected word.
+The idea is really straightforward: if bias is present in the embedding,
+distances to associated attribute words should be systematically lower
+than to other attribute words.
 
@@
 \(\mathsf{different}, \mathsf{associated},\mathsf{human}\) and
-\(\mathsf{neutral}\) are binary variables.
-
-This completes our description of the underlying mechanism. Now the
-priors and the hierarchy. We assume all the \(a\) parameters come from
-one distribution, that is normal around a higher-level parameter
-\(\bar{a}\) and so on for the other three groups of parameters. That is,
-\(a_{\mathsf{pw[i]}}\) is the average distance of a given particular
-protected word to attributes stereotypically associated with it, while
-\(\bar{a}\) is the overall average distance of protected words to
-attributes associated with them {[}10,13{]}. \begin{align*}
-d_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{d}, \overline{\sigma_d}) \\
-a_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{a}, \overline{\sigma_a}) \\
-h_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{h}, \overline{\sigma_h}) \\
-n_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{n}, \overline{\sigma_n}) 
+\(\mathsf{neutral}\) are binary variables. This completes our
+description of the simple underlying process that we would like to
+investigate.
+
+Now the priors and the hierarchy. We assume all the \(a\) parameters
+come from one distribution, that is normal around a higher-level
+parameter \(\bar{a}\) and so on for the other three groups of
+parameters. That is, \(a_{\mathsf{pw[i]}}\) is the average distance of a
+given particular protected word to attributes stereotypically associated
+with it, while \(\bar{a}\) is the overall average distance of protected
+words to attributes associated with them.\footnote{For a thorough
+  introduction to the concepts we're using, see {[}10,14{]}.}
+\begin{align*}
+d_{\mathsf{pw[i]}} \sim \mathsf{Norm}(\bar{d}, \overline{\sigma_d}) &  & 
+a_{\mathsf{pw[i]}} \sim \mathsf{Norm}(\bar{a}, \overline{\sigma_a}) \\
+h_{\mathsf{pw[i]}} \sim \mathsf{Norm}(\bar{h}, \overline{\sigma_h}) & & 
+n_{\mathsf{pw[i]}} \sim \mathsf{Norm}(\bar{n}, \overline{\sigma_n}) 
 \end{align*}
@@
 deviations \(\bar{\sigma_a}\), \(\bar{\sigma_d}\), \(\bar{\sigma_h}\)
-and \(\bar{sigma_n}\) to be estimated, according to our prior, come from
-one distribution, exponential with rate parameter equal to \(2\). Our
-priors are slightly skeptical. They do reflect our knowledge and
+and \(\bar{\sigma_n}\) to be estimated, according to our prior, come
+from one distribution, exponential with rate parameter equal to \(2\).
+Our priors are slightly skeptical. They do reflect our knowledge and
 intuition on the probable distribution of the cosine distances in the
@@
 to be biased). Our priors for the standard deviations are also fairly
-weak.\footnote{To build a model one may use original \textit{Stan} which is a state-of-the-art platform for statistical modeling and high-performance statistical computation (https://mc-stan.org/).
-The simpler option is to use \textit{Rethinking} package from the \textit{Statistical rethinking} course.
-With one of the functions \textit{ulam} one may build RStan models from the formulas in an easy manner}
-
-\begin{align*}
+weak. \begin{align*}
 \bar{d}, \bar{a}, \bar{h}, \bar{n} &\sim \mathsf{Norm}(1, .3)\\ 
@@
 data. If the model is a good fit to the data, the simulated data should
-resemble the observed data. A posterior predictive check is crucial
-because it provides a way to assess the validity of a Bayesian model.
-This is important because a model that fits the observed data well may
-still fail to generalize to new data. In the figure below we can see
-that the model fits the the data well.
+resemble the observed data. In Figure \ref{fig:posteriorCheck1} we
+illustrate a posterior predictive check for one corpus (Reddit) and one
+word list. The remaining posterior predictive checks are in section
+\ref{appendix:posterior}.
 
@@
   predicates are systematically further than associated and different
-  attributes - but even then, there is no clear difference between
+  attributes---but even then, there is no clear difference between
   associated and different attributes.
@@
 leaving the visual inspection of particular results for other protected
-words to the reader. One general phenomenon is that--as we already
-pointed out--the word lists are quite short, which contributes to large
+words to the reader. One general phenomenon is that---as we already
+pointed out---the word lists are quite short, which contributes to large
 uncertainty involved in some cases.
@@
 
-These visualizations can be also handy when it comes to the
-investigation of the effect that debiasing has on the embedding space.
-In Figures \ref{fig:empiricalPriorToDebiasing} and
-\ref{fig:empiricalDebiased} we see chosen pair depicting the difference
-in means with 89\% highest posterior density intervals before and after
-applying debiasing.
+Bayesian analyses and visualizations thereof can be also handy when it
+comes to the investigation of the effect that debiasing has on the
+embedding space. In Figures \ref{fig:empiricalPriorToDebiasing} and
+\ref{fig:empiricalDebiased} we see an example of two visualizations
+depicting the difference in means with 89\% highest posterior density
+intervals before and after applying debiasing (the remaining
+visualizations are in the Appendix).
+
+\begin{itemize}
+\item
+  In \emph{Gender (Reddit)}, minor differences between different and
+  associated predicates end up being smaller. However, this is not
+  achieved by any major change in the relative positions of associated
+  and different predicates with respect to protected words, but rather
+  by shifting them jointly together. The only protected word for which a
+  major difference is noticeable is \emph{hers}.
+\item
+  In \emph{Religion (Reddit)} debiasing aligns general coefficients for
+  all groups together, all of them getting closer to where neutral words
+  were prior to debiasing (this is true also for human predicates in
+  general, which intuitively did not require debiasing). For some
+  protected words such as \emph{christian}, \emph{jew}, the proximity
+  ordering between associated and different predicates have been
+  reversed, and most of the distances shifted a bit towards 1 (sometimes
+  even beyond, such as predicates associated with the word
+  \emph{quran}), but for most protected words, the relative differences
+  between the coefficient did not change much (for instance, there is no
+  change in the way the protected word \emph{muslim} is mistreated).
+\item
+  For \emph{Race (Reddit)}, general coefficients for different and
+  associated predicates became aligned. However, most of the changes
+  roughly preserve the structure of bias for particular protected words
+  with minor exceptions, such as making the proximities of different
+  predicates for protected words \emph{asian} and \emph{asia} much lower
+  than associated predicates, which is the main factor responsible for
+  the alignment of the general level coefficients.
+\end{itemize}
+
+In general, debiasing might end up leading to lower differences between
+general level coefficients for associated and different attributes. But
+that usually happens without any major change to the structure of the
+coefficients for protected words, sporadic extreme and undesirable
+changes for some protected words, usually with the side-effect of
+changing what happens with neutral and human predicates.
+
+We wouldn't be even able to notice these phenomena had we restricted our
+attention to \textsf{MAC} or \textsf{WEAT} scores only. To be able to
+diagnose and remove biases at the right level of granularity, we need to
+go beyond single metric chasing.
 
@@
 
-\hypertarget{related-work-and-conclusions}{%
-\section{Related work and
-conclusions}\label{related-work-and-conclusions}}
+\hypertarget{loose-ends-and-conclusions}{%
+\section{Loose ends and conclusions}\label{loose-ends-and-conclusions}}
 
@@
 
-It has been suggested {[}2{]} to use Bernstein bounds to state the
-uncertainty of the sample bias in a form of confidence interval. The
-interpretation of this interval is as follows, 95\% confidence interval
-contains such a range of values, that one can be 95\% certain that it
-contains the population mean. As we will prove later in the text,
+{[}20{]} employ Bayesian methods to estimate uncertainty in NLP tasks,
+but they apply their Bayesian Neural Networks-based method to sentiment
+analysis or named entity recognition, not to bias. {[}2{]} correctly
+argues that bias estimate should not be expressed as a single number
+without taking into account that the estimate is made using a sample of
+data and therefore has intrinsic uncertainty. The authors suggest to use
+Bernstein bounds to gauge the uncertainty in terms of confidence
+intervals. We do not discuss this approach extensively, as we think that
 confidence intervals are quite problematic for several reasons, among
-others the confusing interpretability. In {[}2{]} the current problem of
-bias estimation was correctly identified. Indeed, the bias estimate
-should not be expressed as a single number without taking into account
-that the estimate is made on a sample of data and therefore has
-intrinsic uncertainty. However, we argue that Bernstein bounds do not
-provide the best solution to this problem. Applying their method to a
-popular WinoBias dataset leads to the conclusion that more than 11903
+others the confusing interpretability. We do not think that Bernstein
+bounds provide the best solution to the problem. Applying this method to
+a popular WinoBias dataset leads to the conclusion that more than 11903
 samples are needed to claim a 95\% confidence interval for a bias
-estimate. This amount exceeds existing datasets for bias measurement. We
-propose a more realistic method that can be applied to commonly used
-datasets to estimate the uncertainty of the bias measurement.
+estimate. This amount vastly exceeds the existing word lists for bias
+estimation. We propose a more realistic Bayesian method. Our conclusion
+is still that the word lists are sometimes too small, but at least
+allows for gauging uncertainty as we go on to improve our methodology
+and extend the lists gradually.
+
+{[}19{]} criticize some exisisting bias metrics such as \textsf{MAC} or
+\textsf{WEAT} on the grounds of them not satisfying some general formal
+principles, such as magnitude-comparability, and they propose a
+modification, called \textsf{SAME}. {[}13{]} develop a generalization of
+\textsf{WEAT} meant to apply to sets of sentences, \textsf{SEAT}, which
+basically applies \textsf{WEAT} to vector representations of sentences.
+The authors, however, still pre-average and play the game of finding a
+single-number metric, so our remarks apply.
 
@@
 
+The short general lesson here is this: it's complicated. Instead of
+pretending it isn't and chasing single-number metrics, we should rather
+devote attention to more nuanced and methodologically sound analysis.
+
 \newpage
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-may-etal-2019-measuring}{}}%
+\CSLLeftMargin{{[}13{]} }%
+\CSLRightInline{Chandler May, Alex Wang, Shikha Bordia, Samuel R.
+Bowman, and Rachel Rudinger. 2019. On measuring social biases in
+sentence encoders. In \emph{Proceedings of the 2019 conference of the
+north {A}merican chapter of the association for computational
+linguistics: Human language technologies, volume 1 (long and short
+papers)}, Association for Computational Linguistics, Minneapolis,
+Minnesota, 622--628.
+DOI:https://doi.org/\href{https://doi.org/10.18653/v1/N19-1063}{10.18653/v1/N19-1063}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-statrethinkingbook2020}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{Richard McElreath. 2020. \emph{Statistical rethinking: A
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Mikolov2013efficient}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{Tomas Mikolov, Kai Chen, Greg Corrado, and Jeffrey Dean.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Morey2015confidenceFallacy}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Richard Morey, Rink Hoekstra, Jeffrey Rouder, Michael
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nissim2020fair}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Malvina Nissim, Rik van Noord, and Rob van der Goot.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nosek2002harvesting}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Brian A. Nosek, Mahzarin R. Banaji, and Anthony G.
@@
 DOI:https://doi.org/\href{https://doi.org/10.1037/1089-2699.6.1.101}{10.1037/1089-2699.6.1.101}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schruxf6der2021evaluating}{}}%
+\CSLLeftMargin{{[}19{]} }%
+\CSLRightInline{Sarah Schröder, Alexander Schulz, Philip Kenneweg,
+Robert Feldhans, Fabian Hinder, and Barbara Hammer. 2021. Evaluating
+metrics for bias in word embeddings. Retrieved from
+\url{https://arxiv.org/abs/2111.07864}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-DBLP:journalsux2fcorrux2fabs-1811-07253}{}}%
+\CSLLeftMargin{{[}20{]} }%
+\CSLRightInline{Yijun Xiao and William Yang Wang. 2018. Quantifying
+uncertainties in natural language processing tasks. \emph{CoRR}
+abs/1811.07253, (2018). Retrieved from
+\url{http://arxiv.org/abs/1811.07253}}
 
@@
 \hypertarget{custom-lists-used-in-this-paper}{%
-\subsubsection{Custom lists used in this
-paper}\label{custom-lists-used-in-this-paper}}
+\subsubsection{\texorpdfstring{Custom lists used in this paper
+\label{app:custom}}{Custom lists used in this paper }}\label{custom-lists-used-in-this-paper}}
 

### 2023-06-15 paper/arxivSubmission/BayesBias.tex
paper sent to arxiv

--- 2023-06-01 paper/paperDraft.tex
@@
 \hypersetup{
-  pdftitle={Taking uncertainty in word embedding bias estimation seriously: a bayesian approach},
+  pdftitle={A Bayesian approach to uncertainty in word embedding bias estimation},
   pdfauthor={Alicja Dobrzeniecka and Rafal Urbaniak},
@@
 
-\title{Taking uncertainty in word embedding bias estimation seriously: a
-bayesian approach}
+\title{A Bayesian approach to uncertainty in word embedding bias
+estimation}
 \author{Alicja Dobrzeniecka and Rafal Urbaniak}
@@
 
+\textbf{Abstract.} Multiple measures, such as WEAT or MAC, attempt to
+quantify the magnitude of bias present in word embeddings in terms of a
+single-number metric. However, such metrics and the related statistical
+significance calculations rely on treating pre-averaged data as
+individual data points and employing bootstrapping techniques with low
+sample sizes. We show that similar results can be easily obtained using
+such methods even if the data are generated by a null model lacking the
+intended bias. Consequently, we argue that this approach generates false
+confidence. To address this issue, we propose a Bayesian alternative:
+hierarchical Bayesian modeling, which enables a more
+uncertainty-sensitive inspection of bias in word embeddings at different
+levels of granularity. To showcase our method, we apply it to
+\emph{Religion, Gender}, and \emph{Race} word lists from the original
+research, together with our control neutral word lists. We deploy the
+method using \emph{Google, Glove}, and \emph{Reddit} embeddings.
+Further, we utilize our approach to evaluate a debiasing technique
+applied to the \emph{Reddit} word embedding. Our findings reveal a more
+complex landscape than suggested by the proponents of single-number
+metrics. The datasets and source code for the paper are publicly
+available.\footnote{\url{https://github.com/efemeryds/Bayesian-analysis-for-NLP-bias}}
+
 \hypertarget{introduction}{%
@@
 
-It has been suggested\footnote{See for instance {[}1,4,5,8,11,12{]}.}
+It has been suggested\footnote{See for instance {[}1,4,5,9,12,13{]}.}
 that language models can learn implicit biases that reflect harmful
@@
 downstream tasks. Roughly speaking, cosine similarity is an imperfect
-mathematical proxy for semantic similarity {[}15{]}.
+mathematical proxy for semantic similarity {[}16{]}.
 
@@
 line with the arguments to the effect that learning algorithms are
-always value-laden {[}9{]}: they employ inductive methods that require
+always value-laden {[}10{]}: they employ inductive methods that require
 design-, data-, or risk-related decisions that have to be guided by
@@
 \emph{Reddit Corpus}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
-\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}8{]}, and
-\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}12{]}. We first
+\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}9{]}, and
+\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}13{]}. We first
 explain how these measures are supposed to work. Then we argue that they
@@
 sample sizes used in the studies are usually small,\footnote{Depending
-  on a list for {[}8{]} the range for protected words is between 13 and
-  100, and for attributes between 16 and 25; for {[}12{]} the range for
+  on a list for {[}9{]} the range for protected words is between 13 and
+  100, and for attributes between 16 and 25; for {[}13{]} the range for
   protected words is between 14 and 18, and for attributes between 11
@@
   do not discuss this method in this paper, see its criticism in
-  {[}17{]}.}
+  {[}18{]}.}
 
@@
 Embedding Association Test (\textsf{WEAT}), has been proposed in
-{[}8{]}. The idea here is that the bias between two sets of target
+{[}9{]}. The idea here is that the bias between two sets of target
 words, \(X\) and \(Y\) (we call them protected words), should be
@@
 \textsf{WEAT} is inspired by the Implicit Association Test (IAT)
-{[}18{]} used in psychology, and in some applications it uses almost the
+{[}19{]} used in psychology, and in some applications it uses almost the
 same word sets, allowing for a \emph{prima facie} sensible comparison
-with bias in humans. In {[}8{]} the authors argue that significant
+with bias in humans. In {[}9{]} the authors argue that significant
 biases---thus measured--- similar to the ones discovered by IAT can be
-discovered in word embeddings. In {[}11{]} the methodology is extended
+discovered in word embeddings. In {[}12{]} the methodology is extended
 to a multilingual and cross-lingual setting, arguing that using
@@
 to more than two stereotypical groups are to be measured. Such a
-generalization can be found in {[}12{]}. The authors introduce Mean
+generalization can be found in {[}13{]}. The authors introduce Mean
 Average Cosine distance (\textsf{MAC}) as a measure of bias. Let
@@
 debiasing methods to the Reddit embedding, where the score is calculated
-using the Religion word lists from {[}12{]}. For our purpose the details
+using the Religion word lists from {[}13{]}. For our purpose the details
 of the debiasing method are not important: what matters is that
@@
 case of religion-related stereotypes to illustrate. The full lists from
-{[}12{]} can be found in Appendix \ref{appendix:manzini_word_lists}. In
+{[}13{]} can be found in Appendix \ref{appendix:manzini_word_lists}. In
 the original paper, words from all three religions were compared against
@@
 
-\noindent The \textsf{WEAT} scores reported by {[}8{]} for lists of
+\noindent The \textsf{WEAT} scores reported by {[}9{]} for lists of
 words for which the embeddings are supposedly biased range from 2.06 to
@@
   sample size used in the WEAT7 word list, which is not much different
-  from the other sample sizes in word lists used by {[}8{]}).} Each such
+  from the other sample sizes in word lists used by {[}9{]}).} Each such
 a collection of samples, as far as our question is involved, is
@@
 attributes, and \(8\) distances from hypothetical \(B\) attributes.
-Calculate the test statistic and effect size the way {[}8{]} did. Do
+Calculate the test statistic and effect size the way {[}9{]} did. Do
 this 10000 times, each time calculating \textsf{WEAT} and \textsf{s}
@@
 section we discussed the performance of the \textsf{WEAT} measure, but
-since the {[}12{]} one is a generalization thereof, including the method
+since the {[}13{]} one is a generalization thereof, including the method
 of running statistical tests on pre-averaged data, our remarks,
@@
   procedure is performed a fixed amount of times. For a nice survey and
-  explanation of these misinterpretations, see {[}16{]}. For a
+  explanation of these misinterpretations, see {[}17{]}. For a
   psychological study of the occurrence of such misinterpretations, see
-  REF. In this study, 120 researchers and 442 students were asked to
+  {[}8{]}. In this study, 120 researchers and 442 students were asked to
   assess the truth value of six false statements involving different
   interpretations of a CI. Both researchers and students endorsed, on
-  average, more than three of these statements. See also {[}7{]} for a
-  list of problems with the classical approach.}
+  average, more than three of these statements.}
 
@@
 words to attributes associated with them.\footnote{For a thorough
-  introduction to the concepts we're using, see {[}10,14{]}.}
+  introduction to the concepts we're using, see {[}11,15{]}.}
 \begin{align*}
@@
 
-\hypertarget{loose-ends-and-conclusions}{%
-\section{Loose ends and conclusions}\label{loose-ends-and-conclusions}}
+\hypertarget{related-works-and-conclusions}{%
+\section{Related works and
+conclusions}\label{related-works-and-conclusions}}
 
@@
 
-{[}20{]} employ Bayesian methods to estimate uncertainty in NLP tasks,
-but they apply their Bayesian Neural Networks-based method to sentiment
-analysis or named entity recognition, not to bias. {[}2{]} correctly
-argues that bias estimate should not be expressed as a single number
-without taking into account that the estimate is made using a sample of
-data and therefore has intrinsic uncertainty. The authors suggest to use
-Bernstein bounds to gauge the uncertainty in terms of confidence
-intervals. We do not discuss this approach extensively, as we think that
-confidence intervals are quite problematic for several reasons, among
-others the confusing interpretability. We do not think that Bernstein
-bounds provide the best solution to the problem. Applying this method to
-a popular WinoBias dataset leads to the conclusion that more than 11903
-samples are needed to claim a 95\% confidence interval for a bias
-estimate. This amount vastly exceeds the existing word lists for bias
-estimation. We propose a more realistic Bayesian method. Our conclusion
-is still that the word lists are sometimes too small, but at least
-allows for gauging uncertainty as we go on to improve our methodology
-and extend the lists gradually.
-
-{[}19{]} criticize some exisisting bias metrics such as \textsf{MAC} or
-\textsf{WEAT} on the grounds of them not satisfying some general formal
-principles, such as magnitude-comparability, and they propose a
-modification, called \textsf{SAME}. {[}13{]} develop a generalization of
-\textsf{WEAT} meant to apply to sets of sentences, \textsf{SEAT}, which
-basically applies \textsf{WEAT} to vector representations of sentences.
-The authors, however, still pre-average and play the game of finding a
+There are a few related papers, whose discussion goes beyond the scope
+of this paper:
+
+\begin{itemize}
+\item
+  {[}21{]} employ Bayesian methods to estimate uncertainty in NLP tasks,
+  but they apply their Bayesian Neural Networks-based method to
+  sentiment analysis or named entity recognition, not to bias.
+\item
+  {[}2{]} correctly argues that a bias estimate should not be expressed
+  as a single number without taking into account that the estimate is
+  made using a sample of data and therefore has intrinsic uncertainty.
+  The authors suggest using Bernstein bounds to gauge the uncertainty in
+  terms of confidence intervals. We do not discuss this approach
+  extensively, as we think that confidence intervals are quite
+  problematic for several reasons, among others the confusing
+  interpretability. We do not think that Bernstein bounds provide the
+  best solution to the problem. Applying this method to a popular
+  WinoBias dataset leads to the conclusion that more than 11903 samples
+  are needed to claim a 95\% confidence interval for a bias estimate.
+  This amount vastly exceeds the existing word lists for bias
+  estimation. We propose a more realistic Bayesian method. Our
+  conclusion is still that the word lists are sometimes too small, but
+  at least they allow for gauging uncertainty as we go on to improve our
+  methodology and extend the lists gradually.
+\item
+  {[}20{]} criticize some existing bias metrics such as \textsf{MAC} or
+  \textsf{WEAT} on the grounds of them not satisfying some general
+  formal principles, such as magnitude-comparability, and they propose a
+  modification, called \textsf{SAME}.
+\item
+  {[}14{]} develop a generalization of \textsf{WEAT} meant to apply to
+  sets of sentences, \textsf{SEAT}, which basically applies
+  \textsf{WEAT} to vector representations of sentences. The authors,
+  however, still pre-average and play the game of finding a
   single-number metric, so our remarks apply.
-
-\todo{More references to similar papers}
-
-A Bayesian data analysis with hierarchical models of cosine distances
-between protected words, control group words, and stereotypical
-attributes provides more modest and realistic assessment of the
-uncertainty involved. It reveals that much complexity is hidden when one
-instead chases single bias metrics present in the literature.
-
-After introducing the method, we apply it to multiple word embeddings
+\item
+  {[}7{]} introduce the Contextualized Embedding Association Test
+  Intersectional, meant to apply to dynamic word embeddings and,
+  importantly develop methods for intersectional bias detection. The
+  measure is a generalization of the WEAT method. The authors do inspect
+  a distribution of effect sizes that arises from the consideration of
+  various possible contexts, but they continue to standardize the
+  difference in averaged means and use a single-number summary: the
+  weighted mean of the effect sizes thus understood. The method,
+  admittedly, deserves further evaluation which goes beyond the scope of
+  this paper.
+\end{itemize}
+
+To summarize, a Bayesian data analysis with hierarchical models of
+cosine distances between protected words, control group words, and
+stereotypical attributes provides more modest and realistic assessment
+of the uncertainty involved. It reveals that much complexity is hidden
+when one instead chases single bias metrics present in the literature.
+
+After introducing the method, we applied it to multiple word embeddings
 and results of supposed debiasing, putting forward some general
@@
 word list sizes and sample sizes used in the studies are usually small.
-posterior density intervals are fairly wide, often the differences
+Posterior density intervals are fairly wide. Often the differences
 between associated, different and neutral human predicates, are not very
@@
 
-The short general lesson here is this: it's complicated. Instead of
-pretending it isn't and chasing single-number metrics, we should rather
-devote attention to more nuanced and methodologically sound analysis.
+The short general and somewhat disappointing lesson here is this: things
+are complicated. Instead of chasing single-number metrics, we should
+rather devote attention to more nuanced analysis.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Guo2021CEAT}{}}%
+\CSLLeftMargin{{[}7{]} }%
+\CSLRightInline{Wei Guo and Aylin Caliskan. 2021. Detecting emergent
+intersectional biases: Contextualized word embeddings contain a
+distribution of human-like biases. In \emph{Proceedings of the 2021
+{AAAI}/{ACM} conference on {AI}, ethics, and society}, {ACM}.
+DOI:https://doi.org/\href{https://doi.org/10.1145/3461702.3462536}{10.1145/3461702.3462536}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Hoekstra2014Misinterpretation}{}}%
-\CSLLeftMargin{{[}7{]} }%
+\CSLLeftMargin{{[}8{]} }%
 \CSLRightInline{Rink Hoekstra, Richard D. Morey, Jeffrey N. Rouder, and
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Caliskan2017semanticsBiases}{}}%
-\CSLLeftMargin{{[}8{]} }%
+\CSLLeftMargin{{[}9{]} }%
 \CSLRightInline{Aylin Caliskan Islam, Joanna J. Bryson, and Arvind
@@
 \leavevmode\vadjust pre{\hypertarget{ref-JohnsonValueFree}{}}%
-\CSLLeftMargin{{[}9{]} }%
+\CSLLeftMargin{{[}10{]} }%
 \CSLRightInline{Gabbrielle Johnson. forthcoming. Are algorithms
@@
 \leavevmode\vadjust pre{\hypertarget{ref-kruschke2015bayesian}{}}%
-\CSLLeftMargin{{[}10{]} }%
+\CSLLeftMargin{{[}11{]} }%
 \CSLRightInline{John Kruschke. 2015. \emph{Doing bayesian data analysis
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Lauscher2019multidimensional}{}}%
-\CSLLeftMargin{{[}11{]} }%
+\CSLLeftMargin{{[}12{]} }%
 \CSLRightInline{Anne Lauscher and Goran Glavas. 2019. Are we
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Manzini2019blackToCriminal}{}}%
-\CSLLeftMargin{{[}12{]} }%
+\CSLLeftMargin{{[}13{]} }%
 \CSLRightInline{Thomas Manzini, Yao Chong Lim, Yulia Tsvetkov, and Alan
@@
 \leavevmode\vadjust pre{\hypertarget{ref-may-etal-2019-measuring}{}}%
-\CSLLeftMargin{{[}13{]} }%
+\CSLLeftMargin{{[}14{]} }%
 \CSLRightInline{Chandler May, Alex Wang, Shikha Bordia, Samuel R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-statrethinkingbook2020}{}}%
-\CSLLeftMargin{{[}14{]} }%
+\CSLLeftMargin{{[}15{]} }%
 \CSLRightInline{Richard McElreath. 2020. \emph{Statistical rethinking: A
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Mikolov2013efficient}{}}%
-\CSLLeftMargin{{[}15{]} }%
+\CSLLeftMargin{{[}16{]} }%
 \CSLRightInline{Tomas Mikolov, Kai Chen, Greg Corrado, and Jeffrey Dean.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Morey2015confidenceFallacy}{}}%
-\CSLLeftMargin{{[}16{]} }%
+\CSLLeftMargin{{[}17{]} }%
 \CSLRightInline{Richard Morey, Rink Hoekstra, Jeffrey Rouder, Michael
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nissim2020fair}{}}%
-\CSLLeftMargin{{[}17{]} }%
+\CSLLeftMargin{{[}18{]} }%
 \CSLRightInline{Malvina Nissim, Rik van Noord, and Rob van der Goot.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nosek2002harvesting}{}}%
-\CSLLeftMargin{{[}18{]} }%
+\CSLLeftMargin{{[}19{]} }%
 \CSLRightInline{Brian A. Nosek, Mahzarin R. Banaji, and Anthony G.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-schruxf6der2021evaluating}{}}%
-\CSLLeftMargin{{[}19{]} }%
+\CSLLeftMargin{{[}20{]} }%
 \CSLRightInline{Sarah Schröder, Alexander Schulz, Philip Kenneweg,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-DBLP:journalsux2fcorrux2fabs-1811-07253}{}}%
-\CSLLeftMargin{{[}20{]} }%
+\CSLLeftMargin{{[}21{]} }%
 \CSLRightInline{Yijun Xiao and William Yang Wang. 2018. Quantifying
@@
 
-\label{appendix:manzini_word_lists} The lists from {[}12{]} are
+\label{appendix:manzini_word_lists} The lists from {[}13{]} are
 available here:

### 2023-06-15 paper/paperDraft.Rmd
paper sent to arxiv

@@
 ---
-title: 'Taking uncertainty in word embedding bias estimation seriously: a  bayesian
-  approach'
+title: 'A Bayesian approach to  uncertainty in word embedding bias estimation'
 author: "Alicja Dobrzeniecka and Rafal Urbaniak"
@@
 ```
+
+\textbf{Abstract.} Multiple measures, such as WEAT or MAC, attempt to quantify the magnitude of  bias present in  word embeddings  in  terms of a single-number  metric.
+However, such  metrics and the related statistical significance calculations rely on treating pre-averaged data as individual  data points and employing bootstrapping techniques with low sample sizes.
+We show that similar results can be easily obtained using such methods even if the data are generated by a null model lacking the intended bias.
+Consequently, we argue that this approach generates false confidence.
+To address this issue, we propose a Bayesian alternative: hierarchical Bayesian modeling, which enables a  more uncertainty-sensitive inspection of bias in  word embeddings at different levels of granularity.
+To showcase our method, we apply it to  \emph{Religion, Gender}, and \emph{Race} word lists from the original research,  together with our control neutral word lists.
+We   deploy the method using  \emph{Google, Glove}, and \emph{Reddit} embeddings.
+Further, we utilize our approach to evaluate a debiasing technique applied to the \emph{Reddit} word embedding.
+Our findings reveal a more complex landscape than suggested by the proponents of single-number metrics.
+The datasets and source code for the paper are publicly available.^[\url{https://github.com/efemeryds/Bayesian-analysis-for-NLP-bias}]
 
@@
 For a nice survey and explanation of these misinterpretations, see  [@Morey2015confidenceFallacy].
-For a psychological study of the occurrence of such  misinterpretations, see REF.
+For a psychological study of the occurrence of such  misinterpretations, see [@Hoekstra2014Misinterpretation].
 In this study, 120 researchers and 442 students were asked to assess the truth value of six false statements involving different interpretations of a CI.
-Both researchers and students endorsed, on average, more than three of these statements.
-See also [@Hoekstra2014Misinterpretation] for a list of problems with the classical approach.]
+Both researchers and students endorsed, on average, more than three of these statements.]
 
@@
 
-# Loose ends and conclusions
+# Related works and conclusions
 \label{sec:related}
 
-[@DBLP:journals/corr/abs-1811-07253] employ Bayesian methods to estimate uncertainty in NLP tasks, but they apply their Bayesian Neural Networks-based method to sentiment analysis or named entity recognition, not to bias.
-[@Ethayarajh2020Bernstein] correctly argues that bias estimate should not be expressed as a single number without taking into account that the estimate is made using a sample of data and therefore has intrinsic uncertainty.
-The authors   suggest to use Bernstein bounds to gauge  the uncertainty in terms of  confidence intervals.
+There are a few related papers, whose discussion goes beyond the scope of this paper:
+
+-  [@DBLP:journals/corr/abs-1811-07253] employ Bayesian methods to estimate uncertainty in NLP tasks, but they apply their Bayesian Neural Networks-based method to sentiment analysis or named entity recognition, not to bias.
+
+- [@Ethayarajh2020Bernstein] correctly argues that a bias estimate should not be expressed as a single number without taking into account that the estimate is made using a sample of data and therefore has intrinsic uncertainty.
+The authors   suggest using Bernstein bounds to gauge  the uncertainty in terms of  confidence intervals.
 We do not discuss this approach extensively, as we think that confidence intervals are quite problematic for several reasons, among others the confusing interpretability.
@@
 We propose a more realistic Bayesian method.
-Our conclusion is still that the word lists are sometimes too small, but at least allows for gauging uncertainty as we go on to improve our methodology and  extend the lists gradually.
-
- [@schröder2021evaluating]  criticize some exisisting bias metrics such as \textsf{MAC} or \textsf{WEAT} on the grounds of them not satisfying some general formal principles, such as magnitude-comparability, and they propose a modification, called \textsf{SAME}.
-[@may-etal-2019-measuring] develop a generalization of \textsf{WEAT} meant to apply to sets of sentences, \textsf{SEAT}, which basically applies \textsf{WEAT} to vector representations of sentences.
+Our conclusion is still that the word lists are sometimes too small, but at least they allow for gauging uncertainty as we go on to improve our methodology and  extend the lists gradually.
+
+-  [@schröder2021evaluating]  criticize some existing bias metrics such as \textsf{MAC} or \textsf{WEAT} on the grounds of them not satisfying some general formal principles, such as magnitude-comparability, and they propose a modification, called \textsf{SAME}.
+
+- [@may-etal-2019-measuring] develop a generalization of \textsf{WEAT} meant to apply to sets of sentences, \textsf{SEAT}, which basically applies \textsf{WEAT} to vector representations of sentences.
 The authors, however, still pre-average and play the game of finding a single-number metric, so our remarks  apply.
 
-\todo{More references to similar papers}
-
-A Bayesian data analysis with hierarchical models of cosine distances between protected words, control group words, and stereotypical attributes  provides more modest and realistic assessment of the uncertainty involved.
+- [@Guo2021CEAT] introduce the Contextualized Embedding Association Test Intersectional, meant to apply to dynamic word embeddings and, importantly develop methods for intersectional bias detection.
+The measure is a generalization of the WEAT method.
+The authors do inspect a distribution of effect sizes that arises from the consideration of various possible contexts, but they continue to standardize the difference in averaged means and  use a single-number summary: the weighted mean of the effect sizes thus understood.
+The method, admittedly, deserves further evaluation which goes beyond the scope of this paper.
+
+To summarize, a Bayesian data analysis with hierarchical models of cosine distances between protected words, control group words, and stereotypical attributes  provides more modest and realistic assessment of the uncertainty involved.
 It reveals that much complexity is hidden when one instead chases single bias metrics present in the literature.
 
-After introducing the method, we apply it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of \textsf{WEAT} or \textsf{MAC} (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small.
-posterior density intervals are fairly wide,   often the differences between associated, different and neutral human predicates, are not very impressive.
+After introducing the method, we applied it to multiple word embeddings and results of supposed debiasing, putting forward some general observations that are not exactly in line with the usual picture painted in terms of \textsf{WEAT} or \textsf{MAC} (and the problem generalizes to any approach that focuses on chasing a single numeric metric): the word list sizes and sample sizes used in the studies are usually small.
+Posterior density intervals are fairly wide.
+Often the differences between associated, different and neutral human predicates, are not very impressive.
 Also, a preliminary inspection suggests that  the desirability of changes obtained by the usual debiasing methods is debatable.
@@
 
-The short general lesson here is this: it's complicated.
-Instead of pretending it isn't and chasing single-number metrics, we should rather devote attention to more nuanced and methodologically sound analysis.
+The short general and somewhat disappointing lesson here is this: things are complicated. Instead of chasing single-number metrics, we should rather devote attention to more nuanced analysis.
 

### 2023-06-15 paper/paperDraft.tex
paper sent to arxiv

@@
 
-It has been suggested\footnote{See for instance {[}1,4,5,8,11,12{]}.}
+It has been suggested\footnote{See for instance {[}1,2,5,6,12,13{]}.}
 that language models can learn implicit biases that reflect harmful
@@
 do complete justice to learning it in the model used to study it. One
-example of this approach is {[}3{]}, where the authors use language
+example of this approach is {[}4{]}, where the authors use language
 models to study the shape of certain biases across a century.
@@
 a corpus actually greatly exaggerates the actual lower disproportion in
-the real distribution of occupations {[}6{]}.
+the real distribution of occupations {[}7{]}.
 
@@
 \emph{Reddit Corpus}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
-\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}8{]}, and
-\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}12{]}. We first
+\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}2{]}, and
+\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}13{]}. We first
 explain how these measures are supposed to work. Then we argue that they
@@
 sample sizes used in the studies are usually small,\footnote{Depending
-  on a list for {[}8{]} the range for protected words is between 13 and
-  100, and for attributes between 16 and 25; for {[}12{]} the range for
+  on a list for {[}2{]} the range for protected words is between 13 and
+  100, and for attributes between 16 and 25; for {[}13{]} the range for
   protected words is between 14 and 18, and for attributes between 11
@@
 complaining'' or ``macho''.\footnote{However, for some research-related
-  purposes, such as the study of stereotypes across history {[}3{]},
+  purposes, such as the study of stereotypes across history {[}4{]},
   embeddings which do not protect certain classes may also be useful.}
@@
 The use of projections in bias estimation has been criticized for
-instance in {[}5{]}, where it is pointed out that while a higher average
+instance in {[}6{]}, where it is pointed out that while a higher average
 similarity to the gender direction might be an indicator of bias with
@@
 stereotypical attribute words (such as \emph{scientific} or
-\emph{caring}). Further, it is observed in {[}5{]} that most word pairs
+\emph{caring}). Further, it is observed in {[}6{]} that most word pairs
 retain similarity under debiasing meant to minimize projection-based
@@
 Embedding Association Test (\textsf{WEAT}), has been proposed in
-{[}8{]}. The idea here is that the bias between two sets of target
+{[}2{]}. The idea here is that the bias between two sets of target
 words, \(X\) and \(Y\) (we call them protected words), should be
@@
 \textsf{WEAT} is inspired by the Implicit Association Test (IAT)
-{[}18{]} used in psychology, and in some applications it uses almost the
+{[}19{]} used in psychology, and in some applications it uses almost the
 same word sets, allowing for a \emph{prima facie} sensible comparison
-with bias in humans. In {[}8{]} the authors argue that significant
+with bias in humans. In {[}2{]} the authors argue that significant
 biases---thus measured--- similar to the ones discovered by IAT can be
-discovered in word embeddings. In {[}11{]} the methodology is extended
+discovered in word embeddings. In {[}12{]} the methodology is extended
 to a multilingual and cross-lingual setting, arguing that using
@@
   embeddings being less biased than those based on Wikipedia.} A similar
-methodology is employed in {[}4{]}. The authors employ word embeddings
+methodology is employed in {[}5{]}. The authors employ word embeddings
 trained on corpora from different decades to study the shifts in various
@@
 
-\noindent The \textsf{WEAT} scores reported by {[}8{]} for lists of
+\noindent The \textsf{WEAT} scores reported by {[}2{]} for lists of
 words for which the embeddings are supposedly biased range from 2.06 to
@@
   sample size used in the WEAT7 word list, which is not much different
-  from the other sample sizes in word lists used by {[}8{]}).} Each such
+  from the other sample sizes in word lists used by {[}2{]}).} Each such
 a collection of samples, as far as our question is involved, is
@@
 attributes, and \(8\) distances from hypothetical \(B\) attributes.
-Calculate the test statistic and effect size the way {[}8{]} did. Do
+Calculate the test statistic and effect size the way {[}2{]} did. Do
 this 10000 times, each time calculating \textsf{WEAT} and \textsf{s}
@@
   procedure is performed a fixed amount of times. For a nice survey and
-  explanation of these misinterpretations, see {[}16{]}. For a
+  explanation of these misinterpretations, see {[}17{]}. For a
   psychological study of the occurrence of such misinterpretations, see
-  REF. In this study, 120 researchers and 442 students were asked to
+  {[}9{]}. In this study, 120 researchers and 442 students were asked to
   assess the truth value of six false statements involving different
   interpretations of a CI. Both researchers and students endorsed, on
-  average, more than three of these statements. See also {[}7{]} for a
-  list of problems with the classical approach.}
+  average, more than three of these statements.}
 
@@
 
-{[}20{]} employ Bayesian methods to estimate uncertainty in NLP tasks,
-but they apply their Bayesian Neural Networks-based method to sentiment
-analysis or named entity recognition, not to bias. {[}2{]} correctly
-argues that bias estimate should not be expressed as a single number
-without taking into account that the estimate is made using a sample of
-data and therefore has intrinsic uncertainty. The authors suggest to use
-Bernstein bounds to gauge the uncertainty in terms of confidence
-intervals. We do not discuss this approach extensively, as we think that
-confidence intervals are quite problematic for several reasons, among
-others the confusing interpretability. We do not think that Bernstein
-bounds provide the best solution to the problem. Applying this method to
-a popular WinoBias dataset leads to the conclusion that more than 11903
-samples are needed to claim a 95\% confidence interval for a bias
-estimate. This amount vastly exceeds the existing word lists for bias
-estimation. We propose a more realistic Bayesian method. Our conclusion
-is still that the word lists are sometimes too small, but at least
-allows for gauging uncertainty as we go on to improve our methodology
-and extend the lists gradually.
-
-{[}19{]} criticize some exisisting bias metrics such as \textsf{MAC} or
-\textsf{WEAT} on the grounds of them not satisfying some general formal
-principles, such as magnitude-comparability, and they propose a
-modification, called \textsf{SAME}. {[}13{]} develop a generalization of
-\textsf{WEAT} meant to apply to sets of sentences, \textsf{SEAT}, which
-basically applies \textsf{WEAT} to vector representations of sentences.
-The authors, however, still pre-average and play the game of finding a
+There are a few related papers, whose discussion goes beyond the scope
+of this paper:
+
+\begin{itemize}
+\item
+  {[}21{]} employ Bayesian methods to estimate uncertainty in NLP tasks,
+  but they apply their Bayesian Neural Networks-based method to
+  sentiment analysis or named entity recognition, not to bias.
+\item
+  {[}3{]} correctly argues that a bias estimate should not be expressed
+  as a single number without taking into account that the estimate is
+  made using a sample of data and therefore has intrinsic uncertainty.
+  The authors suggest using Bernstein bounds to gauge the uncertainty in
+  terms of confidence intervals. We do not discuss this approach
+  extensively, as we think that confidence intervals are quite
+  problematic for several reasons, among others the confusing
+  interpretability. We do not think that Bernstein bounds provide the
+  best solution to the problem. Applying this method to a popular
+  WinoBias dataset leads to the conclusion that more than 11903 samples
+  are needed to claim a 95\% confidence interval for a bias estimate.
+  This amount vastly exceeds the existing word lists for bias
+  estimation. We propose a more realistic Bayesian method. Our
+  conclusion is still that the word lists are sometimes too small, but
+  at least they allow for gauging uncertainty as we go on to improve our
+  methodology and extend the lists gradually.
+\item
+  {[}20{]} criticize some existing bias metrics such as \textsf{MAC} or
+  \textsf{WEAT} on the grounds of them not satisfying some general
+  formal principles, such as magnitude-comparability, and they propose a
+  modification, called \textsf{SAME}.
+\item
+  {[}14{]} develop a generalization of \textsf{WEAT} meant to apply to
+  sets of sentences, \textsf{SEAT}, which basically applies
+  \textsf{WEAT} to vector representations of sentences. The authors,
+  however, still pre-average and play the game of finding a
   single-number metric, so our remarks apply.
-
-\todo{More references to similar papers}
-
-A Bayesian data analysis with hierarchical models of cosine distances
-between protected words, control group words, and stereotypical
-attributes provides more modest and realistic assessment of the
-uncertainty involved. It reveals that much complexity is hidden when one
-instead chases single bias metrics present in the literature.
-
-After introducing the method, we apply it to multiple word embeddings
+\item
+  {[}8{]} introduce the Contextualized Embedding Association Test
+  Intersectional, meant to apply to dynamic word embeddings and,
+  importantly develop methods for intersectional bias detection. The
+  measure is a generalization of the WEAT method. The authors do inspect
+  a distribution of effect sizes that arises from the consideration of
+  various possible contexts, but they continue to standardize the
+  difference in averaged means and use a single-number summary: the
+  weighted mean of the effect sizes thus understood. The method,
+  admittedly, deserves further evaluation which goes beyond the scope of
+  this paper.
+\end{itemize}
+
+To summarize, a Bayesian data analysis with hierarchical models of
+cosine distances between protected words, control group words, and
+stereotypical attributes provides more modest and realistic assessment
+of the uncertainty involved. It reveals that much complexity is hidden
+when one instead chases single bias metrics present in the literature.
+
+After introducing the method, we applied it to multiple word embeddings
 and results of supposed debiasing, putting forward some general
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Caliskan2017semanticsBiases}{}}%
+\CSLLeftMargin{{[}2{]} }%
+\CSLRightInline{Aylin Caliskan, Joanna J. Bryson, and Arvind Narayanan.
+2016. Semantics derived automatically from language corpora necessarily
+contain human biases. \emph{CoRR} abs/1608.07187, (2016). Retrieved from
+\url{http://arxiv.org/abs/1608.07187}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Ethayarajh2020Bernstein}{}}%
-\CSLLeftMargin{{[}2{]} }%
+\CSLLeftMargin{{[}3{]} }%
 \CSLRightInline{Kawin Ethayarajh. 2020. Is your classifier actually
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Garg2017hundredYears}{}}%
-\CSLLeftMargin{{[}3{]} }%
+\CSLLeftMargin{{[}4{]} }%
 \CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Garg2018years}{}}%
-\CSLLeftMargin{{[}4{]} }%
+\CSLLeftMargin{{[}5{]} }%
 \CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Gonen2019lipstick}{}}%
-\CSLLeftMargin{{[}5{]} }%
+\CSLLeftMargin{{[}6{]} }%
 \CSLRightInline{Hila Gonen and Yoav Goldberg. 2019. Lipstick on a pig:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gordon2012reporting}{}}%
-\CSLLeftMargin{{[}6{]} }%
+\CSLLeftMargin{{[}7{]} }%
 \CSLRightInline{Jonathan Gordon and Benjamin Durme. 2013. Reporting bias
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Guo2021CEAT}{}}%
+\CSLLeftMargin{{[}8{]} }%
+\CSLRightInline{Wei Guo and Aylin Caliskan. 2021. Detecting emergent
+intersectional biases: Contextualized word embeddings contain a
+distribution of human-like biases. In \emph{Proceedings of the 2021
+{AAAI}/{ACM} conference on {AI}, ethics, and society}, {ACM}.
+DOI:https://doi.org/\href{https://doi.org/10.1145/3461702.3462536}{10.1145/3461702.3462536}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Hoekstra2014Misinterpretation}{}}%
-\CSLLeftMargin{{[}7{]} }%
+\CSLLeftMargin{{[}9{]} }%
 \CSLRightInline{Rink Hoekstra, Richard D. Morey, Jeffrey N. Rouder, and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Caliskan2017semanticsBiases}{}}%
-\CSLLeftMargin{{[}8{]} }%
-\CSLRightInline{Aylin Caliskan Islam, Joanna J. Bryson, and Arvind
-Narayanan. 2016. Semantics derived automatically from language corpora
-necessarily contain human biases. \emph{CoRR} abs/1608.07187, (2016).
-Retrieved from \url{http://arxiv.org/abs/1608.07187}}
-
 \leavevmode\vadjust pre{\hypertarget{ref-JohnsonValueFree}{}}%
-\CSLLeftMargin{{[}9{]} }%
+\CSLLeftMargin{{[}10{]} }%
 \CSLRightInline{Gabbrielle Johnson. forthcoming. Are algorithms
@@
 \leavevmode\vadjust pre{\hypertarget{ref-kruschke2015bayesian}{}}%
-\CSLLeftMargin{{[}10{]} }%
-\CSLRightInline{John Kruschke. 2015. \emph{Doing bayesian data analysis
-(second edition)}. Academic Press, Boston.}
+\CSLLeftMargin{{[}11{]} }%
+\CSLRightInline{John Kruschke. 2015. \emph{Doing {B}ayesian data
+analysis (second edition)}. Academic Press, Boston.}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Lauscher2019multidimensional}{}}%
-\CSLLeftMargin{{[}11{]} }%
+\CSLLeftMargin{{[}12{]} }%
 \CSLRightInline{Anne Lauscher and Goran Glavas. 2019. Are we

### 2023-06-22 paper/CLsubmission/COLI-manual3.tex
CLsubmission

% if your latex compiler failed to compile, uncomment the command below:
% \RequirePackage[2020-02-02]{latexrelease}
\documentclass{clv3}

\usepackage{hyperref}
\usepackage{xcolor}
\definecolor{darkblue}{rgb}{0, 0, 0.5}
\hypersetup{colorlinks=true,citecolor=darkblue, linkcolor=darkblue, urlcolor=darkblue}

\bibliographystyle{compling}

% test compatibility with algorithmic.sty
%\usepackage{algorithmic}

\begin{document}
\issue{1}{1}{2016}

%Document Head
\dochead{CLV3 Class File Manual}

\runningtitle{How to Use CLV3 \LaTeX\ Class File}

\runningauthor{Odi\'e N. Gementera}

\pageonefooter{Action editor: \{action editor name\}. Submission received: DD Month YYYY; revised version received: DD Month YYYY; accepted for publication: DD Month YYYY.}

\title{How to Use CLV3 \LaTeX\ Class File}

\author{Odi\'e N. Gementera\thanks{PITC Building, Pascor Drive, Sto. Ni\~no, Para\~naque City, 1700 Philippines. E-mail: o.gementera@spi-bpo.com.}}
\affil{Publishing / SPi}

\author{Another Author\thanks{PITC Building}}
\affil{Publishing / SPi}

\author{And Another Author}
\affil{Publishing / SPi}

\author{And Yet Another}
\affil{Publishing / SPi}

\maketitle

\begin{abstract}
This article describes how to use the ``CLV3'' class file, developed by {\sf SPi}, 
to produce typeset papers based on Computer Linguistics typesetting specifications 
for submission to MIT. This document was produced using ``{\tt clv3}'' class file 
and will be modified whenever there is an update in the layout specifications.
\end{abstract}

\section{Introduction}

This document applies to version 3 of CL class file. Prior style files such as 
``{\tt cl.sty}'' and ``{\tt coli.sty}'' do not have all of the features 
described here. It is assumed that the user has a basic knowledge of \LaTeX\ 
typesetting commands.

\section{Class File Options}

There are several options that can be used to switch the mode of MIT2 from normal 
article to manuscript style, or to different layout styles. This is specified in 
the usual \LaTeX\ way by declaring:

\verb|\documentclass[bookreview,manuscript]{clv3}|

%\subsection{Default Option}

\begin{deflist}
\item[bookreview] Sets the article layout for Book Review.
\item[brief] Sets the article layout for Briefly Noted.
\item[discussion] Sets the article layout for Squibs and Discussions.
\item[pubrec] Sets the article layout for Publication Received.
\item[shortpaper] Sets the article layout for Short Paper.
\item[manuscript] Sets the baseline spacing to double space. This 
option can be used in combination with other options.
\end{deflist}

By not declaring any option in the \verb|\documentclass| command the class file
will automatically set to standard article layout.

\section{Title Page}

The title page is created using the standard \LaTeX\ command \verb|\maketitle|. 
Before this command is declared, the author must declare all the data which are 
to appear in the title area.\footnote{$\backslash$maketitle is the command to execute all the title page information.}

\subsection{Volume, Number and Year}

The command \verb|\issue{vv}{nn}{yyyy}| is used in declaring the volume, number
and year of the article. The first argument is for the volume, the second argument 
is for the issue number. Volume and Issue number will appear on the even page 
running head opposite the journal name. The third argument is for the Year which 
will appear in the copyright line at the bottom of the title page.

\subsection{Document Head}

Document head is produced with the command \verb|\dochead{Document Head}|. Doc head 
will output differently, or may not appear at all, depending on the option used in the 
documentclass.

\subsection{Paper Title}

The paper title is declared like: \verb|\title{Computer Linguistic Article}|
in the usual \LaTeX manner. Line breaks may be inserted with (\verb|\\|) to equalize 
the length of the title lines.

\subsection{Authors}
The name and related information for authors is declared with the \verb|\author{}| command.

The \verb|\thanks{}| command produces the ``first footnotes.''. \LaTeX\ \verb|\thanks| 
cannot accommodate multiple paragraphs, author will have to use a separate \verb|\thanks|
for each paragraph.

The \verb|\affil{}| command produces the author affiliations that appears right under 
the author's name.

\subsection{Running Headers}
The running heads are declared with the \verb|\runningtitle{Running Title}| for the 
journal name and \verb|\runningauthor{Author's Surname}| for author. These information 
will appear on the odd pages. For {\tt bookreview} option, odd page running head is 
automatically set to "Book Reviews". Even page running head is default to Computational 
Linguistics opposite volume and issue number.

\subsection{Action Editor and Dates of Submission, Revision, and Acceptance}

For regular papers and survey papers, the name of the action editor and the dates of submission, revision, and acceptance must appear in the footnote area of the title page, and they are declared with the following command:

\noindent \verb|\pageonefooter{Action editor: XXX. Submission received: XXX;| \newline \hspace*{1em} \verb|revised version received: XXX; accepted for publication: XXX}|. % don't actually separate it into two lines when using this command.

\noindent For squib papers, there is no action editor and only the three dates need to be included.

\section{Abstract}

Abstract is the first part of a paper after \verb|\maketitle|. Abstract text is 
placed within the abstract environment:

\begin{verbatim}
\begin{abstract}
This is the abstract text . . .
\end{abstract}
\end{verbatim}

\section{Section Headings}

Section headings are declared in the usual \LaTeX\ way via \verb|\section{}|, 
\verb|\subsection{}|, \verb|\subsubsection{}|, and \verb|\paragraph{}|. The 
first 3 levels of section head will have Arabic numbering separated 
by period. The \verb|\paragraph{}| section will have the title head in Italics 
and at the same line with the first line of succeeding paragraph.

\section{Citations}
Citations in parentheses are declared using the \verb|\cite{}|
command, and appear in the text as follows: 
This technique is widely used \cite{woods}.
The command \verb|\citep{}| (cite parenthetical) is a synonym of \verb|\cite{}|.

Citations used in the sentence are declared using the \verb|\namecite{}|
commands, and appear in the text as follows: 
\namecite{woods} first described this technique.
The command \verb|\citet{}| (cite textual) is a synonym of \verb|\namecite{}|.

This style file is designed to be used with the BibTeX
style file \verb|compling.bst|.  Include the command
\verb|\bibliographystyle{compling}| in your source file.

Citation commands are based on the \verb|natbib| package;
for details on options and further variants of the commands,
see the \verb|natbib| documentation.  In particular, options 
exist to add extra text and page numbers.  For example, 
\verb|\cite[cf.][ch.\ 1]{winograd}| yields: \cite[cf.][ch.\ 1]{winograd}.

The following examples illustrate how citations appear both in the text
and in the references section at the end of this document.
\begin{enumerate}
\item Article in journal: 
 \namecite{akmajian};
 \namecite{woods}.
\item Book: 
  \namecite{altenberg};
  \namecite{winograd}.
\item Article in edited collection/Chapter in book: 
  \namecite{cutler};
  \namecite{sgall};
  \namecite{jurafsky}.
\item Technical report:
  \namecite{appelt};
  \namecite{robinson}.
\item Thesis or dissertation: 
  \namecite{baart};
  \namecite{spaerckjones};
  \namecite{cahn}.
\item Unpublished item: 
  \namecite{ayers}.
\item Conference proceedings: 
  \namecite{benoit}.
\item Paper published in conference proceedings: 
  \namecite{krahmer};
  \namecite{Copestake2001}.
\end{enumerate}

\section{Definition with Head}

Definition with head is declared by using the environment:
\\
\begin{verbatim}
\begin{definition}
Definition text. . .
\end{definition}
\end{verbatim}

This environment will generate the word {\bf ``Definition 1''} in bold on separate
line. The sequence number is generated for every definition environment. Definition 
data will have no indention on the first line while succeeding lines will have hang 
indention.

\section{Lists}

The usual \LaTeX\ itemize, enumerate and definition list environments are used 
in CLV3 style.

To produce Numbered List use the environment:

\begin{verbatim}
\begin{enumerate}
\item First numbered list item
\item Second numbered list item
\item Third numbered list item
\end{enumerate}
\end{verbatim}

To produce Bulleted List use the environment:

\begin{verbatim}
\begin{itemize}
\item First bulleted list item
\item Second bulleted list item
\item Third bulleted list item
\end{itemize}
\end{verbatim}

To produce Definition List use the environment:

\begin{verbatim}
\begin{deflist}
\item[First]  Definition list item. . .
\item[Second] Definition list item. . .
\item[Third]  Definition list item. . .
\end{deflist}
\end{verbatim}

Additional list environment were also defined such as Unnumbered, Arabic and Alpha lists.

Unnumbered List is the list where item labels are not generated. To produce Unnumbered List use the environment:

\begin{verbatim}
\begin{unenumerate}
\item First list item
\item Second list item
\item Third list item
\end{unenumerate}
\end{verbatim}

To produce Arabic List use the environment:

\begin{verbatim}
\begin{arabiclist}
\item First arabic list item
\item Second arabic list item
\item Third arabic list item
\end{arabiclist}
\end{verbatim}

To produce Alpha List use the environment:

\begin{verbatim}
\begin{alphalist}
\item First alpha list item
\item Second alpha list item
\item Third alpha list item
\end{alphalist}
\end{verbatim}

All the list environments mentioned above can be nested with each other.

\subsection{Other List Types}

\subsubsection{Outline List or Example List}

\begin{verbatim}
\begin{exlist}
\item First outline list item. . .
\item Second outline list item. . .
\item Third outline list item. . .
\end{exlist}
\end{verbatim}

\subsubsection{Output Formula or Algorithm}

\begin{verbatim}
\begin{algorithm}
\item[Step 1] First item. . . 
\item[Step 2] Second item. . .
\end{algorithm}
\end{verbatim}

% test compatibility with algorithmic.sty
%\begin{algorithmic}
%\STATE i
%\end{algorithmic}

See sample on the {\tt COLI-template.pdf}.

\section{Word Formula or Displayed Text}

Word formula and displayed text are text that should be displayed in a 
separate line without indention. This are achieved by using the environment:

\begin{verbatim}
\begin{displaytext}
This is a sample of displayed text . . .
\end{displaytext}
\end{verbatim}

\section{Dialogue}

Dialogue text are presentation of people's conversation. These will be presented 
on a separate line where each dialogue starts with the name of speaker, followed by 
colon. Succeeding lines will be hang indented. To produce Dialogue use the environment:
\\
\begin{verbatim}
\begin{dialogue}
Speaker 1: dialogue. . .

Speaker 2: dialogue. . .
\end{dialogue}
\end{verbatim}

\noindent Please make sure to insert an empty line between dialogues.

\section{Extracts}

Extract text acts like quote, where left and right margins are indented.
To produce Extract use the environment:

\begin{verbatim}
\begin{extract}
This is an example of Extract text. . .
\end{extract}
\end{verbatim}

\noindent See sample on the {\tt COLI-template.pdf}.

\section{Theorem-like Environments}

There are several theorem-like environments defined in CLV3 class file. Theorem-like
environments generate the name of the theorem as label, and counter number in bold.

\subsection{Example}

To produce Example use the environment:

\begin{verbatim}
\begin{example}
This is Example text. . .
\end{example}
\end{verbatim}

\subsection{Lemma}

To produce Lemma use the environment:

\begin{verbatim}
\begin{lemma}
Lemma text. . .
\end{lemma}
\end{verbatim}

This produces the following output:
\begin{lemma}\label{lem}
Lemma text.
\end{lemma}
A small vertical space separates the end of the lemma
from the following text.

\subsection{Theorem}

To produce Theorem use the environment:

\begin{verbatim}
\begin{theorem}
Theorem text. . .
\end{theorem}
\end{verbatim}

This produces the following output:
\begin{theorem}\label{thm}
Theorem text.
\end{theorem}
\noindent
A small vertical space separates the end of the theorem
from the following text.

\subsection{Proof}

The proof environment produces a square at the end of the text. To produce Proof 
use the environment:

\begin{verbatim}
\begin{proof}
Proof text. . .
\end{proof}
\end{verbatim}

This produces the following output:
\begin{proof}\label{proof}
Proof text.
\end{proof}
A small vertical space separates the end of the lemma
from the following text.

\subsection{Unnumbered Theorem-like Environments}

There are also unnumbered version of some of the theorem-like environments. 
These are declared by using its asterisked version. Here are the three 
unnumbered version of theorem-like environments:

\begin{verbatim}
\begin{theorem*}
Unnumbered theorem text. . .
\end{theorem*}
\end{verbatim}

\section{Appendix}

Appendix is declared by issuing the command \verb|\appendix|. This will set 
the necessary labels to appendix's rule (i.e. (A.1) for equation number).

Sections inside Appendix are declared using \verb|\appendixsection{}|, which 
will produce {\bf Appendix A: Section Title} for first section.

Equation numbers are automatically set to (A.1), (A.2) and (A.3). Where the letters 
follow the current level of Appendix section. So equations on {\bf Appendix B} 
will have equation numbers as follow: (B.1), (B.2) and (B.3).

\section{Acknowledgments}

Acknowledgments are produce by using the environment:
\\
\begin{verbatim}
\begin{acknowledgments}
Acknowledgments text. . .
\end{acknowledgments}
\end{verbatim}

\section{Others}

Other items such as Equations, Figures, Tables and References are produced in
the standard \LaTeX\ typesetting.

\starttwocolumn
\bibliography{compling_style}

\end{document}

### 2023-06-22 paper/CLsubmission/bayesBiasCL.tex
CLsubmission

--- 2023-06-15 paper/arxivSubmission/BayesBias.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  12pt,
-  dvipsnames,enabledeprecatedfontcommands]{scrartcl}
+\documentclass{clv3}
+
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
+\usepackage{float}
+\usepackage{nicefrac}
+
 \usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-%\documentclass{article}
-
-% %packages
+
 \usepackage{booktabs}
@@
 
-\usepackage{graphicx}
-%\usepackage{longtable}
-\usepackage{supertabular, booktabs}
-\usepackage{ragged2e}
-\usepackage{etex}
-%\usepackage{yfonts}
-\usepackage{marvosym}
-\usepackage[notextcomp]{kpfonts}
-\usepackage{nicefrac}
-\newcommand*{\QED}{\hfill \footnotesize {\sc Q.e.d.}}
-\usepackage{floatrow}
-
-\usepackage[textsize=footnotesize]{todonotes}
-
-%\linespread{1.5}
-\newcommand{\indep}{\!\perp \!\!\! \perp\!}
-
-\setlength{\parindent}{5mm}
-\setlength{\parskip}{1pt}
-
-%language
-%\usepackage{times}
-\usepackage{mathptmx}
-\usepackage[scaled=0.86]{helvet}
-\usepackage{t1enc}
-%\usepackage[utf8x]{inputenc}
-%\usepackage[polish]{babel}
-%\usepackage{polski}
-
-%AMS
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{amsmath}
-\usepackage{mathtools}
-
-\usepackage{geometry}
- \geometry{a4paper,left=35mm,top=20mm,}
-
-%environments
-\newtheorem{fact}{Fact}
-
-%abbreviations
-\newcommand{\ra}{\rangle}
-\newcommand{\la}{\langle}
-\newcommand{\n}{\neg}
-\newcommand{\et}{\wedge}
-\newcommand{\jt}{\rightarrow}
-\newcommand{\ko}[1]{\forall  #1\,}
-\newcommand{\ro}{\leftrightarrow}
-\newcommand{\exi}[1]{\exists\, {_{#1}}}
-\newcommand{\pr}[1]{\ensuremath{\mathsf{P}(#1)}}
-\newcommand{\cost}{\mathsf{cost}}
-\newcommand{\benefit}{\mathsf{benefit}}
-\newcommand{\ut}{\mathsf{ut}}
-
-\newcommand{\odds}{\mathsf{Odds}}
-\newcommand{\ind}{\mathsf{Ind}}
-\newcommand{\nf}[2]{\nicefrac{#1\,}{#2}}
-\newcommand{\R}[1]{\texttt{#1}}
-\newcommand{\prr}[1]{\mbox{$\mathtt{P}_{prior}(#1)$}}
-\newcommand{\prp}[1]{\mbox{$\mathtt{P}_{posterior}(#1)$}}
-
-\newtheorem{q}{\color{blue}Question}
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}[fact]
-
-%technical intermezzo
-%---------------------
-
-\newcommand{\intermezzoa}{
-	\begin{minipage}[c]{13cm}
-	\begin{center}\rule{10cm}{0.4pt}
-
-	\tiny{\sc Optional Content Starts}
-
-	\vspace{-1mm}
-
-	\rule{10cm}{0.4pt}\end{center}
-	\end{minipage}\nopagebreak 
-	}
-
-\newcommand{\intermezzob}{\nopagebreak 
-	\begin{minipage}[c]{13cm}
-	\begin{center}\rule{10cm}{0.4pt}
-
-	\tiny{\sc Optional Content Ends}
-
-	\vspace{-1mm}
-
-	\rule{10cm}{0.4pt}\end{center}
-	\end{minipage}
-	}
-
-%--------------------
-
-\newtheorem*{reply*}{Reply}
-\usepackage{enumitem}
-\newcommand{\question}[1]{\begin{enumerate}[resume,leftmargin=0cm,labelsep=0cm,align=left]
-\item #1
-\end{enumerate}}
-
-\usepackage{float}
-
-% \usepackage{sectsty}
-% \sectionfont{\fontsize{12}{11}\selectfont}
-% \subsectionfont{\fontsize{11}{11}\selectfont}
-
-% \setbeamertemplate{blocks}[rounded][shadow=true]
-% \setbeamertemplate{itemize items}[ball]
-% \AtBeginPart{}
-% \AtBeginSection{}
-% \AtBeginSubsection{}
-% \AtBeginSubsubsection{}
-% \setlength{\emergencystretch}{0em}
-% \setlength{\parskip}{0pt}
-
-\usepackage[authoryear]{natbib}
-
-%\bibliographystyle{apalike}
-
-\usepackage{tikz}
-\usetikzlibrary{positioning,shapes,arrows}
-
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={A Bayesian approach to uncertainty in word embedding bias estimation},
-  pdfauthor={Alicja Dobrzeniecka and Rafal Urbaniak},
-  colorlinks=true,
-  linkcolor={Maroon},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={blue},
-  pdfcreator={LaTeX via pandoc}}
+\usepackage{url}
+\usepackage{hyperref}
+\usepackage[table]{xcolor}
+\definecolor{darkblue}{rgb}{0, 0, 0.5}
+\hypersetup{colorlinks=true,citecolor=darkblue, linkcolor=darkblue, urlcolor=darkblue}
+
+\bibliographystyle{compling}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\begin{document}
+
+%\issue{1}{1}{2016}
+%Document Head
+%\dochead{CLV3 Class File Manual}
+
+\runningtitle{Bayesian word embedding bias estimation}
+
+\runningauthor{Dobrzeniecka A. and Urbaniak, R.}
+
+%\pageonefooter{Action editor: \{action editor name\}. Submission received: DD Month YYYY; revised version received: DD Month YYYY; accepted for publication: DD Month YYYY.}
 
@@
 estimation}
-\author{Alicja Dobrzeniecka and Rafal Urbaniak}
-\date{}
-
-\begin{document}
+
+\author{Alicja Dobrzeniecka}
+
+\affil{Vrije Universiteit Amsterdam \\ The Netherlands}
+
+\author{Rafal Urbaniak%\thanks{PITC Building}
+}
+\affil{Basis Research Institute, NYC, USA  \\  The University of Gdansk,  Poland}
+
+%\author{And Another Author}
+%\affil{Publishing / SPi}
+
+%author{And Yet Another}
+%\affil{Publishing / SPi}
+
 \maketitle
 
-\textbf{Abstract.} Multiple measures, such as WEAT or MAC, attempt to
+\begin{abstract}
+Multiple measures, such as WEAT or MAC, attempt to
 quantify the magnitude of bias present in word embeddings in terms of a
@@
 levels of granularity. To showcase our method, we apply it to
-\emph{Religion, Gender}, and \emph{Race} word lists from the original
+Religion, Gender, and Race word lists from the original
 research, together with our control neutral word lists. We deploy the
-method using \emph{Google, Glove}, and \emph{Reddit} embeddings.
+method using Google, Glove, and Reddit embeddings.
 Further, we utilize our approach to evaluate a debiasing technique
-applied to the \emph{Reddit} word embedding. Our findings reveal a more
+applied to the Reddit word embedding. Our findings reveal a more
 complex landscape than suggested by the proponents of single-number
 metrics. The datasets and source code for the paper are publicly
-available.\footnote{\url{https://github.com/efemeryds/Bayesian-analysis-for-NLP-bias}}
+available.\footnote{The datasets and source code for the paper are publicly
+available at \url{https://github.com/efemeryds/Bayesian-analysis-for-NLP-bias}.}
+\end{abstract}
 
@@
 
-It has been suggested\footnote{See for instance {[}1,4,5,9,12,13{]}.}
+It has been suggested\footnote{See for instance \citep{Bolukbasi2016man, Caliskan2017semanticsBiases, Gonen2019lipstick, Lauscher2019multidimensional, Garg2018years, Manzini2019blackToCriminal}.}
 that language models can learn implicit biases that reflect harmful
@@
 downstream tasks. Roughly speaking, cosine similarity is an imperfect
-mathematical proxy for semantic similarity {[}16{]}.
+mathematical proxy for semantic similarity \citep{Mikolov2013efficient}.
 
 One response to the raising of the issue of bias in natural language
-models might be to say that there is not much point to reflecting on
+models might be to say that there is not much point in reflecting on
 such biases, as they are unavoidable. This unavoidability might seem in
 line with the arguments to the effect that learning algorithms are
-always value-laden {[}10{]}: they employ inductive methods that require
+always value-laden \citep{JohnsonValueFree}: they employ inductive methods that require
 design-, data-, or risk-related decisions that have to be guided by
@@
 Another response might be to insist that there is no bias introduced by
-the use of machine learning methods here, since the algorithm is simply
+the use of machine learning methods here since the algorithm is simply
 learning to correctly predict co-occurrences based on what ``reality''
@@
 do complete justice to learning it in the model used to study it. One
-example of this approach is {[}3{]}, where the authors use language
+example of this approach is \citep{Garg2017hundredYears}, where the authors use language
 models to study the shape of certain biases across a century.
@@
 a corpus actually greatly exaggerates the actual lower disproportion in
-the real distribution of occupations {[}6{]}.
+the real distribution of occupations \citep{gordon2012reporting}.
 
@@
 \emph{Reddit Corpus}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
-\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}9{]}, and
-\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}13{]}. We first
+\emph{Word Embedding Association Test} (\textsf{WEAT}) \citep{Caliskan2017semanticsBiases}, and
+\emph{Mean Average Cosine Distance} (\textsf{MAC}) \citep{Manzini2019blackToCriminal}. We first
 explain how these measures are supposed to work. Then we argue that they
@@
 sample sizes used in the studies are usually small,\footnote{Depending
-  on a list for {[}9{]} the range for protected words is between 13 and
-  100, and for attributes between 16 and 25; for {[}13{]} the range for
+  on a list for \citep{Caliskan2017semanticsBiases} the range for protected words is between 13 and
+  100, and for attributes between 16 and 25; for \citep{Manzini2019blackToCriminal} the range for
   protected words is between 14 and 18, and for attributes between 11
@@
 complaining'' or ``macho''.\footnote{However, for some research-related
-  purposes, such as the study of stereotypes across history {[}3{]},
+  purposes, such as the study of stereotypes across history \citep{Garg2018years},
   embeddings which do not protect certain classes may also be useful.}
@@
 
-One of the first measures of bias has been developed in {[}1{]}. The
+One of the first measures of bias has been developed in \citep{Bolukbasi2016man}. The
 general idea is that a certain topic is associated with a vector of real
@@
 investigated by considering the projection of its corresponding vector
-on this direction. For instance, in {[}1{]}, the gender direction
+on this direction. For instance, in \citep{Bolukbasi2016man}, the gender direction
 \(\mathsf{gd}\) is obtained by taking the differences of the vectors
@@
 The use of projections in bias estimation has been criticized for
-instance in {[}5{]}, where it is pointed out that while a higher average
+instance in \citep{Gonen2019lipstick}, where it is pointed out that while a higher average
 similarity to the gender direction might be an indicator of bias with
@@
 stereotypical attribute words (such as \emph{scientific} or
-\emph{caring}). Further, it is observed in {[}5{]} that most word pairs
+\emph{caring}). Further, it is observed in \citep{Gonen2019lipstick} that most word pairs
 retain similarity under debiasing meant to minimize projection-based
-bias.\footnote{In {[}1{]} another method which involves analogies and
+bias.\footnote{In \citep{Bolukbasi2016man} another method that involves analogies and
   their evaluations by human users on Mechanical Turk is also used. We
   do not discuss this method in this paper, see its criticism in
-  {[}18{]}.}
-
-A measure of bias in word embeddings which does not proceed by
+  \citep{Nissim2020fair}.}
+
+A measure of bias in word embeddings that does not proceed by
 identifying bias directions (such as a gender vector), the Word
 Embedding Association Test (\textsf{WEAT}), has been proposed in
-{[}9{]}. The idea here is that the bias between two sets of target
+\citep{Caliskan2017semanticsBiases}. The idea here is that the bias between two sets of target
 words, \(X\) and \(Y\) (we call them protected words), should be
@@
 \textsf{WEAT} is inspired by the Implicit Association Test (IAT)
-{[}19{]} used in psychology, and in some applications it uses almost the
+\citep{Nosek2002harvesting} used in psychology, and in some applications it uses almost the
 same word sets, allowing for a \emph{prima facie} sensible comparison
-with bias in humans. In {[}9{]} the authors argue that significant
+with bias in humans. In \citep{Caliskan2017semanticsBiases} the authors argue that significant
 biases---thus measured--- similar to the ones discovered by IAT can be
-discovered in word embeddings. In {[}12{]} the methodology is extended
+discovered in word embeddings. In \citep{Lauscher2019multidimensional} the methodology is extended
 to a multilingual and cross-lingual setting, arguing that using
@@
   embeddings being less biased than those based on Wikipedia.} A similar
-methodology is employed in {[}4{]}. The authors employ word embeddings
+methodology is employed in \citep{Garg2018years}. The authors employ word embeddings
 trained on corpora from different decades to study the shifts in various
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.55\linewidth]{paperDraft_files/figure-latex/WEATexample5-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.55\linewidth]{figures/WEATexample5} \end{center}
 \caption{A simple example of a two-class set-up. Two groups of protected words, $X$ and $Y$ with two stereotypical attribute sets. An example of \textsf{WEAT} calculations follows.}
@@
 to more than two stereotypical groups are to be measured. Such a
-generalization can be found in {[}13{]}. The authors introduce Mean
+generalization can be found in \citep{Manzini2019blackToCriminal}. The authors introduce Mean
 Average Cosine distance (\textsf{MAC}) as a measure of bias. Let
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.5\linewidth]{paperDraft_files/figure-latex/MACexample-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{figures/MACexample} \end{center}
 \caption{A small subset of the religion dataset.
@@
 
-Notably, the inutitive distinction between different attribute sets
+Notably, the intuitive distinction between different attribute sets
 plays no real role in the \textsf{MAC} calculations. Equally well one
@@
 predicates, mean distance of \emph{christian} to all the predicates,
-mean distance of \emph{jew} to all the predicates, and then to take the
+the mean distance of \emph{jew} to all the predicates, and then to take the
 mean of these three means.
@@
 debiasing methods to the Reddit embedding, where the score is calculated
-using the Religion word lists from {[}13{]}. For our purpose the details
+using the Religion word lists from \citep{Manzini2019blackToCriminal}. For our purpose the details
 of the debiasing method are not important: what matters is that
@@
 
-\begin{table}[H]
+\begin{table}
 \footnotesize
@@
 Thinking abstractly, 1 is the ideal distance for unrelated words. But in
-fact there is some variation in distances, which might lead to
+fact, there is some variation in distances, which might lead to
 non-biased lists also having \textsf{MAC} scores smaller than 1. How
@@
 to 1) even before debiasing. High cosine distance indicates low cosine
-similarity between values. One could think that average cosine
+similarity between values. One could think that the average cosine
 similarity equal to approximately 0.141 is not large enough to claim the
@@
 calculations even larger. The question is, on what basis is this small
-similarity still considered as a proof of the presence of bias, and
+similarity still considered as proof of the presence of bias, and
 whether these small changes are meaningful.
@@
 \textsf{MAC} to the abstractly ideal value of 1 is a methodologically
-sound idea. For this reason we also use a second list with intuitively
+sound idea. For this reason, we also use a second list with intuitively
 non-stereotypical human attributes.\footnote{See Appendix
@@
 case of religion-related stereotypes to illustrate. The full lists from
-{[}13{]} can be found in Appendix \ref{appendix:manzini_word_lists}. In
+\citep{Manzini2019blackToCriminal} can be found in Appendix \ref{appendix:manzini_word_lists}. In
 the original paper, words from all three religions were compared against
@@
 to the final \textsf{MAC} score. This is problematic, as not all of the
-stereotypical words have to be considered as harmful for all of the
+stereotypical words have to be considered harmful for all 
 religions. To avoid the masking effect, one should pay attention to how
-protected words and attributes are paired by stereotypes.
+protected words and attributes are paired with stereotypes.
 
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8d-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.7\linewidth]{figures/figcosineDistributions8d} \end{center}
 
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8df-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.7\linewidth]{figures/figcosineDistributions8df} \end{center}
 
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8dg-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.7\linewidth]{figures/figcosineDistributions8dg} \end{center}
 
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paperDraft_files/figure-latex/visMuslimd-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{figures/visMuslimd} \end{center}
 \caption{Actual distances for the protected word \textsf{muslim}. }
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paperDraft_files/figure-latex/visPriestd-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{figures/visPriestd} \end{center}
 
@@
 
-\noindent The \textsf{WEAT} scores reported by {[}9{]} for lists of
+\noindent The \textsf{WEAT} scores reported by \citep{Caliskan2017semanticsBiases} for lists of
 words for which the embeddings are supposedly biased range from 2.06 to
@@
   sample size used in the WEAT7 word list, which is not much different
-  from the other sample sizes in word lists used by {[}9{]}).} Each such
+  from the other sample sizes in word lists used by \citep{Caliskan2017semanticsBiases}).} Each such
 a collection of samples, as far as our question is involved, is
 equivalent to a sample of \(16^2\) cosine distances. Further, imagine
-there really is no difference between these groups of word and the model
+there really is no difference between these groups of words and the model
 is in fact null. That is, we draw the cosine distances from the
@@
 
-On purpose we picked a rather unusual one: the observed test statistic
+On purpose, we picked a rather unusual one: the observed test statistic
 is 0.39 and 1.27. The bootstrapped distributions of the test statistics
@@
 \(p\) values for our example are rather low (Figure
-\ref{fig:caliskanCalc}). These facts might suggested that we ended up
+\ref{fig:caliskanCalc}). These facts might suggest that we ended up
 with a situation where ``bias'' is present (albeit, due to random
@@
 list that ends up with relatively low \(p\)-value and a relatively
-unusual effect size, but nevertheless its closer inspection shows that
+unusual effect size, but nevertheless, its closer inspection shows that
 even for a word list with such properties there is no clear reason to
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanCalcB2D-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{figures/figcaliskanCalcB2D} \end{center}
 
 \caption{Bootstrapped distributions of test statistics and effect sizes in a random sample given the null hypothesis.
-We used a sample from the null  model with N(0,.08) and 16 protected words, and then boostrapped from it, following the original methodology.
-One particular boostrapped sample is highlighted, and discussed further in the text.
+We used a sample from the null  model with N(0,.08) and 16 protected words, and then bootstrapped from it, following the original methodology.
+One particular bootstrapped sample is highlighted, and discussed further in the text.
 }
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanDistC2K-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{figures/figcaliskanDistC2K} \end{center}
 
@@
 In the calculations of \textsf{WEAT} means are taken twice. The
-\textsf{s}-values themselves are means, and then means of
+\textsf{s}-values themselves are means and then means of
 \textsf{s}-values are compared between groups. Statistical troubles
@@
 
-\item When we pre-average, we disregard variation, and therefore pre-averaging  tend to manufacture false confidence.
-Group means display less variation than the raw data points, and the standard deviation of a set of means of sets of means is bound to be lower than the original standard deviation in the row data.
-Now, if you calculate your effect size by dividing by the pre-averaged standard deviation, your quite likely to get something that looks like a strong effect size, but the results of your calculations might not track anything interesting.
+\item When we pre-average, we disregard variation, and therefore pre-averaging  tends to manufacture false confidence.
+Group means display less variation than the raw data points and the standard deviation of a set of means of sets of means is bound to be lower than the original standard deviation in the row data.
+Now, if you calculate your effect size by dividing by the pre-averaged standard deviation, you are quite likely to get something that looks like a strong effect size, but the results of your calculations might not track anything interesting.
 \end{enumerate}
@@
 \(p\)-value is \(0.02\) (more than ten times higher than the
-\(p\)-valued obtained by the boostratpping procedure. So the sample is
+\(p\)-valued obtained by the bootstrapping procedure. So the sample is
 in some sense unusual. But the 95\% confidence interval for the
@@
 attributes, and \(8\) distances from hypothetical \(B\) attributes.
-Calculate the test statistic and effect size the way {[}9{]} did. Do
+Calculate the test statistic and effect size the way \citep{Caliskan2017semanticsBiases} did. Do
 this 10000 times, each time calculating \textsf{WEAT} and \textsf{s}
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2f-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{figures/figourDistancesPlot2f} \end{center}
 
@@
 underlying processes actually have the same mean. The uncertainty
-resulting from including the raw data point variance into considerations
+resulting from including the raw data point variance in considerations
 is more extensive than the one suggested by the low \(p\)-values
@@
 section we discussed the performance of the \textsf{WEAT} measure, but
-since the {[}13{]} one is a generalization thereof, including the method
+since the \citep{Manzini2019blackToCriminal} one is a generalization thereof, including the method
 of running statistical tests on pre-averaged data, our remarks,
@@
 What is the alternative? As we already emphasized: focusing on what the
-real underlying question is, and trying to answer it using a statistical
+real underlying question is and trying to answer it using a statistical
 analysis of the raw data using meaningful control groups, to ensure
@@
 capturing our uncertainty about their actual values. One important
-difference between such a result and the result of a classical
+difference between such a result and the result of classical
 statistical analysis is that classical confidence intervals (CIs) have a
@@
   regard to precision, it is a common mistake to interpret narrow
-  intervals as the ones corresponding to a more precise knowledge.
-  Another fallacy is to associate CIs with likelihood and stating that
+  intervals as the ones corresponding to more precise knowledge.
+  Another fallacy is to associate CIs with likelihood and to state that
   values within a given interval are more probable than the ones outside
@@
   interpretations. CIs should be plainly interpreted as a result of
-  certain procedure (there are many ways to obtain CIs from a given set
+  a certain procedure (there are many ways to obtain CIs from a given set
   of data) that will in the long run contain the true value if the
   procedure is performed a fixed amount of times. For a nice survey and
-  explanation of these misinterpretations, see {[}17{]}. For a
+  explanation of these misinterpretations, see \citep{Morey2015confidenceFallacy}. For a
   psychological study of the occurrence of such misinterpretations, see
-  {[}8{]}. In this study, 120 researchers and 442 students were asked to
+  \citep{Hoekstra2014Misinterpretation}. In this study, 120 researchers and 442 students were asked to
   assess the truth value of six false statements involving different
@@
 posterior distributions are easily interpretable and have direct
-relevance for the question at hand. Moreover, Bayesian data analysis is
+relevance to the question at hand. Moreover, Bayesian data analysis is
 better at handling hierarchical models and small datasets, which is
@@
 
-In a standard Bayesian analysis, the first step is to understand the
-data, think hard about the underlying process and select potential
+In standard Bayesian analysis, the first step is to understand the
+data, think hard about the underlying process, and select potential
 predictors and the outcome variable. The next step is to formulate a
@@
 
-In our analysis the outcome variable is the cosine distances between the
+In our analysis, the outcome variable is the cosine distances between the
 protected words and attribute words. The predictor is a factor
 determining whether a given attribute word is a neutral word, a human
-predicate, is stereotypically associated with the protected word, or
+predicate is stereotypically associated with the protected word, or
 comes from a different stereotype connected with another protected word.
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/dagVis-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{figures/dagVis} \end{center}
 
@@
 words to attributes associated with them.\footnote{For a thorough
-  introduction to the concepts we're using, see {[}11,15{]}.}
+  introduction to the concepts we're using, see \citep{kruschke2015bayesian,statrethinkingbook2020}.}
 \begin{align*}
@@
 illustrate a posterior predictive check for one corpus (Reddit) and one
-word list. The remaining posterior predictive checks are in section
+word list. The remaining posterior predictive checks are in Appendix 
 \ref{appendix:posterior}.
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/figposteriorPrCheckA6b-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{figures/figposteriorPrCheckA6b} \end{center}
 \caption{Example of a posterior predictive check.
@@
 Notice that 90\% of actual values fall within the 89\% HPDI and 55\% of actual values fall into 50\% HPDI, which indicates appropriate performance of the model.
-The left-right alignment of differrent colors coressponds to the fact that cosine differences between elements of different categories differ, to some extent systematically (this will be studied in the results section).
+The left-right alignment of different colors corresponds to the fact that cosine differences between elements of different categories differ, to some extent systematically (this will be studied in the results section).
 (Bottom) Densities of predicted and observed distances.}
@@
 (illustrated in the top parts of the plots). Depending on the corpus
-used and word class, there is a large variety as to the posterior
+used and word class, there is a large variety as to posterior
 densities. Quite aware of this being a crude approximation, let's
@@
   associated attributes are not systematically closer than attributes
-  belonging to different stereotypes, and the difference from neutral
+  belonging to different stereotypes, and the difference between neutral
   and human predicates is rather low, if noticeable. The situation is
@@
   and opposite attributes are much closer than neutral and human (around
-  .6 vs.~.9), but even then, there seems to be no reason to think than
+  .6 vs.~.9), but even then, there seems to be no reason to think that
   cosine distances to associated predicates are much different from
@@
   cosine similarity as compared to the theoretically expected value of
-  1, instead of running comparison to neutral and human attributes is
+  1, instead of running a comparison to neutral and human attributes is
   misleading.
@@
   (Google) and in Gender (Reddit) --- note however that in the latter,
-  for some words the opposite attributes seem to be a little bit closer
+  for some words, the opposite attributes seem to be a little bit closer
   than the associated ones.
@@
   protected words such as \emph{christian}, \emph{jew}, the proximity
-  ordering between associated and different predicates have been
+  ordering between associated and different predicates has been
   reversed, and most of the distances shifted a bit towards 1 (sometimes
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics{paperDraft_files/figure-latex/fig:debiasedCosinePair2e-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics{figures/figdebiasedCosinePair2e} \end{center}
 \caption{Mean cosine distances with 89\% highest posterior density intervals for the  gender dataset before debiasing.}
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics{paperDraft_files/figure-latex/fig:debiasedCosinePair2_e-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics{figures/figdebiasedCosinePair2_e} \end{center}
 \caption{Mean with 89\% highest posterior density intervals for gender after debiasing.}
@@
 
-In Figures \ref{fig:empiricalDebiased}-\ref{fig:empiricalDebiased3} we
+In Figures \ref{fig:empiricalDebiased1}-\ref{fig:empiricalDebiased3} we
 inspect the empirical distributions for the debiased embeddings.
 Comparing the results to the original embedding, one may notice that for
-the Religion group the neutral and human distribution has changed
-slightly. Before within the ``correct'' cosine similarity boundaries
+the Religion group, the neutral and human distribution has changed
+slightly. Before within the ``correct'' cosine similarity boundaries,
 there were 56\% of neutral and 55\% of human word lists. After the
-debiasing the values changed to 59\% (for neutral) and 59\% (for human).
+debiasing, the values changed to 59\% (for neutral) and 59\% (for human).
 The different and associated word lists were more influenced. The
@@
 43\% of the different word lists and 35\% of the associated word lists
-were within the accepted boundaries. After the embedding manipulation
+were within the accepted boundaries. After the embedding manipulation,
 the percentage has increased for both lists to 63\%. Visualization for
@@
 modification. In the Race group, the percentage within the boundaries
-for neutral and associated word lists have increased. The opposite
+for neutral and associated word lists has increased. The opposite
 happened for human and different word lists, where the percentage of
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine21c-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.95\linewidth]{figures/figdebiasedCosine21c} \end{center}
 \caption{Empirical distributions of cosine similarities before and after  debiasing  for  the Religion word list  used in  the original paper.}
-\label{fig:empiricalDebiased}
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine22c-1} \end{center}
+\label{fig:empiricalDebiased1}
+\end{figure}
+
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.95\linewidth]{figures/figdebiasedCosine22c} \end{center}
 \caption{Empirical distributions of cosine similarities before and after  debiasing for  the Gender word list used in  the original paper.}
@@
 
-\begin{figure}[H]
-
-\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine23c-1} \end{center}
+\begin{figure}
+
+\begin{center}\includegraphics[width=0.95\linewidth]{figures/figdebiasedCosine23c} \end{center}
 \caption{Empirical distributions of cosine similarities before and after  debiasing for the Race word list used in  the original paper.}
@@
 \begin{itemize}
-\item
-  {[}21{]} employ Bayesian methods to estimate uncertainty in NLP tasks,
+\item  \citet{DBLP:journals/corr/abs-1811-07253} employ Bayesian methods to estimate uncertainty in NLP tasks,
   but they apply their Bayesian Neural Networks-based method to
   sentiment analysis or named entity recognition, not to bias.
-\item
-  {[}2{]} correctly argues that a bias estimate should not be expressed
+
+\item \citet{Ethayarajh2020Bernstein} correctly argues that a bias estimate should not be expressed
   as a single number without taking into account that the estimate is
@@
   methodology and extend the lists gradually.
-\item
-  {[}20{]} criticize some existing bias metrics such as \textsf{MAC} or
+
+\item    \citet{schroder2021evaluating} criticize some existing bias metrics such as \textsf{MAC} or
   \textsf{WEAT} on the grounds of them not satisfying some general
@@
   modification, called \textsf{SAME}.
-\item
-  {[}14{]} develop a generalization of \textsf{WEAT} meant to apply to
+
+\item \citet{may-etal-2019-measuring} develop a generalization of \textsf{WEAT} meant to apply to
   sets of sentences, \textsf{SEAT}, which basically applies
@@
   single-number metric, so our remarks apply.
-\item
-  {[}7{]} introduce the Contextualized Embedding Association Test
+
+\item \citet{Guo2021CEAT} introduce the Contextualized Embedding Association Test
   Intersectional, meant to apply to dynamic word embeddings and,
@@
 
+\vspace{2mm}
+
 To summarize, a Bayesian data analysis with hierarchical models of
@@
 of the uncertainty involved. It reveals that much complexity is hidden
-when one instead chases single bias metrics present in the literature.
-
-After introducing the method, we applied it to multiple word embeddings
+when one instead chases single bias metrics present in the literature. After introducing the method, we applied it to multiple word embeddings
 and results of supposed debiasing, putting forward some general
@@
 fine-grained and multi-level evaluation of bias and debiasing in
-language models without losing modesty about the uncertainties involved.
-
-The short general and somewhat disappointing lesson here is this: things
+language models without losing modesty about the uncertainties involved. The short, general, and somewhat disappointing lesson here is this: things
 are complicated. Instead of chasing single-number metrics, we should
@@
 
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Bolukbasi2016man}{}}%
-\CSLLeftMargin{{[}1{]} }%
-\CSLRightInline{Tolga Bolukbasi, Kai-Wei Chang, James Y. Zou, Venkatesh
-Saligrama, and Adam Kalai. 2016. Man is to computer programmer as woman
-is to homemaker? Debiasing word embeddings. \emph{CoRR} abs/1607.06520,
-(2016). Retrieved from \url{http://arxiv.org/abs/1607.06520}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ethayarajh2020Bernstein}{}}%
-\CSLLeftMargin{{[}2{]} }%
-\CSLRightInline{Kawin Ethayarajh. 2020. Is your classifier actually
-biased? Measuring fairness under uncertainty with bernstein bounds.
-\emph{CoRR} abs/2004.12332, (2020). Retrieved from
-\url{https://arxiv.org/abs/2004.12332}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Garg2017hundredYears}{}}%
-\CSLLeftMargin{{[}3{]} }%
-\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
-Zou. 2017. Word embeddings quantify 100 years of gender and ethnic
-stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
-(November 2017).
-DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Garg2018years}{}}%
-\CSLLeftMargin{{[}4{]} }%
-\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
-Zou. 2018. Word embeddings quantify 100 years of gender and ethnic
-stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
-16 (April 2018), E3635--E3644.
-DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gonen2019lipstick}{}}%
-\CSLLeftMargin{{[}5{]} }%
-\CSLRightInline{Hila Gonen and Yoav Goldberg. 2019. Lipstick on a pig:
-{D}ebiasing methods cover up systematic gender biases in word embeddings
-but do not remove them. In \emph{Proceedings of the 2019 conference of
-the north {A}merican chapter of the association for computational
-linguistics: Human language technologies, volume 1 (long and short
-papers)}, Association for Computational Linguistics, Minneapolis,
-Minnesota, 609--614.
-DOI:https://doi.org/\href{https://doi.org/10.18653/v1/N19-1061}{10.18653/v1/N19-1061}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gordon2012reporting}{}}%
-\CSLLeftMargin{{[}6{]} }%
-\CSLRightInline{Jonathan Gordon and Benjamin Durme. 2013. Reporting bias
-and knowledge acquisition. In \emph{AKBC 2013 - Proceedings of the 2013
-Workshop on Automated Knowledge Base Construction, Co-located with CIKM
-2013}, 25--30.
-DOI:https://doi.org/\href{https://doi.org/10.1145/2509558.2509563}{10.1145/2509558.2509563}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Guo2021CEAT}{}}%
-\CSLLeftMargin{{[}7{]} }%
-\CSLRightInline{Wei Guo and Aylin Caliskan. 2021. Detecting emergent
-intersectional biases: Contextualized word embeddings contain a
-distribution of human-like biases. In \emph{Proceedings of the 2021
-{AAAI}/{ACM} conference on {AI}, ethics, and society}, {ACM}.
-DOI:https://doi.org/\href{https://doi.org/10.1145/3461702.3462536}{10.1145/3461702.3462536}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hoekstra2014Misinterpretation}{}}%
-\CSLLeftMargin{{[}8{]} }%
-\CSLRightInline{Rink Hoekstra, Richard D. Morey, Jeffrey N. Rouder, and
-Eric-Jan Wagenmakers. 2014. Robust misinterpretation of confidence
-intervals. \emph{Psychonomic Bulletin \& Review} 21, 5 (October 2014),
-1157--1164.
-DOI:https://doi.org/\href{https://doi.org/10.3758/s13423-013-0572-3}{10.3758/s13423-013-0572-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Caliskan2017semanticsBiases}{}}%
-\CSLLeftMargin{{[}9{]} }%
-\CSLRightInline{Aylin Caliskan Islam, Joanna J. Bryson, and Arvind
-Narayanan. 2016. Semantics derived automatically from language corpora
-necessarily contain human biases. \emph{CoRR} abs/1608.07187, (2016).
-Retrieved from \url{http://arxiv.org/abs/1608.07187}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-JohnsonValueFree}{}}%
-\CSLLeftMargin{{[}10{]} }%
-\CSLRightInline{Gabbrielle Johnson. forthcoming. Are algorithms
-value-free? Feminist theoretical virtues in machine learning.
-\emph{Journal Moral Philosophy} (forthcoming).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruschke2015bayesian}{}}%
-\CSLLeftMargin{{[}11{]} }%
-\CSLRightInline{John Kruschke. 2015. \emph{Doing bayesian data analysis
-(second edition)}. Academic Press, Boston.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lauscher2019multidimensional}{}}%
-\CSLLeftMargin{{[}12{]} }%
-\CSLRightInline{Anne Lauscher and Goran Glavas. 2019. Are we
-consistently biased? Multidimensional analysis of biases in
-distributional word vectors. \emph{CoRR} abs/1904.11783, (2019).
-Retrieved from \url{http://arxiv.org/abs/1904.11783}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Manzini2019blackToCriminal}{}}%
-\CSLLeftMargin{{[}13{]} }%
-\CSLRightInline{Thomas Manzini, Yao Chong Lim, Yulia Tsvetkov, and Alan
-W Black. 2019. Black is to criminal as caucasian is to police: Detecting
-and removing multiclass bias in word embeddings. Retrieved from
-\url{https://arxiv.org/abs/1904.04047}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-may-etal-2019-measuring}{}}%
-\CSLLeftMargin{{[}14{]} }%
-\CSLRightInline{Chandler May, Alex Wang, Shikha Bordia, Samuel R.
-Bowman, and Rachel Rudinger. 2019. On measuring social biases in
-sentence encoders. In \emph{Proceedings of the 2019 conference of the
-north {A}merican chapter of the association for computational
-linguistics: Human language technologies, volume 1 (long and short
-papers)}, Association for Computational Linguistics, Minneapolis,
-Minnesota, 622--628.
-DOI:https://doi.org/\href{https://doi.org/10.18653/v1/N19-1063}{10.18653/v1/N19-1063}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-statrethinkingbook2020}{}}%
-\CSLLeftMargin{{[}15{]} }%
-\CSLRightInline{Richard McElreath. 2020. \emph{Statistical rethinking: A
-bayesian course with examples in r and stan, 2nd edition} (2nd ed.). CRC
-Press. Retrieved from
-\url{http://xcelab.net/rm/statistical-rethinking/}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Mikolov2013efficient}{}}%
-\CSLLeftMargin{{[}16{]} }%
-\CSLRightInline{Tomas Mikolov, Kai Chen, Greg Corrado, and Jeffrey Dean.
-2013. Efficient estimation of word representations in vector space.
-DOI:https://doi.org/\href{https://doi.org/10.48550/ARXIV.1301.3781}{10.48550/ARXIV.1301.3781}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Morey2015confidenceFallacy}{}}%
-\CSLLeftMargin{{[}17{]} }%
-\CSLRightInline{Richard Morey, Rink Hoekstra, Jeffrey Rouder, Michael
-Lee, and EJ Wagenmakers. 2015. The fallacy of placing confidence in
-confidence intervals. \emph{Psychonomic Bulletin \& Review} (September
-2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nissim2020fair}{}}%
-\CSLLeftMargin{{[}18{]} }%
-\CSLRightInline{Malvina Nissim, Rik van Noord, and Rob van der Goot.
-2020. Fair is better than sensational: Man is to doctor as woman is to
-doctor. \emph{Computational Linguistics} 46, 2 (June 2020), 487--497.
-DOI:https://doi.org/\href{https://doi.org/10.1162/coli_a_00379}{10.1162/coli\_a\_00379}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nosek2002harvesting}{}}%
-\CSLLeftMargin{{[}19{]} }%
-\CSLRightInline{Brian A. Nosek, Mahzarin R. Banaji, and Anthony G.
-Greenwald. 2002. Harvesting implicit group attitudes and beliefs from a
-demonstration web site. \emph{Group Dynamics: Theory, Research, and
-Practice} 6, 1 (2002), 101--115.
-DOI:https://doi.org/\href{https://doi.org/10.1037/1089-2699.6.1.101}{10.1037/1089-2699.6.1.101}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schruxf6der2021evaluating}{}}%
-\CSLLeftMargin{{[}20{]} }%
-\CSLRightInline{Sarah Schröder, Alexander Schulz, Philip Kenneweg,
-Robert Feldhans, Fabian Hinder, and Barbara Hammer. 2021. Evaluating
-metrics for bias in word embeddings. Retrieved from
-\url{https://arxiv.org/abs/2111.07864}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-DBLP:journalsux2fcorrux2fabs-1811-07253}{}}%
-\CSLLeftMargin{{[}21{]} }%
-\CSLRightInline{Yijun Xiao and William Yang Wang. 2018. Quantifying
-uncertainties in natural language processing tasks. \emph{CoRR}
-abs/1811.07253, (2018). Retrieved from
-\url{http://arxiv.org/abs/1811.07253}}
-
-\end{CSLReferences}
-
-\newpage
+%\newpage
 \appendix
@@
 
-\begin{center}\includegraphics{paperDraft_files/figure-latex/weatjoint2z-1} \end{center}
+\begin{center}\includegraphics{figures/weatjoint2z} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/weat7glovh-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat7Reddita-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Googlea-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Glovea-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Reddita-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGooglea-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGlovea-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionReddita-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGooglea-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGlovea-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceReddita-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGooglea-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGlovea-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderReddita-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedReligionReddita-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedRaceReddita-1} \end{center}
-
-\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedGenderReddita-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{figures/weat7glovh} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsWeat7Reddita} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsWeat1Googlea} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsWeat1Glovea} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsWeat1Reddita} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsReligionGooglea} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsReligionGlovea} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsReligionReddita} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsRaceGooglea} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsRaceGlovea} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsRaceReddita} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsGenderGooglea} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsGenderGlovea} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsGenderReddita} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsDebiasedReligionReddita} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsDebiasedRaceReddita} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/resultsDebiasedGenderReddita} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.85\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixa-1} \end{center}
-
-\begin{center}\includegraphics[width=0.85\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixb-1} \end{center}
-
-\begin{center}\includegraphics[width=0.87\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixc-1} \end{center}
-
-\begin{center}\includegraphics[width=0.89\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixd-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth ]{figures/figposteriorPrCheckAppendixa} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth ]{figures/figposteriorPrCheckAppendixb} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/figposteriorPrCheckAppendixc} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{figures/figposteriorPrCheckAppendixd} \end{center}
+
+\pagebreak
 
@@
 
+\bibliography{cosineReferences}
+
 \end{document}

### 2023-09-06 paper/CLsubmission/bayesBiasCL.tex

@@
 
-One response to the raising of the issue of bias in natural language
-models might be to say that there is not much point in reflecting on
-such biases, as they are unavoidable. This unavoidability might seem in
-line with the arguments to the effect that learning algorithms are
-always value-laden \citep{JohnsonValueFree}: they employ inductive methods that require
-design-, data-, or risk-related decisions that have to be guided by
-extra-algorithmic considerations. Such choices necessarily involve value
-judgments and have to do, for instance, with what simplifications or
-risks one finds acceptable. Admittedly, algorithmic decision making
-cannot fulfill the value-free ideal, but this only means that even more
-attention needs to be paid to the values underlying different techniques
-and decisions, and to the values being pursued in a particular use of an
-algorithm.
-
-Another response might be to insist that there is no bias introduced by
-the use of machine learning methods here since the algorithm is simply
-learning to correctly predict co-occurrences based on what ``reality''
-looks like. However, this objection overlooks the fact that we, humans,
-are the ones who construct this linguistic reality, which is shaped in
-part by the natural language processing tools we use on a massive scale.
-Sure, if there is unfairness and our goal is to diagnose it, we should
-do complete justice to learning it in the model used to study it. One
-example of this approach is \citep{Garg2017hundredYears}, where the authors use language
-models to study the shape of certain biases across a century.
-
-However, if our goal is to develop downstream tools that perform tasks
-that we care about without further perpetuating or exacerbating harmful
-stereotypes, we still have good reasons to try to minimize the negative
-impact. Moreover, it is often not the case that the corpora mirror
-reality---to give a trivial example, heads are spoken of more often than
-kidneys, but this does not mean that kidneys occur much less often in
-reality than heads. To give a more relevant example, the
-disproportionate association of female words with female occupations in
-a corpus actually greatly exaggerates the actual lower disproportion in
-the real distribution of occupations \citep{gordon2012reporting}.
-
 In what follows, we focus on two popular measures of bias applicable to
@@
   does not recognize non-binary categories, and yet we use it without
-  claiming that such categories should be ignored.}
+  claiming that such categories should be ignored.}$^{, }$ \footnote{A few more philosophical comments on the enterprise of reflecting on bias in language models can be found in Appendix \ref{appendix:philosophical}.}
 
@@
 %\newpage
+
 \appendix
@@
 \label{sec:appendix}
+
+\hypertarget{philosophical}{%
+	\subsection{A philosophical commentary}}
+
+\label{appendix:philosophical}
+
+One response to the raising of the issue of bias in natural language
+	models might be to say that there is not much point in reflecting on
+	such biases, as they are unavoidable. This unavoidability might seem in
+	line with the arguments to the effect that learning algorithms are
+	always value-laden \citep{JohnsonValueFree}: they employ inductive methods that require
+	design-, data-, or risk-related decisions that have to be guided by
+	extra-algorithmic considerations. Such choices necessarily involve value
+	judgments and have to do, for instance, with what simplifications or
+	risks one finds acceptable. Admittedly, algorithmic decision making
+	cannot fulfill the value-free ideal, but this only means that even more
+	attention needs to be paid to the values underlying different techniques
+	and decisions, and to the values being pursued in a particular use of an
+	algorithm.
+
+	Another response might be to insist that there is no bias introduced by
+	the use of machine learning methods here since the algorithm is simply
+	learning to correctly predict co-occurrences based on what ``reality''
+	looks like. However, this objection overlooks the fact that we, humans,
+	are the ones who construct this linguistic reality, which is shaped in
+	part by the natural language processing tools we use on a massive scale.
+	Sure, if there is unfairness and our goal is to diagnose it, we should
+	do complete justice to learning it in the model used to study it. One
+	example of this approach is \citep{Garg2017hundredYears}, where the authors use language
+	models to study the shape of certain biases across a century.
+
+	However, if our goal is to develop downstream tools that perform tasks
+	that we care about without further perpetuating or exacerbating harmful
+	stereotypes, we still have good reasons to try to minimize the negative
+	impact. Moreover, it is often not the case that the corpora mirror
+	reality---to give a trivial example, heads are spoken of more often than
+	kidneys, but this does not mean that kidneys occur much less often in
+	reality than heads. To give a more relevant example, the
+	disproportionate association of female words with female occupations in
+	a corpus actually greatly exaggerates the actual lower disproportion in
+	the real distribution of occupations \citep{gordon2012reporting}.
 

### 2023-09-06 paper/paperDraft.Rmd

@@
 Such vectors can then be compared in terms of their similarity---the usual measure is cosine similarity---and the results of such comparisons can be used in downstream tasks.
-Roughly speaking, cosine similarity is an imperfect mathematical proxy for semantic similarity [@Mikolov2013efficient].
-
-One response to the raising of the issue of bias in natural language models might be to say that there is not much point to reflecting on such biases, as they are  unavoidable.
+Roughly speaking, cosine similarity is an imperfect mathematical proxy for semantic similarity [@Mikolov2013efficient].^[One response to the raising of the issue of bias in natural language models might be to say that there is not much point to reflecting on such biases, as they are  unavoidable.
 This unavoidability might seem in line with the arguments to the effect that learning algorithms are always value-laden [@JohnsonValueFree]: they employ inductive methods that require design-, data-, or risk-related decisions that have to be guided by extra-algorithmic considerations.
@@
 Moreover, it is often not the case that the corpora mirror reality---to give a trivial example, heads are spoken of more often than kidneys, but this does not mean that kidneys occur much less often in reality than heads.
-To give a more relevant example, the disproportionate association of female words with female occupations in a corpus actually greatly exaggerates the actual lower disproportion in the real distribution of occupations [@gordon2012reporting].
+To give a more relevant example, the disproportionate association of female words with female occupations in a corpus actually greatly exaggerates the actual lower disproportion in the real distribution of occupations [@gordon2012reporting].]
 

### 2023-09-06 paper/CLsubmission/bayesBiasCL.tex
Adds review changes

@@
 \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.}
-\emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.}
+\emph{Glove}\citet{Pennington2014Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.}
 and
-\emph{Reddit Corpus}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
+\emph{Reddit Corpus}\citet{Rabinovich2018Reddit}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
 \emph{Word Embedding Association Test} (\textsf{WEAT}) \citep{Caliskan2017semanticsBiases}, and
@@
 \end{align} \normalsize \noindent then, the association difference
-between \(A\) a \(B\) is:
+between \(A\) and \(B\) is:
 
@@
 measure the uncertainty involved and do make claims of statistical
-significance. Yet, we will later on argue that these method are not
+significance. Yet, we will later on argue that these methods are not
 proper for the goal at hand. By using Bayesian methods we will show that
@@
 comes to the investigation of the effect that debiasing has on the
-embedding space. In Figures \ref{fig:empiricalPriorToDebiasing} and
+embedding space. We used the embeddings that were debiased using \textit{hard} mode debiasing from \citet{Manzini2019blackToCriminal}. In Figures \ref{fig:empiricalPriorToDebiasing} and
 \ref{fig:empiricalDebiased} we see an example of two visualizations
@@
   this paper.
+
+\item \citet{Lum2022Debiasing} employ ..
+
+\item \citet{Ethayarajh2019understanding} points out known limitations to the use of WEAT, namely its sensitivity to the word frequency of the given data and the effect of the choice of word lists.
+As a result, WEAT systematically overestimates the bias present in the data.
+The authors propose Relational Inner Product Association (RIPA) to measure association in word embeddings.
+This method uses the scalar projection of a word vector onto a one-dimensional bias subspace (which is found by employing approach from \citet{Bolukbasi2016man}).
+RIPA method can be used for those embedding models that do implicitly matrix factorization and contain a co-occurrence statistic.
+Unlike RIPA, our Bayesian approach can be used for contextual models.
+
+\item \citet{zhang2020robustness} shows the limitations of methods that use gender word pairs to detect bias in the embeddings.
+They claim that using analogies to detect bias may not necessarily reflect societal bias, but rather the similarity between the words.
+They conduct experiments where they evaluate four popular bias mesures: Direct Bias (DB), Word Association (WA), Neighbourhood Bias Metric (NBM), and Relational Inner Product Association (RIPA).
+They show that these measures are not robust to changing either the base pair or the form of a word used.
+It is also difficult to compare measures because, although they are often presented as better alternatives, they all have similar limitations.
+
+\item \citet{Du2021Assessing} investigates the reliability of bias measures.
+They find that the bias measure can depend on the way the embedding was obtained.
+The words that are not frequent in the training data may not have a reliable vector representation that can be used for bias estimation.
+
+\item \citet{Goldfarb2021BiasNotCorrelate} conducts research to understand the relationship between intrinsic bias, which concerns the representation of embeddings, and extrinsic bias, which concerns downstream tasks.
+The authors seek to answer the question of whether intrinsic bias correlates with extrinsic bias.
+They examine the intrinsic bias metric WEAT against the extrinsic bias metrics Equality of Opportunity and Predictive Parity and conclude that the correlation is limited.
+This may further confirm that WEAT is not the most appropriate approach to account for bias in a model.
+
+\item \citet{Spliethover2021BiasSilhouette} propose Bias Silhouette Analysis (BSA), a method for assessing the quality of metrics that measure bias in word embedding models based on word lists.
+BSA quantifies how robust the word lists used are and how accurate the metrics are with respect to an unbiased and a biased reference model.
+They examine the Embedding Coherence Test (ECT), the Relative Negative Sentiment Bias (RNSB), and the Word Embedding Association Test (WEAT).
+Their conclusion is that none of these metrics can reliably discriminate between biased and non-biased models in all cases.
+
 \end{itemize}
@@
 
-\label{appendix:manzini_word_lists} The lists from {[}13{]} are
+\label{appendix:manzini_word_lists} The lists from \citet{Manzini2019blackToCriminal} are
 available here:

### 2023-09-07 paper/CLsubmission/bayesBiasCL.tex
reviewed works related

--- 2023-09-06 paper/CLsubmission/bayesBiasCL.tex
@@
 
-\item \citet{Ethayarajh2020Bernstein} correctly argues that a bias estimate should not be expressed
-  as a single number without taking into account that the estimate is
-  made using a sample of data and therefore has intrinsic uncertainty.
-  The authors suggest using Bernstein bounds to gauge the uncertainty in
-  terms of confidence intervals. We do not discuss this approach
-  extensively, as we think that confidence intervals are quite
-  problematic for several reasons, among others the confusing
-  interpretability. We do not think that Bernstein bounds provide the
-  best solution to the problem. Applying this method to a popular
-  WinoBias dataset leads to the conclusion that more than 11903 samples
-  are needed to claim a 95\% confidence interval for a bias estimate.
-  This amount vastly exceeds the existing word lists for bias
-  estimation. We propose a more realistic Bayesian method. Our
-  conclusion is still that the word lists are sometimes too small, but
-  at least they allow for gauging uncertainty as we go on to improve our
-  methodology and extend the lists gradually.
-
 \item    \citet{schroder2021evaluating} criticize some existing bias metrics such as \textsf{MAC} or
@@
   this paper.
+
+\item \citet{Lum2022Debiasing} observe that many group-wise performance meta-metrics used in algorithmic fairness consideration are biased estimators of disparities and propose a double-corrected variance estimator, which provides unbiased estimates and uncertainty quantification of the variance of model performance.
+This is certaintly valuable.
+Our approach differs in the following dimensions: it is not clear how bias variance estimator for model performance should be used in the context of group-wise cosine similarity evaluation, where we are not dealing with performance, but with cosine similarities, and where we are not dealing with a binomial distribution.
+Moreover, while having an unbiased estimate is valuable, if we have one this only means that \emph{in the long run} our estimates would tend to the true value.
+In the current situation we were looking at, the datasets are relatively small and we want to focus on what can be said before the long run has passed.
+Moreover, our Bayesian approach allows for regularization, inspection at values levels of granularity, and more straightforward interpretability, as it results in posterior distributions.
+
+\item \citet{Ethayarajh2019understanding} point out that WEAT will be overblown in degenerate cases: if the word groups are singletons, the effect size is always maximal in one direction.
+They also bring up  sensitivity to the word frequency in a given corpus and to the word list choice.
+The authors propose a measure they call Relational Inner Product Association (RIPA) to measure the association between a word vector and a relation vector in  a  word embedding.
+Roughly, this is  the scalar projection of a word vector onto a one-dimensional bias subspace (which is found by employing approach from \citet{Bolukbasi2016man}).
+They do not attempt any statistical analysis or generalization to a bias measure that would apply to an embedding space as a whole (as opposed to assigning RIPA to one word vector with respect to a relation vector), or a particular protected word list.
+In contrast, our analysis allows for a more general evaluation, with uncertainty estimates.
+Moreover, RIPA can be used only for those embedding models that implicitly employ matrix factorization and contain co-occurrence statistic.
+In contrast, such technicalities do not prevent the Bayesian approach from being deployed for contextual models.
+
+\item \citet{Ethayarajh2020Bernstein} argues that a bias estimate should not be expressed as a single number without taking into account that the estimate is
+made using a sample of data and therefore has intrinsic uncertainty.
+The author suggest using Bernstein bounds to gauge the uncertainty in
+terms of confidence intervals. We do not discuss this approach
+extensively, as we think that confidence intervals are quite
+problematic for several reasons, among others the confusing
+interpretability. We do not think that Bernstein bounds provide the
+best solution to the problem. Applying this method to a popular
+WinoBias dataset leads to the conclusion that more than 11903 data points of protected words 
+are needed to claim a 95\% confidence interval for a bias estimate.
+This amount vastly exceeds the existing word lists for bias
+estimation. We propose a more realistic Bayesian method. Our
+conclusion is still that the word lists are sometimes too small, but
+at least they allow for gauging uncertainty as we go on to improve our
+methodology and extend the lists gradually.
+
+\item \citet{zhang2020robustness} shows the limitations of methods that use gender word pairs to detect bias in an embedding space.
+They claim that using analogies to detect bias may not necessarily reflect societal bias, but rather simply co-occurrence frequency between words.
+They conduct experiments where they evaluate four popular bias mesures: Direct Bias (DB), Word Association (WA), Neighbourhood Bias Metric (NBM), and Relational Inner Product Association (RIPA).
+They show that these measures are not robust to changing either the base pair or the form of a word used.
+This is  a valid point, to some extent related to the limitations on non-contextual models, and to some extent suggesting the inclusion of various word forms in the word list, and not constructing base direction using small word lists.
+As one of our observations is that the uncertainty resulting from the use of currently existing protected word lists is too large to justify sweeping statements, this is in line with our criticism.
+
+\item \citet{Du2021Assessing} investigate the reliability of bias measures.
+They find that key existing candidates for bias measure often fail to agree with one another on particular pairs and protected words, and are sensitive to the word embedding algorithm and the corpus used.
+While some of these sensitivities are not necessarily signs of failure, we agree that the more abstract a measure is the more degrees of freedom for particular ways it is constructed, and the more ways such measures can disagree without a clear and principled reason.
+This is partially why we propose a less abstract approach that estimates expected cosine distances directly using raw data points.
+
+\item \citet{Goldfarb2021BiasNotCorrelate} further our  understanding of  the relationship between intrinsic bias measured as a property of an embedding space, and extrinsic bias, measured in terms of  downstream task performance.
+They compare  WEAT as an intrinsic measure with Equality of Opportunity and Predictive Parity as external metrics.
+The conclusion is such a  correlation is very limited.
+We consider this to be a point that matches our criticism of WEAT.
+Running a similar analysis for the Bayesian approach we discussed in this paper would be a useful task, which remains beyond the scope of this paper.
+One interesting general difference is that our method often claims that there are no sufficient  reasons to claim that an embedding is biased, so one would be on the lookout for such cases in which extrinsic bias measures nevertheless suggest the presence of bias.
+Such a presence, however, would not necessarily have to mean that the Bayesian approach to estimating potential systematic differences in cosine similarity is wrong, but rather suggest that external bias in downstream performance is not a function thereof.
+
+\item \citet{Spliethover2021BiasSilhouette} propose Bias Silhouette Analysis (BSA), a method for assessing the quality of metrics that measure bias in word embedding models based on word lists.
+The core idea here is to quantify how much the bias values of a metric vary depending on what words from the lists are actually observed, where the computations result in  values for each model obtained  using word list subsets of increasing length.
+This allows for an inspection of bias metric convergence and sensitivity to word list choice, with a biased (GloVe) and an explicitly debiased model whose lower bias has been confirmed empirically (NBatch).
+They examine the Embedding Coherence Test (ECT), the Relative Negative Sentiment Bias (RNSB), and the Word Embedding Association Test (WEAT), concluding  that none of these metrics can reliably discriminate between biased and non-biased models in all cases.
+This is not unexpected.
+An interesting question is what would happen if a similar test was applied to our method.
+However, our point is that the existing word lists are too short to provide reliable estimate of bias.
+
 \end{itemize}

### 2023-09-29 paper/CLsubmission/bayesBiasCL.tex

@@
 \emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.}
-\emph{Glove}\citet{Pennington2014Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.}
+\emph{Glove} \citep{Pennington2014Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.}
 and
-\emph{Reddit Corpus}\citet{Rabinovich2018Reddit}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
+\emph{Reddit Corpus} \citep{Rabinovich2018Reddit}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
 \emph{Word Embedding Association Test} (\textsf{WEAT}) \citep{Caliskan2017semanticsBiases}, and
@@
   combination in a way that maximizes the variance of the projections.}
-The gender bias of a word \(w\) is then understood as \(w\)'s projection
-on the gender direction: \(\vec{w} \cdot gd\) (which, after normalizing
-by dividing by \(\lVert w \rVert \,\lVert \mathsf{gd} \rVert\), is the
-same as cosine similarity). Given a list \(N\) of supposedly gender
+The gender bias of a word \(W\) is then understood as \(W\)'s projection
+on the gender direction: \(\vec{W} \cdot GD\) (which, after normalizing
+by dividing by \(\lVert W \rVert \,\lVert \mathsf{GD} \rVert\), is the
+same as cosine similarity). Given a list \(\mathbf{N}\) of supposedly gender
 neutral words,\footnote{We follow the methodology used in the debate in
@@
   discussion in Subsection \ref{subsec:interpretability}.} and the
-gender direction \(\mathsf{gd}\), the direct gender bias is defined as
-the average cosine similarity of the words in \(N\) from \(\mathsf{gd}\)
+gender direction \(\mathsf{GD}\), the direct gender bias is defined as
+the average cosine similarity of the words in \(\mathbf{N}\) from \(\mathsf{GD}\)
 (\(c\) is a parameter determining how strict we want to be):
 \begin{align*}
-\mathsf{directBias_c(N,gd)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},\mathsf{gd})\vert^c}{\vert N \vert }
+\mathsf{directBias_c(\mathbf{N},GD)} & = \frac{\sum_{W\in \mathbb{N}}\vert \mathsf{cos}(\vec{W},\mathsf{GD})\vert^c}{\vert \mathbf{N} \vert }
 \end{align*} \normalsize
@@
 curve) and CIs end up being numerically the same only if the prior
-probabilities are uniform. This illustrates that (1) classical analysis
+distributions are completely uninformative. This illustrates that (1) classical analysis
 is unable to incorporate non-trivial priors, and (2) is therefore more
@@
 \end{align*}
+
+The governing principles guiding our prior selection are:
+
+\begin{enumerate}
+	\item  We come with the same level of agnosticism about how biased a given embedding is with respect to a particular word list, and so we want to use 
+	the same priors across all word lists. Prior to the analysis there seem to be no particular reasons to have strong beliefs about the outcomes being different for different word lists. 
+	\item We want the priors to provide some level of regularization to avoid over-fitting.
+Taking coefficient priors centered around zeros is a fairly standard technique equivalent to other machine-learning regularization techniques (see Chapter 6 of \citep{james2013introduction}, especially the section on Bayesian interpretation of ridge regression and the lasso).
+Hence centering around the no-effect value (in our case, cosine distance equal to 1, or in other words, cosine similarity equal to 0) and non-uniformity.
+
+	\item We also want to build in conceptual information that we already have, such as there are mathematical restrictions on the range of cosine similarities, and therefore on the expected values.
+If we choose sd = .3 that means that prior to seeing the data we expect the mean distances in groups to be with 99.7\% prior confidence to be within $1 \pm 3sd$, that is, $(-.2, 2.2)$ which is more than plenty and does not really exclude any mathematically possible value.
+
+\end{enumerate}
+
+One useful thing about a Bayesian analysis is that the reader is welcome to re-run it with their own priors, if they find another choice more compelling, which we invite the reader to do.
 
@@
 \end{figure}
+
+\hypertarget{potential-objections}{%
+	\subsection{Potential objections}\label{potential-objections}}
+
+A worry that one might have is that the original challenge was to develop a single-number bias metric, and that our approach does not reach this goal.
+Essentially, this is correct: we argue that chasing a single metric is the wrong game to play to start with.
+In line with general Bayesian methodology, we grant epistemic priority to the inspection of posterior distributions, of which particular single-number summaries are just that.
+Imperfect summaries that can be easily misinterpreted if no attention is paid to the uncertainties surronding it.
+Having said this, if one really needs a single-number summary, then mean posterior contrasts between associated and human attributes are closer to adequacy than the single number metrics proposed so  (but they still should come accompanied with some description of the uncertainty involved).
+
+One of our recommendations is that the wordlists should be larger.
+A possible concern here is that one might ask: \emph{But wouldn't larger wordlists solve the problem with the existing measures of bias?} This is in some sense true: the overconfidence resulting from using aggregated data will be less damaging if more data is used.
+However, as \citet{Ethayarajh2020Bernstein} prove, the list sizes needed for useful uncertainty estimation from this perspective obtained using theoretical bounds tend to be prohibitive.
+Bayesian uncertainty estimates, however, scale with size and can be improved and evaluated as the list sizes increase, even if the size improvement is far from the one required by classical theoretical considerations.
 

### 2023-10-19 paper/CLsubmission/bayesBiasCL.tex
added discussion of contextual embeddings

@@
 Bayesian uncertainty estimates, however, scale with size and can be improved and evaluated as the list sizes increase, even if the size improvement is far from the one required by classical theoretical considerations.
+
+Another concern may be that this research focuses primarily on traditional uncontextualized embeddings, which may be considered outdated given recent advances in the field.
+Our focus on WEAT and similar methods is motivated by the fact that WEAT is the main bias measure that is claimed to be supported by data and findings from the Implicit Association Test in psychology, and by the fact that its analogues have seen the most research contributions.
+
+Having said that, here is a generalization of the WEAT for contextual models,  The Sentence Encoder Association Test (SEAT) \citep{may-etal-2019-measuring}.
+It works by injecting words into the context of template sentences, which are then embedded: the bias is computed from the sentence embeddings.
+A number of similar generalizations have been reviewed and studied by \citet{husse-spitz-2022-mind}.
+The authors' main observation is that the reported results are heterogeneous, inconsistent, and ultimately inconclusive, as minor design and implementation choices or errors have a substantial and often significant impact on the derived bias scores.
+This is consistent with our critique of the use of such scores.
+
+However, just as WEAT can be generalized to contextual models, so can our method.
+The fact that it has not been applied in this way is due to a practical rather than a conceptual or technical limitation.
+The results of such generalization remain to be produced and studied.
 

### 2023-10-19 codeTutorial.Rmd

---
title: "R Notebook"
output: html_notebook
---

# Goal

Preparing a Bayesian model for custom data.

# Steps

Load libraries.

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')
library("viridis") 
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(dplyr)
library(dagitty)
library(tidyverse)
library(magrittr)
library(kableExtra)
library(ggpubr)
library(ggExtra)
library(ggthemes)
library(ggforce)
library(rethinking)
library(latex2exp)
library(Hmisc)
library(grid)

```

Load datasets.

```{r cosineDistributions6d,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}

genderGoogle <- read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1]
genderReddit <-  read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1]
genderGlove <- read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1]

raceGlove <-  read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1]
raceGoogle <- read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1]
raceReddit <- read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1]

religionGlove <- read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1]
religionGoogle <- read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1]
religionReddit <- read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1]

debiasedGenderReddit <- read.csv("datasets/macWeatDatasets/debiased_gender_reddit.csv")[-1]
debiasedRaceReddit <- read.csv("datasets/macWeatDatasets/debiased_race_reddit.csv")[-1]
debiasedReligionReddit <- read.csv("datasets/macWeatDatasets/debiased_religion_reddit.csv")[-1]
```

Define a function to clean the data.

```{r}
cleanDataset <- function (dataset) {

colnames(dataset) <- c("pw","word","stereotype", "distance", "similarity", "connection")
dataset$pw <- as.factor(dataset$pw)
dataset$word <- as.factor(dataset$word)
dataset$stereotype <- as.factor(dataset$stereotype)
dataset$connection <- as.factor(dataset$connection)

ifelse(sum( dataset$similarity < -1 | dataset$distance < -1) != 0, 
       print(
         paste("WARNING:  ",  sum( dataset$similarity < -1 | dataset$distance < -1), 
               " out of ", nrow(dataset), " (", sum( dataset$similarity < -1 | dataset$distance < -1)/nrow(dataset) * 100, 
               "%) missing comparisons have been removed!", sep = "")
       ),
       print("No word removal needed.")
)

dataset <- dataset[  dataset$similarity >= -1 | dataset$distance >= -1,]

dataset$associated <- as.integer(dataset$connection == "associated")
dataset$different <- as.integer(dataset$connection == "different")
dataset$human <- as.integer(dataset$connection == "human")
dataset$none <- as.integer(dataset$connection == "none")
dataset$pwi <- as.integer(dataset$pw)
return(dataset)
}
```

Define a function to build models.

```{r}
buildModel <- function(dataset){
options(buildtools.check = function(action) TRUE )
modelResult <- ulam(
  alist(
    distance ~ dnorm(mu,sigma),
    mu <- m +  d[pwi] * different + a[pwi] * associated + h[pwi] * human + n[pw] * none,
    m ~ dnorm(1,.5),
    d[pwi] ~ dnorm(dbar, dsigmabar),
    a[pwi] ~ dnorm(abar, asigmabar),
    h[pwi] ~ dnorm(hbar, hsigmabar),
    n[pwi] ~ dnorm(nbar, nsigmabar),
    dbar ~ dnorm(0,.2),
    abar ~ dnorm(0,.2),
    hbar ~ dnorm(0,.2),
    nbar ~ dnorm(0,.2),
    dsigmabar ~ dexp(5),
    asigmabar ~ dexp(5),
    hsigmabar ~ dexp(5),
    nsigmabar ~ dexp(5),
    sigma ~ dexp(5)
  ),   data = dataset, chains=2 , iter=8000 , warmup=1000,  log_lik = TRUE, cores = 4
)
return(modelResult)
}
```

Define helper functions.

```{r}
savePrecis <- function(modelResult,name){
precisResult <- precis(modelResult, depth = 2)
saveRDS(precisResult,file = paste("resultsDFs/", name, "DF.rds", sep = ""))
}

plotFromPrecis <- function (precis, dataset){

  DFoverall <- rbind(head(precis,1),head(tail(precis, 9),4))
  DFindividual <- head(precis[-1,], -9)

  DFoverall <- DFoverall[-1,]
  DFoverall$type <- c("different", "associated", "human", "none")
  colnames(DFoverall) <- c("mean", "sd", "low", "high", "neff", "rhat", "type")

  ifelse(nrow(DFindividual)/4 != nlevels(dataset$pw),
         print("WARNING: dataset and precis don't match!"),
         print("Dataset and precis match in size"))

  DFindividual$pw <- as.factor(levels(dataset$pw)[rep((1:(nrow(DFindividual)/4)),4)])
  DFindividual$type <- as.factor(c(rep("different", nrow(DFindividual)/4),rep("associated", 
                                                                              nrow(DFindividual)/4),rep("human", nrow(DFindividual)/4), 
                                   rep("none", nrow(DFindividual)/4)))

  pwPlot <- ggplot() + geom_point(data = DFindividual, aes( x = reorder(pw, desc(pw)), y = mean, color = type),
                                  position = position_nudge((as.integer(DFindividual$type)-1)*0.15), size = 1)+
    coord_flip()+theme_tufte()+expand_limits(x= c(-1, nrow(DFindividual)/4 +1))+
    scale_color_manual(values = c("associated" ="orangered4", 
                                  "different" = "chartreuse4",
                                  "human" = "skyblue",
                                  "none" = "grey")) +xlab("protected word")+
    ylab("distance from overall baseline")+
    geom_segment(data = DFindividual,aes(x=reorder(pw, desc(pw)),xend=reorder(pw, desc(pw)), y  = `5.5%`, yend= `94.5%`,
                                         color = type),
                 position = position_nudge((as.integer(DFindividual$type)-1)*0.15), size = .4, alpha = .4)+
    geom_hline(yintercept = 0, lty = 2, size = .2, alpha = .3)+ylim(-.5,.5) +ggtitle("Protected-word-relative coefficients")+ 
    theme(legend.position='none')

  overallPlot <- ggplot() + geom_point(data = DFoverall, aes( x = type, y = mean, color = type), size = 1, position = position_dodge2(width = -1))+
    coord_flip()+theme_tufte() +  
    geom_segment(data = DFoverall,aes(x=type,xend = type, y  = low, yend= high,
                                      color = type))+ylim(-.5,.5)+
    scale_color_manual(values = c("associated" ="orangered4", 
                                  "different" = "chartreuse4",
                                  "human" = "skyblue",
                                  "none" = "grey"))+ theme(legend.position = c(0.2, 0.6))+ggtitle("Overall coefficients")+
    ylab("distance from overall baseline")+
    geom_hline(yintercept = 0, lty = 2, size = .2, alpha = .3)

  plotModel <- grid.arrange(overallPlot,pwPlot, ncol = 1)

  return(plotModel)
}

```

Clean datasets.

```{r}
genderGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1])
genderReddit <-  cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1])

raceGlove <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1])
raceGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1])
raceReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1])

religionGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1])
religionGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1])
religionReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1])
```

Build exemplary model and save it.

```{r}
modelGenderGoogle <- buildModel(genderGoogle)
saveRDS(modelGenderGoogle, file = "models/modelGenderGoogle.RDS")
```

### 2023-10-19 contextualModelsPOC.Rmd

---
title: "contextualModelsPOC"
output: html_document
date: "2023-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Applying Bayesian model for contextual language models.

## List of sentences

When using contextual models, we need to provide the model with a sentence in order to get embeddings for each token depending on a context. Therefore, to perform such a bias analysis on the model's embeddings, we would need a new list of sentences representing the bias and a list of neutral words.

## Limitations

As embedding depends on the context of a word, the use of methods that focus on specific embeddings may be limited. Such lists of phrases to be used should be prepared with care and the limitations should be known and listed in advance.

## Application

Once the lists have been created, the analysis can be carried out in the same way as for the uncontextual models.

### 2023-10-19 paper/CLsubmission/bayesBiasCL.tex

@@
 \usepackage{nicefrac}
-
+\usepackage{enumitem}
 \usepackage{graphicx}
@@
 
-It has been suggested\footnote{See for instance \citep{Bolukbasi2016man, Caliskan2017semanticsBiases, Gonen2019lipstick, Lauscher2019multidimensional, Garg2018years, Manzini2019blackToCriminal}.}
-that language models can learn implicit biases that reflect harmful
-stereotypical thinking---for example, the (vector corresponding to the)
+It has been suggested  that language models can learn implicit biases that reflect harmful
+stereotypical thinking (see for instance \citealt{Bolukbasi2016man, Caliskan2017semanticsBiases, Gonen2019lipstick, Lauscher2019multidimensional, Garg2018years, Manzini2019blackToCriminal}).
+For example, the (vector corresponding to the)
 word \textit{she} might be much closer in the vector space to the word
@@
   on a list for \citep{Caliskan2017semanticsBiases} the range for protected words is between 13 and
-  100, and for attributes between 16 and 25; for \citep{Manzini2019blackToCriminal} the range for
+%  100, and for attributes between 16 and 25; for \citep{Manzini2019blackToCriminal} the range for
   protected words is between 14 and 18, and for attributes between 11
@@
 results are, roughly, as follows. (A) Posterior density intervals are
-fairly wide and the average differences between associated, different
-and neutral human predicates, are not very large. (B) A preliminary
-inspection suggests that the desirability of changes obtained by the
+fairly wide and the average differences in cosine distances between stereotypically associated, stereotypically different,  
+random neutral, and regular human-related predicates are not very large. (B) A preliminary inspection suggests that the desirability of changes obtained by the
 usual debiasing methods is debatable.
@@
 \ref{sec:results} we elaborate on the results that it leads to,
-including a somewhat skeptical view of the efficiency of debiasing
-methods, discussed in Subsection \ref{subsec:rethinking}. Finally, in
+including low  efficiency of debiasing methods, discussed in Subsection \ref{subsec:rethinking}. Finally, in
 Section \ref{sec:related} we spend some time placing our results in the
@@
 on this direction. For instance, in \citep{Bolukbasi2016man}, the gender direction
-\(\mathsf{gd}\) is obtained by taking the differences of the vectors
+\(\mathsf{GD}\) is obtained by taking the differences of the vectors
 corresponding to ten different gendered pairs (such as
@@
 of it, and reducing the cosine similarity to such a projection may not
-be sufficient to eliminate bias. For instance, ``math'' and ``delicate''
+be sufficient to eliminate bias. For instance, \emph{math} and \emph{delicate} 
 might be equally similar to a pair of opposed explicitly gendered words
@@
 \(A_j\in \mathcal{A}\) be a set of attributes stereotypically associated
-with a protected word where \(\mathcal{A}\). For instance, when biases
+with a protected word. For instance, when biases
 related to religion are to be investigated, they use a dataset of the
@@
 
-\noindent That is, for each protected word \(t\in T\), and each
+\noindent For each protected word \(t\in T\), and each
 attribute set \(A_j\), they first take the mean of distances for this
@@
 debiasing methods to the Reddit embedding, where the score is calculated
-using the Religion word lists from \citep{Manzini2019blackToCriminal}. For our purpose the details
-of the debiasing method are not important: what matters is that
-\textsf{MAC} is used in the evaluation of these methods.
+using the Religion word lists from \citep{Manzini2019blackToCriminal}.
+The main element of the strategy is projection  onto (hard debiasing) or shifting vectors towards (soft debiasing) the bias subspace so that the biased information is no longer linearly accessible, and normalization of the vectors.\footnote{See the authors' GitHub code at \url{{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/biasOps.py}}  for details.}  For our purpose, details of the debiasing methods are not important.
+What matters is that the authors use \textsf{MAC} in the evaluation of these methods.
 
@@
 fact, there is some variation in distances, which might lead to
-non-biased lists also having \textsf{MAC} scores smaller than 1. How
-much smaller? What may attract attention is the fact that the value of
+non-biased lists also having \textsf{MAC} scores smaller than 1.
+ What may attract attention is the fact that the value of
 cosine distance in ``Biased'' category is already quite high (i.e.~close
@@
 empirical distributions, while paying attention to such divisions. The
-horizontal lines represent the values of \(1 - \mathsf{MAC}\) (that is,
-we now talk in terms of cosine similarity rather than cosine distance)
-that the authors considered indicative of bias for stereotypes
-corresponding to given word lists. For instance, in religion,
+horizontal lines represent the values of \(1 - \mathsf{MAC}\) (cosine similarity) that the authors considered indicative of bias for stereotypes corresponding to given word lists.
+For instance, in religion,
 \textsf{MAC} was .859, which was considered a sign of bias, so we plot
@@
 
-\caption{Empirical distributions of cosine similarities  for the Religion word list  used in  the original paper. }
+\caption{Empirical distributions of cosine similarities  for the Religion word list  used in  the original paper.  The
+	horizontal lines represent the values of \(1 - \mathsf{MAC}\) (cosine similarity) that the authors considered indicative of bias for stereotypes corresponding to given word lists. }
 
@@
 
-\caption{Empirical distributions of cosine similarities for the Gender word list used in  the original paper.  }
+\caption{Empirical distributions of cosine similarities for the Gender word list used in  the original paper.  The
+	horizontal lines represent the values of \(1 - \mathsf{MAC}\) (cosine similarity) that the authors considered indicative of bias for stereotypes corresponding to given word lists.}
 
@@
 
-\caption{Empirical distributions of cosine similarities  for the Race word list used in  the original paper.  }
+\caption{Empirical distributions of cosine similarities  for the Race word list used in  the original paper.  The
+	horizontal lines represent the values of \(1 - \mathsf{MAC}\) (cosine similarity) that the authors considered indicative of bias for stereotypes corresponding to given word lists. }
 
@@
 the protected word \textsf{priest}, the situation is even less as
-expected: the nearest attributes are human attributes, and all there
-seems to be no clear pattern to the distances to other attributes.
+expected: the nearest attributes are human attributes, and there
+seems to be no clear pattern when it comes to the distances between attributes.
 
@@
 calculate the \textsf{s} values and \textsf{WEAT}. What are the
-resulting distributions of \(\textsf{s}\) scores and what \(p\)-values
-do they lead to? What are the resulting effect sizes for each
-bootstrapped sample, and how often can we get an effect size as large as
-the ones reported in the original paper?
+resulting distributions of \(\textsf{s}\) scores,  what \(p\)-values and effect sizes
+do they lead to?
 
@@
 and low \textsf{WEAT}: after all, these are just random permutations of
-random distances all of which are drawn from the same null distribution.
-Let's take a look at one such a bootstrapped sample.
-
-On purpose, we picked a rather unusual one: the observed test statistic
-is 0.39 and 1.27. The bootstrapped distributions of the test statistics
-and effect sizes are illustrated in Figure \ref{fig:caliskanCalc},
-together with this particular example. Quite notably both (two-sided)
-\(p\) values for our example are rather low (Figure
-\ref{fig:caliskanCalc}). These facts might suggest that we ended up
-with a situation where ``bias'' is present (albeit, due to random
-noise). The reason why we picked it is that it is an example of a word
-list that ends up with relatively low \(p\)-value and a relatively
-unusual effect size, but nevertheless, its closer inspection shows that
-even for a word list with such properties there is no clear reason to
-think that the bias is present.
+random distances all of which are drawn from the same null distribution. Let's take a look at one such a bootstrapped sample. For the sake of illustration,
+ we picked a rather unusual one: the observed test statistic in this sample is 0.39 and 1.27.
+
+ The bootstrapped distributions of the test statistics
+and effect sizes across multiple samples are illustrated in Figure \ref{fig:caliskanCalc} (we marked the location of our particular example). 
+Notably, both (two-sided) \(p\) values for our example are rather low.
+This  might suggest that we ended up with a situation where ``bias'' is present (albeit, due to random
+noise). After all the observed statistic is unusual enough for the \(p\) values to pass the traditional significance threshold.
 
@@
 
-At this point, we might think that we just stumbled into a bootstrapped
-sample that randomly happened to display strong bias. We decide to
-double-check this by visual inspection expecting exactly this: a strong,
-clearly visible bias (Figure \ref{fig:caliskanDistances}).
+The reason why we picked the example that we did is that  while it leads to a relatively low \(p\)-value, and a relatively unusual effect size.
+Yet, a visualization of the sample reveals no interesting patterns (see Figure \ref{fig:caliskanDistances}), which strongly suggests that the way we calculated effect sizes and \(p\) values overestimates the impact of random noise, in line with our more theoretical comment, which will follow.
 
@@
 saying that a clear bias on which one group is systematically closer to
-\(A\)s than another is definitely a stretch. What happened?
-
-In the calculations of \textsf{WEAT} means are taken twice. The
-\textsf{s}-values themselves are means and then means of
-\textsf{s}-values are compared between groups. Statistical troubles
+one word group  than another lacks support.
+Crucially, in the calculations of \textsf{WEAT} means are taken twice.
+The \textsf{s}-values themselves are means and then means of \textsf{s}-values are compared between groups.
+Statistical troubles
 start when we run statistical tests on sets of means, for at least two
@@
 Let us think again about the question that we are ultimately interested
-in. Are the \(X\) terms systematically closer to (further from) the
-\(A\) attributes (\(B\) attributes) than the \(Y\) words? But now let's
-use the raw data points to try to answer these questions.
-
+in.  Are the \(X\) terms systematically closer to or further from the
+\(A\)  or \(B\) attributes than the \(Y\) words?
+But this time, instead of pre-averaging, let's use the raw data points to answer the question.
 To start with, let us run two quick \(t\)-tests to gauge what the raw
 data illustrated in Figure \ref{fig:caliskanDistances} tell us. First,
-distances to \(A\) attributes for \(X\) words and \(Y\) words. Well, the
-result is---strictly speaking---statistically significant. The
+distances to \(A\) attributes for \(X\) words and \(Y\) words. The
+result is statistically significant. The
 \(p\)-value is \(0.02\) (more than ten times higher than the
-\(p\)-valued obtained by the bootstrapping procedure. So the sample is
+\(p\)-valued obtained by the bootstrapping procedure). So the sample is
 in some sense unusual. But the 95\% confidence interval for the
-difference in means is \([.0052, .061]\), clearly nothing that a reader
-would expect given that the calculated effect size seemed quite large.
-How about the distances to the \(B\) attributes? Here the \(p\)-value is
+difference in means is \([.0052, .061]\), which suggests values that are much smaller  than what a reader  would expect having read that the calculated effect size is quite large.
+Let us inspect the distances to the \(B\) attributes.
+Here the \(p\)-value is
 \(.22\) and the 95\% confidence interval is \([-0.03, .009]\), even less
@@
 
-The difficulties are exacerbated by the fact that statistical tests are
+Another difficulty is that these statistical tests are
 based on bootstrapping from a relatively small data sets, which is quite
@@
 assumption of the null model with realistic empirically motivated raw
-data point standard deviation.
+data point standard deviation (Figure \ref{fig:ourDistances}).
 
@@
 
-What is the alternative? As we already emphasized: focusing on what the
+As an alternative, we will propose focusing on what the
 real underlying question is and trying to answer it using a statistical
@@
 since multiple evaluations are to be made, we will pursue this method
-from the Bayesian perspective. Now we have to describe it.
+from the Bayesian perspective.
 
@@
 statistical analysis is that classical confidence intervals (CIs) have a
-rather complicated and somewhat confusing interpretation, which has
-little to do with the posterior probability distribution.\footnote{Here
+rather complicated and somewhat confusing interpretation not directly related to the posterior probability distribution.\footnote{Here
   are a few usual problems. CIs are often mistakenly interpreted as
@@
 
-In fact, Bayesian highest posterior density intervals (HPDIs, the
+In fact, Bayesian highest posterior density intervals\footnote{HPDIs are  the
 narrowest intervals containing a certain ratio of the area under the
-curve) and CIs end up being numerically the same only if the prior
+curve.} and CIs end up being numerically the same only if the prior
 distributions are completely uninformative. This illustrates that (1) classical analysis
@@
 posterior predictions reflect the data with a posterior predictive
-check.
+check. This will be our evaluation of the methods performance, discussed in Subsection \ref{posterior-predictive-check}, and a more complete collection of such results is available in the appendix.
 
 In our analysis, the outcome variable is the cosine distances between the
-protected words and attribute words. The predictor is a factor
-determining whether a given attribute word is a neutral word, a human
-predicate is stereotypically associated with the protected word, or
-comes from a different stereotype connected with another protected word.
-The idea is really straightforward: if bias is present in the embedding,
+protected words and attribute words. The predictor is a factor related to an attribute word, and has  four levels:
+\begin{itemize}[noitemsep]
+	\item a given attribute word is stereotypically associated with the protected word,
+	\item it comes from a different stereotype connected with another protected word,
+
+	\item it is a neutral word,
+
+	\item it is a human-related predicate.
+
+\end{itemize} 
+\noindent The idea is that if bias is present in the embedding,
 distances to associated attribute words should be systematically lower
@@
 thought of as incorporating adaptive regularization, which avoids
-overfitting and leads to improved estimates for unbalanced datasets (and
-the datasets we need to use are unbalanced).
+overfitting and leads to improved estimates for unbalanced datasets.
 

### 2023-10-20 paper/CLsubmission/bayesBiasCL.tex
final resubmission revisions

@@
 Section \ref{sec:related} we spend some time placing our results in the
-ongoing discussions.\footnote{\textbf{Disclaimer:} throughout the paper
+ongoing discussions. \footnote{\textbf{Disclaimer:} throughout the paper
   we will be mentioning and using word lists and stereotypes we did not
@@
 The underlying intuition is that if a particular harmful stereotype is
-learned in a particular embedding, then certain groups of words will be
+learned in a given embedding, then certain groups of words will be
 systematically closer to (or further from) each other. This gives rise
@@
 biases---thus measured--- similar to the ones discovered by IAT can be
-discovered in word embeddings. In \citep{Lauscher2019multidimensional} the methodology is extended
+found in word embeddings. In \citep{Lauscher2019multidimensional} the methodology is extended
 to a multilingual and cross-lingual setting, arguing that using
@@
 the protected classes.\footnote{The authors' code is available through
-  their github repository at
+  their GitHub repository at
   \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding}.}
@@
 first explain why pre-averaging and chasing single-number metrics is a
-sub-optimal strategy.
+suboptimal strategy.
 
@@
 \textsf{WEAT}. Once a particular \(s(X,Y,A,B)\) is calculated, the
-question arises as to whether a value that high could have arisen by
+question arises whether a value that high could have arisen by
 chance. To address the question, each \(s(X,Y,A,B)\) is used in the
@@
 random distances all of which are drawn from the same null distribution. Let's take a look at one such a bootstrapped sample. For the sake of illustration,
- we picked a rather unusual one: the observed test statistic in this sample is 0.39 and 1.27.
+ we picked a rather unusual one: the observed test statistic in this sample is $0.39$ and $1.27$.
 
@@
 	\caption{Bootstrapped distributions of test statistics and effect sizes in a random sample given the null hypothesis.
-We used a sample from the null  model with N(0,.08) and 16 protected words, and then bootstrapped from it, following the original methodology.
+We used a sample from the null  model with N($0$, $0.08$) and 16 protected words, and then bootstrapped from it, following the original methodology.
 One particular bootstrapped sample is highlighted, and discussed further in the text.
@@
 posterior predictions reflect the data with a posterior predictive
-check. This will be our evaluation of the methods performance, discussed in Subsection \ref{posterior-predictive-check}, and a more complete collection of such results is available in the appendix.
+check. This will be our evaluation of the method's performance, discussed in Subsection \ref{posterior-predictive-check}, and a more complete collection of such results is available in the appendix.
 
@@
 thought of as incorporating adaptive regularization, which avoids
-overfitting and leads to improved estimates for unbalanced datasets.
+over-fitting and leads to improved estimates for unbalanced datasets.
 
@@
 
-Now the priors and the hierarchy. We assume all the \(a\) parameters
+Now let's turn to the the priors and the hierarchy. We assume all the \(a\) parameters
 come from one distribution, that is normal around a higher-level
@@
 Taking coefficient priors centered around zeros is a fairly standard technique equivalent to other machine-learning regularization techniques (see Chapter 6 of \citep{james2013introduction}, especially the section on Bayesian interpretation of ridge regression and the lasso).
-Hence centering around the no-effect value (in our case, cosine distance equal to 1, or in other words, cosine similarity equal to 0) and non-uniformity.
+Hence, centering around the no-effect value (in our case, cosine distance equal to 1, or in other words, cosine similarity equal to 0) and non-uniformity.
 
@@
 
-One useful thing about a Bayesian analysis is that the reader is welcome to re-run it with their own priors, if they find another choice more compelling, which we invite the reader to do.
+If the reader is concerned about the choice of priors, the publicly available code can be used  to re-run it with their own priors.
 
@@
 \ref{appendix:posterior}.
+The general phenomenon is that the frequency of observed values falling within the 89\% and 55\% highest posterior density predictive intervals are close to these percentages -- which illustrates that our model is not systematically incorrect.
+Another observation is that the posterior density intervals are relatively wide, which is not unexpected --- this illustrates that information about what treatment/control group a word belongs to is not very useful in predicting its distance to other words, in line with our general methodological comments.
 
@@
 
-In brief, despite one-number metrics suggesting otherwise, our Bayesian
+Despite one-number metrics suggesting otherwise, our Bayesian
 analysis reveals that insofar as the short word lists usually used in
@@
 between cosine distances between groups) tend to be rather small, with
-few exceptions. Moreover, the choice of protected words is crucial ---
+few exceptions. The choice of protected words is crucial ---
 as there is a lot of variance when it comes to the protected word-level
@@
 
-In a bit more detail, the visualizations in Appendix
-\ref{appendix:visualizations} show that the situation is more
+One example of a visualization of the results can be found in Figure \ref{fig:empiricalPriorToDebiasing}.
+First, it illustrates posterior marginal distributions at the top of the model hierarchy, relating  estimated mean cosine distances by groups to each other.
+Then, it turns to particular protected words and estimates mean cosine distances of attributes to it, also by groups.
+Visualizations for other embeddings and word lists can be found in Appendix \ref{appendix:visualizations}.
+Overall, these show that the situation is more
 complicated than merely looking at one-number summaries might suggest.
@@
 coefficients for protected words, sporadic extreme and undesirable
-changes for some protected words, usually with the side-effect of
+changes for some protected words, usually with the side effect of
 changing what happens with neutral and human predicates.
@@
 Imperfect summaries that can be easily misinterpreted if no attention is paid to the uncertainties surronding it.
-Having said this, if one really needs a single-number summary, then mean posterior contrasts between associated and human attributes are closer to adequacy than the single number metrics proposed so  (but they still should come accompanied with some description of the uncertainty involved).
-
-One of our recommendations is that the wordlists should be larger.
-A possible concern here is that one might ask: \emph{But wouldn't larger wordlists solve the problem with the existing measures of bias?} This is in some sense true: the overconfidence resulting from using aggregated data will be less damaging if more data is used.
+Having said this, if one really needs a single-number summary, then mean posterior contrasts between associated and human attributes are closer to adequacy than the single number metrics proposed so  (but they still should come accompanied by some description of the uncertainty involved).
+
+One of our recommendations is that the word lists should be larger.
+A possible concern here is that one might ask: \emph{But wouldn't larger word lists solve the problem with the existing measures of bias?} This is in some sense true: the overconfidence resulting from using aggregated data will be less damaging if more data is used.
 However, as \citet{Ethayarajh2020Bernstein} prove, the list sizes needed for useful uncertainty estimation from this perspective obtained using theoretical bounds tend to be prohibitive.
@@
 Moreover, while having an unbiased estimate is valuable, if we have one this only means that \emph{in the long run} our estimates would tend to the true value.
-In the current situation we were looking at, the datasets are relatively small and we want to focus on what can be said before the long run has passed.
+In the current situation we were looking at, the datasets are relatively small, and we want to focus on what can be said before the long run has passed.
 Moreover, our Bayesian approach allows for regularization, inspection at values levels of granularity, and more straightforward interpretability, as it results in posterior distributions.
@@
 
-\item \citet{zhang2020robustness} shows the limitations of methods that use gender word pairs to detect bias in an embedding space.
+\item \citet{zhang2020robustness} show the limitations of methods that use gender word pairs to detect bias in an embedding space.
 They claim that using analogies to detect bias may not necessarily reflect societal bias, but rather simply co-occurrence frequency between words.
@@
 
-%\newpage
+\newpage
 
@@
 	judgments and have to do, for instance, with what simplifications or
-	risks one finds acceptable. Admittedly, algorithmic decision making
+	risks one finds acceptable. Admittedly, algorithmic decision-making
 	cannot fulfill the value-free ideal, but this only means that even more
@@
 
-\pagebreak
+%\pagebreak
 

### 2023-10-20 codeTutorial.Rmd
adding code tutorial to readme

@@
 title: "R Notebook"
-output: html_notebook
+output:
+  md_document:
+    variant: markdown_github
 ---
@@
 
-```{r, setup, include=FALSE}
-knitr::opts_knit$set(root.dir = '../')
+```{r, setup, echo = TRUE, eval = FALSE}
 library("viridis") 
@@
 
-```{r cosineDistributions6d,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}
+```{r cosineDistributions6d,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE, results= 'hide'}
 
@@
 
-Define a function to clean the data.
+Clean the data.
 
@@
 
-Define a function to build models.
+Define a function to build a model.
 
@@
 
-Build exemplary model and save it.
+Build exemplary model and save it to an otherwise empty `models` folder.
 
-```{r}
+```{r, eval = FALSE, echo = TRUE}
 modelGenderGoogle <- buildModel(genderGoogle)

### 2023-10-20 paper/earlyDraft/paperDraft.Rmd
cleaning repo

--- 2023-09-06 paper/paperDraft.Rmd
@@
 ```{r, setup, include=FALSE}
-knitr::opts_knit$set(root.dir = '../')
+knitr::opts_knit$set(root.dir = '../../')
 library("viridis") 

### 2023-12-13 paper/CLsubmission/bayesBiasCL.tex
applied final reviewers' suggestions to bayesBiasCL.tex

@@
 mathematical proxy for semantic similarity \citep{Mikolov2013efficient}.
+
+Recent research, however, has criticized the existing methods for assessing bias in word embedding models.
+\citet{zhang2020robustness} highlight the limitations of word pairs in detecting bias, arguing that analogies may not necessarily capture societal bias, but instead reflect co-occurrence frequency.
+\citet{Ethayarajh2019understanding} draw attention to potential problems with the Word Embedding Association Test (WEAT) in degenerate cases, emphasizing its sensitivity to word frequency and list selection.
+Similarly, \citet{schroder2021evaluating} critique existing bias metrics such as \textsf{MAC} and \textsf{WEAT} in light of their failure to satisfy plausible general formal principles.
+\citet{Goldfarb2021BiasNotCorrelate} find only limited correlation between intrinsic bias as measured by WEAT and extrinsic metrics such as equality of opportunity and predictive parity.
+\citet{Du2021Assessing} discover inconsistencies among existing bias measures, indicating sensitivity to embedding algorithms and corpora.
+Finally, \citet{Spliethover2021BiasSilhouette} introduces Bias Silhouette Analysis (BSA) as a method for assessing the quality of metrics that measure bias in word embedding models based on word lists, and concludes that none of the metrics they evaluated can reliably discriminate between biased and non-biased models in all cases.
+\citet{Ethayarajh2020Bernstein} argue against expressing bias as a single number without considering the inherent uncertainty in sample-based estimates.
+\citet{Lum2022Debiasing} proposes a double-corrected variance estimator for unbiased estimates and uncertainty quantification of group-wise performance metrics.
+Both proposals enrich the existing measures with classical statistical uncertainty quantification.
+Our proposal is to get rid of the single-number measures and use Bayesian methods.
 
@@
   \textsf{WEAT} to vector representations of sentences. The authors,
-  however, still pre-average and play the game of finding a
+  however, still pre-average and focus on a
   single-number metric, so our remarks apply.
@@
 made using a sample of data and therefore has intrinsic uncertainty.
-The author suggest using Bernstein bounds to gauge the uncertainty in
+The author suggests using Bernstein bounds to gauge the uncertainty in
 terms of confidence intervals. We do not discuss this approach
@@
 They examine the Embedding Coherence Test (ECT), the Relative Negative Sentiment Bias (RNSB), and the Word Embedding Association Test (WEAT), concluding  that none of these metrics can reliably discriminate between biased and non-biased models in all cases.
-This is not unexpected.
+This is in line with our results.
 An interesting question is what would happen if a similar test was applied to our method.

### 2023-12-23 paper/CLsubmission/bayesBiasCL.tex
Adds info on submission dates
Updates bib

@@
+% \RequirePackage[2020-02-02]{latexrelease}
 \documentclass{clv3}
@@
 
-%\pageonefooter{Action editor: \{action editor name\}. Submission received: DD Month YYYY; revised version received: DD Month YYYY; accepted for publication: DD Month YYYY.}
+%\pageonefooter{Action editor: Rico Sennrich. Submission received: 22 June 2023; revised version received: 13 December 2023; accepted for publication: 19 December 2023.}
 
@@
 \begin{itemize}
-\item  \citet{DBLP:journals/corr/abs-1811-07253} employ Bayesian methods to estimate uncertainty in NLP tasks,
+\item  \citet{xiao2018quantifying} employ Bayesian methods to estimate uncertainty in NLP tasks,
   but they apply their Bayesian Neural Networks-based method to
